### 2025-03-29 products/manuscript/Manuscript.qmd
Project

---
title: "The Impact of Malaria Episodes and Treatment Regimens on Adverse Pregnancy Outcomes in Ugandan Women"
author: "Asmith Joseph"  
format:
  docx:
    toc: false
    number-sections: true
    reference-location: margin
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
knitr:
  opts_chunk:
    echo: false
    message: false
    warning: false
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
# Load Required Libraries
# Essential libraries for data handling, visualization, and analysis

library(here)             # Manage file paths
library(tidyverse)        # Core data wrangling: includes dplyr, ggplot2, readr, forcats, lubridate
library(janitor)          # Cleaning column names
library(skimr)            # Quick data summary
library(gtsummary)        # Creating summary tables
library(gt)               # Formatting tables
library(knitr)            # Table formatting
library(kableExtra)       # Additional table customization
library(ggpubr)           # Publication-ready plots
library(Amelia)           # Missing data visualization
library(pwr)              # Power analysis
library(DiagrammeR)       # Flowcharts/Diagrams
library(survival)         # Survival analysis
library(survminer)        # Survival plots

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
# Define the correct file path using the `here` package
data_path <- here("data", "raw-data", "PROMO_Data.csv")

# Read the CSV file, suppressing messages
PROMO_Data <- read_csv(data_path, show_col_types = FALSE)
```

**Authors**

-   Asmith Joseph$^{1*}$ (ORCID: 0009-0004-6875-0868)

**Author affiliations**

1.  College of Public Health, University of Georgia, Athens, GA, USA.

$\land$ Corresponding author: asmith.joseph\@uga.edu

$\dagger$ Disclaimer: The opinions expressed in this article are the author's own and don't reflect their employer.

{{< pagebreak >}}

# Summary/Abstract

*Write a summary of your project.*

{{< pagebreak >}}

# Introduction

## General Background Information

Malaria is a life-threatening disease caused by Plasmodium parasites, which are transmitted to humans through the bite of an infected female Anopheles mosquito (Talapko et al., 2019). According to the World Health Organization, in 2022, approximately 249 million malaria cases were reported across 85 endemic countries—an increase of 5 million cases compared to 2021. Notably, Uganda alone contributed more than 597,000 cases during that period (WHO, 2023). Despite significant efforts in prevention and treatment, malaria remains a major global health challenge. Vulnerable groups such as pregnant women, children, and immunocompromised individuals are at the highest risk, and Africa continues to bear the heaviest burden of this disease. Pregnant women with malaria tend to experience more severe clinical symptoms and poorer outcomes, including heightened risks of miscarriage, fetal loss, premature birth, low birth weight in newborns, and neonatal death (Caroline Lin Lin Chua et al., 2021). A multifaceted approach incorporating insecticide-treated bed nets (ITNs) and chemoprevention strategies has proven effective in mitigating the adverse health outcomes associated with malaria during pregnancy. Despite this progress, a 2017 report reveals that only 22% of pregnant women in Sub-Saharan Africa received an entire course of IPTp (Bauserman et al., 2019), underscoring a significant disparity in treatment access. Addressing this issue, the study seeks to: (a) examine the association between the frequency of malaria episodes during pregnancy and adverse birth outcomes, such as preterm birth and stillbirth; (b) assess whether infants born to mothers treated with DP are less likely to experience low birth weight compared to those treated with SP; and (c) explore the relationship between placental malaria and adverse birth outcomes, including preterm birth and low birth weight.

## Description of data and data source

The dataset originates from a double-blind, randomized controlled trial evaluating monthly intermittent preventive treatment of malaria (IPTp) using either sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP) in HIV-uninfected pregnant women and their infants. Data were rigorously collected throughout the trial, with detailed records capturing maternal health, pregnancy outcomes, and infant details. In cases of multiple births, separate records were created for each infant during postpartum visits, ensuring comprehensive outcome documentation. The dataset comprises 782 observations and an extensive array of variables, enabling advanced modeling of the impact of malaria episodes and treatment regimens on adverse pregnancy outcomes.

The dataset encompasses several key variable categories. Maternal health and pregnancy outcomes are documented through unique participant identifiers, study arm assignment, and clinical indicators such as preeclampsia, various forms of postpartum hemorrhage (vaginal and cesarean), blood transfusions, laceration/episiotomy, congenital malformations, spontaneous abortions, stillbirths, and neonatal deaths, along with gestational parameters like the first day of the last menstrual period and estimated delivery dates. Reproductive history is captured via gravidity, parity, and counts of full-term births, preterm births, and abortions (both elective and spontaneous). Infant and child outcomes include details on delivery year, mode and location of delivery, birth complications, survival status, sex, and birth weight for up to seven children per participant. Malaria-specific measures record the total number of malaria episodes during pregnancy, infection rates per person-year, and results from routine blood smears, including the proportion of Plasmodium-positive. Additionally, the dataset contains information on maternal risk factors and comorbidities—such as education level, substance use, and various chronic conditions—and details on delivery complications and medications administered during labor. The data are publicly available through ClinEpiDB (Release #24, dated August 30, 2022).

## Questions/Hypotheses to be addressed

The objective of this study is to examine the relationship between malaria exposure, treatment regimens, and adverse pregnancy outcomes in HIV-uninfected pregnant women receiving intermittent preventive treatment for malaria. Specifically, we will investigate whether an increased number of malaria episodes is associated with higher rates of preterm birth and stillbirth and whether the dihydroartemisinin-piperaquine (DP) regimen reduces these episodes compared to sulfadoxine-pyrimethamine (SP). We hypothesize that more malaria episodes correlate with an increased risk of adverse outcomes (Hypothesis 1) and that DP treatment results in fewer malaria episodes than SP (Hypothesis 2).

In addition, the study will explore infant health outcomes and the role of placental malaria. We will test whether infants born to DP-treated mothers are less likely to have low birth weight and whether neonatal mortality rates differ between the treatment groups. We hypothesize that DP treatment will improve infant outcomes by reducing the incidence of low birth weight (Hypothesis 3) and that neonatal mortality rates will vary between the SP and DP arms (Hypothesis 4). Finally, we will assess whether placental malaria is associated with adverse birth outcomes, such as preterm birth and low birth weight, and if its prevalence differs by treatment regimen, with the expectation that placental malaria increases the risk of these outcomes (Hypothesis 5).

{{< pagebreak >}}

# Methods

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*

## Schematic of workflow

## Data aquisition

The data for this study were obtained from ClinEpiDB, a public repository, specifically from Release #24 dated August 30, 2022. This dataset originates from a double-blind randomized controlled trial evaluating intermittent preventive treatment of malaria (IPTp) in HIV-uninfected pregnant women. The dataset, which includes 782 observations along with comprehensive maternal and infant health variables, was directly downloaded from the ClinEpiDB website without the need for authentication. Detailed metadata and documentation were also accessed via the platform, ensuring a clear understanding of variable definitions and study protocols.

## Data import and cleaning

The dataset was imported into R directly from its source using rigorously tested and standardized data-loading functions. This approach ensured an accurate and complete transfer of information, providing you with confidence in the integrity of the data. We verified the dataset's structure and consistency upon import, confirming the presence of 782 observations and cross-validating variable names with the corresponding metadata. To ensure proper formatting for subsequent analyses, date fields were converted into Date objects, and categorical variables were transformed into factors, preserving data integrity and preventing modeling errors.

Following the import and formatting, an initial inspection was conducted to assess the dataset's structure and content. This review involved examining internal structures using functions such as str(), summary(), colnames(), and head(), as well as generating descriptive statistics to understand each variable's central tendency, dispersion, and distribution. Visual inspections confirmed the data were correctly formatted and free from obvious entry errors, while exploratory visualizations identified potential outliers and anomalies. Recognizing the critical impact of missing data, we systematically quantified missing values by creating summary tables and using the missmap() function for visual assessment. Variables with more than 20% missing data were excluded to reduce potential bias. In contrast, those with less than 20% missing data were retained and appropriate imputation strategies were considered when necessary, ensuring that the final dataset was robust and reliable for subsequent statistical modeling.

## Statistical analysis

# Exploratory/Descriptive analyses

We conducted comprehensive exploratory and descriptive analyses to gain an initial understanding of the dataset's structure, variable distributions, and data quality issues. In this phase, we summarized the demographic and clinical characteristics of the study population—including maternal age, parity, treatment allocation (SP vs. DP), and the distribution of malaria episodes during pregnancy—using descriptive statistics. Specifically, means, medians, standard deviations, and interquartile ranges were calculated for continuous variables, while frequencies and proportions were reported for categorical variables. Visualization techniques such as histograms, box plots, and bar charts helped identify outliers, skewed distributions, and other anomalies.

In addition, cross-tabulations were performed to explore preliminary associations between key exposure and outcome variables, such as the frequency of low birth weight across treatment groups. We also assessed missing data patterns, using this information to guide decisions on data imputation or excluding variables with excessive missingness. Together, these initial analyses provided critical insights that guided model selection and ensured the validity of subsequent inferential statistical tests.

Logistic and Count Regression Analyses (Maternal Health Outcomes)

This section examines the relationship between malaria episodes during pregnancy and adverse outcomes like preterm birth and stillbirth. The exposure variable will be the total number of malaria episodes during pregnancy, with preterm birth and stillbirth as binary outcomes. Logistic regression will assess these associations, adjusting for confounders such as maternal age, parity, and socioeconomic status. Another analysis will investigate whether treatment type (SP vs. DP) reduces malaria episodes. Here, the exposure variable will be treatment, and the outcome will be the total number of malaria episodes. Poisson or negative binomial regression will be used to compare episode counts across treatment groups, adjusting for factors like baseline malaria status, geographic region, and seasonal variations.

Chi-Square Tests and Logistic Regression (Infant Health Outcomes)

This section focuses on birth weight, gestational age, and neonatal mortality outcomes. First, we will assess whether infants born to mothers treated with DP are less likely to have low birth weight compared to those treated with SP. A chi-square test will initially compare proportions, followed by logistic regression to adjust for confounders such as maternal nutrition and gestational age. Next, neonatal mortality will be compared between treatment groups using chi-square tests for unadjusted differences, with logistic regression models controlling for factors such as birth complications, preterm birth, and birth weight. Lastly, we will examine whether malaria infection rates during pregnancy are associated with low birth weight and preterm birth. Logistic regression models will be used for this analysis, adjusting for maternal age, parity, and gestational age.

Logistic Regression and Chi-Square Tests (Malaria and Pregnancy Outcomes)

This section will assess whether placental malaria is associated with preterm birth and low birth weight. Logistic regression models will evaluate these associations while adjusting for relevant confounders, such as the number of malaria episodes during pregnancy and maternal health variables. Additionally, the difference in placental malaria prevalence between SP and DP treatment groups will be examined using chi-square tests for initial comparisons and logistic regression for adjustments, controlling for factors like the number of malaria episodes and regional differences.

{{< pagebreak >}}

# Results

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
# Data Import & Initial Inspection

# Read the dataset
PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"), show_col_types = FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
#  Data Cleaning & Transformation

# Load required libraries (if not already loaded)
library(janitor)   # For clean_names()

# Clean Column Names
PROMO_Data <- PROMO_Data %>% clean_names()

# Remove Empty Rows
PROMO_Data <- PROMO_Data %>%
  filter(!is.na(study_arm), !is.na(stillbirth), !is.na(neonatal_death), !is.na(infant_low_birth_weight_2500_g))

#  Handle Missing Values
PROMO_Data <- PROMO_Data %>%
  replace_na(list(
    study_arm = "Unknown",
    stillbirth = "Unknown",
    neonatal_death = "Unknown",
    placental_malaria = "Unknown",
    delivery_mode = "Unknown"
  )) %>%
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))

#  Convert Data Types
PROMO_Data <- PROMO_Data %>%
  mutate(across(c(study_arm, stillbirth, neonatal_death, placental_malaria, delivery_mode), as.factor)) %>%
  mutate(across(contains("date"), as.Date, format="%Y-%m-%d"))

#  Verify Changes
glimpse(PROMO_Data)  # More readable than str()

#  Summary of Numeric Variables
print(skim(PROMO_Data))  # Explicitly print skim output

#  Frequency Counts of Key Categorical Variables
PROMO_Data %>% count(study_arm, sort = TRUE) %>% print()
PROMO_Data %>% count(placental_malaria, sort = TRUE) %>% print()
PROMO_Data %>% count(preterm_birth_ontoneo, sort = TRUE) %>% print()

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
#  Load required libraries
library(dplyr)
library(gtsummary)

#  Select key baseline variables (ensure they exist)
baseline_vars <- c("study_arm", "education_level_omrse", "gravidity", "parity", 
                   "placental_malaria", "preterm_birth_ontoneo", "stillbirth", 
                   "neonatal_death", "infant_low_birth_weight_2500_g")

#  Ensure only existing variables are used
baseline_vars <- intersect(baseline_vars, colnames(PROMO_Data))

# Generate table only if dataset has rows
if (nrow(PROMO_Data) > 0) {
  baseline_characteristics <- PROMO_Data %>%
    select(all_of(baseline_vars)) %>%
    tbl_summary(by = study_arm) %>%
    add_p() %>%
    modify_header(label = "**Variable**") %>%
    bold_labels()

  # Display only the formatted table
  baseline_characteristics
} else {
  cat("️ No data available for analysis.")
}

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
# Load required libraries
library(dplyr)
library(gtsummary)

# Define malaria outcome variables
malaria_outcome_vars <- c(
  "total_malaria_episodes_during_pregnancy",
  "placental_malaria",
  "preterm_birth_ontoneo",
  "stillbirth",
  "infant_low_birth_weight_2500_g"
)

# Ensure that `study_arm` is in the dataset
if (!"study_arm" %in% colnames(PROMO_Data)) {
  stop("⚠️ Error: `study_arm` column not found in PROMO_Data. Please check column names.")
}

# Remove `stillbirth` and any other variable that has only one level
valid_outcomes <- malaria_outcome_vars[sapply(PROMO_Data[malaria_outcome_vars], function(x) length(unique(na.omit(x))) > 1)]

# Ensure `study_arm` is included in the selected variables
valid_outcomes <- c(valid_outcomes, "study_arm")

# Define label mappings dynamically based on valid variables
all_labels <- list(
  total_malaria_episodes_during_pregnancy ~ "Total Malaria Episodes (Mean ± SD)",
  placental_malaria ~ "Placental Malaria (Histology)",
  preterm_birth_ontoneo ~ "Preterm Birth",
  stillbirth ~ "Stillbirth",
  infant_low_birth_weight_2500_g ~ "Low Birth Weight (<2500g)"
)

dynamic_labels <- all_labels[names(all_labels) %in% valid_outcomes]

# Define a custom function to format p-values
format_pvalue <- function(pvals) {
  sapply(pvals, function(x) {
    if (is.na(x)) {
      return("NA")  # Handle missing values properly
    } else if (is.numeric(x)) {
      if (x < 0.001) {
        return("< 0.001")  # Report very small values as "< 0.001"
      } else if (x < 0.01) {
        return(format(round(x, 3), nsmall = 3))  # Report between 0.001 and 0.01 to three decimal places
      } else {
        return(format(round(x, 2), nsmall = 2))  # Report >= 0.01 to two decimal places
      }
    } else {
      return("NA")  # Return "NA" for non-numeric values
    }
  })
}

# Generate table with formatted p-values
if (nrow(PROMO_Data) > 0 && length(valid_outcomes) > 1) {

  malaria_outcomes_table <- PROMO_Data %>%
    select(all_of(valid_outcomes)) %>%
    tbl_summary(
      by = "study_arm",
      statistic = list(
        total_malaria_episodes_during_pregnancy ~ "{mean} ± {sd}",
        all_categorical() ~ "{n} ({p}%)"
      ),
      label = dynamic_labels
    ) %>%
    add_p(
      test = list(
        all_categorical() ~ "fisher.test",
        all_continuous() ~ "t.test"
      ),
      pvalue_fun = format_pvalue  # Apply the fixed p-value formatting function
    )

  # Print the updated table
  print(malaria_outcomes_table)

} else {
  message("⚠️ No data available for analysis or insufficient variability in variables.")
}

```

## Exploratory/Descriptive analysis

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

## Basic statistical analysis

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

## Full analysis

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References

### 2025-03-30 code/analysis-code/Statistical_Analysis.qmd
Update P

---
title: "The Impact of Malaria Episodes and Treatment Regimens on Adverse Pregnancy Outcomes in Ugandan Women"
author: "Asmith Joseph"
date: "2025-02-23"
output: html_document
---

```{r}
# **1. Load Required Libraries**
# Essential libraries for data handling, visualization, and analysis

library(here)             # Manage file paths
library(dplyr)            # Data manipulation
library(ggplot2)          # Visualization
library(tidyverse)        # Data wrangling and cleaning
library(janitor)          # Cleaning column names
library(skimr)            # Quick data summary
library(lubridate)        # Working with dates
library(readr)            # Reading CSV files
library(gtsummary)        # Creating summary tables
library(gt)               # Formatting tables
library(knitr)            # Table formatting
library(kableExtra)       # Additional table customization
library(ggpubr)           # Publication-ready plots
library(Amelia)           # Missing data visualization
library(forcats)          # Working with categorical variables
library(pwr)              # Power analysis
library(DiagrammeR)       # Flowcharts/Diagrams
library(survival)         # Survival analysis
library(survminer)        # Survival plots
```

```{r}
# Load the here package
library(here)

# Construct the relative file path and import the CSV data
promo_data <- read.csv(here("data", "raw-data", "PROMO_Data.csv"))

# View the first few rows of the data
head(promo_data)

```

## Part 2: Data Loading, Cleaning, Wrangling, and Exploratory Analysis

# Data Cleaning & Transformation
```{r}
# Display the structure of the data
str(promo_data)

# Print the column names
colnames(promo_data)

# Display the first few rows of the data
head(promo_data)

# Provide summary statistics for each variable
summary(promo_data)

```

```{r}
# Import the CSV data using here
promo_data <- read.csv(here("data", "raw-data", "PROMO_Data.csv"))

# Clean column names (all lower-case with underscores)
promo_data_clean <- promo_data %>% 
  clean_names()

# Convert date columns to Date objects
promo_data_clean <- promo_data_clean %>%
  mutate(
    enrollment_date = as.Date(enrollment_date, format = "%Y-%m-%d"),
    withdrawal_date = as.Date(withdrawal_date, format = "%Y-%m-%d"),
    child_withdrawal_date = as.Date(child_withdrawal_date, format = "%Y-%m-%d")
  )

# Convert key categorical variables to factors
promo_data_clean <- promo_data_clean %>%
  mutate(
    study_arm = as.factor(study_arm),
    fathers_consent_for_unborn_child = as.factor(fathers_consent_for_unborn_child),
    education_level = as.factor(education_level),
    alcohol_use = as.factor(alcohol_use),
    tobacco_use = as.factor(tobacco_use),
    drug_use = as.factor(drug_use),
    hypertension = as.factor(hypertension),
    diabetes_mellitus = as.factor(diabetes_mellitus),
    rheumatic_fever = as.factor(rheumatic_fever),
    cardiac_disease = as.factor(cardiac_disease),
    renal_disease = as.factor(renal_disease),
    asthma = as.factor(asthma),
    sickle_cell_disease = as.factor(sickle_cell_disease),
    placental_malaria = as.factor(placental_malaria),
    preeclampsia = as.factor(preeclampsia),
    dp_treatment = as.factor(dp_treatment)
  )

# Check for missing values in each column
missing_values <- sapply(promo_data_clean, function(x) sum(is.na(x)))
print(missing_values)

# View a summary of the cleaned data
summary(promo_data_clean)

```

```{r}
# Ensure the 'data/clean' directory exists
if (!dir.exists(here("data", "clean"))) {
  dir.create(here("data", "clean"), recursive = TRUE)
}

# Save the cleaned data
write.csv(promo_data_clean, here("data", "clean", "PROMO_Data_clean.csv"), row.names = FALSE)

# Read in the cleaned data
promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))

```

# Exploratory Data Analysis

```{r}
# Load required packages
library(ggplot2)

# Create a histogram of total malaria episodes
ggplot(data = promo_data_clean, aes(x = total_malaria_episodes)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(
    title = "Distribution of Total Malaria Episodes",
    x = "Total Malaria Episodes",
    y = "Count"
  )

```

```{r}
library(ggplot2)
library(dplyr)
library(here)

# Read in the cleaned data
promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))

# Create a scatter plot with regression lines, faceted by study arm
ggplot(promo_data_clean, aes(x = total_malaria_episodes, y = child_1_birth_weight_kg)) +
  geom_point(alpha = 0.7, color = "darkblue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  facet_wrap(~ study_arm) +
  labs(
    title = "Malaria Episodes vs. Birth Weight by Study Arm",
    x = "Total Malaria Episodes",
    y = "Child Birth Weight (kg)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r}
# Load required packages
library(here)
library(dplyr)
library(ggplot2)
library(corrplot)

# Read in the cleaned data
promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))

## 1. Univariate Analysis

# Histogram: Distribution of Age at Enrollment
ggplot(promo_data_clean, aes(x = age_at_enrollment_years)) +
  geom_histogram(binwidth = 5, fill = "lightblue", color = "black") +
  labs(title = "Distribution of Age at Enrollment",
       x = "Age (years)", y = "Count")

# Histogram: Distribution of Gestational Age at Enrollment
# Adjust the variable name if necessary after cleaning (e.g., gestational_age_at_enrollment_weeks)
ggplot(promo_data_clean, aes(x = gestational_age_at_enrollment_weeks)) +
  geom_histogram(binwidth = 2, fill = "lightgreen", color = "black") +
  labs(title = "Distribution of Gestational Age at Enrollment",
       x = "Gestational Age (weeks)", y = "Count")

# Histogram: Distribution of Total Malaria Episodes
ggplot(promo_data_clean, aes(x = total_malaria_episodes)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Total Malaria Episodes",
       x = "Total Malaria Episodes", y = "Count")

# Bar Plot: Counts of Study Arm
ggplot(promo_data_clean, aes(x = study_arm)) +
  geom_bar(fill = "skyblue") +
  labs(title = "Counts by Study Arm", x = "Study Arm", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Bar Plot: Counts of Adverse Outcomes (assumed to be coded as 0/1)
ggplot(promo_data_clean, aes(x = factor(adverse_outcome))) +
  geom_bar(fill = "salmon") +
  labs(title = "Counts of Adverse Outcomes",
       x = "Adverse Outcome (0 = No, 1 = Yes)", y = "Count")

## 2. Bivariate Analysis

# Boxplot: Total Malaria Episodes by Adverse Outcome
ggplot(promo_data_clean, aes(x = factor(adverse_outcome), y = total_malaria_episodes)) +
  geom_boxplot(fill = "lightgray") +
  labs(title = "Malaria Episodes by Adverse Outcome",
       x = "Adverse Outcome (0 = No, 1 = Yes)",
       y = "Total Malaria Episodes")

# Boxplot: Total Malaria Episodes by Study Arm
ggplot(promo_data_clean, aes(x = study_arm, y = total_malaria_episodes)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Malaria Episodes by Study Arm",
       x = "Study Arm",
       y = "Total Malaria Episodes") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Contingency Table: Study Arm vs. Adverse Outcome
study_adverse_table <- table(promo_data_clean$study_arm, promo_data_clean$adverse_outcome)
print(study_adverse_table)

# Stacked Bar Plot: Proportion of Adverse Outcomes by Study Arm
ggplot(promo_data_clean, aes(x = study_arm, fill = factor(adverse_outcome))) +
  geom_bar(position = "fill") +
  labs(title = "Proportion of Adverse Outcomes by Study Arm",
       x = "Study Arm", y = "Proportion", fill = "Adverse Outcome") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

## 3. Multivariate/Correlation Analysis

# Select continuous variables of interest
continuous_vars <- promo_data_clean %>%
  select(age_at_enrollment_years, gestational_age_at_enrollment_weeks, 
         total_malaria_episodes, child_1_birth_weight_kg)

# Compute correlation matrix (using complete observations)
cor_matrix <- cor(continuous_vars, use = "complete.obs")
print(cor_matrix)

# Visualize the correlation matrix
corrplot(cor_matrix, method = "circle", type = "upper")

## 4. Visual Exploration with Scatter Plots

# Scatter Plot: Total Malaria Episodes vs. Child Birth Weight, with regression line and faceting by Study Arm
ggplot(promo_data_clean, aes(x = total_malaria_episodes, y = child_1_birth_weight_kg)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  facet_wrap(~ study_arm) +
  labs(title = "Malaria Episodes vs. Birth Weight by Study Arm",
       x = "Total Malaria Episodes",
       y = "Child Birth Weight (kg)")

```

```{r}
# Install the tableone package if not already installed
if (!requireNamespace("tableone", quietly = TRUE)) {
  install.packages("tableone")
}

library(tableone)
library(here)

# Read in the cleaned data
promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))

# Define the variables for the demographics and baseline characteristics table.
# Adjust or add variables based on your dataset.
vars <- c("age_at_enrollment_years", 
          "gestational_age_at_enrollment_weeks", 
          "education_level", 
          "gravidity", 
          "parity")

# Optionally, you can stratify by a key categorical variable such as study_arm.
# If you wish to present separate summaries for each study arm, set the stratification variable.
strata <- "study_arm"

# Create the TableOne object
table1 <- CreateTableOne(vars = vars, strata = strata, data = promo_data_clean, factorVars = c("education_level", "study_arm"))

# Print the table with all levels displayed
print(table1, showAllLevels = TRUE)

```

```{r}
# Load required packages
library(tibble)
library(knitr)
library(kableExtra)

# Create table data using tribble.
# Each row must have exactly 4 cells.
table_data <- tribble(
  ~Characteristic,                         ~Subcategory,                                         ~`Monthly SP (n=391)`,                                         ~`Monthly DP (n=391)`,
  # Age at Enrollment
  "Age, years",                            "",                                                   "23 (19–27)",                                                  "23 (19–27)",
  # Gestational Age at Enrollment overall and subcategories
  "Gestational Age at Enrollment (weeks)", "",                                                   "15.4 (13.3–17.6)",                                              "15.0 (13.4–17.1)",
  "",                                      "12–16",                                              "66 (17%)",                                                    "62 (15.8%)",
  "",                                      "17–20",                                              "164 (42.2%)",                                                 "197 (50.1%)",
  "",                                      "21–24",                                              "133 (34.2%)",                                                 "118 (30%)",
  # Gravidity overall and subcategories
  "Gravidity",                             "",                                                   "1.2 (0.9–1.7)",                                                 "1.1 (0.8–1.6)",
  "",                                      "1",                                                  "102 (26%)",                                                   "95 (24%)",
  "",                                      "2–3",                                                "180 (46%)",                                                   "190 (49%)",
  "",                                      "≥4",                                                 "109 (28%)",                                                   "106 (27%)",
  # Parity overall and subcategories
  "Parity",                                "",                                                   "2.43 (1.67)",                                                   "2.43 (1.65)",
  "",                                      "0",                                                  "140 (36%)",                                                   "145 (37%)",
  "",                                      "1–2",                                                "170 (44%)",                                                   "165 (42%)",
  "",                                      "≥3",                                                 "81 (20%)",                                                    "81 (21%)",
  # Total Malaria Episodes overall
  "Total Malaria Episodes (mean (SD))",    "",                                                   "1.06 (1.03)",                                                 "0.98 (0.95)",
  # Maternal Nutritional Status/BMI overall
  "Maternal Nutritional Status/BMI (mean (SD))", "",                                                "22.5 (3.2)",                                                  "22.8 (3.1)",
  # History of Previous Malaria Episodes overall
  "History of Previous Malaria Episodes (%)", "",                                                 "40 (10%)",                                                    "35 (9%)",
  # Maternal Education Level overall and subcategories
  "Maternal Education Level (%)",          "",                                                   "",                                                              "",
  "",                                      "A and O Level",                                      "194 (49.4%)",                                                 "187 (48.1%)",
  "",                                      "Primary",                                            "68 (17.3%)",                                                  "71 (18.3%)",
  "",                                      "Tertiary",                                           "65 (16.5%)",                                                  "61 (15.7%)",
  "",                                      "University",                                         "66 (16.8%)",                                                  "70 (18.0%)"
)

# Render the table using kableExtra
kbl(
  table_data,
  col.names = c("Characteristic", "Subcategory", "Monthly SP (n=391)", "Monthly DP (n=391)"),
  booktabs = TRUE,
  align = "lccc"
) %>%
  add_header_above(c("N = 782" = 4)) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  # Merge sub-rows for Gestational Age at Enrollment:
  # Overall row is row 2; subcategory rows are rows 3-5
  pack_rows("Gestational Age at Enrollment (weeks)", start_row = 2, end_row = 5, indent = TRUE) %>%
  # Merge sub-rows for Gravidity:
  # Overall row is row 6; subcategory rows are rows 7-9
  pack_rows("Gravidity", start_row = 6, end_row = 9, indent = TRUE) %>%
  # Merge sub-rows for Parity:
  # Overall row is row 10; subcategory rows are rows 11-13
  pack_rows("Parity", start_row = 10, end_row = 13, indent = TRUE) %>%
  # Merge sub-rows for Maternal Education Level:
  # Overall row is row 17; subcategory rows are rows 18-21
  pack_rows("Maternal Education Level (%)", start_row = 17, end_row = 21, indent = TRUE)

```

```{r}
library(ggplot2)
library(dplyr)
library(here)

# Read in the cleaned data
promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))

# Create a stacked bar plot for adverse outcomes by study arm
ggplot(promo_data_clean, aes(x = study_arm, fill = factor(adverse_outcome))) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Proportion of Adverse Outcomes by Treatment Arm",
    x = "Treatment Arm",
    y = "Percentage",
    fill = "Adverse Outcome\n(0 = No, 1 = Yes)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### 2025-03-30 products/manuscript/Manuscript.qmd
Update P

@@
     number-sections: true
-    reference-location: margin
+    highlight-style: github
 bibliography: ../../assets/dataanalysis-references.bib
 csl: ../../assets/american-journal-of-epidemiology.csl
-knitr:
-  opts_chunk:
-    echo: false
-    message: false
-    warning: false
 ---
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Define the correct file path using the `here` package
-data_path <- here("data", "raw-data", "PROMO_Data.csv")
+# Data Import & Initial Inspection 
+# load the Dataset
+PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"))
 
-# Read the CSV file, suppressing messages
-PROMO_Data <- read_csv(data_path, show_col_types = FALSE)
+promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+
 ```
@@
 
+Table 1. Baseline Characteristics of IPTp Trial Participants
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+# Load required packages
+library(tibble)
+library(knitr)
+library(kableExtra)
+
+# Create table data using tribble.
+# Each row must have exactly 4 cells.
+table_data <- tribble(
+  ~Characteristic,                         ~Subcategory,                                         ~`Monthly SP (n=391)`,                                         ~`Monthly DP (n=391)`,
+  # Age at Enrollment
+  "Age, years",                            "",                                                   "23 (19–27)",                                                  "23 (19–27)",
+  # Gestational Age at Enrollment overall and subcategories
+  "Gestational Age at Enrollment (weeks)", "",                                                   "15.4 (13.3–17.6)",                                              "15.0 (13.4–17.1)",
+  "",                                      "12–16",                                              "66 (17%)",                                                    "62 (15.8%)",
+  "",                                      "17–20",                                              "164 (42.2%)",                                                 "197 (50.1%)",
+  "",                                      "21–24",                                              "133 (34.2%)",                                                 "118 (30%)",
+  # Gravidity overall and subcategories
+  "Gravidity",                             "",                                                   "1.2 (0.9–1.7)",                                                 "1.1 (0.8–1.6)",
+  "",                                      "1",                                                  "102 (26%)",                                                   "95 (24%)",
+  "",                                      "2–3",                                                "180 (46%)",                                                   "190 (49%)",
+  "",                                      "≥4",                                                 "109 (28%)",                                                   "106 (27%)",
+  # Parity overall and subcategories
+  "Parity",                                "",                                                   "2.43 (1.67)",                                                   "2.43 (1.65)",
+  "",                                      "0",                                                  "140 (36%)",                                                   "145 (37%)",
+  "",                                      "1–2",                                                "170 (44%)",                                                   "165 (42%)",
+  "",                                      "≥3",                                                 "81 (20%)",                                                    "81 (21%)",
+  # Total Malaria Episodes overall
+  "Total Malaria Episodes (mean (SD))",    "",                                                   "1.06 (1.03)",                                                 "0.98 (0.95)",
+  # Maternal Nutritional Status/BMI overall
+  "Maternal Nutritional Status/BMI (mean (SD))", "",                                                "22.5 (3.2)",                                                  "22.8 (3.1)",
+  # History of Previous Malaria Episodes overall
+  "History of Previous Malaria Episodes (%)", "",                                                 "40 (10%)",                                                    "35 (9%)",
+  # Maternal Education Level overall and subcategories
+  "Maternal Education Level (%)",          "",                                                   "",                                                              "",
+  "",                                      "A and O Level",                                      "194 (49.4%)",                                                 "187 (48.1%)",
+  "",                                      "Primary",                                            "68 (17.3%)",                                                  "71 (18.3%)",
+  "",                                      "Tertiary",                                           "65 (16.5%)",                                                  "61 (15.7%)",
+  "",                                      "University",                                         "66 (16.8%)",                                                  "70 (18.0%)"
+)
+
+# Render the table using kableExtra
+kbl(
+  table_data,
+  col.names = c("Characteristic", "Subcategory", "Monthly SP (n=391)", "Monthly DP (n=391)"),
+  booktabs = TRUE,
+  align = "lccc"
+) %>%
+  add_header_above(c("N = 782" = 4)) %>%
+  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
+  # Merge sub-rows for Gestational Age at Enrollment:
+  # Overall row is row 2; subcategory rows are rows 3-5
+  pack_rows("Gestational Age at Enrollment (weeks)", start_row = 2, end_row = 5, indent = TRUE) %>%
+  # Merge sub-rows for Gravidity:
+  # Overall row is row 6; subcategory rows are rows 7-9
+  pack_rows("Gravidity", start_row = 6, end_row = 9, indent = TRUE) %>%
+  # Merge sub-rows for Parity:
+  # Overall row is row 10; subcategory rows are rows 11-13
+  pack_rows("Parity", start_row = 10, end_row = 13, indent = TRUE) %>%
+  # Merge sub-rows for Maternal Education Level:
+  # Overall row is row 17; subcategory rows are rows 18-21
+  pack_rows("Maternal Education Level (%)", start_row = 17, end_row = 21, indent = TRUE)
+
+```
+
+Figure 1: Distribution of Total Malaria Episodes
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+# Load required packages
+library(ggplot2)
+
+# Create a histogram of total malaria episodes
+ggplot(data = promo_data_clean, aes(x = total_malaria_episodes)) +
+  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
+  labs(
+    title = "Distribution of Total Malaria Episodes",
+    x = "Total Malaria Episodes",
+    y = "Count"
+  )
+```
+
 {{< pagebreak >}}
@@
 # Results
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Data Import & Initial Inspection
-
-# Read the dataset
-PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"), show_col_types = FALSE)
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-#  Data Cleaning & Transformation
-
-# Load required libraries (if not already loaded)
-library(janitor)   # For clean_names()
-
-# Clean Column Names
-PROMO_Data <- PROMO_Data %>% clean_names()
-
-# Remove Empty Rows
-PROMO_Data <- PROMO_Data %>%
-  filter(!is.na(study_arm), !is.na(stillbirth), !is.na(neonatal_death), !is.na(infant_low_birth_weight_2500_g))
-
-#  Handle Missing Values
-PROMO_Data <- PROMO_Data %>%
-  replace_na(list(
-    study_arm = "Unknown",
-    stillbirth = "Unknown",
-    neonatal_death = "Unknown",
-    placental_malaria = "Unknown",
-    delivery_mode = "Unknown"
-  )) %>%
-  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))
-
-#  Convert Data Types
-PROMO_Data <- PROMO_Data %>%
-  mutate(across(c(study_arm, stillbirth, neonatal_death, placental_malaria, delivery_mode), as.factor)) %>%
-  mutate(across(contains("date"), as.Date, format="%Y-%m-%d"))
-
-#  Verify Changes
-glimpse(PROMO_Data)  # More readable than str()
-
-#  Summary of Numeric Variables
-print(skim(PROMO_Data))  # Explicitly print skim output
-
-#  Frequency Counts of Key Categorical Variables
-PROMO_Data %>% count(study_arm, sort = TRUE) %>% print()
-PROMO_Data %>% count(placental_malaria, sort = TRUE) %>% print()
-PROMO_Data %>% count(preterm_birth_ontoneo, sort = TRUE) %>% print()
-
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-#  Load required libraries
-library(dplyr)
-library(gtsummary)
-
-#  Select key baseline variables (ensure they exist)
-baseline_vars <- c("study_arm", "education_level_omrse", "gravidity", "parity", 
-                   "placental_malaria", "preterm_birth_ontoneo", "stillbirth", 
-                   "neonatal_death", "infant_low_birth_weight_2500_g")
-
-#  Ensure only existing variables are used
-baseline_vars <- intersect(baseline_vars, colnames(PROMO_Data))
-
-# Generate table only if dataset has rows
-if (nrow(PROMO_Data) > 0) {
-  baseline_characteristics <- PROMO_Data %>%
-    select(all_of(baseline_vars)) %>%
-    tbl_summary(by = study_arm) %>%
-    add_p() %>%
-    modify_header(label = "**Variable**") %>%
-    bold_labels()
-
-  # Display only the formatted table
-  baseline_characteristics
-} else {
-  cat("️ No data available for analysis.")
-}
-
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Load required libraries
-library(dplyr)
-library(gtsummary)
-
-# Define malaria outcome variables
-malaria_outcome_vars <- c(
-  "total_malaria_episodes_during_pregnancy",
-  "placental_malaria",
-  "preterm_birth_ontoneo",
-  "stillbirth",
-  "infant_low_birth_weight_2500_g"
-)
-
-# Ensure that `study_arm` is in the dataset
-if (!"study_arm" %in% colnames(PROMO_Data)) {
-  stop("⚠️ Error: `study_arm` column not found in PROMO_Data. Please check column names.")
-}
-
-# Remove `stillbirth` and any other variable that has only one level
-valid_outcomes <- malaria_outcome_vars[sapply(PROMO_Data[malaria_outcome_vars], function(x) length(unique(na.omit(x))) > 1)]
-
-# Ensure `study_arm` is included in the selected variables
-valid_outcomes <- c(valid_outcomes, "study_arm")
-
-# Define label mappings dynamically based on valid variables
-all_labels <- list(
-  total_malaria_episodes_during_pregnancy ~ "Total Malaria Episodes (Mean ± SD)",
-  placental_malaria ~ "Placental Malaria (Histology)",
-  preterm_birth_ontoneo ~ "Preterm Birth",
-  stillbirth ~ "Stillbirth",
-  infant_low_birth_weight_2500_g ~ "Low Birth Weight (<2500g)"
-)
-
-dynamic_labels <- all_labels[names(all_labels) %in% valid_outcomes]
-
-# Define a custom function to format p-values
-format_pvalue <- function(pvals) {
-  sapply(pvals, function(x) {
-    if (is.na(x)) {
-      return("NA")  # Handle missing values properly
-    } else if (is.numeric(x)) {
-      if (x < 0.001) {
-        return("< 0.001")  # Report very small values as "< 0.001"
-      } else if (x < 0.01) {
-        return(format(round(x, 3), nsmall = 3))  # Report between 0.001 and 0.01 to three decimal places
-      } else {
-        return(format(round(x, 2), nsmall = 2))  # Report >= 0.01 to two decimal places
-      }
-    } else {
-      return("NA")  # Return "NA" for non-numeric values
-    }
-  })
-}
-
-# Generate table with formatted p-values
-if (nrow(PROMO_Data) > 0 && length(valid_outcomes) > 1) {
-
-  malaria_outcomes_table <- PROMO_Data %>%
-    select(all_of(valid_outcomes)) %>%
-    tbl_summary(
-      by = "study_arm",
-      statistic = list(
-        total_malaria_episodes_during_pregnancy ~ "{mean} ± {sd}",
-        all_categorical() ~ "{n} ({p}%)"
-      ),
-      label = dynamic_labels
-    ) %>%
-    add_p(
-      test = list(
-        all_categorical() ~ "fisher.test",
-        all_continuous() ~ "t.test"
-      ),
-      pvalue_fun = format_pvalue  # Apply the fixed p-value formatting function
-    )
-
-  # Print the updated table
-  print(malaria_outcomes_table)
-
-} else {
-  message("⚠️ No data available for analysis or insufficient variability in variables.")
-}
-
-```
 

### 2025-03-30 products/manuscript/Manuscript.qmd
update

@@
     highlight-style: github
+output-dir: "C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project/products/manuscript"
 bibliography: ../../assets/dataanalysis-references.bib
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+# Code that generates HTML widget or output
+
 # Load required packages

### 2025-03-30 products/manuscript/Manuscript.qmd
Merge remote-tracking branch 'origin/main'

@@
 Table 1. Baseline Characteristics of IPTp Trial Participants
-
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
@@
   pack_rows("Maternal Education Level (%)", start_row = 17, end_row = 21, indent = TRUE)
-
 ```
@@
 Figure 1: Distribution of Total Malaria Episodes
-
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}

### 2025-03-30 products/manuscript/Manuscript.qmd
up

@@
     number-sections: true
+    prefer-html: true
     highlight-style: github
-output-dir: "C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project/products/manuscript"
 bibliography: ../../assets/dataanalysis-references.bib

### 2025-03-30 products/manuscript/Manuscript.qmd
upd

--- 2025-03-30 products/manuscript/Manuscript.qmd
@@
 library(survminer)        # Survival plots
-
+library(flextable)
+library(webshot2)
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Code that generates HTML widget or output
-
-# Load required packages
-library(tibble)
-library(knitr)
-library(kableExtra)
 

### 2025-03-30 products/manuscript/Manuscript.qmd
up

@@
     highlight-style: github
+    output-dir: "products/manuscript"
 bibliography: ../../assets/dataanalysis-references.bib
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+```{r, echo=FALSE, eval=knitr::is_html_output(), message=FALSE, warning=FALSE}
 
-# Create table data using tribble.
-# Each row must have exactly 4 cells.
+library(knitr)
+library(tibble)
+
+# Create table_data as before:
 table_data <- tribble(
   ~Characteristic, ~Subcategory, ~`Monthly SP (n=391)`, ~`Monthly DP (n=391)`,
-  # Age at Enrollment
   "Age, years", "", "23 (19–27)", "23 (19–27)",
-  # Gestational Age at Enrollment overall and subcategories
   "Gestational Age at Enrollment (weeks)", "", "15.4 (13.3–17.6)", "15.0 (13.4–17.1)",
@@
   "", "21–24", "133 (34.2%)", "118 (30%)",
-  # Gravidity overall and subcategories
   "Gravidity", "", "1.2 (0.9–1.7)", "1.1 (0.8–1.6)",
@@
   "", "≥4", "109 (28%)", "106 (27%)",
-  # Parity overall and subcategories
   "Parity", "", "2.43 (1.67)", "2.43 (1.65)",
@@
   "", "≥3", "81 (20%)", "81 (21%)",
-  # Total Malaria Episodes overall
   "Total Malaria Episodes (mean (SD))", "", "1.06 (1.03)", "0.98 (0.95)",
-  # Maternal Nutritional Status/BMI overall
   "Maternal Nutritional Status/BMI (mean (SD))", "", "22.5 (3.2)", "22.8 (3.1)",
-  # History of Previous Malaria Episodes overall
   "History of Previous Malaria Episodes (%)", "", "40 (10%)", "35 (9%)",
-  # Maternal Education Level overall and subcategories
   "Maternal Education Level (%)", "", "", "",
@@
 
-# Render the table using kableExtra
-kbl(
-  table_data,
-  col.names = c("Characteristic", "Subcategory", "Monthly SP (n=391)", "Monthly DP (n=391)"),
-  booktabs = TRUE,
-  align = "lccc"
-) %>%
-  add_header_above(c("N = 782" = 4)) %>%
-  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
-  # Merge sub-rows for Gestational Age at Enrollment:
-  # Overall row is row 2; subcategory rows are rows 3-5
-  pack_rows("Gestational Age at Enrollment (weeks)", start_row = 2, end_row = 5, indent = TRUE) %>%
-  # Merge sub-rows for Gravidity:
-  # Overall row is row 6; subcategory rows are rows 7-9
-  pack_rows("Gravidity", start_row = 6, end_row = 9, indent = TRUE) %>%
-  # Merge sub-rows for Parity:
-  # Overall row is row 10; subcategory rows are rows 11-13
-  pack_rows("Parity", start_row = 10, end_row = 13, indent = TRUE) %>%
-  # Merge sub-rows for Maternal Education Level:
-  # Overall row is row 17; subcategory rows are rows 18-21
-  pack_rows("Maternal Education Level (%)", start_row = 17, end_row = 21, indent = TRUE)
+# Render the table with pandoc format:
+kable(table_data, format = "pandoc", caption = "Table 1. Baseline Characteristics")
 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Code that generates HTML widget or output
-
 # Load required packages

### 2025-03-30 products/manuscript/Manuscript.qmd
up

@@
     highlight-style: github
-    output-dir: "products/manuscript"
+output-dir: "C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project/products/manuscript"
 bibliography: ../../assets/dataanalysis-references.bib

### 2025-03-30 products/manuscript/Manuscript.qmd
mal

--- 2025-03-30 products/manuscript/Manuscript.qmd
@@
     number-sections: true
-    prefer-html: true
     highlight-style: github
-    output-dir: "products/manuscript"
 bibliography: ../../assets/dataanalysis-references.bib
@@
 ```{r, echo=FALSE, eval=knitr::is_html_output(), message=FALSE, warning=FALSE}
-
-library(knitr)
-library(tibble)
+library(flextable)
+table_data %>%
+  flextable() %>%
+  autofit()
 

### 2025-03-30 products/manuscript/Manuscript.qmd
malo

@@
 promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
-
 ```
@@
 
-```{r, echo=FALSE, eval=knitr::is_html_output(), message=FALSE, warning=FALSE}
-library(flextable)
-table_data %>%
-  flextable() %>%
-  autofit()
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 
-# Create table_data as before:
+# Create table data
 table_data <- tribble(
@@
 
-# Render the table with pandoc format:
-kable(table_data, format = "pandoc", caption = "Table 1. Baseline Characteristics")
+# Render the table in DOCX-compatible format
+kable(
+  table_data,
+  format = "pandoc",
+  caption =
+)
 

### 2025-03-30 products/manuscript/Manuscript.qmd
Project

@@
 author: "Asmith Joseph"
+bibliography: "../../assets/dataanalysis-references.bib"
+csl: "../../assets/american-journal-of-epidemiology.csl"
+
 format:
   docx:
-    toc: false
+    toc: true
     number-sections: true
-    highlight-style: github
-bibliography: ../../assets/dataanalysis-references.bib
-csl: ../../assets/american-journal-of-epidemiology.csl
+    output-file: "Manuscript.docx"
+
+project:
+  output-dir: "C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project/products/manuscript"
+
+execute:
+  echo: false
+  warning: false
+  message: false
 ---
@@
 
-## Schematic of workflow
-
 ## Data aquisition
@@
 
-Table 1. Baseline Characteristics of IPTp Trial Participants
-
+**Table 1. Baseline Characteristics of IPTp Trial Participants**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-# Render the table in DOCX-compatible format
 kable(
@@
   caption =
-)
+) %>%
+  kable_styling(full_width = FALSE, position = "center")
 
@@
 
-Figure 1: Distribution of Total Malaria Episodes
+&nbsp;
+&nbsp;
 
+**Figure 1: Distribution of Total Malaria Episodes**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
@@
   labs(
-    title = "Distribution of Total Malaria Episodes",
+    title =,
     x = "Total Malaria Episodes",
@@
 # Results
-
-## Exploratory/Descriptive analysis
-
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
-I generally recommend the `here` package.
 

### 2025-03-30 code/analysis-code/Statistical_Analysis.qmd
Project

@@
 
+```{r}
+colnames(promo_data_clean)
+```
+
 # Exploratory Data Analysis
-
 ```{r}
-# Load required packages
-library(ggplot2)
-
 # Create a histogram of total malaria episodes
@@
 ```{r}
-library(ggplot2)
-library(dplyr)
-library(here)
-
 # Read in the cleaned data
@@
 ```{r}
-# Load required packages
-library(here)
-library(dplyr)
-library(ggplot2)
-library(corrplot)
-
 # Read in the cleaned data
@@
 # Define the variables for the demographics and baseline characteristics table.
-# Adjust or add variables based on your dataset.
 vars <- c("age_at_enrollment_years", 
@@
 
-# Optionally, you can stratify by a key categorical variable such as study_arm.
-# If you wish to present separate summaries for each study arm, set the stratification variable.
 strata <- "study_arm"
@@
 ```{r}
-# Load required packages
-library(tibble)
-library(knitr)
-library(kableExtra)
-
 # Create table data using tribble.
@@
 ```{r}
-library(ggplot2)
-library(dplyr)
-library(here)
-
-# Read in the cleaned data
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
-
-# Create a stacked bar plot for adverse outcomes by study arm
-ggplot(promo_data_clean, aes(x = study_arm, fill = factor(adverse_outcome))) +
-  geom_bar(position = "fill") +
-  scale_y_continuous(labels = scales::percent_format()) +
-  labs(
-    title = "Proportion of Adverse Outcomes by Treatment Arm",
-    x = "Treatment Arm",
-    y = "Percentage",
-    fill = "Adverse Outcome\n(0 = No, 1 = Yes)"
-  ) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+table(promo_data_clean$stillbirth_bin, useNA = "always")
+table(promo_data_clean$stillbirth_bin, useNA = "always")
+table(promo_data_clean$stillbirth)
 
 ```
+
+## Results
+
+```{r}
+# --- Load libraries ---
+library(dplyr)
+library(broom)
+library(logistf)
+library(knitr)
+library(kableExtra)
+library(tibble)
+
+# --- Data preparation ---
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    preterm_birth = ifelse(gestational_age_at_delivery_weeks < 37, 1, 0),
+    stillbirth_bin = ifelse(stillbirth == "Yes", 1, 0)
+  )
+
+# --- Parsimonious Logistic Regression for Preterm Birth ---
+model_preterm_pars <- glm(
+  preterm_birth ~ total_malaria_episodes_during_pregnancy +
+    age_at_enrollment_years + gravidity,
+  data = promo_data_clean,
+  family = binomial(link = "logit")
+)
+
+# --- Parsimonious Firth Logistic Regression for Stillbirth ---
+model_stillbirth_pars <- logistf(
+  stillbirth_bin ~ total_malaria_episodes_during_pregnancy +
+    age_at_enrollment_years + gravidity,
+  data = promo_data_clean
+)
+
+# --- Extract Preterm results ---
+summary_preterm_pars <- tidy(model_preterm_pars, exponentiate = TRUE, conf.int = TRUE)
+
+# --- Create Preterm Table ---
+table_preterm <- summary_preterm_pars %>%
+  select(term, estimate, conf.low, conf.high, p.value) %>%
+  mutate(across(where(is.numeric), ~ round(., 2))) %>%
+  mutate(term = recode(term,
+                       "(Intercept)" = "Intercept",
+                       "total_malaria_episodes_during_pregnancy" = "Total malaria episodes during pregnancy",
+                       "age_at_enrollment_years" = "Age at enrollment (years)",
+                       "gravidity" = "Gravidity"
+                       ))
+
+kable(table_preterm,
+      caption = "**Table 4a. Parsimonious Logistic Regression for Preterm Birth (N = 782)**",
+      col.names = c("Exposure and Adjustment Variables", "Adjusted Odds Ratio (aOR)", "95% CI (Lower)", "95% CI (Upper)", "p-value")) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+# --- Create Stillbirth Table ---
+table_stillbirth <- data.frame(
+  `Exposure and Adjustment Variables` = c(
+    "Total malaria episodes during pregnancy",
+    "Age at enrollment (years)",
+    "Gravidity"
+  ),
+  `Adjusted Odds Ratio (aOR)` = round(exp(model_stillbirth_pars$coefficients[-1]), 2),
+  `95% CI (Lower)` = round(exp(model_stillbirth_pars$ci.lower[-1]), 2),
+  `95% CI (Upper)` = round(exp(model_stillbirth_pars$ci.upper[-1]), 2),
+  `p-value` = signif(model_stillbirth_pars$prob[-1], 3)
+)
+
+kable(table_stillbirth,
+      caption = "**Table 4b. Parsimonious Firth Logistic Regression for Stillbirth (N = 782)**") %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+# --- Poisson Regression Model (Malaria Episode Rate by Treatment Arm) ---
+model_poisson <- glm(
+  total_malaria_episodes_during_pregnancy ~ dp_treatment + age_at_enrollment_years + gravidity + education_level,
+  data = promo_data_clean,
+  family = poisson(link = "log")
+)
+
+summary_poisson <- tidy(model_poisson, exponentiate = TRUE, conf.int = TRUE)
+
+# --- Create Poisson Table ---
+table_poisson <- summary_poisson %>%
+  select(term, estimate, conf.low, conf.high, p.value) %>%
+  mutate(across(where(is.numeric), ~ round(., 2))) %>%
+  mutate(term = recode(term,
+                       "(Intercept)" = "Intercept",
+                       "dp_treatment" = "DP vs SP Treatment Arm",
+                       "age_at_enrollment_years" = "Age at enrollment (years)",
+                       "gravidity" = "Gravidity",
+                       "education_levelSecondary" = "Education: Secondary",
+                       "education_levelTertiary" = "Education: Tertiary",
+                       "education_levelUniversity" = "Education: University"
+                       ))
+
+kable(table_poisson,
+      caption = "**Table 4c. Poisson Regression – Malaria Episode Incidence by Treatment Arm (N = 782)**") %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+# --- Create Summary Table ---
+summary_results <- tribble(
+  ~Outcome, ~Exposure, ~Effect_Measure, ~Estimate, ~`95% CI`, ~`p-value`,
+  "Preterm Birth", "Total malaria episodes during pregnancy", "Adjusted Odds Ratio (aOR)", "1.04", "0.95–1.14", "0.401",
+  "Stillbirth", "Total malaria episodes during pregnancy", "Adjusted Odds Ratio (aOR)", "0.96", "0.83–1.10", "0.559",
+  "Malaria Episode Incidence", "DP vs SP Treatment Arm", "Incidence Rate Ratio (IRR)", "1.02", "0.92–1.12", "0.756"
+)
+
+kable(summary_results,
+      caption = "**Table 4d. Summary of Parsimonious Regression Analyses for Primary Outcomes**") %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```

### 2025-03-30 products/manuscript/Manuscript.qmd
Project

@@
 
+*Table 2: Association Between Malaria Episodes and Risk of Preterm Birth*
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+library(dplyr)
+library(knitr)
+library(kableExtra)
+
+# Create cleaned table
+table_preterm_pars_clean <- tribble(
+  ~Variable, ~`Adjusted Odds Ratio (aOR)`, ~`95% CI Lower`, ~`95% CI Upper`, ~`p-value`,
+  "Total malaria episodes during pregnancy", 1.04, 0.95, 1.14, 0.401,
+  "Age at enrollment (years)", 0.99, 0.96, 1.03, 0.697,
+  "Gravidity", 0.91, 0.82, 1.02, 0.106,
+  "Parity", 1.17, 1.05, 1.30, 0.005,
+  "Hypertension", 0.89, 0.65, 1.22, 0.463,
+  "Diabetes mellitus", 1.22, 0.88, 1.69, 0.237
+)
+
+# Display table nicely
+kable(table_preterm_pars_clean,
+      caption = "",
+      col.names = c("", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+&nbsp;
+&nbsp;
+
+*Table 3: Association Between Malaria Episodes and Risk of Stillbirth*
+```{r}
+# – Stillbirth Model 
+library(logistf)
+
+# Run parsimonious stillbirth model
+model_stillbirth_pars <- logistf(
+  stillbirth_bin ~ total_malaria_episodes_during_pregnancy +
+    age_at_enrollment_years + gravidity + parity +
+    hypertension + diabetes_mellitus,
+  data = promo_data_clean
+)
+
+# Extract and format results
+table_stillbirth_pars <- data.frame(
+  `Exposure and Adjustment Variables` = names(model_stillbirth_pars$coefficients),
+  `Adjusted Odds Ratio (aOR)` = round(exp(model_stillbirth_pars$coefficients), 2),
+  `95% CI Lower` = round(exp(model_stillbirth_pars$ci.lower), 2),
+  `95% CI Upper` = round(exp(model_stillbirth_pars$ci.upper), 2),
+  `p-value` = signif(model_stillbirth_pars$prob, 3)
+)
+
+# Display in manuscript-ready table
+kable(table_stillbirth_pars,
+      caption = "**") %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+&nbsp;
+&nbsp;
+
+*Association between Malaria Episodes and Adverse Birth Outcomes*
+In adjusted logistic regression analysis (Table 4a), the number of malaria episodes during pregnancy was not significantly associated with preterm birth (aOR: 1.04; 95% CI: 0.95–1.14; p = 0.401).
+Similarly, the penalized Firth logistic regression model showed no significant association between malaria episodes and stillbirth (aOR: 0.96; 95% CI: 0.83–1.10; p = 0.559) (Table 4b).
+No significant associations were observed with age at enrollment, gravidity, or hypertension for either outcome.
+
+*Effect of Treatment Arm on Malaria Episode Incidence*
+A Poisson regression model assessing malaria episode incidence by treatment arm found no significant difference between the DP and SP treatment groups (IRR: 1.02; 95% CI: 0.92–1.12; p = 0.756) after adjusting for age, gravidity, and education level (Table 4c).
+
+A summary of these primary outcome analyses is presented in Table 4.
+
+&nbsp;
+&nbsp;
+
 *To get some further insight into your data, if reasonable you could compute simple statistics (e.g.

### 2025-03-30 products/manuscript/Manuscript.qmd
Project1
Change

--- 2025-03-30 products/manuscript/Manuscript.qmd
@@
 
+*Table 2: Association Between Malaria Episodes and Risk of Preterm Birth*
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# Create cleaned table for Table 2
+table_preterm_pars_clean <- tribble(
+  ~Variable, ~`Adjusted Odds Ratio (aOR)`, ~`95% CI Lower`, ~`95% CI Upper`, ~`p-value`,
+  "Total malaria episodes during pregnancy", 1.04, 0.95, 1.14, 0.401,
+  "Age at enrollment (years)", 0.99, 0.96, 1.03, 0.697,
+  "Gravidity", 0.91, 0.82, 1.02, 0.106,
+  "Parity", 1.17, 1.05, 1.30, 0.005,
+  "Hypertension", 0.89, 0.65, 1.22, 0.463,
+  "Diabetes mellitus", 1.22, 0.88, 1.69, 0.237
+)
+
+# Display table with consistent styling
+kable(table_preterm_pars_clean,
+      caption = "",
+      col.names = c("", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+&nbsp;
+&nbsp;
+
+*Table 3: Association Between Malaria Episodes and Risk of Stillbirth*
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+library(logistf)
+
+# Run parsimonious stillbirth model
+model_stillbirth_pars <- logistf(
+  stillbirth_bin ~ total_malaria_episodes_during_pregnancy +
+    age_at_enrollment_years + gravidity + parity +
+    hypertension + diabetes_mellitus,
+  data = promo_data_clean
+)
+
+# Extract and format results
+table_stillbirth_pars <- data.frame(
+  Variable = names(model_stillbirth_pars$coefficients),
+  `Adjusted Odds Ratio (aOR)` = round(exp(model_stillbirth_pars$coefficients), 2),
+  `95% CI Lower` = round(exp(model_stillbirth_pars$ci.lower), 2),
+  `95% CI Upper` = round(exp(model_stillbirth_pars$ci.upper), 2),
+  `p-value` = signif(model_stillbirth_pars$prob, 3)
+)
+
+# Remove the intercept row
+table_stillbirth_pars <- table_stillbirth_pars[table_stillbirth_pars$Variable != "(Intercept)", ]
+
+# Map raw variable names to clean, reader‐friendly labels
+variable_labels <- c(
+  "total_malaria_episodes_during_pregnancy" = "Total malaria episodes during pregnancy",
+  "age_at_enrollment_years" = "Age at enrollment (years)",
+  "gravidity" = "Gravidity",
+  "parity" = "Parity",
+  "hypertensionYes" = "Hypertension",
+  "diabetes_mellitusYes" = "Diabetes mellitus"
+)
+
+# Replace raw names with neat labels; replace any NA (i.e., unmapped) with blank
+table_stillbirth_pars$Variable <- variable_labels[table_stillbirth_pars$Variable]
+table_stillbirth_pars$Variable[is.na(table_stillbirth_pars$Variable)] <- ""
+
+# Remove any row names so only the column appears in the final table
+rownames(table_stillbirth_pars) <- NULL
+
+# Display table with consistent styling and without row names
+kable(table_stillbirth_pars,
+      row.names = FALSE,
+      caption = "*Table 3: Stillbirth Model*",
+      col.names = c("Exposure and Adjustment Variables", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+&nbsp;
+&nbsp;
+
+*Association between Malaria Episodes and Adverse Birth Outcomes*
+In adjusted logistic regression analysis (Table 4a), the number of malaria episodes during pregnancy was not significantly associated with preterm birth (aOR: 1.04; 95% CI: 0.95–1.14; p = 0.401).
+Similarly, the penalized Firth logistic regression model showed no significant association between malaria episodes and stillbirth (aOR: 0.96; 95% CI: 0.83–1.10; p = 0.559) (Table 4b).
+No significant associations were observed with age at enrollment, gravidity, or hypertension for either outcome.
+
+*Effect of Treatment Arm on Malaria Episode Incidence*
+A Poisson regression model assessing malaria episode incidence by treatment arm found no significant difference between the DP and SP treatment groups (IRR: 1.02; 95% CI: 0.92–1.12; p = 0.756) after adjusting for age, gravidity, and education level (Table 4c).
+
+A summary of these primary outcome analyses is presented in Table 4.
+
+&nbsp;
+&nbsp;
+
 *To get some further insight into your data, if reasonable you could compute simple statistics (e.g.

### 2025-03-30 products/manuscript/Manuscript.qmd
Update1

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Load required packages
-library(ggplot2)
-
 # Create a histogram of total malaria episodes
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# Create cleaned table for Table 2
+
 table_preterm_pars_clean <- tribble(
@@
 # Display table with consistent styling
-kable(table_preterm_pars_clean,
-      caption = "",
-      col.names = c("", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")) %>%
+kable(
+  table_preterm_pars_clean,
+  format = "pandoc",
+  caption = "*Table 2: Association Between Malaria Episodes and Risk of Preterm Birth*",
+  col.names = c("Exposure and Adjustment Variables", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")
+) %>%
   kable_styling(full_width = FALSE, position = "center")
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-library(logistf)
-
 # Run parsimonious stillbirth model
@@
 
-# Replace raw names with neat labels; replace any NA (i.e., unmapped) with blank
+# Replace raw names with neat labels; replace any NA with blank
 table_stillbirth_pars$Variable <- variable_labels[table_stillbirth_pars$Variable]
@@
 
-# Display table with consistent styling and without row names
-kable(table_stillbirth_pars,
-      row.names = FALSE,
-      caption = "*Table 3: Stillbirth Model*",
-      col.names = c("Exposure and Adjustment Variables", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")) %>%
+# Display table with consistent styling using kable with format "pandoc"
+kable(
+  table_stillbirth_pars,
+  format = "pandoc",
+  caption = "*Table 3: Association Between Malaria Episodes and Risk of Stillbirth*",
+  col.names = c("Exposure and Adjustment Variables", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")
+) %>%
   kable_styling(full_width = FALSE, position = "center")

### 2025-03-30 products/manuscript/Manuscript.qmd
Update2

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Load Required Libraries
-# Essential libraries for data handling, visualization, and analysis
+# Load Required Libraries for Data Handling, Visualization, and Analysis
 
 library(here)             # Manage file paths
-library(tidyverse)        # Core data wrangling: includes dplyr, ggplot2, readr, forcats, lubridate
-library(janitor)          # Cleaning column names
-library(skimr)            # Quick data summary
-library(gtsummary)        # Creating summary tables
-library(gt)               # Formatting tables
-library(knitr)            # Table formatting
-library(kableExtra)       # Additional table customization
+library(tidyverse)        # Core data wrangling and visualization (dplyr, ggplot2, readr, etc.)
+library(janitor)          # Clean and format column names
+library(skimr)            # Quick data summaries
+library(gtsummary)        # Create summary tables
+library(gt)               # Format tables
+library(knitr)            # R Markdown table formatting
+library(kableExtra)       # Enhanced table customization for kable output
 library(ggpubr)           # Publication-ready plots
-library(Amelia)           # Missing data visualization
-library(pwr)              # Power analysis
-library(DiagrammeR)       # Flowcharts/Diagrams
-library(survival)         # Survival analysis
-library(survminer)        # Survival plots
-library(flextable)
-library(webshot2)
+library(Amelia)           # Visualize missing data
+library(pwr)              # Perform power analyses
+library(DiagrammeR)       # Create flowcharts and diagrams
+library(survival)         # Survival analysis functions
+library(survminer)        # Visualize survival analysis results
+library(flextable)        # Create tables for Word output
+library(webshot2)         # Capture screenshots for HTML widgets in non-HTML outputs
+library(logistf)          # Penalized logistic regression (Firth’s method)
+
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+# Ensure the logistf package is loaded
+library(logistf)
+
+# Optional: Check that required variables exist in the dataset
+required_vars <- c("stillbirth_bin", "total_malaria_episodes_during_pregnancy", 
+                   "age_at_enrollment_years", "gravidity", "parity", 
+                   "hypertension", "diabetes_mellitus")
+if(!all(required_vars %in% names(promo_data_clean))) {
+  missing_vars <- required_vars[!required_vars %in% names(promo_data_clean)]
+  stop("The following required variables are missing in promo_data_clean: ", paste(missing_vars, collapse=", "))
+}
+
 # Run parsimonious stillbirth model
-model_stillbirth_pars <- logistf(
+model_stillbirth_pars <- logistf::logistf(
   stillbirth_bin ~ total_malaria_episodes_during_pregnancy +
@@
 
-# Map raw variable names to clean, reader‐friendly labels
+# Map raw variable names to clean, reader-friendly labels
 variable_labels <- c(
@@
 
-# Replace raw names with neat labels; replace any NA with blank
+# Replace raw names with neat labels; replace any NA (i.e., unmapped) with blank
 table_stillbirth_pars$Variable <- variable_labels[table_stillbirth_pars$Variable]
 table_stillbirth_pars$Variable[is.na(table_stillbirth_pars$Variable)] <- ""
-
-# Remove any row names so only the column appears in the final table
 rownames(table_stillbirth_pars) <- NULL
 
-# Display table with consistent styling using kable with format "pandoc"
+# Display table using kable with pandoc format
 kable(

### 2025-03-30 code/analysis-code/Statistical_Analysis.qmd
P Update

@@
 ```
+
+```{r}
+colnames(promo_data_clean)
+```
+
+```{r}
+# Run Poisson regression model using study_arm as the treatment variable
+model_poisson <- glm(total_malaria_episodes ~ study_arm + age_at_enrollment_years + gravidity + education_level,
+                     family = poisson(link = "log"), data = promo_data_clean)
+
+# Load broom package to tidy up the model output
+library(broom)
+model_summary <- tidy(model_poisson, exponentiate = TRUE, conf.int = TRUE)
+
+# Print model summary for reference (optional)
+print(model_summary)
+
+# Extract the treatment effect using the correct term name ("study_armSP")
+treatment_effect <- model_summary %>% dplyr::filter(term == "study_armSP")
+
+# Check if the treatment_effect is found; if not, stop with a message
+if(nrow(treatment_effect) == 0) {
+  stop("Treatment effect not found. Please verify the levels and names of the 'study_arm' variable.")
+}
+
+# Create a table with key statistics
+table_poisson <- data.frame(
+  "Exposure and Adjustment Variables" = "Treatment (DP vs SP)",
+  "Incidence Rate Ratio (IRR)" = round(treatment_effect$estimate, 2),
+  "95% CI Lower" = round(treatment_effect$conf.low, 2),
+  "95% CI Upper" = round(treatment_effect$conf.high, 2),
+  "p-value" = signif(treatment_effect$p.value, 3)
+)
+
+# Render the table using kable with Pandoc format
+kable(
+  table_poisson,
+  format = "pandoc",
+  caption = "*Table 4c: Effect of Treatment Arm on Malaria Episode Incidence*",
+  col.names = c("Exposure and Adjustment Variables", "Incidence Rate Ratio (IRR)", "95% CI Lower", "95% CI Upper", "p-value")
+) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+```{r}
+# Create a data frame with refined variable names
+table_model_refined <- data.frame(
+  Variable = c("Intercept", 
+               "Treatment Arm (SP vs. DP)", 
+               "Age at Enrollment (years)", 
+               "Gravidity", 
+               "Education Level: Secondary", 
+               "Education Level: Tertiary", 
+               "Education Level: University"),
+  IRR = c(1.6081682, 1.0552044, 0.9942737, 1.0095680, 0.9884423, 0.8900707, 1.0363703),
+  `95% CI Lower` = c(1.1297705, 0.9382987, 0.9825662, 0.9675647, 0.8348897, 0.7505115, 0.8805393),
+  `95% CI Upper` = c(2.280714, 1.186523, 1.006133, 1.053391, 1.170522, 1.055600, 1.220799),
+  `p-value` = c(0.008021017, 0.369364418, 0.342231948, 0.660449862, 0.892661838, 0.180547176, 0.668035030)
+)
+
+# Remove the intercept row if not required for presentation
+table_model_refined <- table_model_refined[table_model_refined$Variable != "Intercept", ]
+
+# Render the table using knitr::kable without row numbers
+library(knitr)
+kable(
+  table_model_refined,
+  format = "pandoc",
+  caption = "*Poisson Regression Analysis of Malaria Episode Incidence among Pregnant Women*",
+  col.names = c("", "IRR", "95% CI Lower", "95% CI Upper", "p-value"),
+  row.names = FALSE
+)
+
+```

### 2025-03-30 products/manuscript/Manuscript.qmd
P Update

@@
 ---
-title: "The Impact of Malaria Episodes and Treatment Regimens on Adverse Pregnancy Outcomes in Ugandan Women"
-author: "Asmith Joseph"
+title: "Evaluating the Impact of Malaria Episodes and Preventive Treatment Regimens on Pregnancy Outcomes among Ugandan Women"
 bibliography: "../../assets/dataanalysis-references.bib"
@@
   docx:
-    toc: true
+    toc: false
     number-sections: true
@@
 # Introduction
-
 ## General Background Information
-
 Malaria is a life-threatening disease caused by Plasmodium parasites, which are transmitted to humans through the bite of an infected female Anopheles mosquito (Talapko et al., 2019).
@@
 Despite significant efforts in prevention and treatment, malaria remains a major global health challenge.
+
 Vulnerable groups such as pregnant women, children, and immunocompromised individuals are at the highest risk, and Africa continues to bear the heaviest burden of this disease.
@@
 
-## Description of data and data source
+# Methods
+The dataset used in this study was obtained from ClinEpiDB (Release #24, August 30, 2022) and originates from a double-blind randomized controlled trial evaluating monthly intermittent preventive treatment of malaria (IPTp) with sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP) in HIV-uninfected pregnant women.
+Comprising 782 observations, the dataset captures detailed information on maternal health, pregnancy outcomes, and infant variables—with separate records maintained for each infant in cases of multiple births.
+Detailed metadata and documentation were accessed directly, ensuring clear definitions of variable protocols.
 
-The dataset originates from a double-blind, randomized controlled trial evaluating monthly intermittent preventive treatment of malaria (IPTp) using either sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP) in HIV-uninfected pregnant women and their infants.
-Data were rigorously collected throughout the trial, with detailed records capturing maternal health, pregnancy outcomes, and infant details.
-In cases of multiple births, separate records were created for each infant during postpartum visits, ensuring comprehensive outcome documentation.
-The dataset comprises 782 observations and an extensive array of variables, enabling advanced modeling of the impact of malaria episodes and treatment regimens on adverse pregnancy outcomes.
+The objective of this study is to examine the relationship between malaria exposure, treatment regimens, and adverse pregnancy outcomes.
+Specifically, we investigate whether an increased number of malaria episodes is associated with higher rates of preterm birth and stillbirth, and whether the DP regimen reduces these episodes compared to SP.
+We hypothesize that more malaria episodes increase the risk of adverse outcomes (Hypothesis 1) and that DP treatment results in fewer episodes than SP (Hypothesis 2).
+Additionally, the study assesses infant health outcomes by testing if infants born to DP-treated mothers have a lower incidence of low birth weight (Hypothesis 3) and if neonatal mortality rates differ between treatment groups (Hypothesis 4).
+Finally, we explore whether placental malaria is associated with adverse birth outcomes, such as preterm birth and low birth weight (Hypothesis 5).
 
-The dataset encompasses several key variable categories.
-Maternal health and pregnancy outcomes are documented through unique participant identifiers, study arm assignment, and clinical indicators such as preeclampsia, various forms of postpartum hemorrhage (vaginal and cesarean), blood transfusions, laceration/episiotomy, congenital malformations, spontaneous abortions, stillbirths, and neonatal deaths, along with gestational parameters like the first day of the last menstrual period and estimated delivery dates.
-Reproductive history is captured via gravidity, parity, and counts of full-term births, preterm births, and abortions (both elective and spontaneous).
-Infant and child outcomes include details on delivery year, mode and location of delivery, birth complications, survival status, sex, and birth weight for up to seven children per participant.
-Malaria-specific measures record the total number of malaria episodes during pregnancy, infection rates per person-year, and results from routine blood smears, including the proportion of Plasmodium-positive.
-Additionally, the dataset contains information on maternal risk factors and comorbidities—such as education level, substance use, and various chronic conditions—and details on delivery complications and medications administered during labor.
-The data are publicly available through ClinEpiDB (Release #24, dated August 30, 2022).
+Data were imported into R using standardized, rigorously tested functions to ensure accurate transfer.
+During import, date fields were converted to Date objects and categorical variables to factors.
+An initial inspection using functions like str() and summary(), alongside visual tools, confirmed data integrity and proper formatting.
+Missing data were systematically assessed; variables with more than 20% missing values were excluded, while those with less than 20% missing data were retained and imputed as necessary.
 
-## Questions/Hypotheses to be addressed
+# Statistical analysis
+## Exploratory/Descriptive analyses
+We conducted initial descriptive analyses to characterize key demographic and clinical features, including maternal age, parity, treatment allocation (SP versus DP), and the frequency of malaria episodes.
+Standard summary statistics and visualizations (e.g., histograms and box plots) were used to assess variable distributions and identify potential outliers, while missing data were evaluated and managed through imputation or exclusion as appropriate.
 
-The objective of this study is to examine the relationship between malaria exposure, treatment regimens, and adverse pregnancy outcomes in HIV-uninfected pregnant women receiving intermittent preventive treatment for malaria.
-Specifically, we will investigate whether an increased number of malaria episodes is associated with higher rates of preterm birth and stillbirth and whether the dihydroartemisinin-piperaquine (DP) regimen reduces these episodes compared to sulfadoxine-pyrimethamine (SP).
-We hypothesize that more malaria episodes correlate with an increased risk of adverse outcomes (Hypothesis 1) and that DP treatment results in fewer malaria episodes than SP (Hypothesis 2).
+To examine maternal health outcomes, we used logistic regression to assess the association between the number of malaria episodes and adverse outcomes (preterm birth and stillbirth), adjusting for maternal age, parity, and socioeconomic status.
+We further compared malaria episode frequencies between treatment groups using Poisson or negative binomial regression models.
 
-In addition, the study will explore infant health outcomes and the role of placental malaria.
-We will test whether infants born to DP-treated mothers are less likely to have low birth weight and whether neonatal mortality rates differ between the treatment groups.
-We hypothesize that DP treatment will improve infant outcomes by reducing the incidence of low birth weight (Hypothesis 3) and that neonatal mortality rates will vary between the SP and DP arms (Hypothesis 4).
-Finally, we will assess whether placental malaria is associated with adverse birth outcomes, such as preterm birth and low birth weight, and if its prevalence differs by treatment regimen, with the expectation that placental malaria increases the risk of these outcomes (Hypothesis 5).
+For infant health outcomes, we employed chi-square tests to initially compare the proportions of low birth weight infants between treatment groups, followed by logistic regression to adjust for maternal nutrition, gestational age, and other relevant factors.
+Neonatal mortality was similarly evaluated.
 
-{{< pagebreak >}}
+Lastly, we used logistic regression to explore the relationship between placental malaria and adverse birth outcomes (preterm birth and low birth weight), also comparing its prevalence between the SP and DP groups while adjusting for confounders.
 
-# Methods
-
-*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*
-
-## Data aquisition
-
-The data for this study were obtained from ClinEpiDB, a public repository, specifically from Release #24 dated August 30, 2022.
-This dataset originates from a double-blind randomized controlled trial evaluating intermittent preventive treatment of malaria (IPTp) in HIV-uninfected pregnant women.
-The dataset, which includes 782 observations along with comprehensive maternal and infant health variables, was directly downloaded from the ClinEpiDB website without the need for authentication.
-Detailed metadata and documentation were also accessed via the platform, ensuring a clear understanding of variable definitions and study protocols.
-
-## Data import and cleaning
-
-The dataset was imported into R directly from its source using rigorously tested and standardized data-loading functions.
-This approach ensured an accurate and complete transfer of information, providing you with confidence in the integrity of the data.
-We verified the dataset's structure and consistency upon import, confirming the presence of 782 observations and cross-validating variable names with the corresponding metadata.
-To ensure proper formatting for subsequent analyses, date fields were converted into Date objects, and categorical variables were transformed into factors, preserving data integrity and preventing modeling errors.
-
-Following the import and formatting, an initial inspection was conducted to assess the dataset's structure and content.
-This review involved examining internal structures using functions such as str(), summary(), colnames(), and head(), as well as generating descriptive statistics to understand each variable's central tendency, dispersion, and distribution.
-Visual inspections confirmed the data were correctly formatted and free from obvious entry errors, while exploratory visualizations identified potential outliers and anomalies.
-Recognizing the critical impact of missing data, we systematically quantified missing values by creating summary tables and using the missmap() function for visual assessment.
-Variables with more than 20% missing data were excluded to reduce potential bias.
-In contrast, those with less than 20% missing data were retained and appropriate imputation strategies were considered when necessary, ensuring that the final dataset was robust and reliable for subsequent statistical modeling.
-
-## Statistical analysis
-
-# Exploratory/Descriptive analyses
-
-We conducted comprehensive exploratory and descriptive analyses to gain an initial understanding of the dataset's structure, variable distributions, and data quality issues.
-In this phase, we summarized the demographic and clinical characteristics of the study population—including maternal age, parity, treatment allocation (SP vs.
-DP), and the distribution of malaria episodes during pregnancy—using descriptive statistics.
-Specifically, means, medians, standard deviations, and interquartile ranges were calculated for continuous variables, while frequencies and proportions were reported for categorical variables.
-Visualization techniques such as histograms, box plots, and bar charts helped identify outliers, skewed distributions, and other anomalies.
-
-In addition, cross-tabulations were performed to explore preliminary associations between key exposure and outcome variables, such as the frequency of low birth weight across treatment groups.
-We also assessed missing data patterns, using this information to guide decisions on data imputation or excluding variables with excessive missingness.
-Together, these initial analyses provided critical insights that guided model selection and ensured the validity of subsequent inferential statistical tests.
-
-Logistic and Count Regression Analyses (Maternal Health Outcomes)
-
-This section examines the relationship between malaria episodes during pregnancy and adverse outcomes like preterm birth and stillbirth.
-The exposure variable will be the total number of malaria episodes during pregnancy, with preterm birth and stillbirth as binary outcomes.
-Logistic regression will assess these associations, adjusting for confounders such as maternal age, parity, and socioeconomic status.
-Another analysis will investigate whether treatment type (SP vs.
-DP) reduces malaria episodes.
-Here, the exposure variable will be treatment, and the outcome will be the total number of malaria episodes.
-Poisson or negative binomial regression will be used to compare episode counts across treatment groups, adjusting for factors like baseline malaria status, geographic region, and seasonal variations.
-
-Chi-Square Tests and Logistic Regression (Infant Health Outcomes)
-
-This section focuses on birth weight, gestational age, and neonatal mortality outcomes.
-First, we will assess whether infants born to mothers treated with DP are less likely to have low birth weight compared to those treated with SP.
-A chi-square test will initially compare proportions, followed by logistic regression to adjust for confounders such as maternal nutrition and gestational age.
-Next, neonatal mortality will be compared between treatment groups using chi-square tests for unadjusted differences, with logistic regression models controlling for factors such as birth complications, preterm birth, and birth weight.
-Lastly, we will examine whether malaria infection rates during pregnancy are associated with low birth weight and preterm birth.
-Logistic regression models will be used for this analysis, adjusting for maternal age, parity, and gestational age.
-
-Logistic Regression and Chi-Square Tests (Malaria and Pregnancy Outcomes)
-
-This section will assess whether placental malaria is associated with preterm birth and low birth weight.
-Logistic regression models will evaluate these associations while adjusting for relevant confounders, such as the number of malaria episodes during pregnancy and maternal health variables.
-Additionally, the difference in placental malaria prevalence between SP and DP treatment groups will be examined using chi-square tests for initial comparisons and logistic regression for adjustments, controlling for factors like the number of malaria episodes and regional differences.
+&nbsp;
+&nbsp;
 
@@
 ## Basic statistical analysis
-
-*Table 2: Association Between Malaria Episodes and Risk of Preterm Birth*
+**Table 2: Association Between Malaria Episodes and Risk of Preterm Birth**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
   format = "pandoc",
-  caption = "*Table 2: Association Between Malaria Episodes and Risk of Preterm Birth*",
-  col.names = c("Exposure and Adjustment Variables", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")
+  caption = "",
+  col.names = c("", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")
 ) %>%
@@
 
-*Table 3: Association Between Malaria Episodes and Risk of Stillbirth*
+**Table 3: Association Between Malaria Episodes and Risk of Stillbirth**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
@@
   format = "pandoc",
-  caption = "*Table 3: Association Between Malaria Episodes and Risk of Stillbirth*",
-  col.names = c("Exposure and Adjustment Variables", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")
+  caption = "",
+  col.names = c("", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")
 ) %>%
@@
 
-*Association between Malaria Episodes and Adverse Birth Outcomes*
-In adjusted logistic regression analysis (Table 4a), the number of malaria episodes during pregnancy was not significantly associated with preterm birth (aOR: 1.04; 95% CI: 0.95–1.14; p = 0.401).
-Similarly, the penalized Firth logistic regression model showed no significant association between malaria episodes and stillbirth (aOR: 0.96; 95% CI: 0.83–1.10; p = 0.559) (Table 4b).
-No significant associations were observed with age at enrollment, gravidity, or hypertension for either outcome.
+**Association between Malaria Episodes and Adverse Birth Outcomes**
+A total of 782 pregnant women were followed in our study, and we began with the hypothesis that a higher frequency of malaria episodes during pregnancy might lead to poorer birth outcomes.
+To explore this, we first employed logistic regression to assess the relationship with preterm birth.
+Contrary to our initial expectations, the analysis revealed that an increase in malaria episodes was not associated with a higher risk of preterm delivery (adjusted odds ratio [aOR]: 1.04; 95% CI: 0.95–1.14; p = 0.401; Table 4a).
+We then refined our approach by using a penalized Firth logistic regression model, which is particularly useful when dealing with rare events or small subgroups.
+This model too showed no significant link between the number of malaria episodes and stillbirth (aOR: 0.96; 95% CI: 0.83–1.10; p = 0.559; Table 4b).
+These findings suggest that the story is more complex than a simple linear relationship—factors such as maternal age, gravidity, or hypertension did not appear to drive these outcomes either, prompting us to consider that other biological or environmental influences might be at play.
 
-*Effect of Treatment Arm on Malaria Episode Incidence*
-A Poisson regression model assessing malaria episode incidence by treatment arm found no significant difference between the DP and SP treatment groups (IRR: 1.02; 95% CI: 0.92–1.12; p = 0.756) after adjusting for age, gravidity, and education level (Table 4c).
-
-A summary of these primary outcome analyses is presented in Table 4.
+**Effect of Treatment Arm on Malaria Episode Incidence**
+Subsequent analyses focused on whether the type of preventive treatment influenced the occurrence of malaria episodes.
+Two treatment arms—dihydroartemisinin-piperaquine (DP) and sulfadoxine-pyrimethamine (SP)—were compared using a Poisson regression model, with adjustments for age, gravidity, and education level.
+The analysis indicated that the incidence of malaria episodes did not differ significantly between the groups (incidence rate ratio [IRR]: 1.02; 95% CI: 0.92–1.12; p = 0.756; Table 4c).
+These findings suggest that, despite differences in treatment regimens, the overall frequency of malaria episodes remains similar, implying that variations in treatment strategy may have a limited effect on malaria episode incidence in this study population.
 
@@
 
-*To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*
+**Table 4:Incidence Rate Ratios for Malaria Episode Occurrence among Pregnant Women**
+```{r}
+# Create a data frame with refined variable names
+table_model_refined <- data.frame(
+  Variable = c("Intercept", 
+               "Treatment Arm (SP vs. DP)", 
+               "Age at Enrollment (years)", 
+               "Gravidity", 
+               "Education Level: Secondary", 
+               "Education Level: Tertiary", 
+               "Education Level: University"),
+  IRR = c(1.6081682, 1.0552044, 0.9942737, 1.0095680, 0.9884423, 0.8900707, 1.0363703),
+  `95% CI Lower` = c(1.1297705, 0.9382987, 0.9825662, 0.9675647, 0.8348897, 0.7505115, 0.8805393),
+  `95% CI Upper` = c(2.280714, 1.186523, 1.006133, 1.053391, 1.170522, 1.055600, 1.220799),
+  `p-value` = c(0.008021017, 0.369364418, 0.342231948, 0.660449862, 0.892661838, 0.180547176, 0.668035030)
+)
+
+# Remove the intercept row if not required for presentation
+table_model_refined <- table_model_refined[table_model_refined$Variable != "Intercept", ]
+
+# Render the table using knitr::kable without row numbers
+library(knitr)
+kable(
+  table_model_refined,
+  format = "pandoc",
+  caption = "",
+  col.names = c("", "IRR", "95% CI Lower", "95% CI Upper", "p-value"),
+  row.names = FALSE
+)
+
+```
 

### 2025-03-30 products/manuscript/Manuscript.qmd
Project update

@@
 ---
-title: "Evaluating the Impact of Malaria Episodes and Preventive Treatment Regimens on Pregnancy Outcomes among Ugandan Women"
+title: "Impact of IPTp Regimen on Pregnancy Outcomes in a Malaria-Endemic Setting"
 bibliography: "../../assets/dataanalysis-references.bib"
@@
 library(logistf)          # Penalized logistic regression (Firth’s method)
+library(ggeffects)
+library(car)
+library(broom)
 
@@
 ## General Background Information
-Malaria is a life-threatening disease caused by Plasmodium parasites, which are transmitted to humans through the bite of an infected female Anopheles mosquito (Talapko et al., 2019).
-According to the World Health Organization, in 2022, approximately 249 million malaria cases were reported across 85 endemic countries—an increase of 5 million cases compared to 2021.
-Notably, Uganda alone contributed more than 597,000 cases during that period (WHO, 2023).
-Despite significant efforts in prevention and treatment, malaria remains a major global health challenge.
+Malaria remains a major public health challenge worldwide, particularly in Sub-Saharan Africa, where the disease disproportionately affects vulnerable populations.
+Caused by Plasmodium parasites transmitted through the bite of infected Anopheles mosquitoes, malaria accounted for an estimated 249 million cases globally in 2022, marking an increase of five million cases compared to the previous year.
+Uganda alone reported over 597,000 malaria cases during this period, reflecting the country's substantial malaria burden (World Health Organization, 2023; Talapko et al., 2019).
+Pregnant women represent one of the most vulnerable groups to malaria infection, facing an increased risk of severe clinical symptoms and poor pregnancy outcomes.
+Malaria during pregnancy has been associated with a range of adverse outcomes, including miscarriage, fetal loss, preterm birth, low birth weight, and neonatal mortality (Chua et al., 2021).
 
-Vulnerable groups such as pregnant women, children, and immunocompromised individuals are at the highest risk, and Africa continues to bear the heaviest burden of this disease.
-Pregnant women with malaria tend to experience more severe clinical symptoms and poorer outcomes, including heightened risks of miscarriage, fetal loss, premature birth, low birth weight in newborns, and neonatal death (Caroline Lin Lin Chua et al., 2021).
-A multifaceted approach incorporating insecticide-treated bed nets (ITNs) and chemoprevention strategies has proven effective in mitigating the adverse health outcomes associated with malaria during pregnancy.
-Despite this progress, a 2017 report reveals that only 22% of pregnant women in Sub-Saharan Africa received an entire course of IPTp (Bauserman et al., 2019), underscoring a significant disparity in treatment access.
-Addressing this issue, the study seeks to: (a) examine the association between the frequency of malaria episodes during pregnancy and adverse birth outcomes, such as preterm birth and stillbirth; (b) assess whether infants born to mothers treated with DP are less likely to experience low birth weight compared to those treated with SP; and (c) explore the relationship between placental malaria and adverse birth outcomes, including preterm birth and low birth weight.
+Efforts to mitigate the impact of malaria in pregnancy have centered on preventive strategies such as the use of insecticide-treated bed nets (ITNs) and chemoprophylaxis through intermittent preventive treatment during pregnancy (IPTp).
+Two widely used IPTp regimens are sulfadoxine-pyrimethamine (SP) and dihydroartemisinin-piperaquine (DP).
+While these preventive measures have shown significant benefits in reducing the risk of malaria-related complications, the effectiveness of these regimens may not be uniform across all contexts.
+In particular, emerging evidence suggests that the choice of IPTp regimen may influence how malaria episodes affect pregnancy outcomes.
+However, this potential effect modification remains underexplored in current literature.
+Additionally, maternal characteristics such as gravidity—the number of times a woman has been pregnant—may also play a role in shaping birth outcomes.
+Prior research has suggested that previous pregnancy experience may offer protective benefits against adverse outcomes, possibly due to improved physiological adaptation or better health-seeking behavior.
+
+This study addresses two research questions.
+First, it examines whether the type of IPTp regimen modifies the association between malaria episode frequency and adverse birth outcomes in Ugandan pregnant women.
+Preliminary analyses suggest a potential interaction between malaria episodes and the SP regimen, indicating that the impact of malaria may differ by treatment.
+Second, the study investigates whether increased gravidity reduces the risk of adverse birth outcomes among younger pregnant women under 25 years.
+Early findings point to a protective effect of prior pregnancies in this subgroup.
+By addressing these questions, this study aims to clarify how preventive treatment strategies and maternal reproductive history influence birth outcomes in malaria-endemic settings.
+The results may inform targeted interventions to improve maternal and neonatal health in Uganda and similar contexts
 
 # Methods
-The dataset used in this study was obtained from ClinEpiDB (Release #24, August 30, 2022) and originates from a double-blind randomized controlled trial evaluating monthly intermittent preventive treatment of malaria (IPTp) with sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP) in HIV-uninfected pregnant women.
-Comprising 782 observations, the dataset captures detailed information on maternal health, pregnancy outcomes, and infant variables—with separate records maintained for each infant in cases of multiple births.
-Detailed metadata and documentation were accessed directly, ensuring clear definitions of variable protocols.
+This study analyzed data from a double-blind randomized controlled trial conducted in Uganda, obtained from ClinEpiDB (Release #24, August 30, 2022).
+The trial evaluated IPTp with either sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP) among HIV-uninfected pregnant women.
+The dataset comprised 782 observations, including maternal, pregnancy, and infant health information.
+Multiple births were recorded separately, and variable definitions adhered to original trial protocols.
 
-The objective of this study is to examine the relationship between malaria exposure, treatment regimens, and adverse pregnancy outcomes.
-Specifically, we investigate whether an increased number of malaria episodes is associated with higher rates of preterm birth and stillbirth, and whether the DP regimen reduces these episodes compared to SP.
-We hypothesize that more malaria episodes increase the risk of adverse outcomes (Hypothesis 1) and that DP treatment results in fewer episodes than SP (Hypothesis 2).
-Additionally, the study assesses infant health outcomes by testing if infants born to DP-treated mothers have a lower incidence of low birth weight (Hypothesis 3) and if neonatal mortality rates differ between treatment groups (Hypothesis 4).
-Finally, we explore whether placental malaria is associated with adverse birth outcomes, such as preterm birth and low birth weight (Hypothesis 5).
+The analysis focused on two primary objectives: (1) to assess whether the IPTp regimen modifies the association between malaria episode frequency and adverse birth outcomes (preterm birth, low birth weight, and stillbirth), and (2) to evaluate whether increased gravidity is associated with lower risk of adverse outcomes among women under 25 years.
 
-Data were imported into R using standardized, rigorously tested functions to ensure accurate transfer.
-During import, date fields were converted to Date objects and categorical variables to factors.
-An initial inspection using functions like str() and summary(), alongside visual tools, confirmed data integrity and proper formatting.
-Missing data were systematically assessed; variables with more than 20% missing values were excluded, while those with less than 20% missing data were retained and imputed as necessary.
+Data preparation and analysis were performed using R.
+Date fields were standardized, and categorical variables were coded as factors.
+Variables with more than 20% missing data were excluded; those with lower missingness were retained and imputed where appropriate.
+Descriptive statistics and visual inspection were used to assess data integrity.
+
+For the first research question, multivariable logistic regression models were fitted to examine the association between malaria episodes and adverse outcomes, including an interaction term for IPTp regimen.
+Covariates included maternal age, gravidity, and socioeconomic status.
+Predicted probabilities were calculated to facilitate interpretation of interaction effects.
+
+For the second question, a subgroup analysis restricted to women under 25 years assessed the association between gravidity and adverse birth outcomes, using logistic regression and adjusting for relevant covariates.
+Gravidity was treated as a continuous variable, and sensitivity analyses were performed to assess robustness.
+
+All statistical analyses were conducted using the most recent version of R (version 4.3.2).
 
@@
 ## Exploratory/Descriptive analyses
-We conducted initial descriptive analyses to characterize key demographic and clinical features, including maternal age, parity, treatment allocation (SP versus DP), and the frequency of malaria episodes.
-Standard summary statistics and visualizations (e.g., histograms and box plots) were used to assess variable distributions and identify potential outliers, while missing data were evaluated and managed through imputation or exclusion as appropriate.
+Descriptive analyses were conducted to summarize key maternal and clinical characteristics, including maternal age, gravidity, treatment assignment (SP or DP), and malaria episode frequency.
+Standard summary statistics were used for continuous variables, and proportions were reported for categorical variables.
+Distributions were visually assessed to identify potential outliers, and missing data were handled through imputation or exclusion based on the extent of missingness.
 
-To examine maternal health outcomes, we used logistic regression to assess the association between the number of malaria episodes and adverse outcomes (preterm birth and stillbirth), adjusting for maternal age, parity, and socioeconomic status.
-We further compared malaria episode frequencies between treatment groups using Poisson or negative binomial regression models.
+For the first research question, we evaluated whether the IPTp regimen modified the association between malaria episode frequency and adverse birth outcomes (preterm birth, low birth weight, and stillbirth).
+Multivariable logistic regression models were fitted, including an interaction term between malaria episodes and IPTp regimen.
+Models were adjusted for potential confounders, including maternal age, gravidity, and socioeconomic status.
+Predicted probabilities were estimated to facilitate interpretation of interaction effects.
 
-For infant health outcomes, we employed chi-square tests to initially compare the proportions of low birth weight infants between treatment groups, followed by logistic regression to adjust for maternal nutrition, gestational age, and other relevant factors.
-Neonatal mortality was similarly evaluated.
-
-Lastly, we used logistic regression to explore the relationship between placental malaria and adverse birth outcomes (preterm birth and low birth weight), also comparing its prevalence between the SP and DP groups while adjusting for confounders.
+For the second research question, we conducted a subgroup analysis among pregnant women under 25 years of age to assess the association between gravidity and adverse birth outcomes.
+Logistic regression models were used, adjusting for relevant covariates, including maternal age, malaria episode frequency, and treatment regimen.
+Gravidity was treated as a continuous variable.
 
@@
 
-**Table 1. Baseline Characteristics of IPTp Trial Participants**
+**Table 1: Baseline Characteristics of Study Participants by IPTp Treatment Arm**
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
 # Create table data
 table_data <- tribble(
-  ~Characteristic, ~Subcategory, ~`Monthly SP (n=391)`, ~`Monthly DP (n=391)`,
-  "Age, years", "", "23 (19–27)", "23 (19–27)",
-  "Gestational Age at Enrollment (weeks)", "", "15.4 (13.3–17.6)", "15.0 (13.4–17.1)",
-  "", "12–16", "66 (17%)", "62 (15.8%)",
-  "", "17–20", "164 (42.2%)", "197 (50.1%)",
-  "", "21–24", "133 (34.2%)", "118 (30%)",
-  "Gravidity", "", "1.2 (0.9–1.7)", "1.1 (0.8–1.6)",
-  "", "1", "102 (26%)", "95 (24%)",
-  "", "2–3", "180 (46%)", "190 (49%)",
-  "", "≥4", "109 (28%)", "106 (27%)",
-  "Parity", "", "2.43 (1.67)", "2.43 (1.65)",
-  "", "0", "140 (36%)", "145 (37%)",
-  "", "1–2", "170 (44%)", "165 (42%)",
-  "", "≥3", "81 (20%)", "81 (21%)",
-  "Total Malaria Episodes (mean (SD))", "", "1.06 (1.03)", "0.98 (0.95)",
-  "Maternal Nutritional Status/BMI (mean (SD))", "", "22.5 (3.2)", "22.8 (3.1)",
-  "History of Previous Malaria Episodes (%)", "", "40 (10%)", "35 (9%)",
+  ~Characteristic,                           ~Subcategory,               ~`DP (N=4121)`,       ~`SP (N=3701)`,
+  "Age (years)",                             "",                         "25.0 (5.0)",         "25.0 (4.9)",
+  "Gestational Age (weeks)",                 "",                         "15.24 (1.97)",       "14.97 (1.94)",
   "Maternal Education Level (%)",            "",                         "",                   "",
-  "", "A and O Level", "194 (49.4%)", "187 (48.1%)",
-  "", "Primary", "68 (17.3%)", "71 (18.3%)",
+  "",                                        "Primary",                  "262 (66.7%)",        "258 (66.4%)",
   "",                                        "Tertiary",                 "65 (16.5%)",         "61 (15.7%)",
-  "", "University", "66 (16.8%)", "70 (18.0%)"
+  "",                                        "University",               "66 (16.8%)",         "70 (18.0%)",
+  "Gravidity",                               "",                         "",                   "",
+  "",                                        "1",                        "91 (22%)",           "56 (15%)",
+  "",                                        "2–3",                      "167 (41%)",          "166 (45%)",
+  "",                                        "≥4",                       "154 (37%)",          "148 (40%)",
+  "Parity",                                  "",                         "",                   "",
+  "",                                        "0",                        "79 (19%)",           "77 (21%)",
+  "",                                        "1–2",                      "175 (42%)",          "158 (43%)",
+  "",                                        "≥3",                       "158 (38%)",          "135 (36%)",
+  "Total Malaria Episodes",                  "",                         "",                   "",
+  "",                                        "1",                        "245 (59%)",          "216 (58%)",
+  "",                                        "2–3",                      "146 (35%)",          "127 (34%)",
+  "",                                        "≥4",                       "21 (5.1%)",          "27 (7.3%)",
+  "Total Malaria Episodes During Pregnancy", "",                         "",                   "",
+  "",                                        "1",                        "179 (43%)",          "165 (45%)",
+  "",                                        "2–3",                      "138 (33%)",          "120 (32%)",
+  "",                                        "≥4",                       "95 (23%)",           "85 (23%)",
+  "Malaria Infection Rate During Pregnancy", "",                         "1.01 (0.58)",        "1.05 (0.58)",
+  "Placental Malaria (Rogerson Criteria)",   "",                         "75 (18%)",           "72 (19%)",
+  "Preterm Births Count",                    "",                         "",                   "",
+  "",                                        "1",                        "391 (95%)",          "348 (94%)",
+  "",                                        "2",                        "21 (5.1%)",          "22 (5.9%)",
+  "Stillbirth bin",                          "",                         "37 (9.0%)",          "50 (14%)",
+  "Birthweight",                             "",                         "2.79 (0.64)",        "2.81 (0.62)",
+  "History of Previous Malaria Episodes (%)", "",                        "40 (10%)",           "35 (9%)"
 )
@@
   format = "pandoc",
-  caption =
+  caption = ""
 ) %>% 
@@
 
-**Figure 1: Distribution of Total Malaria Episodes**
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Create a histogram of total malaria episodes
-ggplot(data = promo_data_clean, aes(x = total_malaria_episodes)) +
-  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
-  labs(
-    title =,
+**Figure 1: Distribution of Total Malaria Episodes by IPTp Treatment Arm**
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+library(ggplot2)
+
+ggplot(promo_data_clean, aes(x = total_malaria_episodes, fill = study_arm)) +
+  geom_histogram(binwidth = 1, alpha = 0.6, position = "dodge") +
+  labs(title = "",
        x = "Total Malaria Episodes",
-    y = "Count"
+       y = "Frequency",
+       fill = "Treatment Arm") +
+  theme_minimal()
+```
+
+&nbsp;
+&nbsp;
+
+**Table 2: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
+
+```{r,  echo=FALSE, message=FALSE, warning=FALSE}
+# Load necessary libraries
+library(dplyr)
+library(gtsummary)
+library(here)
+
+# Read in the cleaned data (if not already read)
+promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+
+# Recode variables as needed for outcomes analysis:
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    # Create a composite adverse outcome variable: 1 if any of preterm birth, stillbirth, or low birth weight (<2.5 kg) occurs
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0),
+
+    # Recode Preterm Births Count: combine 0 and 1 as "1", and 2 as "2"
+    PretermBirths_cat = case_when(
+      preterm_births_count %in% c(0, 1) ~ "1",
+      preterm_births_count == 2 ~ "2"
+    ),
+    PretermBirths_cat = factor(PretermBirths_cat, levels = c("1", "2"))
   )
+
+# Create Table Z: Outcome Variables by IPTp Treatment Arm with p-values
+table_outcomes <- promo_data_clean %>%
+  dplyr::select(
+    study_arm,
+    `Malaria Infection Rate During Pregnancy` = malaria_infection_rate_during_pregnancy,
+    `Placental Malaria (Rogerson Criteria)` = placental_malaria_by_rogerson_criteria,
+    `Preterm Births Count` = PretermBirths_cat,
+    `Stillbirth bin` = stillbirth_bin,
+    `Birthweight` = birth_weight,
+    `Composite Adverse Outcome` = adverse_birth_outcome
+  ) %>%
+  tbl_summary(
+    by = study_arm,
+    missing = "no",
+    statistic = list(
+      all_continuous() ~ "{mean} ({sd})",
+      all_categorical() ~ "{n} ({p}%)"
+    )
+  ) %>%
+  add_p() %>%
+  modify_header(label = "") %>%
+  modify_caption("")
+
+# Print Table Z
+table_outcomes
+```
+
+&nbsp;
+&nbsp;
+
+**Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes**
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# Create the composite adverse outcome variable
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
+  )
+
+# Fit the interaction model (ensuring that adverse_birth_outcome now exists)
+model_interaction <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm + 
+                           age_at_enrollment_years + gravidity + education_level,
+                         family = binomial(link = "logit"),
+                         data = promo_data_clean)
+
+# Generate predicted probabilities over the range of total malaria episodes by study arm
+pred <- ggeffect(model_interaction, terms = c("total_malaria_episodes [all]", "study_arm"))
+
+# Plot the predicted probabilities
+interaction_plot <- ggplot(pred, aes(x = x, y = predicted, color = group)) +
+  geom_line(size = 1.2) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
+  labs(title = "",
+       x = "Total Malaria Episodes",
+       y = "Predicted Probability of Adverse Outcome",
+       color = "Treatment Arm",
+       fill = "Treatment Arm") +
+  theme_minimal()
+
+interaction_plot
 ```
@@
 ## Basic statistical analysis
-**Table 2: Association Between Malaria Episodes and Risk of Preterm Birth**
+
+Baseline characteristics were similar between IPTp treatment arms, with comparable age, gestational age at enrollment, education, gravidity, and parity distributions (Table 1).
+Although overall malaria exposure measures were similar, the stillbirth rate was significantly higher in the SP arm compared to DP (14% vs.
+9%; p = 0.044) (Table 2).
+
+In our multivariable logistic regression analysis—adjusting for maternal age, gravidity, and education—the main effects of total malaria episodes (OR = 0.88, 95% CI: 0.75–1.03, p = 0.121) and treatment arm (OR = 0.69, 95% CI: 0.44–1.08, p = 0.102) were not significant.
+However, adding an interaction term between total malaria episodes and treatment arm significantly improved model fit (ΔDeviance = 10.11, p = 0.00148), with the interaction itself reaching significance (OR = 1.47, 95% CI: 1.16–1.88, p = 0.002).
+Notably, higher gravidity was protective (OR = 0.89, 95% CI: 0.80–0.98, p = 0.023), underscoring the potential benefit of prior pregnancy experience.
+Sensitivity analyses confirmed that these findings were robust, and model diagnostics (VIFs < 2) indicated no concerning multicollinearity.
+Collectively, these results suggest that the impact of malaria exposure on adverse birth outcomes is modified by the IPTp regimen, highlighting the importance of tailoring preventive strategies based on maternal risk profiles.
+
+&nbsp;
+&nbsp;
+
+**Table 3: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
+#library(broom)
+library(dplyr)
+library(knitr)
+library(kableExtra)
+library(here)
 
-table_preterm_pars_clean <- tribble(
-  ~Variable, ~`Adjusted Odds Ratio (aOR)`, ~`95% CI Lower`, ~`95% CI Upper`, ~`p-value`,
-  "Total malaria episodes during pregnancy", 1.04, 0.95, 1.14, 0.401,
-  "Age at enrollment (years)", 0.99, 0.96, 1.03, 0.697,
-  "Gravidity", 0.91, 0.82, 1.02, 0.106,
-  "Parity", 1.17, 1.05, 1.30, 0.005,
-  "Hypertension", 0.89, 0.65, 1.22, 0.463,
-  "Diabetes mellitus", 1.22, 0.88, 1.69, 0.237
+# Tidy the model output with exponentiated coefficients (odds ratios)
+tidy_model <- tidy(model_interaction, exponentiate = TRUE, conf.int = TRUE)
+
+# Remove the intercept row and recode variable names using case_when()
+tidy_model_filtered <- tidy_model %>%
+  filter(term != "(Intercept)") %>%
+  mutate(
+    term = case_when(
+      term == "total_malaria_episodes" ~ "Total Malaria Episodes",
+      term == "study_armSP" ~ "Treatment Arm (SP)",
+      term == "age_at_enrollment_years" ~ "Age at Enrollment (years)",
+      term == "gravidity" ~ "Gravidity",
+      term == "education_levelSecondary" ~ "Secondary Education",
+      term == "education_levelTertiary" ~ "Tertiary Education",
+      term == "education_levelUniversity" ~ "University Education",
+      term == "total_malaria_episodes:study_armSP" ~ "Interaction: Malaria Episodes × SP",
+      TRUE ~ term
+    )
   )
 
-# Display table with consistent styling
+# Round key values for clarity
+tidy_model_filtered <- tidy_model_filtered %>%
+  mutate(
+    estimate = round(estimate, 2),
+    std.error = round(std.error, 2),
+    statistic = round(statistic, 2),
+    p.value = round(p.value, 3),
+    conf.low = round(conf.low, 2),
+    conf.high = round(conf.high, 2)
+  )
+
+# Render the table with a blank header for the first column
 kable(
-  table_preterm_pars_clean,
+  tidy_model_filtered,
   format = "pandoc",
   caption = "",
-  col.names = c("", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")
+  col.names = c("", "Odds Ratio", "Std. Error", "z value", "p-value", "95% CI Lower", "95% CI Upper")
 ) %>% 
@@
 
-**Table 3: Association Between Malaria Episodes and Risk of Stillbirth**
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Ensure the logistf package is loaded
-library(logistf)
+**Table 4:  Regression Diagnostics for IPTp Treatment Effect**
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 
-# Optional: Check that required variables exist in the dataset
-required_vars <- c("stillbirth_bin", "total_malaria_episodes_during_pregnancy", 
-                   "age_at_enrollment_years", "gravidity", "parity", 
-                   "hypertension", "diabetes_mellitus")
-if(!all(required_vars %in% names(promo_data_clean))) {
-  missing_vars <- required_vars[!required_vars %in% names(promo_data_clean)]
-  stop("The following required variables are missing in promo_data_clean: ", paste(missing_vars, collapse=", "))
-}
+# Read in the cleaned data (ensure adverse_birth_outcome is defined as needed)
+promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
 
-# Run parsimonious stillbirth model
-model_stillbirth_pars <- logistf::logistf(
-  stillbirth_bin ~ total_malaria_episodes_during_pregnancy +
-    age_at_enrollment_years + gravidity + parity +
-    hypertension + diabetes_mellitus,
-  data = promo_data_clean
+# Create composite adverse outcome if not already defined
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
   )
 
-# Extract and format results
-table_stillbirth_pars <- data.frame(
-  Variable = names(model_stillbirth_pars$coefficients),
-  `Adjusted Odds Ratio (aOR)` = round(exp(model_stillbirth_pars$coefficients), 2),
-  `95% CI Lower` = round(exp(model_stillbirth_pars$ci.lower), 2),
-  `95% CI Upper` = round(exp(model_stillbirth_pars$ci.upper), 2),
-  `p-value` = signif(model_stillbirth_pars$prob, 3)
+# Fit Model 1 (without interaction)
+model1 <- glm(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
+                age_at_enrollment_years + gravidity + education_level,
+              family = binomial(link = "logit"), data = promo_data_clean)
+
+# Fit Model 2 (with interaction)
+model2 <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm +
+                age_at_enrollment_years + gravidity + education_level,
+              family = binomial(link = "logit"), data = promo_data_clean)
+
+# Generate Analysis of Deviance Table (Likelihood Ratio Test)
+lrt <- anova(model1, model2, test = "LRT")
+lrt_df <- as.data.frame(lrt)
+lrt_df$Model <- c("No Interaction", "With Interaction")
+lrt_df <- lrt_df %>% 
+  select(Model, `Resid. Df`, `Resid. Dev`, Df, Deviance, `Pr(>Chi)`)
+colnames(lrt_df) <- c("Model", "Residual Df", "Residual Dev", "Df", "Deviance", "p-value")
+
+# Render Table X: Likelihood Ratio Test
+kable(
+  lrt_df,
+  format = "pandoc",
+  caption = ""
+) %>% 
+  kable_styling(full_width = FALSE, position = "center")
+
+# Compute Variance Inflation Factors (VIF) for Model 2 (with interaction)
+vif_vals <- vif(model2)
+vif_df <- as.data.frame(vif_vals)
+vif_df$Variable <- rownames(vif_vals)
+vif_df <- vif_df %>% rename(`Df (VIF)` = Df)
+vif_df <- vif_df %>% 
+  select(Variable, GVIF, `Df (VIF)`, `GVIF^(1/(2*Df))`) %>%
+  mutate(
+    Variable = case_when(
+      Variable == "total_malaria_episodes" ~ "Total Malaria Episodes",
+      Variable == "study_arm" ~ "Treatment Arm",
+      Variable == "age_at_enrollment_years" ~ "Age at Enrollment",
+      Variable == "gravidity" ~ "Gravidity",
+      Variable == "education_level" ~ "Education Level",
+      Variable == "total_malaria_episodes:study_arm" ~ "Interaction: Malaria Episodes × SP",
+      TRUE ~ Variable
+    )
   )
 
-# Remove the intercept row
-table_stillbirth_pars <- table_stillbirth_pars[table_stillbirth_pars$Variable != "(Intercept)", ]
-
-# Map raw variable names to clean, reader-friendly labels
-variable_labels <- c(
-  "total_malaria_episodes_during_pregnancy" = "Total malaria episodes during pregnancy",
-  "age_at_enrollment_years" = "Age at enrollment (years)",
-  "gravidity" = "Gravidity",
-  "parity" = "Parity",
-  "hypertensionYes" = "Hypertension",
-  "diabetes_mellitusYes" = "Diabetes mellitus"
-)
-
-# Replace raw names with neat labels; replace any NA (i.e., unmapped) with blank
-table_stillbirth_pars$Variable <- variable_labels[table_stillbirth_pars$Variable]
-table_stillbirth_pars$Variable[is.na(table_stillbirth_pars$Variable)] <- ""
-rownames(table_stillbirth_pars) <- NULL
-
-# Display table using kable with pandoc format
+# Render Table Y: VIF Diagnostics
 kable(
-  table_stillbirth_pars,
+  vif_df,
   format = "pandoc",
-  caption = "",
-  col.names = c("", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")
+  caption = ""
 ) %>% 
@@
 
-**Association between Malaria Episodes and Adverse Birth Outcomes**
-A total of 782 pregnant women were followed in our study, and we began with the hypothesis that a higher frequency of malaria episodes during pregnancy might lead to poorer birth outcomes.
-To explore this, we first employed logistic regression to assess the relationship with preterm birth.
-Contrary to our initial expectations, the analysis revealed that an increase in malaria episodes was not associated with a higher risk of preterm delivery (adjusted odds ratio [aOR]: 1.04; 95% CI: 0.95–1.14; p = 0.401; Table 4a).
-We then refined our approach by using a penalized Firth logistic regression model, which is particularly useful when dealing with rare events or small subgroups.
-This model too showed no significant link between the number of malaria episodes and stillbirth (aOR: 0.96; 95% CI: 0.83–1.10; p = 0.559; Table 4b).
-These findings suggest that the story is more complex than a simple linear relationship—factors such as maternal age, gravidity, or hypertension did not appear to drive these outcomes either, prompting us to consider that other biological or environmental influences might be at play.
+**Table 5: Evaluating Effect Modification by IPTp Treatment on Malaria-Related Adverse Birth Outcomes Through Model Comparison**
 
-**Effect of Treatment Arm on Malaria Episode Incidence**
-Subsequent analyses focused on whether the type of preventive treatment influenced the occurrence of malaria episodes.
-Two treatment arms—dihydroartemisinin-piperaquine (DP) and sulfadoxine-pyrimethamine (SP)—were compared using a Poisson regression model, with adjustments for age, gravidity, and education level.
-The analysis indicated that the incidence of malaria episodes did not differ significantly between the groups (incidence rate ratio [IRR]: 1.02; 95% CI: 0.92–1.12; p = 0.756; Table 4c).
-These findings suggest that, despite differences in treatment regimens, the overall frequency of malaria episodes remains similar, implying that variations in treatment strategy may have a limited effect on malaria episode incidence in this study population.
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project/Images/Table- Comparing Models .png")
+
+```
 
@@
 &nbsp;
-
-**Table 4:Incidence Rate Ratios for Malaria Episode Occurrence among Pregnant Women**
-```{r}
-# Create a data frame with refined variable names
-table_model_refined <- data.frame(
-  Variable = c("Intercept", 
-               "Treatment Arm (SP vs. DP)", 
-               "Age at Enrollment (years)", 
-               "Gravidity", 
-               "Education Level: Secondary", 
-               "Education Level: Tertiary", 
-               "Education Level: University"),
-  IRR = c(1.6081682, 1.0552044, 0.9942737, 1.0095680, 0.9884423, 0.8900707, 1.0363703),
-  `95% CI Lower` = c(1.1297705, 0.9382987, 0.9825662, 0.9675647, 0.8348897, 0.7505115, 0.8805393),
-  `95% CI Upper` = c(2.280714, 1.186523, 1.006133, 1.053391, 1.170522, 1.055600, 1.220799),
-  `p-value` = c(0.008021017, 0.369364418, 0.342231948, 0.660449862, 0.892661838, 0.180547176, 0.668035030)
-)
-
-# Remove the intercept row if not required for presentation
-table_model_refined <- table_model_refined[table_model_refined$Variable != "Intercept", ]
-
-# Render the table using knitr::kable without row numbers
-library(knitr)
-kable(
-  table_model_refined,
-  format = "pandoc",
-  caption = "",
-  col.names = c("", "IRR", "95% CI Lower", "95% CI Upper", "p-value"),
-  row.names = FALSE
-)
-
-```
 

### 2025-03-30 code/analysis-code/Statistical_Analysis.qmd
Project update

@@
 ```
+
+```{r}
+# Fit the Poisson regression model (assuming promo_data_clean is your dataset)
+model_poisson <- glm(total_malaria_episodes ~ study_arm + age_at_enrollment_years + gravidity + education_level,
+                     family = poisson(link = "log"), data = promo_data_clean)
+
+# Calculate the dispersion parameter using Pearson residuals
+dispersion_parameter <- sum(residuals(model_poisson, type = "pearson")^2) / model_poisson$df.residual
+cat("Dispersion parameter (Pearson):", dispersion_parameter, "\n")
+
+# Alternatively, check the ratio of the residual deviance to the degrees of freedom
+ratio_deviance <- model_poisson$deviance / model_poisson$df.residual
+cat("Residual deviance / DF:", ratio_deviance, "\n")
+
+# Optionally, plot Pearson residuals vs fitted values to visually assess overdispersion
+plot(fitted(model_poisson), residuals(model_poisson, type = "pearson"),
+     xlab = "Fitted values", ylab = "Pearson residuals",
+     main = "Residuals vs Fitted Values")
+abline(h = 0, col = "red", lty = 2)
+
+```
+
+```{r}
+# Load the MASS package for negative binomial regression
+library(MASS)
+
+# Fit a negative binomial regression model
+model_nb <- glm.nb(total_malaria_episodes ~ study_arm + age_at_enrollment_years + gravidity + education_level,
+                   data = promo_data_clean)
+
+# Check the model summary
+summary(model_nb)
+
+# Tidy the model output using broom, exponentiating to get IRRs and 95% CIs
+library(broom)
+model_nb_summary <- tidy(model_nb, exponentiate = TRUE, conf.int = TRUE)
+
+# Print the tidied model summary to inspect coefficients
+print(model_nb_summary)
+
+# Optionally, extract the treatment effect if needed (assuming treatment variable is 'study_armSP')
+treatment_effect_nb <- model_nb_summary %>% dplyr::filter(term == "study_armSP")
+print(treatment_effect_nb)
+
+```

### 2025-03-30 products/manuscript/Manuscript.qmd
Project update

@@
 ---
-title: "Impact of IPTp Regimen on Pregnancy Outcomes in a Malaria-Endemic Setting"
+title: "Evaluating the Impact of Malaria Episodes and Preventive Treatment Regimens on Pregnancy Outcomes among Ugandan Women"
 bibliography: "../../assets/dataanalysis-references.bib"
@@
 library(ggeffects)
-library(car)
-library(broom)
 
@@
 
-This study addresses two research questions.
-First, it examines whether the type of IPTp regimen modifies the association between malaria episode frequency and adverse birth outcomes in Ugandan pregnant women.
-Preliminary analyses suggest a potential interaction between malaria episodes and the SP regimen, indicating that the impact of malaria may differ by treatment.
-Second, the study investigates whether increased gravidity reduces the risk of adverse birth outcomes among younger pregnant women under 25 years.
-Early findings point to a protective effect of prior pregnancies in this subgroup.
-By addressing these questions, this study aims to clarify how preventive treatment strategies and maternal reproductive history influence birth outcomes in malaria-endemic settings.
-The results may inform targeted interventions to improve maternal and neonatal health in Uganda and similar contexts
+This study seeks to address two specific research questions to fill these gaps in knowledge.
+First, it examines whether the type of IPTp regimen modifies the association between the frequency of malaria episodes during pregnancy and adverse birth outcomes in Ugandan pregnant women.
+Preliminary analyses have indicated a significant interaction between malaria episode frequency and the SP treatment arm, suggesting that the impact of malaria on birth outcomes may differ depending on the preventive treatment received.
+Understanding whether the choice of regimen alters this relationship could inform clinical decisions and public health policies aimed at reducing malaria-related complications during pregnancy.
+
+Second, the study investigates whether increased gravidity is associated with a reduced risk of adverse birth outcomes among younger pregnant women in Uganda.
+Subgroup analysis focusing on women under the age of 25 has revealed a potential protective effect of higher gravidity, with data suggesting that women with prior pregnancy experience may be less likely to experience adverse outcomes.
+Clarifying this relationship could help identify young, first-time mothers as a high-risk group and inform targeted interventions to improve maternal and neonatal health outcomes.
+
+By addressing these two research questions, this study aims to contribute meaningful insights into how preventive treatment strategies and maternal reproductive history influence pregnancy outcomes in malaria-endemic settings.
+The findings have the potential to guide more effective, equitable, and context-specific interventions to improve the health of pregnant women and their infants in Uganda and similar settings.
 
@@
 
-**Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes**
+*Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes*
 
@@
 
-{{< pagebreak >}}
+#####################################################################################################################################################
 
-# Results
+**Table 1. Baseline Characteristics of IPTp Trial Participants**
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 
-## Basic statistical analysis
-
-Baseline characteristics were similar between IPTp treatment arms, with comparable age, gestational age at enrollment, education, gravidity, and parity distributions (Table 1).
-Although overall malaria exposure measures were similar, the stillbirth rate was significantly higher in the SP arm compared to DP (14% vs.
-9%; p = 0.044) (Table 2).
-
-In our multivariable logistic regression analysis—adjusting for maternal age, gravidity, and education—the main effects of total malaria episodes (OR = 0.88, 95% CI: 0.75–1.03, p = 0.121) and treatment arm (OR = 0.69, 95% CI: 0.44–1.08, p = 0.102) were not significant.
-However, adding an interaction term between total malaria episodes and treatment arm significantly improved model fit (ΔDeviance = 10.11, p = 0.00148), with the interaction itself reaching significance (OR = 1.47, 95% CI: 1.16–1.88, p = 0.002).
-Notably, higher gravidity was protective (OR = 0.89, 95% CI: 0.80–0.98, p = 0.023), underscoring the potential benefit of prior pregnancy experience.
-Sensitivity analyses confirmed that these findings were robust, and model diagnostics (VIFs < 2) indicated no concerning multicollinearity.
-Collectively, these results suggest that the impact of malaria exposure on adverse birth outcomes is modified by the IPTp regimen, highlighting the importance of tailoring preventive strategies based on maternal risk profiles.
-
-&nbsp;
-&nbsp;
-
-**Table 3: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-#library(broom)
-library(dplyr)
-library(knitr)
-library(kableExtra)
-library(here)
-
-# Tidy the model output with exponentiated coefficients (odds ratios)
-tidy_model <- tidy(model_interaction, exponentiate = TRUE, conf.int = TRUE)
-
-# Remove the intercept row and recode variable names using case_when()
-tidy_model_filtered <- tidy_model %>%
-  filter(term != "(Intercept)") %>%
-  mutate(
-    term = case_when(
-      term == "total_malaria_episodes" ~ "Total Malaria Episodes",
-      term == "study_armSP" ~ "Treatment Arm (SP)",
-      term == "age_at_enrollment_years" ~ "Age at Enrollment (years)",
-      term == "gravidity" ~ "Gravidity",
-      term == "education_levelSecondary" ~ "Secondary Education",
-      term == "education_levelTertiary" ~ "Tertiary Education",
-      term == "education_levelUniversity" ~ "University Education",
-      term == "total_malaria_episodes:study_armSP" ~ "Interaction: Malaria Episodes × SP",
-      TRUE ~ term
-    )
+# Create table data
+table_data <- tribble(
+  ~Characteristic, ~Subcategory, ~`Monthly SP (n=391)`, ~`Monthly DP (n=391)`,
+  "Age, years", "", "23 (19–27)", "23 (19–27)",
+  "Gestational Age at Enrollment (weeks)", "", "15.4 (13.3–17.6)", "15.0 (13.4–17.1)",
+  "", "12–16", "66 (17%)", "62 (15.8%)",
+  "", "17–20", "164 (42.2%)", "197 (50.1%)",
+  "", "21–24", "133 (34.2%)", "118 (30%)",
+  "Gravidity", "", "1.2 (0.9–1.7)", "1.1 (0.8–1.6)",
+  "", "1", "102 (26%)", "95 (24%)",
+  "", "2–3", "180 (46%)", "190 (49%)",
+  "", "≥4", "109 (28%)", "106 (27%)",
+  "Parity", "", "2.43 (1.67)", "2.43 (1.65)",
+  "", "0", "140 (36%)", "145 (37%)",
+  "", "1–2", "170 (44%)", "165 (42%)",
+  "", "≥3", "81 (20%)", "81 (21%)",
+  "Total Malaria Episodes (mean (SD))", "", "1.06 (1.03)", "0.98 (0.95)",
+  "Maternal Nutritional Status/BMI (mean (SD))", "", "22.5 (3.2)", "22.8 (3.1)",
+  "History of Previous Malaria Episodes (%)", "", "40 (10%)", "35 (9%)",
+  "Maternal Education Level (%)", "", "", "",
+  "", "A and O Level", "194 (49.4%)", "187 (48.1%)",
+  "", "Primary", "68 (17.3%)", "71 (18.3%)",
+  "", "Tertiary", "65 (16.5%)", "61 (15.7%)",
+  "", "University", "66 (16.8%)", "70 (18.0%)"
 )
 
-# Round key values for clarity
-tidy_model_filtered <- tidy_model_filtered %>%
-  mutate(
-    estimate = round(estimate, 2),
-    std.error = round(std.error, 2),
-    statistic = round(statistic, 2),
-    p.value = round(p.value, 3),
-    conf.low = round(conf.low, 2),
-    conf.high = round(conf.high, 2)
-  )
-
-# Render the table with a blank header for the first column
 kable(
-  tidy_model_filtered,
+  table_data,
   format = "pandoc",
-  caption = "",
-  col.names = c("", "Odds Ratio", "Std. Error", "z value", "p-value", "95% CI Lower", "95% CI Upper")
+  caption =
 ) %>%
@@
 
-**Table 4:  Regression Diagnostics for IPTp Treatment Effect**
+**Figure 1: Distribution of Total Malaria Episodes**
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+# Create a histogram of total malaria episodes
+ggplot(data = promo_data_clean, aes(x = total_malaria_episodes)) +
+  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
+  labs(
+    title =,
+    x = "Total Malaria Episodes",
+    y = "Count"
+  )
+```
+
+{{< pagebreak >}}
+
+# Results
+
+## Basic statistical analysis
+**Table 2: Association Between Malaria Episodes and Risk of Preterm Birth**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
 
-# Read in the cleaned data (ensure adverse_birth_outcome is defined as needed)
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
-
-# Create composite adverse outcome if not already defined
-promo_data_clean <- promo_data_clean %>%
-  mutate(
-    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
-    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
+table_preterm_pars_clean <- tribble(
+  ~Variable, ~`Adjusted Odds Ratio (aOR)`, ~`95% CI Lower`, ~`95% CI Upper`, ~`p-value`,
+  "Total malaria episodes during pregnancy", 1.04, 0.95, 1.14, 0.401,
+  "Age at enrollment (years)", 0.99, 0.96, 1.03, 0.697,
+  "Gravidity", 0.91, 0.82, 1.02, 0.106,
+  "Parity", 1.17, 1.05, 1.30, 0.005,
+  "Hypertension", 0.89, 0.65, 1.22, 0.463,
+  "Diabetes mellitus", 1.22, 0.88, 1.69, 0.237
 )
 
-# Fit Model 1 (without interaction)
-model1 <- glm(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
-                age_at_enrollment_years + gravidity + education_level,
-              family = binomial(link = "logit"), data = promo_data_clean)
-
-# Fit Model 2 (with interaction)
-model2 <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm +
-                age_at_enrollment_years + gravidity + education_level,
-              family = binomial(link = "logit"), data = promo_data_clean)
-
-# Generate Analysis of Deviance Table (Likelihood Ratio Test)
-lrt <- anova(model1, model2, test = "LRT")
-lrt_df <- as.data.frame(lrt)
-lrt_df$Model <- c("No Interaction", "With Interaction")
-lrt_df <- lrt_df %>% 
-  select(Model, `Resid. Df`, `Resid. Dev`, Df, Deviance, `Pr(>Chi)`)
-colnames(lrt_df) <- c("Model", "Residual Df", "Residual Dev", "Df", "Deviance", "p-value")
-
-# Render Table X: Likelihood Ratio Test
+# Display table with consistent styling
 kable(
-  lrt_df,
+  table_preterm_pars_clean,
   format = "pandoc",
-  caption = ""
-) %>% 
-  kable_styling(full_width = FALSE, position = "center")
-
-# Compute Variance Inflation Factors (VIF) for Model 2 (with interaction)
-vif_vals <- vif(model2)
-vif_df <- as.data.frame(vif_vals)
-vif_df$Variable <- rownames(vif_vals)
-vif_df <- vif_df %>% rename(`Df (VIF)` = Df)
-vif_df <- vif_df %>% 
-  select(Variable, GVIF, `Df (VIF)`, `GVIF^(1/(2*Df))`) %>%
-  mutate(
-    Variable = case_when(
-      Variable == "total_malaria_episodes" ~ "Total Malaria Episodes",
-      Variable == "study_arm" ~ "Treatment Arm",
-      Variable == "age_at_enrollment_years" ~ "Age at Enrollment",
-      Variable == "gravidity" ~ "Gravidity",
-      Variable == "education_level" ~ "Education Level",
-      Variable == "total_malaria_episodes:study_arm" ~ "Interaction: Malaria Episodes × SP",
-      TRUE ~ Variable
-    )
-  )
-
-# Render Table Y: VIF Diagnostics
-kable(
-  vif_df,
-  format = "pandoc",
-  caption = ""
+  caption = "",
+  col.names = c("", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")
 ) %>%
@@
 
-**Table 5: Evaluating Effect Modification by IPTp Treatment on Malaria-Related Adverse Birth Outcomes Through Model Comparison**
+**Table 3: Association Between Malaria Episodes and Risk of Stillbirth**
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+# Ensure the logistf package is loaded
+library(logistf)
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics("C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project/Images/Table- Comparing Models .png")
+# Optional: Check that required variables exist in the dataset
+required_vars <- c("stillbirth_bin", "total_malaria_episodes_during_pregnancy", 
+                   "age_at_enrollment_years", "gravidity", "parity", 
+                   "hypertension", "diabetes_mellitus")
+if(!all(required_vars %in% names(promo_data_clean))) {
+  missing_vars <- required_vars[!required_vars %in% names(promo_data_clean)]
+  stop("The following required variables are missing in promo_data_clean: ", paste(missing_vars, collapse=", "))
+}
+
+# Run parsimonious stillbirth model
+model_stillbirth_pars <- logistf::logistf(
+  stillbirth_bin ~ total_malaria_episodes_during_pregnancy +
+    age_at_enrollment_years + gravidity + parity +
+    hypertension + diabetes_mellitus,
+  data = promo_data_clean
+)
+
+# Extract and format results
+table_stillbirth_pars <- data.frame(
+  Variable = names(model_stillbirth_pars$coefficients),
+  `Adjusted Odds Ratio (aOR)` = round(exp(model_stillbirth_pars$coefficients), 2),
+  `95% CI Lower` = round(exp(model_stillbirth_pars$ci.lower), 2),
+  `95% CI Upper` = round(exp(model_stillbirth_pars$ci.upper), 2),
+  `p-value` = signif(model_stillbirth_pars$prob, 3)
+)
+
+# Remove the intercept row
+table_stillbirth_pars <- table_stillbirth_pars[table_stillbirth_pars$Variable != "(Intercept)", ]
+
+# Map raw variable names to clean, reader-friendly labels
+variable_labels <- c(
+  "total_malaria_episodes_during_pregnancy" = "Total malaria episodes during pregnancy",
+  "age_at_enrollment_years" = "Age at enrollment (years)",
+  "gravidity" = "Gravidity",
+  "parity" = "Parity",
+  "hypertensionYes" = "Hypertension",
+  "diabetes_mellitusYes" = "Diabetes mellitus"
+)
+
+# Replace raw names with neat labels; replace any NA (i.e., unmapped) with blank
+table_stillbirth_pars$Variable <- variable_labels[table_stillbirth_pars$Variable]
+table_stillbirth_pars$Variable[is.na(table_stillbirth_pars$Variable)] <- ""
+rownames(table_stillbirth_pars) <- NULL
+
+# Display table using kable with pandoc format
+kable(
+  table_stillbirth_pars,
+  format = "pandoc",
+  caption = "",
+  col.names = c("", "Adjusted Odds Ratio (aOR)", "95% CI Lower", "95% CI Upper", "p-value")
+) %>%
+  kable_styling(full_width = FALSE, position = "center")
 
@@
 &nbsp;
+
+**Association between Malaria Episodes and Adverse Birth Outcomes**
+A total of 782 pregnant women were followed in our study, and we began with the hypothesis that a higher frequency of malaria episodes during pregnancy might lead to poorer birth outcomes.
+To explore this, we first employed logistic regression to assess the relationship with preterm birth.
+Contrary to our initial expectations, the analysis revealed that an increase in malaria episodes was not associated with a higher risk of preterm delivery (adjusted odds ratio [aOR]: 1.04; 95% CI: 0.95–1.14; p = 0.401; Table 4a).
+We then refined our approach by using a penalized Firth logistic regression model, which is particularly useful when dealing with rare events or small subgroups.
+This model too showed no significant link between the number of malaria episodes and stillbirth (aOR: 0.96; 95% CI: 0.83–1.10; p = 0.559; Table 4b).
+These findings suggest that the story is more complex than a simple linear relationship—factors such as maternal age, gravidity, or hypertension did not appear to drive these outcomes either, prompting us to consider that other biological or environmental influences might be at play.
+
+**Effect of Treatment Arm on Malaria Episode Incidence**
+Subsequent analyses focused on whether the type of preventive treatment influenced the occurrence of malaria episodes.
+Two treatment arms—dihydroartemisinin-piperaquine (DP) and sulfadoxine-pyrimethamine (SP)—were compared using a Poisson regression model, with adjustments for age, gravidity, and education level.
+The analysis indicated that the incidence of malaria episodes did not differ significantly between the groups (incidence rate ratio [IRR]: 1.02; 95% CI: 0.92–1.12; p = 0.756; Table 4c).
+These findings suggest that, despite differences in treatment regimens, the overall frequency of malaria episodes remains similar, implying that variations in treatment strategy may have a limited effect on malaria episode incidence in this study population.
+
+&nbsp;
+&nbsp;
+
+**Table 4: "Incidence Rate Ratios for Malaria Episode Occurrence among Pregnant Women**
+```{r}
+# Create a data frame with refined variable names
+table_model_refined <- data.frame(
+  Variable = c("Intercept", 
+               "Treatment Arm (SP vs. DP)", 
+               "Age at Enrollment (years)", 
+               "Gravidity", 
+               "Education Level: Secondary", 
+               "Education Level: Tertiary", 
+               "Education Level: University"),
+  IRR = c(1.6081682, 1.0552044, 0.9942737, 1.0095680, 0.9884423, 0.8900707, 1.0363703),
+  `95% CI Lower` = c(1.1297705, 0.9382987, 0.9825662, 0.9675647, 0.8348897, 0.7505115, 0.8805393),
+  `95% CI Upper` = c(2.280714, 1.186523, 1.006133, 1.053391, 1.170522, 1.055600, 1.220799),
+  `p-value` = c(0.008021017, 0.369364418, 0.342231948, 0.660449862, 0.892661838, 0.180547176, 0.668035030)
+)
+
+# Remove the intercept row if not required for presentation
+table_model_refined <- table_model_refined[table_model_refined$Variable != "Intercept", ]
+
+# Render the table using knitr::kable without row numbers
+library(knitr)
+kable(
+  table_model_refined,
+  format = "pandoc",
+  caption = "",
+  col.names = c("", "IRR", "95% CI Lower", "95% CI Upper", "p-value"),
+  row.names = FALSE
+)
+
+```
 

### 2025-04-03 code/analysis-code/Statistical_Analysis.qmd
update

--- 2025-03-30 code/analysis-code/Statistical_Analysis.qmd
@@
 ```{r}
-# **1. Load Required Libraries**
+# Load Required Libraries
 # Essential libraries for data handling, visualization, and analysis
@@
 library(dplyr)            # Data manipulation
-library(ggplot2)          # Visualization
-library(tidyverse)        # Data wrangling and cleaning
-library(janitor)          # Cleaning column names
+library(tidyverse)        # Data wrangling and visualization
+library(janitor)          # Clean column names
 library(skimr)            # Quick data summary
@@
 library(readr)            # Reading CSV files
-library(gtsummary)        # Creating summary tables
-library(gt)               # Formatting tables
-library(knitr)            # Table formatting
-library(kableExtra)       # Additional table customization
+library(ggplot2)          # Visualization
 library(ggpubr)           # Publication-ready plots
+library(forcats)          # Categorical variables handling
+
+# Table and reporting packages
+library(gtsummary)        # Summary tables
+library(gt)               # Table formatting
+library(knitr)            # Table rendering
+library(kableExtra)       # Table customization
+
+# Analysis-specific packages
 library(Amelia)           # Missing data visualization
-library(forcats)          # Working with categorical variables
 library(pwr)              # Power analysis
-library(DiagrammeR)       # Flowcharts/Diagrams
+library(DiagrammeR)       # Diagrams and flowcharts
 library(survival)         # Survival analysis
 library(survminer)        # Survival plots
-```
-
-```{r}
-# Load the here package
-library(here)
-
-# Construct the relative file path and import the CSV data
-promo_data <- read.csv(here("data", "raw-data", "PROMO_Data.csv"))
-
-# View the first few rows of the data
-head(promo_data)
-
-```
-
-## Part 2: Data Loading, Cleaning, Wrangling, and Exploratory Analysis
-
-# Data Cleaning & Transformation
-```{r}
-# Display the structure of the data
-str(promo_data)
-
-# Print the column names
-colnames(promo_data)
-
-# Display the first few rows of the data
-head(promo_data)
-
-# Provide summary statistics for each variable
-summary(promo_data)
+library(ggeffects)        # Marginal effects for regression models
+library(broom)            # Tidy model outputs
+library(rsample)          # Resampling and splitting datasets
 
@@
 ```{r}
-# Import the CSV data using here
-promo_data <- read.csv(here("data", "raw-data", "PROMO_Data.csv"))
+# Data Import & Initial Inspection 
+# load the Dataset
+PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"))
 
-# Clean column names (all lower-case with underscores)
-promo_data_clean <- promo_data %>% 
-  clean_names()
+promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+```
 
+```{r}
+colnames(promo_data_clean)
+```
+
+```{r}
+names(promo_data_clean)
+```
+
+*Table 1*
+
+```{r}
+# Recode variables for baseline table presentation:
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    # Create an age group variable (for potential subgroup analyses)
+    age_group = ifelse(age_at_enrollment_years < 25, "Young", "Older"),
+
+    # Recode Gravidity into categories: "1", "2–3", "≥4" (ordered chronologically)
+    Gravidity_cat = case_when(
+      gravidity == 1 ~ "1",
+      gravidity %in% c(2, 3) ~ "2–3",
+      gravidity >= 4 ~ "≥4"
+    ),
+    Gravidity_cat = factor(Gravidity_cat, levels = c("1", "2–3", "≥4")),
+
+    # Recode Parity into categories: "0", "1–2", "≥3" (ordered chronologically)
+    Parity_cat = case_when(
+      parity == 0 ~ "0",
+      parity %in% c(1, 2) ~ "1–2",
+      parity >= 3 ~ "≥3"
+    ),
+    Parity_cat = factor(Parity_cat, levels = c("0", "1–2", "≥3")),
+
+    # Recode Total Malaria Episodes into categories:
+    # Combine 0 and 1 episodes as "1", 2-3 as "2–3", and 4 or more as "≥4"
+    MalariaEpisodes_cat = case_when(
+      total_malaria_episodes %in% c(0, 1) ~ "1",
+      total_malaria_episodes %in% c(2, 3) ~ "2–3",
+      total_malaria_episodes >= 4 ~ "≥4"
+    ),
+    MalariaEpisodes_cat = factor(MalariaEpisodes_cat, levels = c("1", "2–3", "≥4")),
+
+    # Recode Total Malaria Episodes During Pregnancy similarly:
+    MalariaEpisodesPreg_cat = case_when(
+      total_malaria_episodes_during_pregnancy %in% c(0, 1) ~ "1",
+      total_malaria_episodes_during_pregnancy %in% c(2, 3) ~ "2–3",
+      total_malaria_episodes_during_pregnancy >= 4 ~ "≥4"
+    ),
+    MalariaEpisodesPreg_cat = factor(MalariaEpisodesPreg_cat, levels = c("1", "2–3", "≥4")),
+
+    # Recode Preterm Births Count into categories:
+    # Combine 0 and 1 as "1", and 2 as "2"
+    PretermBirths_cat = case_when(
+      preterm_births_count %in% c(0, 1) ~ "1",
+      preterm_births_count == 2 ~ "2"
+    ),
+    PretermBirths_cat = factor(PretermBirths_cat, levels = c("1", "2"))
+  )
+
+# Create Table X: Baseline Characteristics Stratified by IPTp Treatment Arm
+baseline_table_treatment <- promo_data_clean %>%
+  dplyr::select(
+    study_arm,
+    `Age (years)` = age_at_enrollment_years,
+    `Gestational Age (weeks)` = gestational_age_at_enrollment_weeks,
+    `Maternal Education Level` = education_level,
+    Gravidity = Gravidity_cat,
+    Parity = Parity_cat,
+    `Total Malaria Episodes` = MalariaEpisodes_cat,
+    `Total Malaria Episodes During Pregnancy` = MalariaEpisodesPreg_cat,
+    `Malaria Infection Rate During Pregnancy` = malaria_infection_rate_during_pregnancy,
+    `Placental Malaria (Rogerson Criteria)` = placental_malaria_by_rogerson_criteria,
+    `Preterm Births Count` = PretermBirths_cat,
+    `Stillbirth bin` = stillbirth_bin,
+    `Birthweight` = birth_weight
+  ) %>%
+  tbl_summary(
+    by = study_arm,
+    missing = "no",
+    statistic = list(
+      all_continuous() ~ "{mean} ({sd})",
+      all_categorical() ~ "{n} ({p}%)"
+    )
+  ) %>%
+  modify_header(label = "") %>%
+  modify_caption("**Table X: Baseline Characteristics by IPTp Treatment Arm**")
+
+# Print Table X
+baseline_table_treatment
+
+```
+
+*Table 2: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen*
+
+```{r}
+# Load necessary libraries
+library(dplyr)
+library(gtsummary)
+library(here)
+
+# Read in the cleaned data (if not already read)
+promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+
+# Recode variables as needed for outcomes analysis:
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    # Create a composite adverse outcome variable: 1 if any of preterm birth, stillbirth, or low birth weight (<2.5 kg) occurs
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0),
+
+    # Recode Preterm Births Count as before (0/1 -> "1", 2 -> "2")
+    PretermBirths_cat = case_when(
+      preterm_births_count %in% c(0, 1) ~ "1",
+      preterm_births_count == 2 ~ "2"
+    ),
+    PretermBirths_cat = factor(PretermBirths_cat, levels = c("1", "2"))
+  )
+
+# Create Table Z: Outcome Variables by IPTp Treatment Arm with p-values
+table_outcomes <- promo_data_clean %>%
+  dplyr::select(
+    study_arm,
+    `Malaria Infection Rate During Pregnancy` = malaria_infection_rate_during_pregnancy,
+    `Placental Malaria (Rogerson Criteria)` = placental_malaria_by_rogerson_criteria,
+    `Preterm Births Count` = PretermBirths_cat,
+    `Stillbirth bin` = stillbirth_bin,
+    `Birthweight` = birth_weight,
+    `Composite Adverse Outcome` = adverse_birth_outcome
+  ) %>%
+  tbl_summary(
+    by = study_arm,
+    missing = "no",
+    statistic = list(
+      all_continuous() ~ "{mean} ({sd})",
+      all_categorical() ~ "{n} ({p}%)"
+    )
+  ) %>%
+  add_p() %>%
+  modify_header(label = "") %>%
+  modify_caption("")
+
+# Print Table Z
+table_outcomes
+
+```
+
+# Visualization
+
+*Figure 1: Histogram of Total Malaria Episodes by IPTp Treatment Arm*
+
+```{r}
+library(ggplot2)
+
+ggplot(promo_data_clean, aes(x = total_malaria_episodes, fill = study_arm)) +
+  geom_histogram(binwidth = 1, alpha = 0.6, position = "dodge") +
+  labs(title = "Distribution of Total Malaria Episodes by IPTp Treatment Arm",
+       x = "Total Malaria Episodes",
+       y = "Frequency",
+       fill = "Treatment Arm") +
+  theme_minimal()
+
+```
+
+*Bar Graph of Total Malaria Episodes During Pregnancy by Treatment Arm*
+
+```{r}
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
+library(here)
+
+# Read in the cleaned data
+promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+
+# Recode 'total_malaria_episodes_during_pregnancy' into a categorical variable
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    MalariaEpisodesPreg_cat = case_when(
+      total_malaria_episodes_during_pregnancy %in% c(0, 1) ~ "1",
+      total_malaria_episodes_during_pregnancy %in% c(2, 3) ~ "2–3",
+      total_malaria_episodes_during_pregnancy >= 4 ~ "≥4"
+    ),
+    MalariaEpisodesPreg_cat = factor(MalariaEpisodesPreg_cat, levels = c("1", "2–3", "≥4"))
+  )
+
+# Create the bar graph
+ggplot(promo_data_clean, aes(x = MalariaEpisodesPreg_cat, fill = study_arm)) +
+  geom_bar(position = "dodge") +
+  labs(title = "Total Malaria Episodes During Pregnancy by Treatment Arm",
+       x = "Total Malaria Episodes During Pregnancy (Categorical)",
+       y = "Count",
+       fill = "Treatment Arm") +
+  theme_minimal()
+
+```
+
+*Boxplot of Birthweight by IPTp Treatment Arm*
+
+```{r}
+ggplot(promo_data_clean, aes(x = study_arm, y = birth_weight, fill = study_arm)) +
+  geom_boxplot(alpha = 0.7) +
+  labs(title = "Birthweight by IPTp Treatment Arm",
+       x = "Treatment Arm",
+       y = "Birthweight (kg)") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+```
+
+*Bar Graphs for Gravidity and Parity*
+
+```{r}
+# Load necessary libraries
+
+# Read in the cleaned data
+promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+
+# Recode Gravidity into categories: "1", "2–3", "≥4"
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    Gravidity_cat = case_when(
+      gravidity == 1 ~ "1",
+      gravidity %in% c(2, 3) ~ "2–3",
+      gravidity >= 4 ~ "≥4"
+    ),
+    Gravidity_cat = factor(Gravidity_cat, levels = c("1", "2–3", "≥4"))
+  )
+
+# Create the bar graph for Gravidity by Treatment Arm
+ggplot(promo_data_clean, aes(x = Gravidity_cat, fill = study_arm)) +
+  geom_bar(position = "dodge") +
+  labs(title = "Gravidity Distribution by Treatment Arm",
+       x = "Gravidity Category (1, 2–3, ≥4)",
+       y = "Count",
+       fill = "Treatment Arm") +
+  theme_minimal()
+
+```
+
+*Parity Distribution by Treatment Arm*
+
+```{r}
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
+library(here)
+
+# Read in the cleaned data
+promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+
+# Recode Parity into categories: "0", "1–2", "≥3"
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    Parity_cat = case_when(
+      parity == 0 ~ "0",
+      parity %in% c(1, 2) ~ "1–2",
+      parity >= 3 ~ "≥3"
+    ),
+    Parity_cat = factor(Parity_cat, levels = c("0", "1–2", "≥3"))
+  )
+
+# Create the bar graph for Parity by Treatment Arm
+ggplot(promo_data_clean, aes(x = Parity_cat, fill = study_arm)) +
+  geom_bar(position = "dodge") +
+  labs(title = "Parity Distribution by Treatment Arm",
+       x = "Parity Category (0, 1–2, ≥3)",
+       y = "Count",
+       fill = "Treatment Arm") +
+  theme_minimal()
+
+```
+
+*Figure 1: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes*
+
+```{r}
+# Create the composite adverse outcome variable
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
+  )
+
+# Fit the interaction model (ensuring that adverse_birth_outcome now exists)
+model_interaction <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm + 
+                           age_at_enrollment_years + gravidity + education_level,
+                         family = binomial(link = "logit"),
+                         data = promo_data_clean)
+
+# Generate predicted probabilities over the range of total malaria episodes by study arm
+pred <- ggeffect(model_interaction, terms = c("total_malaria_episodes [all]", "study_arm"))
+
+# Plot the predicted probabilities
+interaction_plot <- ggplot(pred, aes(x = x, y = predicted, color = group)) +
+  geom_line(size = 1.2) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
+  labs(title = "Differential Impact of IPTp Treatment on the Relationship\nBetween Malaria Episodes and Adverse Birth Outcomes",
+       x = "Total Malaria Episodes",
+       y = "Predicted Probability of Adverse Outcome",
+       color = "Treatment Arm",
+       fill = "Treatment Arm") +
+  theme_minimal()
+
+interaction_plot
+
+```
+
+*Research questions 1& 2*
+
+```{r}
 # Convert date columns to Date objects
@@
 
-# View a summary of the cleaned data
-summary(promo_data_clean)
-
-```
-
-```{r}
 # Ensure the 'data/clean' directory exists
@@
 ```{r}
-colnames(promo_data_clean)
+# Save the cleaned data (if not already saved)
+write.csv(promo_data_clean, here::here("data", "clean", "PROMO_Data_clean.csv"), row.names = FALSE)
+
+# Read in the cleaned data
+promo_data_clean <- read.csv(here::here("data", "clean", "PROMO_Data_clean.csv"))
+
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
+library(broom)
+
+### Data Preparation
+
+# 1. Create a composite adverse birth outcome variable
+#    Adverse outcome is defined as: 
+#    - Preterm birth: preterm_births_count > 0 
+#    OR stillbirth: stillbirth_bin == 1 
+#    OR low birth weight: birth_weight < 2.5 kg
+promo_data_clean <- promo_data_clean %>%
+  mutate(low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+         adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0))
+
+# 2. Create a subgroup variable for maternal age (<25 years vs. >=25 years)
+promo_data_clean <- promo_data_clean %>%
+  mutate(age_group = ifelse(age_at_enrollment_years < 25, "Young", "Older"))
+
+### Research Question 1: 
+### "Does the type of IPTp regimen modify the association between malaria episode frequency and adverse birth outcomes in Ugandan pregnant women?"
+
+# Fit a logistic regression model with an interaction between total malaria episodes and study_arm
+model_interaction <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm + 
+                           age_at_enrollment_years + gravidity + education_level,
+                         family = binomial(link = "logit"),
+                         data = promo_data_clean)
+
+# Summarize and tidy the model output (exponentiating coefficients to yield odds ratios)
+summary(model_interaction)
+tidy_model_interaction <- tidy(model_interaction, exponentiate = TRUE, conf.int = TRUE)
+print(tidy_model_interaction)
+
+### Research Question 2: 
+### "Among younger Ugandan pregnant women, is increased gravidity associated with a reduced risk of adverse birth outcomes?"
+
+# Subset the data for young mothers (age < 25)
+young_mothers <- promo_data_clean %>% filter(age_group == "Young")
+
+# Fit a logistic regression model within the young mothers subgroup 
+# to examine the association between gravidity and adverse birth outcomes
+model_gravidity_young <- glm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + education_level,
+                             family = binomial(link = "logit"),
+                             data = young_mothers)
+
+# Summarize and tidy the subgroup model output
+summary(model_gravidity_young)
+tidy_model_gravidity_young <- tidy(model_gravidity_young, exponentiate = TRUE, conf.int = TRUE)
+print(tidy_model_gravidity_young)
+
+### Additional Exploratory Plots (Optional)
+
+# Plot the distribution of total malaria episodes by study arm
+ggplot(promo_data_clean, aes(x = study_arm, y = total_malaria_episodes)) +
+  geom_boxplot(fill = "lightblue") +
+  labs(title = "Malaria Episode Frequency by Treatment Arm", 
+       x = "Treatment Arm", y = "Total Malaria Episodes")
+
+# Plot the proportion of adverse outcomes by gravidity among young mothers
+ggplot(young_mothers, aes(x = gravidity, fill = factor(adverse_birth_outcome))) +
+  geom_bar(position = "fill") +
+  labs(title = "Proportion of Adverse Outcomes by Gravidity (Young Mothers)",
+       x = "Gravidity", y = "Proportion", fill = "Adverse Outcome")
+
 ```
 
-# Exploratory Data Analysis
-```{r}
-# Create a histogram of total malaria episodes
-ggplot(data = promo_data_clean, aes(x = total_malaria_episodes)) +
-  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
-  labs(
-    title = "Distribution of Total Malaria Episodes",
-    x = "Total Malaria Episodes",
-    y = "Count"
-  )
+##################################################################################################################################################################
 
-```
+*Question 1: "Does the type of IPTp regimen modify the association between malaria episode frequency and adverse birth outcomes in Ugandan pregnant women?"*
 
@@
 
-# Create a scatter plot with regression lines, faceted by study arm
-ggplot(promo_data_clean, aes(x = total_malaria_episodes, y = child_1_birth_weight_kg)) +
-  geom_point(alpha = 0.7, color = "darkblue") +
-  geom_smooth(method = "lm", se = FALSE, color = "red") +
-  facet_wrap(~ study_arm) +
-  labs(
-    title = "Malaria Episodes vs. Birth Weight by Study Arm",
+# Create the composite adverse outcome variable:
+# adverse_birth_outcome = 1 if any of the following occur:
+# preterm birth (preterm_births_count > 0), stillbirth (stillbirth_bin == 1), or low birth weight (<2.5 kg)
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
+  )
+
+# Fit the logistic regression model with an interaction term
+model_interaction <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm +
+                           age_at_enrollment_years + gravidity + education_level,
+                         family = binomial(link = "logit"),
+                         data = promo_data_clean)
+
+# Display the model summary
+summary(model_interaction)
+
+# Tidy the model output (exponentiating coefficients to yield odds ratios)
+tidy_model <- tidy(model_interaction, exponentiate = TRUE, conf.int = TRUE)
+print(tidy_model)
+
+# Generate predicted probabilities over the range of total malaria episodes by treatment arm
+pred <- ggeffect(model_interaction, terms = c("total_malaria_episodes [all]", "study_arm"))
+
+# Create the interaction plot
+interaction_plot <- ggplot(pred, aes(x = x, y = predicted, color = group)) +
+  geom_line(size = 1.2) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
+  labs(title = "Differential Impact of IPTp Treatment on the Relationship\nBetween Malaria Episodes and Adverse Birth Outcomes",
        x = "Total Malaria Episodes",
-    y = "Child Birth Weight (kg)"
-  ) +
-  theme_minimal() +
-  theme(legend.position = "none")
+       y = "Predicted Probability of Adverse Outcome",
+       color = "Treatment Arm",
+       fill = "Treatment Arm") +
+  theme_minimal()
+
+interaction_plot
+
+```
+
+*Table 3: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes*
+
+```{r}
+# Tidy the model output with exponentiated coefficients (odds ratios)
+tidy_model <- tidy(model_interaction, exponentiate = TRUE, conf.int = TRUE)
+
+# Remove the intercept row and recode variable names using case_when()
+tidy_model_filtered <- tidy_model %>%
+  filter(term != "(Intercept)") %>%
+  mutate(
+    term = case_when(
+      term == "total_malaria_episodes" ~ "Total Malaria Episodes",
+      term == "study_armSP" ~ "Treatment Arm (SP)",
+      term == "age_at_enrollment_years" ~ "Age at Enrollment (years)",
+      term == "gravidity" ~ "Gravidity",
+      term == "education_levelSecondary" ~ "Secondary Education",
+      term == "education_levelTertiary" ~ "Tertiary Education",
+      term == "education_levelUniversity" ~ "University Education",
+      term == "total_malaria_episodes:study_armSP" ~ "Interaction: Malaria Episodes × SP",
+      TRUE ~ term
+    )
+  )
+
+# Round key values for clarity
+tidy_model_filtered <- tidy_model_filtered %>%
+  mutate(
+    estimate = round(estimate, 2),
+    std.error = round(std.error, 2),
+    statistic = round(statistic, 2),
+    p.value = round(p.value, 3),
+    conf.low = round(conf.low, 2),
+    conf.high = round(conf.high, 2)
+  )
+
+# Render the table with a blank header for the first column
+kable(
+  tidy_model_filtered,
+  format = "pandoc",
+  caption = "*Table X: Regression Results with VIFs*",
+  col.names = c("", "Odds Ratio", "Std. Error", "z value", "p-value", "95% CI Lower", "95% CI Upper")
+) %>% 
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+*Fit a logistic regression model without the interaction term*
+
+```{r}
+# Fit a logistic regression model without the interaction term
+model_no_interaction <- glm(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
+                              age_at_enrollment_years + gravidity + education_level,
+                            family = binomial(link = "logit"),
+                            data = promo_data_clean)
+
+# Perform a Likelihood Ratio Test comparing models with and without the interaction
+lrt <- anova(model_no_interaction, model_interaction, test = "LRT")
+print(lrt)
+
+# Check for multicollinearity using the VIF (Variance Inflation Factor)
+library(car)
+vif_values <- vif(model_interaction)
+print(vif_values)
+
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+library(dplyr)
+library(knitr)
+library(kableExtra)
+library(here)
+
+# Generate Analysis of Deviance Table (Likelihood Ratio Test)
+anova(model_no_interaction, model_interaction, test = "LRT")
+anova_df <- as.data.frame(anova_out)
+anova_df$Model <- c("Model 1 (No Interaction)", "Model 2 (With Interaction)")
+anova_df <- anova_df %>% 
+  select(Model, `Resid. Df`, `Resid. Dev`, Df, Deviance, `Pr(>Chi)`)
+colnames(anova_df) <- c("Model", "Residual Df", "Residual Dev", "Df", "Deviance", "p-value")
+
+kable(
+  anova_df,
+  format = "pandoc",
+  caption = "*Table X: Analysis of Deviance Comparing Models With and Without Interaction*"
+) %>% 
+  kable_styling(full_width = FALSE, position = "center")
 
@@
 ```{r}
-# Read in the cleaned data
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+library(dplyr)
+library(knitr)
+library(kableExtra)
+library(car)
+library(here)
 
-## 1. Univariate Analysis
+# Compute Variance Inflation Factors (VIF) for Model 2 (with interaction)
+vif_vals <- vif(model2)
+vif_df <- as.data.frame(vif_vals)
+vif_df$Variable <- rownames(vif_vals)
+vif_df <- vif_df %>% rename(`Df (VIF)` = Df)
+vif_df <- vif_df %>% 
+  select(Variable, GVIF, `Df (VIF)`, `GVIF^(1/(2*Df))`) %>%
+  mutate(
+    Variable = case_when(
+      Variable == "total_malaria_episodes" ~ "Total Malaria Episodes",
+      Variable == "study_arm" ~ "Treatment Arm",
+      Variable == "age_at_enrollment_years" ~ "Age at Enrollment",
+      Variable == "gravidity" ~ "Gravidity",
+      Variable == "education_level" ~ "Education Level",
+      Variable == "total_malaria_episodes:study_arm" ~ "Interaction: Malaria Episodes × SP",
+      TRUE ~ Variable
+    )
+  )
 
-# Histogram: Distribution of Age at Enrollment
-ggplot(promo_data_clean, aes(x = age_at_enrollment_years)) +
-  geom_histogram(binwidth = 5, fill = "lightblue", color = "black") +
-  labs(title = "Distribution of Age at Enrollment",
-       x = "Age (years)", y = "Count")
-
-# Histogram: Distribution of Gestational Age at Enrollment
-# Adjust the variable name if necessary after cleaning (e.g., gestational_age_at_enrollment_weeks)
-ggplot(promo_data_clean, aes(x = gestational_age_at_enrollment_weeks)) +
-  geom_histogram(binwidth = 2, fill = "lightgreen", color = "black") +
-  labs(title = "Distribution of Gestational Age at Enrollment",
-       x = "Gestational Age (weeks)", y = "Count")
-
-# Histogram: Distribution of Total Malaria Episodes
-ggplot(promo_data_clean, aes(x = total_malaria_episodes)) +
-  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
-  labs(title = "Distribution of Total Malaria Episodes",
-       x = "Total Malaria Episodes", y = "Count")
-
-# Bar Plot: Counts of Study Arm
-ggplot(promo_data_clean, aes(x = study_arm)) +
-  geom_bar(fill = "skyblue") +
-  labs(title = "Counts by Study Arm", x = "Study Arm", y = "Count") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-# Bar Plot: Counts of Adverse Outcomes (assumed to be coded as 0/1)
-ggplot(promo_data_clean, aes(x = factor(adverse_outcome))) +
-  geom_bar(fill = "salmon") +
-  labs(title = "Counts of Adverse Outcomes",
-       x = "Adverse Outcome (0 = No, 1 = Yes)", y = "Count")
-
-## 2. Bivariate Analysis
-
-# Boxplot: Total Malaria Episodes by Adverse Outcome
-ggplot(promo_data_clean, aes(x = factor(adverse_outcome), y = total_malaria_episodes)) +
-  geom_boxplot(fill = "lightgray") +
-  labs(title = "Malaria Episodes by Adverse Outcome",
-       x = "Adverse Outcome (0 = No, 1 = Yes)",
-       y = "Total Malaria Episodes")
-
-# Boxplot: Total Malaria Episodes by Study Arm
-ggplot(promo_data_clean, aes(x = study_arm, y = total_malaria_episodes)) +
-  geom_boxplot(fill = "lightblue") +
-  labs(title = "Malaria Episodes by Study Arm",
-       x = "Study Arm",
-       y = "Total Malaria Episodes") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-# Contingency Table: Study Arm vs. Adverse Outcome
-study_adverse_table <- table(promo_data_clean$study_arm, promo_data_clean$adverse_outcome)
-print(study_adverse_table)
-
-# Stacked Bar Plot: Proportion of Adverse Outcomes by Study Arm
-ggplot(promo_data_clean, aes(x = study_arm, fill = factor(adverse_outcome))) +
-  geom_bar(position = "fill") +
-  labs(title = "Proportion of Adverse Outcomes by Study Arm",
-       x = "Study Arm", y = "Proportion", fill = "Adverse Outcome") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-## 3. Multivariate/Correlation Analysis
-
-# Select continuous variables of interest
-continuous_vars <- promo_data_clean %>%
-  select(age_at_enrollment_years, gestational_age_at_enrollment_weeks, 
-         total_malaria_episodes, child_1_birth_weight_kg)
-
-# Compute correlation matrix (using complete observations)
-cor_matrix <- cor(continuous_vars, use = "complete.obs")
-print(cor_matrix)
-
-# Visualize the correlation matrix
-corrplot(cor_matrix, method = "circle", type = "upper")
-
-## 4. Visual Exploration with Scatter Plots
-
-# Scatter Plot: Total Malaria Episodes vs. Child Birth Weight, with regression line and faceting by Study Arm
-ggplot(promo_data_clean, aes(x = total_malaria_episodes, y = child_1_birth_weight_kg)) +
-  geom_point(alpha = 0.7) +
-  geom_smooth(method = "lm", se = FALSE, color = "red") +
-  facet_wrap(~ study_arm) +
-  labs(title = "Malaria Episodes vs. Birth Weight by Study Arm",
-       x = "Total Malaria Episodes",
-       y = "Child Birth Weight (kg)")
+kable(
+  vif_df,
+  format = "pandoc",
+  caption = "*Table Y: Variance Inflation Factors for Model 2 (With Interaction)*"
+) %>% 
+  kable_styling(full_width = FALSE, position = "center")
 
@@
 
+##################################################################################################################################################################
+
+## Research Question 2:
+
+*"Among younger (women under 25 years old) Ugandan pregnant women, is increased gravidity associated with a reduced risk of adverse birth outcomes?"*
+
 ```{r}
-# Install the tableone package if not already installed
-if (!requireNamespace("tableone", quietly = TRUE)) {
-  install.packages("tableone")
-}
+# Subset the data to include only women under 25 years of age
+promo_data_young <- promo_data_clean %>%
+  filter(age_at_enrollment_years < 25)
 
-library(tableone)
-library(here)
+# Create the composite adverse birth outcome variable if not already done
+promo_data_young <- promo_data_young %>%
+  mutate(
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
+  )
 
-# Read in the cleaned data
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+# Fit the logistic regression model
+model_gravidity <- glm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + study_arm + education_level,
+                       family = binomial(link = "logit"),
+                       data = promo_data_young)
 
-# Define the variables for the demographics and baseline characteristics table.
-vars <- c("age_at_enrollment_years", 
-          "gestational_age_at_enrollment_weeks", 
-          "education_level", 
-          "gravidity", 
-          "parity")
+# Display the model summary
+summary(model_gravidity)
 
-strata <- "study_arm"
+# Tidy the model output, exponentiating coefficients to yield odds ratios
+tidy_model_gravidity <- tidy(model_gravidity, exponentiate = TRUE, conf.int = TRUE)
+print(tidy_model_gravidity)
 
-# Create the TableOne object
-table1 <- CreateTableOne(vars = vars, strata = strata, data = promo_data_clean, factorVars = c("education_level", "study_arm"))
-
-# Print the table with all levels displayed
-print(table1, showAllLevels = TRUE)
+# Optionally, plot predicted probabilities for gravidity
+library(ggeffects)
+predicted_probs <- ggeffect(model_gravidity, terms = "gravidity")
+ggplot(predicted_probs, aes(x = x, y = predicted)) +
+  geom_line(size = 1.2) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "blue") +
+  labs(title = "Predicted Probability of Adverse Outcome by Gravidity (Age < 25)",
+       x = "Gravidity",
+       y = "Predicted Probability") +
+  theme_minimal()
 
@@
 
+*Logistic Regression Model Fitting and Summary*
+
 ```{r}
-# Create table data using tribble.
-# Each row must have exactly 4 cells.
-table_data <- tribble(
-  ~Characteristic,                         ~Subcategory,                                         ~`Monthly SP (n=391)`,                                         ~`Monthly DP (n=391)`,
-  # Age at Enrollment
-  "Age, years",                            "",                                                   "23 (19–27)",                                                  "23 (19–27)",
-  # Gestational Age at Enrollment overall and subcategories
-  "Gestational Age at Enrollment (weeks)", "",                                                   "15.4 (13.3–17.6)",                                              "15.0 (13.4–17.1)",
-  "",                                      "12–16",                                              "66 (17%)",                                                    "62 (15.8%)",
-  "",                                      "17–20",                                              "164 (42.2%)",                                                 "197 (50.1%)",
-  "",                                      "21–24",                                              "133 (34.2%)",                                                 "118 (30%)",
-  # Gravidity overall and subcategories
-  "Gravidity",                             "",                                                   "1.2 (0.9–1.7)",                                                 "1.1 (0.8–1.6)",
-  "",                                      "1",                                                  "102 (26%)",                                                   "95 (24%)",
-  "",                                      "2–3",                                                "180 (46%)",                                                   "190 (49%)",
-  "",                                      "≥4",                                                 "109 (28%)",                                                   "106 (27%)",
-  # Parity overall and subcategories
-  "Parity",                                "",                                                   "2.43 (1.67)",                                                   "2.43 (1.65)",
-  "",                                      "0",                                                  "140 (36%)",                                                   "145 (37%)",
-  "",                                      "1–2",                                                "170 (44%)",                                                   "165 (42%)",
-  "",                                      "≥3",                                                 "81 (20%)",                                                    "81 (21%)",
-  # Total Malaria Episodes overall
-  "Total Malaria Episodes (mean (SD))",    "",                                                   "1.06 (1.03)",                                                 "0.98 (0.95)",
-  # Maternal Nutritional Status/BMI overall
-  "Maternal Nutritional Status/BMI (mean (SD))", "",                                                "22.5 (3.2)",                                                  "22.8 (3.1)",
-  # History of Previous Malaria Episodes overall
-  "History of Previous Malaria Episodes (%)", "",                                                 "40 (10%)",                                                    "35 (9%)",
-  # Maternal Education Level overall and subcategories
-  "Maternal Education Level (%)",          "",                                                   "",                                                              "",
-  "",                                      "A and O Level",                                      "194 (49.4%)",                                                 "187 (48.1%)",
-  "",                                      "Primary",                                            "68 (17.3%)",                                                  "71 (18.3%)",
-  "",                                      "Tertiary",                                           "65 (16.5%)",                                                  "61 (15.7%)",
-  "",                                      "University",                                         "66 (16.8%)",                                                  "70 (18.0%)"
+# Subset the data to include only women under 25 years of age and create the composite adverse outcome variable
+promo_data_young <- promo_data_clean %>%
+  filter(age_at_enrollment_years < 25) %>%
+  mutate(
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
   )
 
-# Render the table using kableExtra
-kbl(
-  table_data,
-  col.names = c("Characteristic", "Subcategory", "Monthly SP (n=391)", "Monthly DP (n=391)"),
-  booktabs = TRUE,
-  align = "lccc"
-) %>%
-  add_header_above(c("N = 782" = 4)) %>%
-  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
-  # Merge sub-rows for Gestational Age at Enrollment:
-  # Overall row is row 2; subcategory rows are rows 3-5
-  pack_rows("Gestational Age at Enrollment (weeks)", start_row = 2, end_row = 5, indent = TRUE) %>%
-  # Merge sub-rows for Gravidity:
-  # Overall row is row 6; subcategory rows are rows 7-9
-  pack_rows("Gravidity", start_row = 6, end_row = 9, indent = TRUE) %>%
-  # Merge sub-rows for Parity:
-  # Overall row is row 10; subcategory rows are rows 11-13
-  pack_rows("Parity", start_row = 10, end_row = 13, indent = TRUE) %>%
-  # Merge sub-rows for Maternal Education Level:
-  # Overall row is row 17; subcategory rows are rows 18-21
-  pack_rows("Maternal Education Level (%)", start_row = 17, end_row = 21, indent = TRUE)
+# Fit the logistic regression model for Research Question 2
+model_gravidity <- glm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + study_arm + education_level,
+                       family = binomial(link = "logit"),
+                       data = promo_data_young)
+
+# Display the model summary
+summary(model_gravidity)
+
+# Tidy the model output with exponentiated coefficients (odds ratios)
+tidy_model_gravidity <- tidy(model_gravidity, exponentiate = TRUE, conf.int = TRUE)
+print(tidy_model_gravidity)
 
@@
 
+*Predicted Probability of Adverse Outcome by Gravidity (Age \< 25)*
+
 ```{r}
-table(promo_data_clean$stillbirth_bin, useNA = "always")
-table(promo_data_clean$stillbirth_bin, useNA = "always")
-table(promo_data_clean$stillbirth)
+# Load necessary package for generating predicted effects
+library(ggeffects)
+
+# Generate predicted probabilities over the range of gravidity
+predicted_probs <- ggeffect(model_gravidity, terms = "gravidity")
+
+# Create a scatter/line plot of predicted probabilities by gravidity
+ggplot(predicted_probs, aes(x = x, y = predicted)) +
+  geom_line(size = 1.2) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "blue") +
+  labs(title = "Predicted Probability of Adverse Outcome by Gravidity (Age < 25)",
+       x = "Gravidity",
+       y = "Predicted Probability") +
+  theme_minimal()
 
@@
 
-## Results
+In my logistic regression model, I found that gravidity has a statistically significant protective effect (OR = 0.857, 95% CI: 0.737–0.995, p = 0.044), suggesting that each additional pregnancy among young women under 25 reduces the odds of adverse birth outcomes.
+In contrast, total malaria episodes, treatment arm, and education level were not statistically significant predictors.
+This indicates that, within this subgroup, prior pregnancy experience is the key factor influencing birth outcomes, while other factors seem to have little effect.
+
+In my model, gravidity has an odds ratio of 0.857 (95% CI: 0.737–0.995, p = 0.044), indicating that each additional pregnancy among women under 25 reduces the odds of adverse birth outcomes by roughly 14%.
+Other predictors (total malaria episodes, treatment arm, and education level) were not statistically significant, suggesting that gravidity is the key protective factor in this subgroup.
 
 ```{r}
-# --- Load libraries ---
-library(dplyr)
-library(broom)
-library(logistf)
-library(knitr)
-library(kableExtra)
-library(tibble)
+# Load necessary package for ROC analysis
+library(pROC)
 
-# --- Data preparation ---
-promo_data_clean <- promo_data_clean %>%
-  mutate(
-    preterm_birth = ifelse(gestational_age_at_delivery_weeks < 37, 1, 0),
-    stillbirth_bin = ifelse(stillbirth == "Yes", 1, 0)
+# Generate predicted probabilities using the logistic regression model (model_gravidity)
+promo_data_young$predicted_prob <- predict(model_gravidity, type = "response")
+
+# Create the ROC curve
+roc_obj <- roc(promo_data_young$adverse_birth_outcome, promo_data_young$predicted_prob)
+plot(roc_obj, col = "blue", lwd = 2, main = "ROC Curve for Model: Gravidity in Women < 25")
+auc_value <- auc(roc_obj)
+print(paste("AUC:", round(auc_value, 2)))
+
+# Load necessary package for calibration plot
+library(caret)
+
+# Create a calibration plot
+calibration_data <- data.frame(
+  observed = factor(promo_data_young$adverse_birth_outcome, levels = c(0,1)),
+  predicted = promo_data_young$predicted_prob
 )
 
-# --- Parsimonious Logistic Regression for Preterm Birth ---
-model_preterm_pars <- glm(
-  preterm_birth ~ total_malaria_episodes_during_pregnancy +
-    age_at_enrollment_years + gravidity,
-  data = promo_data_clean,
-  family = binomial(link = "logit")
+# Use the calibration function from the caret package
+cal_plot <- calibration(observed ~ predicted, data = calibration_data, class = "1")
+plot(cal_plot, main = "Calibration Plot for Model: Gravidity in Women < 25")
+
+```
+
+I see that the ROC curve for my model is close to the diagonal, with an AUC only slightly above 0.5.
+This tells me that the model doesn't have strong discriminative ability for predicting adverse outcomes in women under 25.
+Additionally, the calibration plot shows that my predicted probabilities often stray from the ideal diagonal—especially in the mid-range—indicating that my model’s risk estimates don't consistently match the observed rates.
+Overall, while gravidity is statistically significant, my model as a whole isn't very effective at distinguishing between those who experience adverse outcomes and those who don't, and its probability estimates need improvement.
+
+*Train/Test Split and Logistic Regression (Interaction Model)* *Research Question 1*
+
+ML (Research Question 1.
+I built and evaluated logistic regression, random forest, and boosting models (using cross-validation and train/test splits) to assess how IPTp regimen modifies the impact of malaria episodes on adverse birth outcomes.)
+
+Research Question 1, which examines whether the IPTp regimen modifies the association between malaria episodes and adverse birth outcomes.
+
+Below is an example of how I can extend my analysis by performing a train/test split and comparing a couple of machine learning classification models.
+For Research Question 1 (the logistic regression with the interaction term), I'll split the data into training (70%) and test (30%) sets and then fit both a logistic regression and a random forest classifier.
+This lets me assess model performance on unseen data using metrics like ROC AUC.
+
+```{r}
+# Set seed and perform a 70/30 train/test split (stratified on adverse outcomes)
+# Load necessary package for data splitting
+library(rsample)
+
+# Set seed and perform a 70/30 train/test split (stratified on adverse outcomes)
+set.seed(1234)
+split_data <- initial_split(promo_data_clean, prop = 0.7, strata = adverse_birth_outcome)
+train_data <- training(split_data)
+test_data  <- testing(split_data)
+
+# Fit the logistic regression model with an interaction term on training data
+model_interaction <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm +
+                           age_at_enrollment_years + gravidity + education_level,
+                         family = binomial(link = "logit"),
+                         data = train_data)
+
+# Summarize the model
+summary(model_interaction)
+
+# Generate predictions on the test set and compute the ROC curve
+library(pROC)
+pred_test_logit <- predict(model_interaction, newdata = test_data, type = "response")
+roc_logit <- roc(test_data$adverse_birth_outcome, pred_test_logit)
+plot(roc_logit, col = "blue", lwd = 2, main = "ROC Curve: Logistic Regression")
+auc_logit <- auc(roc_logit)
+print(paste("Logistic Regression AUC:", round(auc_logit, 2)))
+
+```
+
+*Train/Test Split and Random Forest Model* which fits a random forest model and plots its ROC curve, is also used for Research Question 1 as an alternative ML approach
+
+```{r}
+# Ensure the outcome is a factor in train_data
+train_data <- train_data %>%
+  mutate(adverse_birth_outcome = factor(adverse_birth_outcome))
+
+# Build a recipe for the random forest model (no inline interaction here)
+rf_recipe <- recipe(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
+                      age_at_enrollment_years + gravidity + education_level,
+                    data = train_data) %>%
+  step_dummy(all_nominal_predictors())
+
+# Create a workflow for the random forest model
+rf_workflow <- workflow() %>%
+  add_model(rf_model) %>%
+  add_recipe(rf_recipe)
+
+# Fit the random forest model on training data
+rf_fit <- rf_workflow %>% fit(data = train_data)
+
+# Generate predicted probabilities on the test set
+rf_preds <- predict(rf_fit, new_data = test_data, type = "prob")
+
+# Compute ROC for the random forest using the probability for class "1"
+library(pROC)
+roc_rf <- roc(test_data$adverse_birth_outcome, rf_preds$.pred_1)
+plot(roc_rf, col = "red", lwd = 2, main = "ROC Curve: Random Forest")
+auc_rf <- auc(roc_rf)
+print(paste("Random Forest AUC:", round(auc_rf, 2)))
+
+```
+
+-   Model Comparison and Interpretation\*
+
+```{r}
+# Compare ROC curves
+plot(roc_logit, col = "blue", lwd = 2, main = "ROC Comparison: Logistic Regression vs. Random Forest")
+lines(roc_rf, col = "red", lwd = 2)
+legend("bottomright", legend = c("Logistic Regression", "Random Forest"),
+       col = c("blue", "red"), lwd = 2)
+
+# Optionally, you could also compare other metrics like accuracy or calibration on the test set.
+
+```
+
+I started by performing a 70/30 train/test split, stratified on the adverse birth outcome.
+For Research Question 1, I fit a logistic regression model with an interaction term on the training set, generated predictions on the test set, and plotted the ROC curve (with an AUC computed to assess discriminative performance).
+Then, I fit a random forest classifier (converting factor predictors to dummies in the recipe) and similarly evaluated its performance on the test set using ROC analysis.
+Finally, I overlaid both ROC curves to compare the models directly.
+This approach provides an "honest" assessment of model performance on unseen data and allows me to explore additional machine learning methods for addressing my research question.
+
+From these plots, I see that both the logistic regression (blue) and random forest (red) models produce ROC curves only slightly above the diagonal, indicating that neither model has strong discriminative ability for predicting adverse outcomes in this dataset.
+In the combined plot, the logistic regression curve sits just above the random forest curve, suggesting it might be marginally better, but the difference is small—both are near an AUC of around 0.55–0.60.
+Overall, these results tell me that, with my current predictors and data, neither model reliably distinguishes between those who experience adverse birth outcomes and those who do not.
+
+```{r}
+# Load necessary packages (if not already loaded)
+library(tidymodels)
+library(xgboost)
+
+# Define a boosting model for classification (using boost_tree)
+boost_model <- boost_tree(
+  mode = "classification",
+  trees = 500,
+  learn_rate = tune(),
+  tree_depth = tune()
+) %>%
+  set_engine("xgboost")
+
+# Build a recipe (convert factor predictors to dummy variables)
+boost_recipe <- recipe(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
+                         age_at_enrollment_years + gravidity + education_level,
+                       data = train_data) %>%
+  step_dummy(all_nominal_predictors())
+
+# Create a workflow
+boost_workflow <- workflow() %>%
+  add_model(boost_model) %>%
+  add_recipe(boost_recipe)
+
+# Create a tuning grid for learning rate and tree depth
+boost_grid <- grid_regular(
+  learn_rate(range = c(0.01, 0.3)),
+  tree_depth(range = c(3, 8)),
+  levels = 5
 )
 
-# --- Parsimonious Firth Logistic Regression for Stillbirth ---
-model_stillbirth_pars <- logistf(
-  stillbirth_bin ~ total_malaria_episodes_during_pregnancy +
-    age_at_enrollment_years + gravidity,
-  data = promo_data_clean
+# Create 10-fold cross-validation folds from the training data
+cv_folds <- vfold_cv(train_data, v = 10)
+
+# Perform 10-fold cross-validation tuning using the boosting model workflow
+boost_tune <- tune_grid(
+  boost_workflow,
+  resamples = cv_folds,
+  grid = boost_grid
 )
 
-# --- Extract Preterm results ---
-summary_preterm_pars <- tidy(model_preterm_pars, exponentiate = TRUE, conf.int = TRUE)
+# Select best parameters based on ROC AUC
+best_boost <- select_best(boost_tune, metric = "roc_auc")
 
-# --- Create Preterm Table ---
-table_preterm <- summary_preterm_pars %>%
-  select(term, estimate, conf.low, conf.high, p.value) %>%
-  mutate(across(where(is.numeric), ~ round(., 2))) %>%
-  mutate(term = recode(term,
-                       "(Intercept)" = "Intercept",
-                       "total_malaria_episodes_during_pregnancy" = "Total malaria episodes during pregnancy",
-                       "age_at_enrollment_years" = "Age at enrollment (years)",
-                       "gravidity" = "Gravidity"
-                       ))
+# Finalize the workflow with best parameters
+final_boost <- finalize_workflow(boost_workflow, best_boost)
 
-kable(table_preterm,
-      caption = "**Table 4a. Parsimonious Logistic Regression for Preterm Birth (N = 782)**",
-      col.names = c("Exposure and Adjustment Variables", "Adjusted Odds Ratio (aOR)", "95% CI (Lower)", "95% CI (Upper)", "p-value")) %>%
-  kable_styling(full_width = FALSE, position = "center")
+# Fit the final boosting model on the training data
+boost_fit <- final_boost %>% fit(data = train_data)
 
-# --- Create Stillbirth Table ---
-table_stillbirth <- data.frame(
-  `Exposure and Adjustment Variables` = c(
-    "Total malaria episodes during pregnancy",
-    "Age at enrollment (years)",
-    "Gravidity"
-  ),
-  `Adjusted Odds Ratio (aOR)` = round(exp(model_stillbirth_pars$coefficients[-1]), 2),
-  `95% CI (Lower)` = round(exp(model_stillbirth_pars$ci.lower[-1]), 2),
-  `95% CI (Upper)` = round(exp(model_stillbirth_pars$ci.upper[-1]), 2),
-  `p-value` = signif(model_stillbirth_pars$prob[-1], 3)
+# Generate predicted probabilities on the test set
+boost_preds <- predict(boost_fit, new_data = test_data, type = "prob")
+
+# Compute ROC for the boosting model using the probability for class "1"
+roc_boost <- roc(test_data$adverse_birth_outcome, boost_preds$.pred_1)
+plot(roc_boost, col = "green", lwd = 2, main = "ROC Curve: Boosting Model")
+auc_boost <- auc(roc_boost)
+print(paste("Boosting Model AUC:", round(auc_boost, 2)))
+
+```
+
+The ROC curve for the boosting model rises more clearly above the diagonal line than the previous models, suggesting it does a better job distinguishing between those who experience adverse birth outcomes and those who do not.
+Although it’s still not perfect, the curve indicates an improvement in predictive performance compared to the logistic regression and random forest models, meaning the boosting model likely has a higher AUC and provides a more accurate risk estimate for adverse outcomes in this dataset.
+
+*ML for research question 2*
+
+*Subset Data and Create Outcome Variable & Fit Logistic Regression Model*
+
+```{r}
+# Subset data to include only women under 25 years
+promo_data_young <- promo_data_clean %>%
+  filter(age_at_enrollment_years < 25)
+
+# Create the composite adverse outcome variable:
+# adverse_birth_outcome = 1 if any of the following occur:
+# preterm birth (preterm_births_count > 0), stillbirth (stillbirth_bin == 1), or low birth weight (<2.5 kg)
+promo_data_young <- promo_data_young %>%
+  mutate(
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
   )
 
-kable(table_stillbirth,
-      caption = "**Table 4b. Parsimonious Firth Logistic Regression for Stillbirth (N = 782)**") %>%
-  kable_styling(full_width = FALSE, position = "center")
+# Fit the logistic regression model for women under 25
+model_gravidity <- glm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + study_arm + education_level,
+                       family = binomial(link = "logit"),
+                       data = promo_data_young)
 
-# --- Poisson Regression Model (Malaria Episode Rate by Treatment Arm) ---
-model_poisson <- glm(
-  total_malaria_episodes_during_pregnancy ~ dp_treatment + age_at_enrollment_years + gravidity + education_level,
-  data = promo_data_clean,
-  family = poisson(link = "log")
-)
-
-summary_poisson <- tidy(model_poisson, exponentiate = TRUE, conf.int = TRUE)
-
-# --- Create Poisson Table ---
-table_poisson <- summary_poisson %>%
-  select(term, estimate, conf.low, conf.high, p.value) %>%
-  mutate(across(where(is.numeric), ~ round(., 2))) %>%
-  mutate(term = recode(term,
-                       "(Intercept)" = "Intercept",
-                       "dp_treatment" = "DP vs SP Treatment Arm",
-                       "age_at_enrollment_years" = "Age at enrollment (years)",
-                       "gravidity" = "Gravidity",
-                       "education_levelSecondary" = "Education: Secondary",
-                       "education_levelTertiary" = "Education: Tertiary",
-                       "education_levelUniversity" = "Education: University"
-                       ))
-
-kable(table_poisson,
-      caption = "**Table 4c. Poisson Regression – Malaria Episode Incidence by Treatment Arm (N = 782)**") %>%
-  kable_styling(full_width = FALSE, position = "center")
-
-# --- Create Summary Table ---
-summary_results <- tribble(
-  ~Outcome, ~Exposure, ~Effect_Measure, ~Estimate, ~`95% CI`, ~`p-value`,
-  "Preterm Birth", "Total malaria episodes during pregnancy", "Adjusted Odds Ratio (aOR)", "1.04", "0.95–1.14", "0.401",
-  "Stillbirth", "Total malaria episodes during pregnancy", "Adjusted Odds Ratio (aOR)", "0.96", "0.83–1.10", "0.559",
-  "Malaria Episode Incidence", "DP vs SP Treatment Arm", "Incidence Rate Ratio (IRR)", "1.02", "0.92–1.12", "0.756"
-)
-
-kable(summary_results,
-      caption = "**Table 4d. Summary of Parsimonious Regression Analyses for Primary Outcomes**") %>%
-  kable_styling(full_width = FALSE, position = "center")
+# Display the model summary and tidy output with exponentiated coefficients (odds ratios)
+summary(model_gravidity)
+tidy_model_gravidity <- tidy(model_gravidity, exponentiate = TRUE, conf.int = TRUE)
+print(tidy_model_gravidity)
 
 ```
+
+-   Plot Predicted Probabilities by Gravidity\*
+
+```{r}
+# Load necessary package for generating predicted effects
+library(ggeffects)
+
+# Generate predicted probabilities over the range of gravidity
+predicted_probs <- ggeffect(model_gravidity, terms = "gravidity")
+
+# Create a line plot of predicted probabilities with confidence intervals
+ggplot(predicted_probs, aes(x = x, y = predicted)) +
+  geom_line(size = 1.2) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "blue") +
+  labs(title = "Predicted Probability of Adverse Outcome by Gravidity (Age < 25)",
+       x = "Gravidity", y = "Predicted Probability") +
+  theme_minimal()
+
+```
+
+In this subgroup of women under 25, gravidity is the only statistically significant predictor: each additional pregnancy lowers the odds of an adverse birth outcome (OR ≈ 0.86, 95% CI: 0.74–0.995, p = 0.044).
+As shown in the predicted probability plot, the likelihood of an adverse outcome steadily drops from roughly 60% at gravidity = 1 to about 45% at gravidity = 5.
+The other predictors—total malaria episodes, treatment arm, and education level—do not significantly change the risk in this subgroup, suggesting that prior pregnancy experience is the key factor driving improved outcomes for younger women.

### 2025-04-03 code/eda-code/eda_manuscript.qmd
update

---
title: " EDA_ Manuscript"
date: "2025-02-23"
output: html_document
---

```{r}
# Load Required Libraries
# Essential libraries for data handling, visualization, and analysis

library(here)             # Manage file paths
library(dplyr)            # Data manipulation
library(tidyverse)        # Data wrangling and visualization
library(janitor)          # Clean column names
library(skimr)            # Quick data summary
library(lubridate)        # Working with dates
library(readr)            # Reading CSV files
library(ggplot2)          # Visualization
library(ggpubr)           # Publication-ready plots
library(forcats)          # Categorical variables handling

# Table and reporting packages
library(gtsummary)        # Summary tables
library(gt)               # Table formatting
library(knitr)            # Table rendering
library(kableExtra)       # Table customization

# Analysis-specific packages
library(Amelia)           # Missing data visualization
library(pwr)              # Power analysis
library(DiagrammeR)       # Diagrams and flowcharts
library(survival)         # Survival analysis
library(survminer)        # Survival plots
library(ggeffects)        # Marginal effects for regression models
library(broom)            # Tidy model outputs
library(rsample)          # Resampling and splitting datasets

```

*Data Import*
```{r}
# Data Import & Initial Inspection 
# load the Dataset
PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"))

promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
```

*Initial Inspection*

```{r}
# Check column names
colnames(promo_data_clean)

# Basic structure
str(promo_data_clean)

# Quick look at first few rows
head(promo_data_clean)

# Overview of missing data
skimr::skim(promo_data_clean)

# Count total missing values per column
promo_data_clean %>% 
  summarise(across(everything(), ~sum(is.na(.))))

```

```{r}
colnames(promo_data_clean)
```

```{r}
names(promo_data_clean)
```

*Summary Statistics*
Numeric Variables
```{r}
# Summary of numeric variables
promo_data_clean %>% 
  select(where(is.numeric)) %>% 
  summary()

```

*Categorical Variables*
```{r}
# Frequency tables for categorical variables
promo_data_clean %>% 
  select(where(is.character)) %>% 
  map(~table(.))

```

*Missing Data Visualization*
```{r}
# Visualizing missing data pattern
naniar::vis_miss(promo_data_clean)

```

*Univariate Plots*
Numeric Variables
```{r}
# Histograms of numeric variables
promo_data_clean %>% 
  select(where(is.numeric)) %>% 
  gather() %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "black") +
  facet_wrap(~key, scales = "free") +
  theme_minimal()

```

*Categorical Variables*
```{r}
# Bar plots of categorical variables
promo_data_clean %>% 
  select(where(is.character)) %>% 
  gather() %>% 
  ggplot(aes(x = value)) +
  geom_bar(fill = "lightblue", color = "black") +
  facet_wrap(~key, scales = "free") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

*Correlation Matrix (Numeric Variables)*
```{r}
# Correlation plot
numeric_vars <- promo_data_clean %>% 
  select(where(is.numeric))

cor_matrix <- cor(numeric_vars, use = "pairwise.complete.obs")

corrplot::corrplot(cor_matrix, method = "circle", type = "lower")

```

*Group-wise Summary*
```{r}
# Grouped summary by treatment group (replace 'treatment_group' with actual variable)
promo_data_clean %>% 
  group_by(treatment_group) %>% 
  summarise(across(where(is.numeric), list(mean = mean, sd = sd), na.rm = TRUE))

```

### 2025-04-05 code/analysis-code/Statistical_Analysis.qmd
Project update

@@
 
-*Boxplot of Birthweight by IPTp Treatment Arm*
-
-```{r}
-ggplot(promo_data_clean, aes(x = study_arm, y = birth_weight, fill = study_arm)) +
-  geom_boxplot(alpha = 0.7) +
-  labs(title = "Birthweight by IPTp Treatment Arm",
-       x = "Treatment Arm",
-       y = "Birthweight (kg)") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-```
-
 *Bar Graphs for Gravidity and Parity*
@@
 promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
-
-```
-
-```{r}
-# Save the cleaned data (if not already saved)
-write.csv(promo_data_clean, here::here("data", "clean", "PROMO_Data_clean.csv"), row.names = FALSE)
-
-# Read in the cleaned data
-promo_data_clean <- read.csv(here::here("data", "clean", "PROMO_Data_clean.csv"))
-
-# Load necessary libraries
-library(dplyr)
-library(ggplot2)
-library(broom)
-
-### Data Preparation
-
-# 1. Create a composite adverse birth outcome variable
-#    Adverse outcome is defined as: 
-#    - Preterm birth: preterm_births_count > 0 
-#    OR stillbirth: stillbirth_bin == 1 
-#    OR low birth weight: birth_weight < 2.5 kg
-promo_data_clean <- promo_data_clean %>%
-  mutate(low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
-         adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0))
-
-# 2. Create a subgroup variable for maternal age (<25 years vs. >=25 years)
-promo_data_clean <- promo_data_clean %>%
-  mutate(age_group = ifelse(age_at_enrollment_years < 25, "Young", "Older"))
-
-### Research Question 1: 
-### "Does the type of IPTp regimen modify the association between malaria episode frequency and adverse birth outcomes in Ugandan pregnant women?"
-
-# Fit a logistic regression model with an interaction between total malaria episodes and study_arm
-model_interaction <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm + 
-                           age_at_enrollment_years + gravidity + education_level,
-                         family = binomial(link = "logit"),
-                         data = promo_data_clean)
-
-# Summarize and tidy the model output (exponentiating coefficients to yield odds ratios)
-summary(model_interaction)
-tidy_model_interaction <- tidy(model_interaction, exponentiate = TRUE, conf.int = TRUE)
-print(tidy_model_interaction)
-
-### Research Question 2: 
-### "Among younger Ugandan pregnant women, is increased gravidity associated with a reduced risk of adverse birth outcomes?"
-
-# Subset the data for young mothers (age < 25)
-young_mothers <- promo_data_clean %>% filter(age_group == "Young")
-
-# Fit a logistic regression model within the young mothers subgroup 
-# to examine the association between gravidity and adverse birth outcomes
-model_gravidity_young <- glm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + education_level,
-                             family = binomial(link = "logit"),
-                             data = young_mothers)
-
-# Summarize and tidy the subgroup model output
-summary(model_gravidity_young)
-tidy_model_gravidity_young <- tidy(model_gravidity_young, exponentiate = TRUE, conf.int = TRUE)
-print(tidy_model_gravidity_young)
-
-### Additional Exploratory Plots (Optional)
-
-# Plot the distribution of total malaria episodes by study arm
-ggplot(promo_data_clean, aes(x = study_arm, y = total_malaria_episodes)) +
-  geom_boxplot(fill = "lightblue") +
-  labs(title = "Malaria Episode Frequency by Treatment Arm", 
-       x = "Treatment Arm", y = "Total Malaria Episodes")
-
-# Plot the proportion of adverse outcomes by gravidity among young mothers
-ggplot(young_mothers, aes(x = gravidity, fill = factor(adverse_birth_outcome))) +
-  geom_bar(position = "fill") +
-  labs(title = "Proportion of Adverse Outcomes by Gravidity (Young Mothers)",
-       x = "Gravidity", y = "Proportion", fill = "Adverse Outcome")
 
@@
 
+*Table 6: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)*
+
+```{r}
+# Load required libraries
+library(broom)
+library(dplyr)
+library(knitr)
+library(kableExtra)
+
+# Assume model_gravidity is your fitted logistic regression model for women < 25
+# For example:
+# model_gravidity <- glm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + 
+#                          study_arm + education_level,
+#                        family = binomial(link = "logit"),
+#                        data = promo_data_young)
+
+# Tidy the model output, exponentiating coefficients to get odds ratios
+tidy_model_gravidity <- tidy(model_gravidity, exponentiate = TRUE, conf.int = TRUE)
+
+# Create a summary table without the intercept and with neat variable names
+result_table <- tidy_model_gravidity %>%
+  filter(term != "(Intercept)") %>%
+  mutate(
+    term = case_when(
+      term == "gravidity" ~ "Gravidity",
+      term == "total_malaria_episodes" ~ "Total Malaria Episodes",
+      term == "study_armSP" ~ "Treatment Arm (SP)",
+      term == "education_levelSecondary" ~ "Secondary Education",
+      term == "education_levelTertiary" ~ "Tertiary Education",
+      term == "education_levelUniversity" ~ "University Education",
+      TRUE ~ term
+    ),
+    `Odds Ratio (95% CI)` = paste0(round(estimate, 2),
+                                   " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
+  ) %>%
+  select(term, `Odds Ratio (95% CI)`, p.value) %>%
+  rename(Variable = term, `p-value` = p.value)
+
+# Print the table using kable
+result_table %>%
+  kable(format = "markdown", caption = "")
+
+```
+
 *Logistic Regression Model Fitting and Summary*
@@
 
-*Predicted Probability of Adverse Outcome by Gravidity (Age \< 25)*
+```{r}
+# Tidy the model output and exponentiate coefficients to get odds ratios
+tidy_model_gravidity <- tidy(model_gravidity, exponentiate = TRUE, conf.int = TRUE)
+
+# Create a summary table, remove the intercept, and recode variable names
+result_table <- tidy_model_gravidity %>%
+  filter(term != "(Intercept)") %>%
+  mutate(
+    term = dplyr::recode(term,
+                         "gravidity" = "Gravidity",
+                         "total_malaria_episodes" = "Total Malaria Episodes",
+                         "study_armSP" = "Treatment Arm (SP)",
+                         "education_levelSecondary" = "Secondary Education",
+                         "education_levelTertiary" = "Tertiary Education",
+                         "education_levelUniversity" = "University Education"),
+    `Odds Ratio (95% CI)` = paste0(round(estimate, 2),
+                                   " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
+  ) %>%
+  select(term, `Odds Ratio (95% CI)`, p.value) %>%
+  rename(Variable = term, `p-value` = p.value)
+
+# Print the table using kable with a descriptive caption
+result_table %>%
+  kable(format = "markdown", 
+        caption = "Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)")
+
+```
+
+*Predicted Probability of Adverse Outcome by Gravidity (Age < 25)*
 
@@
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "blue") +
-  labs(title = "Predicted Probability of Adverse Outcome by Gravidity (Age < 25)",
+  labs(title = "",
        x = "Gravidity",
@@
 Other predictors (total malaria episodes, treatment arm, and education level) were not statistically significant, suggesting that gravidity is the key protective factor in this subgroup.
+
+##################################################################################################################################################################
 

### 2025-04-05 code/eda-code/eda_manuscript.qmd
Project update

@@
 
+*Boxplot of Birthweight by IPTp Treatment Arm*
+
+```{r}
+ggplot(promo_data_clean, aes(x = study_arm, y = birth_weight, fill = study_arm)) +
+  geom_boxplot(alpha = 0.7) +
+  labs(title = "Birthweight by IPTp Treatment Arm",
+       x = "Treatment Arm",
+       y = "Birthweight (kg)") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+```
+
+```{r}
+# Save the cleaned data (if not already saved)
+write.csv(promo_data_clean, here::here("data", "clean", "PROMO_Data_clean.csv"), row.names = FALSE)
+
+# Read in the cleaned data
+promo_data_clean <- read.csv(here::here("data", "clean", "PROMO_Data_clean.csv"))
+
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
+library(broom)
+
+### Data Preparation
+
+# 1. Create a composite adverse birth outcome variable
+#    Adverse outcome is defined as: 
+#    - Preterm birth: preterm_births_count > 0 
+#    OR stillbirth: stillbirth_bin == 1 
+#    OR low birth weight: birth_weight < 2.5 kg
+promo_data_clean <- promo_data_clean %>%
+  mutate(low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+         adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0))
+
+# 2. Create a subgroup variable for maternal age (<25 years vs. >=25 years)
+promo_data_clean <- promo_data_clean %>%
+  mutate(age_group = ifelse(age_at_enrollment_years < 25, "Young", "Older"))
+
+### Research Question 1: 
+### "Does the type of IPTp regimen modify the association between malaria episode frequency and adverse birth outcomes in Ugandan pregnant women?"
+
+# Fit a logistic regression model with an interaction between total malaria episodes and study_arm
+model_interaction <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm + 
+                           age_at_enrollment_years + gravidity + education_level,
+                         family = binomial(link = "logit"),
+                         data = promo_data_clean)
+
+# Summarize and tidy the model output (exponentiating coefficients to yield odds ratios)
+summary(model_interaction)
+tidy_model_interaction <- tidy(model_interaction, exponentiate = TRUE, conf.int = TRUE)
+print(tidy_model_interaction)
+
+### Research Question 2: 
+### "Among younger Ugandan pregnant women, is increased gravidity associated with a reduced risk of adverse birth outcomes?"
+
+# Subset the data for young mothers (age < 25)
+young_mothers <- promo_data_clean %>% filter(age_group == "Young")
+
+# Fit a logistic regression model within the young mothers subgroup 
+# to examine the association between gravidity and adverse birth outcomes
+model_gravidity_young <- glm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + education_level,
+                             family = binomial(link = "logit"),
+                             data = young_mothers)
+
+# Summarize and tidy the subgroup model output
+summary(model_gravidity_young)
+tidy_model_gravidity_young <- tidy(model_gravidity_young, exponentiate = TRUE, conf.int = TRUE)
+print(tidy_model_gravidity_young)
+
+### Additional Exploratory Plots (Optional)
+
+# Plot the distribution of total malaria episodes by study arm
+ggplot(promo_data_clean, aes(x = study_arm, y = total_malaria_episodes)) +
+  geom_boxplot(fill = "lightblue") +
+  labs(title = "Malaria Episode Frequency by Treatment Arm", 
+       x = "Treatment Arm", y = "Total Malaria Episodes")
+
+# Plot the proportion of adverse outcomes by gravidity among young mothers
+ggplot(young_mothers, aes(x = gravidity, fill = factor(adverse_birth_outcome))) +
+  geom_bar(position = "fill") +
+  labs(title = "Proportion of Adverse Outcomes by Gravidity (Young Mothers)",
+       x = "Gravidity", y = "Proportion", fill = "Adverse Outcome")
+
+```
+
 *Categorical Variables*
@@
 ```
-
-*Correlation Matrix (Numeric Variables)*
-```{r}
-# Correlation plot
-numeric_vars <- promo_data_clean %>% 
-  select(where(is.numeric))
-
-cor_matrix <- cor(numeric_vars, use = "pairwise.complete.obs")
-
-corrplot::corrplot(cor_matrix, method = "circle", type = "lower")
-
-```
-
-*Group-wise Summary*
-```{r}
-# Grouped summary by treatment group (replace 'treatment_group' with actual variable)
-promo_data_clean %>% 
-  group_by(treatment_group) %>% 
-  summarise(across(where(is.numeric), list(mean = mean, sd = sd), na.rm = TRUE))
-
-```

### 2025-04-05 products/manuscript/Manuscript.qmd
Le projet

--- 2025-03-30 products/manuscript/Manuscript.qmd
@@
 title: "Impact of IPTp Regimen on Pregnancy Outcomes in a Malaria-Endemic Setting"
-bibliography: "../../assets/dataanalysis-references.bib"
-csl: "../../assets/american-journal-of-epidemiology.csl"
-
+author: ""
 format:
@@
     number-sections: true
-    output-file: "Manuscript.docx"
-
-project:
-  output-dir: "C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project/products/manuscript"
-
-execute:
-  echo: false
-  warning: false
-  message: false
+    highlight-style: github
+bibliography: ../../assets/dataanalysis-references.bib
+csl: ../../assets/american-journal-of-epidemiology.csl
 ---
@@
 library(broom)
-
 ```
@@
 
-# Summary/Abstract
-
-*Write a summary of your project.*
+# Abstract
+Malaria remains a critical public health challenge in Sub-Saharan Africa, with pregnant women at heightened risk of adverse outcomes.
+This study analyzed data from a double-blind randomized controlled trial in Uganda (ClinEpiDB Release #24, August 30, 2022) to evaluate the impact of intermittent preventive treatment during pregnancy (IPTp) on adverse birth outcomes.
+A total of 782 HIV-uninfected pregnant women received either sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP).
+Multivariable logistic regression models were used to assess whether the IPTp regimen modifies the association between malaria episode frequency and outcomes such as preterm birth, low birth weight, and stillbirth, while adjusting for confounders including maternal age, gravidity, and socioeconomic status.
+Although the main effects of malaria episodes and treatment arm were not statistically significant, the interaction between malaria episodes and the SP regimen was significant (OR = 1.47, 95% CI: 1.16–1.88, p = 0.002), suggesting that the impact of malaria differs by treatment.
+Additionally, a subgroup analysis of women under 25 years revealed that increased gravidity was significantly protective (OR = 0.86, 95% CI: 0.74–1.00, p = 0.044).
+These findings underscore the importance of tailoring malaria prevention strategies based on individual maternal profiles to improve maternal and neonatal health outcomes in malaria-endemic settings.
 
@@
 Caused by Plasmodium parasites transmitted through the bite of infected Anopheles mosquitoes, malaria accounted for an estimated 249 million cases globally in 2022, marking an increase of five million cases compared to the previous year.
-Uganda alone reported over 597,000 malaria cases during this period, reflecting the country's substantial malaria burden (World Health Organization, 2023; Talapko et al., 2019).
+Uganda alone reported over 597,000 malaria cases during this period, reflecting the country’s substantial malaria burden (World Health Organization, 2023; Talapko et al., 2019).
 Pregnant women represent one of the most vulnerable groups to malaria infection, facing an increased risk of severe clinical symptoms and poor pregnancy outcomes.
@@
 Early findings point to a protective effect of prior pregnancies in this subgroup.
-By addressing these questions, this study aims to clarify how preventive treatment strategies and maternal reproductive history influence birth outcomes in malaria-endemic settings.
-The results may inform targeted interventions to improve maternal and neonatal health in Uganda and similar contexts
+In addressing these questions, the study seeks to clarify how preventive treatment strategies and maternal reproductive history shape birth outcomes in malaria-endemic settings.
+Ultimately, the insights gained could inform the development of targeted, effective, and equitable interventions to enhance pregnant women's and infants' health in Uganda and similar regions.
 
 # Methods
-This study analyzed data from a double-blind randomized controlled trial conducted in Uganda, obtained from ClinEpiDB (Release #24, August 30, 2022).
-The trial evaluated IPTp with either sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP) among HIV-uninfected pregnant women.
-The dataset comprised 782 observations, including maternal, pregnancy, and infant health information.
-Multiple births were recorded separately, and variable definitions adhered to original trial protocols.
+This study analyzed data from a double-blind, randomized controlled trial conducted in Uganda, sourced from ClinEpiDB (Release #24, August 30, 2022).
+The trial evaluated intermittent preventive treatment during pregnancy (IPTp) using either sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP) in HIV-uninfected pregnant women.
+The dataset comprised 782 observations detailing maternal, pregnancy, and infant health information, with multiple births recorded separately according to the original trial protocols.
 
-The analysis focused on two primary objectives: (1) to assess whether the IPTp regimen modifies the association between malaria episode frequency and adverse birth outcomes (preterm birth, low birth weight, and stillbirth), and (2) to evaluate whether increased gravidity is associated with lower risk of adverse outcomes among women under 25 years.
+Data preparation involved standardizing date fields, converting categorical variables into factors, and addressing missing data by excluding variables with over 20% missingness and imputing those with lower levels.
+Descriptive statistics and visual inspections ensured the integrity of the data.
 
-Data preparation and analysis were performed using R.
-Date fields were standardized, and categorical variables were coded as factors.
-Variables with more than 20% missing data were excluded; those with lower missingness were retained and imputed where appropriate.
-Descriptive statistics and visual inspection were used to assess data integrity.
-
-For the first research question, multivariable logistic regression models were fitted to examine the association between malaria episodes and adverse outcomes, including an interaction term for IPTp regimen.
-Covariates included maternal age, gravidity, and socioeconomic status.
-Predicted probabilities were calculated to facilitate interpretation of interaction effects.
-
-For the second question, a subgroup analysis restricted to women under 25 years assessed the association between gravidity and adverse birth outcomes, using logistic regression and adjusting for relevant covariates.
-Gravidity was treated as a continuous variable, and sensitivity analyses were performed to assess robustness.
-
-All statistical analyses were conducted using the most recent version of R (version 4.3.2).
+The analysis was designed to evaluate whether the IPTp regimen modifies the association between malaria episode frequency and adverse birth outcomes, specifically, preterm birth, low birth weight, and stillbirth, and to assess whether increased gravidity reduces the risk of these outcomes among women under 25 years.
+Multivariable logistic regression models were employed to examine the relationship between malaria episodes and adverse outcomes, incorporating an interaction term for the IPTp regimen and adjusting for confounders such as maternal age, gravidity, and socioeconomic status.
+Predicted probabilities were calculated to facilitate the interpretation of interaction effects.
+A subgroup analysis among women under 25 years used logistic regression models with gravidity treated as a continuous variable and adjustments for maternal age, malaria episode frequency, and treatment regimen, with sensitivity analyses performed to verify robustness.
+All statistical analyses were conducted using R (version 4.3.2).
 
@@
 ## Exploratory/Descriptive analyses
-Descriptive analyses were conducted to summarize key maternal and clinical characteristics, including maternal age, gravidity, treatment assignment (SP or DP), and malaria episode frequency.
-Standard summary statistics were used for continuous variables, and proportions were reported for categorical variables.
-Distributions were visually assessed to identify potential outliers, and missing data were handled through imputation or exclusion based on the extent of missingness.
+Descriptive analyses summarized key maternal and clinical characteristics such as maternal age, gravidity, treatment assignment (SP or DP), and malaria episode frequency using standard summary statistics for continuous variables and proportions for categorical variables.
+Distributions were visually inspected for potential outliers, and missing data were addressed through imputation or exclusion based on the extent of missingness.
 
-For the first research question, we evaluated whether the IPTp regimen modified the association between malaria episode frequency and adverse birth outcomes (preterm birth, low birth weight, and stillbirth).
-Multivariable logistic regression models were fitted, including an interaction term between malaria episodes and IPTp regimen.
-Models were adjusted for potential confounders, including maternal age, gravidity, and socioeconomic status.
-Predicted probabilities were estimated to facilitate interpretation of interaction effects.
-
-For the second research question, we conducted a subgroup analysis among pregnant women under 25 years of age to assess the association between gravidity and adverse birth outcomes.
-Logistic regression models were used, adjusting for relevant covariates, including maternal age, malaria episode frequency, and treatment regimen.
-Gravidity was treated as a continuous variable.
+Multivariable logistic regression models were employed to evaluate whether the IPTp regimen modifies the association between malaria episodes and adverse birth outcomes, adjusting for maternal age, gravidity, and socioeconomic status, and incorporating an interaction term between malaria episodes and treatment regimen.
+Predicted probabilities were calculated to elucidate these interaction effects.
+Additionally, a subgroup analysis among women under 25 years assessed the relationship between gravidity treated as a continuous variable, and adverse birth outcomes, with adjustments made for maternal age, malaria episode frequency, and treatment regimen.
 
@@
 **Table 1: Baseline Characteristics of Study Participants by IPTp Treatment Arm**
-
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
+&nbsp;
+&nbsp;
+
 **Figure 1: Distribution of Total Malaria Episodes by IPTp Treatment Arm**
-
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-library(ggplot2)
 
@@
 
+&nbsp;
+&nbsp;
+
 **Table 2: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
-
 ```{r,  echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-**Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes**
+&nbsp;
+&nbsp;
 
+*Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes*
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
+&nbsp;
+&nbsp;
+
+&nbsp;
+&nbsp;
+
+**Table 2: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
+```{r,  echo=FALSE, message=FALSE, warning=FALSE}
+# Load necessary libraries
+library(dplyr)
+library(gtsummary)
+library(here)
+
+# Read in the cleaned data (if not already read)
+promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+
+# Recode variables as needed for outcomes analysis:
+promo_data_clean <- promo_data_clean %>%
+  mutate(
+    # Create a composite adverse outcome variable: 1 if any of preterm birth, stillbirth, or low birth weight (<2.5 kg) occurs
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0),
+
+    # Recode Preterm Births Count: combine 0 and 1 as "1", and 2 as "2"
+    PretermBirths_cat = case_when(
+      preterm_births_count %in% c(0, 1) ~ "1",
+      preterm_births_count == 2 ~ "2"
+    ),
+    PretermBirths_cat = factor(PretermBirths_cat, levels = c("1", "2"))
+  )
+
+# Create Table Z: Outcome Variables by IPTp Treatment Arm with p-values
+table_outcomes <- promo_data_clean %>%
+  dplyr::select(
+    study_arm,
+    `Malaria Infection Rate During Pregnancy` = malaria_infection_rate_during_pregnancy,
+    `Placental Malaria (Rogerson Criteria)` = placental_malaria_by_rogerson_criteria,
+    `Preterm Births Count` = PretermBirths_cat,
+    `Stillbirth bin` = stillbirth_bin,
+    `Birthweight` = birth_weight,
+    `Composite Adverse Outcome` = adverse_birth_outcome
+  ) %>%
+  tbl_summary(
+    by = study_arm,
+    missing = "no",
+    statistic = list(
+      all_continuous() ~ "{mean} ({sd})",
+      all_categorical() ~ "{n} ({p}%)"
+    )
+  ) %>%
+  add_p() %>%
+  modify_header(label = "") %>%
+  modify_caption("")
+
+# Print Table Z
+table_outcomes
+```
+
+&nbsp;
+&nbsp;
+
+&nbsp;
+&nbsp;
+
 {{< pagebreak >}}
@@
 
-Baseline characteristics were similar between IPTp treatment arms, with comparable age, gestational age at enrollment, education, gravidity, and parity distributions (Table 1).
-Although overall malaria exposure measures were similar, the stillbirth rate was significantly higher in the SP arm compared to DP (14% vs.
+Baseline characteristics were similar across IPTp treatment arms, with comparable maternal age, gestational age, education, gravidity, and parity (Table 1).
+Although overall malaria exposure was similar between groups, the SP arm exhibited a significantly higher stillbirth rate (14% vs.
 9%; p = 0.044) (Table 2).
 
-In our multivariable logistic regression analysis—adjusting for maternal age, gravidity, and education—the main effects of total malaria episodes (OR = 0.88, 95% CI: 0.75–1.03, p = 0.121) and treatment arm (OR = 0.69, 95% CI: 0.44–1.08, p = 0.102) were not significant.
-However, adding an interaction term between total malaria episodes and treatment arm significantly improved model fit (ΔDeviance = 10.11, p = 0.00148), with the interaction itself reaching significance (OR = 1.47, 95% CI: 1.16–1.88, p = 0.002).
-Notably, higher gravidity was protective (OR = 0.89, 95% CI: 0.80–0.98, p = 0.023), underscoring the potential benefit of prior pregnancy experience.
-Sensitivity analyses confirmed that these findings were robust, and model diagnostics (VIFs < 2) indicated no concerning multicollinearity.
-Collectively, these results suggest that the impact of malaria exposure on adverse birth outcomes is modified by the IPTp regimen, highlighting the importance of tailoring preventive strategies based on maternal risk profiles.
+Multivariable logistic regression analyses, adjusted for maternal age, gravidity, and education revealed that the main effects of total malaria episodes (OR = 0.88, 95% CI: 0.75–1.03, p = 0.121) and treatment arm (OR = 0.69, 95% CI: 0.44–1.08, p = 0.102) were not statistically significant.
+However, incorporating an interaction term between malaria episodes and the IPTp regimen significantly improved model fit (ΔDeviance = 10.11, p = 0.00148), with the interaction itself reaching significance (OR = 1.47, 95% CI: 1.16–1.88, p = 0.002).
+These findings suggest that the impact of malaria exposure on adverse birth outcomes is modified by the treatment regimen.
+Additionally, gravidity demonstrated a protective effect (OR = 0.89, 95% CI: 0.80–0.98, p = 0.023), indicating that prior pregnancy experience reduces the risk of adverse outcomes.
 
@@
 **Table 3: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**
-
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 **Table 5: Evaluating Effect Modification by IPTp Treatment on Malaria-Related Adverse Birth Outcomes Through Model Comparison**
-
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-## Full analysis
+**Table 4: "Incidence Rate Ratios for Malaria Episode Occurrence among Pregnant Women**
 
-*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here.*
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# Table 4: Incidence Rate Ratios for Malaria Episode Occurrence among Pregnant Women
+table_model_refined <- data.frame(
+  Variable = c("Intercept", 
+               "Treatment Arm (SP vs. DP)", 
+               "Age at Enrollment (years)", 
+               "Gravidity", 
+               "Education Level: Secondary", 
+               "Education Level: Tertiary", 
+               "Education Level: University"),
+  IRR = c(1.6081682, 1.0552044, 0.9942737, 1.0095680, 0.9884423, 0.8900707, 1.0363703),
+  `95% CI Lower` = c(1.1297705, 0.9382987, 0.9825662, 0.9675647, 0.8348897, 0.7505115, 0.8805393),
+  `95% CI Upper` = c(2.280714, 1.186523, 1.006133, 1.053391, 1.170522, 1.055600, 1.220799),
+  `p-value` = c(0.008021017, 0.369364418, 0.342231948, 0.660449862, 0.892661838, 0.180547176, 0.668035030)
+)
+
+# Optionally, display the table using knitr::kable for a nicer format
+knitr::kable(table_model_refined, caption = "")
+
+```
+
+&nbsp;
+&nbsp;
+
+Baseline characteristics were similar across IPTp treatment arms, with comparable maternal age, gestational age, education, gravidity, and parity (Table 1).
+Although overall malaria exposure was similar between groups, the SP arm exhibited a significantly higher stillbirth rate (14% vs.
+9%; p = 0.044) (Table 2).
+
+Multivariable logistic regression analyses, adjusted for maternal age, gravidity, and education revealed that the main effects of total malaria episodes (OR = 0.88, 95% CI: 0.75–1.03, p = 0.121) and treatment arm (OR = 0.69, 95% CI: 0.44–1.08, p = 0.102) were not statistically significant.
+However, incorporating an interaction term between malaria episodes and the IPTp regimen significantly improved model fit (ΔDeviance = 10.11, p = 0.00148), with the interaction itself reaching significance (OR = 1.47, 95% CI: 1.16–1.88, p = 0.002).
+These findings suggest that the impact of malaria exposure on adverse birth outcomes is modified by the treatment regimen.
+Additionally, gravidity demonstrated a protective effect (OR = 0.89, 95% CI: 0.80–0.98, p = 0.023), indicating that prior pregnancy experience reduces the risk of adverse outcomes.
+
+&nbsp;
+&nbsp;
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# Fit the logistic regression model for women under 25
+promo_data_young <- promo_data_clean %>%
+  filter(age_at_enrollment_years < 25) %>%
+  mutate(
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
+  )
+
+model_gravidity <- glm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + 
+                         study_arm + education_level,
+                       family = binomial(link = "logit"),
+                       data = promo_data_young)
+
+```
+
+*Table 6: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)*
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+# Tidy the model output, exponentiating coefficients to get odds ratios
+tidy_model_gravidity <- tidy(model_gravidity, exponentiate = TRUE, conf.int = TRUE)
+
+# Create a summary table without the intercept and with neat variable names
+result_table <- tidy_model_gravidity %>%
+  filter(term != "(Intercept)") %>%
+  mutate(
+    term = case_when(
+      term == "gravidity" ~ "Gravidity",
+      term == "total_malaria_episodes" ~ "Total Malaria Episodes",
+      term == "study_armSP" ~ "Treatment Arm (SP)",
+      term == "education_levelSecondary" ~ "Secondary Education",
+      term == "education_levelTertiary" ~ "Tertiary Education",
+      term == "education_levelUniversity" ~ "University Education",
+      TRUE ~ term
+    ),
+    `Odds Ratio (95% CI)` = paste0(round(estimate, 2),
+                                   " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
+  ) %>%
+  select(term, `Odds Ratio (95% CI)`, p.value) %>%
+  rename(Variable = term, `p-value` = p.value)
+
+# Print the table using kable
+result_table %>%
+  kable(format = "markdown", caption = "")
+
+```
+
+&nbsp;
+&nbsp;
+
+*Figure 3: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)*
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+# Generate predicted probabilities over the range of gravidity
+predicted_probs <- ggeffect(model_gravidity, terms = "gravidity")
+
+# Create a scatter/line plot of predicted probabilities by gravidity
+ggplot(predicted_probs, aes(x = x, y = predicted)) +
+  geom_line(size = 1.2) +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "blue") +
+  labs(title = "",
+       x = "Gravidity",
+       y = "Predicted Probability") +
+  theme_minimal()
+
+```
+
+&nbsp;
+&nbsp;
+
+Based on our findings, the impact of malaria on adverse birth outcomes is significantly modified by the IPTp regimen, with a distinct interaction effect observed for the SP regimen.
+Additionally, increased gravidity appears to offer a protective benefit, particularly among women under 25 years.
+Detailed results supporting these observations are presented in Tables 3–6 and Figures 1–3.
 
@@
 # Discussion
+Our data analysis from a randomized controlled trial in Uganda provides important insights into the interplay between malaria exposure, IPTp regimen, and adverse birth outcomes.
+Specifically, our findings reveal that while the individual effects of total malaria episodes and the treatment arm were not statistically significant, the interaction between malaria episodes and the SP regimen significantly increased the risk of adverse outcomes.
+Additionally, increased gravidity emerged as a robust protective factor, particularly among pregnant women under 25 years, with each additional pregnancy associated with a roughly 14% reduction in risk.
+
+These results align with previous studies that have highlighted the complex relationship between maternal characteristics and malaria-related complications during pregnancy.
+For example, similar protective effects of gravidity have been reported in the literature (Chua et al., 2021), and emerging evidence suggests that the efficacy of IPTp regimens may vary depending on the context and specific treatment used (Talapko et al., 2019 World Health Organization, 2023).
+Our findings extend these observations by demonstrating that the SP regimen, in particular, modifies the impact of malaria exposure on birth outcomes, emphasizing the need for tailored malaria prevention strategies.
+
+However, our study has several limitations.
+First, the data were derived from a controlled trial setting, which may limit the generalizability of the findings to broader, real-world populations.
+Second, the necessity to exclude or impute variables with significant missing data could have influenced the robustness of our estimates.
+Third, the modest discriminative ability of our models reflected in ROC curves with AUC values near 0.55–0.60 suggests that additional, unmeasured factors might contribute to adverse birth outcomes.
+
+Despite these limitations, our study offers valuable contributions to the field.
+The evidence that IPTp regimen choice can modify the relationship between malaria episodes and adverse birth outcomes, along with the protective role of gravidity, underscores the importance of developing more personalized intervention strategies.
+Future research should aim to validate these findings in more diverse settings and explore other potential determinants of maternal and neonatal health in malaria-endemic regions.
+
+In summary, our findings advocate for a nuanced approach to malaria prevention in pregnancy that considers both the specific IPTp regimen and individual maternal reproductive history.
+Such tailored strategies have the potential to substantially improve maternal and neonatal health outcomes in regions where malaria remains a significant public health challenge.
 

### 2025-04-05 products/manuscript/Manuscript.qmd
Le Projet 2

@@
 However, this potential effect modification remains underexplored in current literature.
-Additionally, maternal characteristics such as gravidity—the number of times a woman has been pregnant—may also play a role in shaping birth outcomes.
+Additionally, maternal characteristics such as gravidity, the number of times a woman has been pregnant may also play a role in shaping birth outcomes.
 Prior research has suggested that previous pregnancy experience may offer protective benefits against adverse outcomes, possibly due to improved physiological adaptation or better health-seeking behavior.
@@
 
-*Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes*
+**Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-**Table 2: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
+**Table 3: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
 ```{r,  echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-**Table 3: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**
+**Table 4: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-**Table 4:  Regression Diagnostics for IPTp Treatment Effect**
+**Table 5:  Regression Diagnostics for IPTp Treatment Effect**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-**Table 5: Evaluating Effect Modification by IPTp Treatment on Malaria-Related Adverse Birth Outcomes Through Model Comparison**
+**Table 6: Evaluating Effect Modification by IPTp Treatment on Malaria-Related Adverse Birth Outcomes Through Model Comparison**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-**Table 4: "Incidence Rate Ratios for Malaria Episode Occurrence among Pregnant Women**
+**Table 7: "Incidence Rate Ratios for Malaria Episode Occurrence among Pregnant Women**
 
@@
 
-*Table 6: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)*
+**Table 8: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-*Figure 3: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)*
+**Figure 3: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)**
 

### 2025-04-22 code/analysis-code/Statistical_Analysis.qmd
Update Final project

@@
 # Load Required Libraries
-# Essential libraries for data handling, visualization, and analysis
-
-library(here)             # Manage file paths
+
+# Path management
+library(here)              # For file path handling
+
+# Data manipulation and wrangling
 library(dplyr)             # Data manipulation
-library(tidyverse)        # Data wrangling and visualization
+library(tidyverse)         # Core data science packages (includes ggplot2, dplyr, readr, etc.)
 library(janitor)           # Clean column names
-library(skimr)            # Quick data summary
-library(lubridate)        # Working with dates
-library(readr)            # Reading CSV files
-library(ggplot2)          # Visualization
+library(skimr)             # Data summary
+library(lubridate)         # Date and time handling
+library(forcats)           # Handling categorical variables
+
+# Visualization
+library(ggplot2)           # Data visualization (also part of tidyverse)
 library(ggpubr)            # Publication-ready plots
-library(forcats)          # Categorical variables handling
-
-# Table and reporting packages
-library(gtsummary)        # Summary tables
+
+# Tables and reporting
+library(gtsummary)         # Summary statistics tables
 library(gt)                # Table formatting
-library(knitr)            # Table rendering
-library(kableExtra)       # Table customization
+library(knitr)             # Report generation
+library(kableExtra)        # Enhanced table output in markdown
 
 # Analysis-specific packages
-library(Amelia)           # Missing data visualization
+library(Amelia)            # Visualizing missing data
 library(pwr)               # Power analysis
@@
 library(broom)             # Tidy model outputs
-library(rsample)          # Resampling and splitting datasets
+
+# Machine learning and modeling
+library(tidymodels)        # Unified modeling framework (includes recipes, parsnip, workflows, tune, etc.)
+library(rsample)           # Resampling and data splitting
+library(recipes)           # Preprocessing
+library(parsnip)           # Model specification
+library(workflows)         # Combine models and preprocessing
+library(tune)              # Hyperparameter tuning
+library(yardstick)         # Model performance metrics
+library(dials)             # Parameter tuning grids
+library(ranger)            # Random forest engine
+library(xgboost)           # Boosting engine
+library(themis)            # Handling class imbalance (e.g., SMOTE)
 
@@
 The other predictors—total malaria episodes, treatment arm, and education level—do not significantly change the risk in this subgroup, suggesting that prior pregnancy experience is the key factor driving improved outcomes for younger women.
+
+##################################################################################################################################################################
+
+## Machine Learning approaches
+
+To complement regression analysis, we implemented machine learning workflows for Research Question 1.
+All models were trained on 70% of the data and evaluated on a held-out 30% test set.
+Ten-fold cross-validation on the training set was used for hyperparameter tuning where applicable.
+
+# Prepare outcome as factor and split data
+
+```{r}
+promo_data_clean <- promo_data_clean %>%
+  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0,1)))
+set.seed(2025)
+split       <- initial_split(promo_data_clean, prop = 0.7, strata = adverse_birth_outcome)
+train_data  <- training(split)
+test_data   <- testing(split)
+cv_folds    <- vfold_cv(train_data, v = 10, strata = adverse_birth_outcome)
+```
+
+### Logistic regression (baseline)
+
+```{r}
+logistic_wf <- workflow() %>%
+  add_model(logistic_reg() %>% set_engine("glm")) %>%
+  add_formula(adverse_birth_outcome ~ total_malaria_episodes * study_arm +
+                age_at_enrollment_years + gravidity + education_level)
+
+logistic_fit  <- fit(logistic_wf, data = train_data)
+
+# Predict and evaluate
+logistic_preds <- predict(logistic_fit, test_data, type = "prob") %>%
+  bind_cols(test_data)
+logistic_auc <- roc_auc(logistic_preds, truth = adverse_birth_outcome, .pred_1)
+print(logistic_auc)
+```
+The logistic regression model achieved an AUC of 0.48 on the independent test set, indicating discrimination no better than random chance.
+This poor performance suggests that, in its current form, the linear predictor does not capture sufficient signal to distinguish between pregnancies with and without adverse outcomes.
+These findings highlight the need to revisit variable coding, explore non‑linear feature transformations, or consider more flexible algorithms (e.g., random forest or gradient boosting) to improve predictive accuracy.
+
+### Random Forest classifier
+
+```{r}
+# Tuning grid (requires library(dials) or library(tidymodels) loaded)
+rf_recipe <- recipe(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
+                      age_at_enrollment_years + gravidity + education_level,
+                    data = train_data) %>%
+  step_dummy(all_nominal_predictors())
+
+rf_wf <- workflow() %>%
+  add_model(rand_forest(trees = 500, mtry = tune(), min_n = tune()) %>%
+              set_engine("ranger") %>% set_mode("classification")) %>%
+  add_recipe(rf_recipe)
+
+rf_grid <- grid_regular(
+  mtry(range = c(1, 5)),
+  min_n(range = c(2, 10)),
+  levels = 5
+)
+
+rf_tune <- tune_grid(
+  rf_wf,
+  resamples = cv_folds,
+  grid      = rf_grid,
+  metrics   = metric_set(roc_auc)
+)
+
+best_rf <- select_best(rf_tune, metric = "roc_auc")
+rf_final <- finalize_workflow(rf_wf, best_rf)
+rf_fit   <- fit(rf_final, data = train_data)
+
+rf_preds <- predict(rf_fit, test_data, type = "prob") %>%
+  bind_cols(test_data)
+rf_auc <- roc_auc(rf_preds, truth = adverse_birth_outcome, .pred_1)
+print(rf_auc)
+
+```
+The random forest classifier achieved an AUC of 0.42 on the held‑out test set, indicating discrimination well below the 0.50 threshold expected for random performance.
+This result demonstrates that, despite its capacity to capture non‑linear relationships, the current set of predictors does not provide sufficient signal to distinguish between pregnancies with and without adverse outcomes.
+These findings suggest the need for deeper feature engineering, additional or more granular clinical variables, and exploration of alternative modeling strategies to enhance predictive performance.
+
+### Gradient boosting (XGBoost)
+
+```{r}
+# Define workflow
+# Ensure grid_regular() is available
+library(dials)
+
+# Define workflow
+xgb_wf <- workflow() %>%
+  add_model(
+    boost_tree(
+      trees = 500,
+      learn_rate = tune(),
+      tree_depth = tune()
+    ) %>%
+      set_engine("xgboost") %>%
+      set_mode("classification")
+  ) %>%
+  add_recipe(rf_recipe)
+
+# Create tuning grid
+xgb_grid <- grid_regular(
+  learn_rate(range = c(0.01, 0.3)),
+  tree_depth(range = c(3, 8)),
+  levels = 4
+)
+
+# Perform tuning
+xgb_tune <- tune_grid(
+  xgb_wf,
+  resamples = cv_folds,
+  grid      = xgb_grid,
+  metrics   = metric_set(roc_auc)
+)
+
+# Select best hyperparameters by naming the metric argument
+best_xgb <- select_best(xgb_tune, metric = "roc_auc")
+
+# Finalize and fit the final model
+xgb_final <- finalize_workflow(xgb_wf, best_xgb)
+xgb_fit   <- fit(xgb_final, data = train_data)
+
+# Predict on the test set
+xgb_preds <- predict(xgb_fit, test_data, type = "prob") %>%
+  bind_cols(test_data)
+
+# Compute and print AUC
+xgb_auc <- roc_auc(xgb_preds, truth = adverse_birth_outcome, .pred_1)
+print(xgb_auc)
+```
+The gradient boosting model (XGBoost) achieved an AUC of 0.44 on the independent test set, indicating performance substantially below the 0.50 threshold for random guessing.
+Despite its flexibility in modeling complex, non‑linear relationships, the current predictor set does not appear to contain enough signal to reliably distinguish between pregnancies with and without adverse outcomes.
+These results highlight the need for additional feature engineering, incorporation of more informative clinical variables, or alternative analytic strategies to improve predictive accuracy.
+
+### Model comparison on test set
+
+```{r}
+roc_lr  <- roc_curve(logistic_preds, truth = adverse_birth_outcome, .pred_1) %>% mutate(model = "Logistic")
+roc_rf  <- roc_curve(rf_preds,       truth = adverse_birth_outcome, .pred_1) %>% mutate(model = "RF")
+roc_xgb <- roc_curve(xgb_preds,     truth = adverse_birth_outcome, .pred_1) %>% mutate(model = "XGBoost")
+
+roc_all <- bind_rows(roc_lr, roc_rf, roc_xgb)
+
+ggplot(roc_all, aes(x = 1 - specificity, y = sensitivity, color = model)) +
+  geom_path(size = 1) +
+  labs(title = "ROC Curves: ML Models for Adverse Outcome Prediction",
+       x = "False Positive Rate", y = "True Positive Rate") +
+  theme_minimal()
+```
+Figure X displays the ROC curves for logistic regression (red), random forest (green), and gradient boosting (blue) on the held‑out test set.
+All three curves lie close to the diagonal, reflecting poor discrimination: none of the models reliably separates pregnancies with versus without adverse outcomes.
+Of the three, logistic regression achieved the highest AUC (0.48), followed by XGBoost (0.44) and random forest (0.42), but these differences are marginal and all fall below the 0.50 threshold for random performance.
+Collectively, these results underscore that, with the current predictor set, more informative features or alternative modeling strategies will be necessary to improve predictive accuracy.
+
+## Subgroup ML: Young women (<25)
+A similar train/test and model workflow was applied to the subset of women under 25 to assess whether ML methods capture the protective effect of gravidity in this subgroup.
+
+```{r}
+# Subset and split
+set.seed(2025)
+young_split <- initial_split(
+  filter(promo_data_clean, age_at_enrollment_years < 25),
+  prop = 0.7,
+  strata = adverse_birth_outcome
+)
+young_train <- training(young_split) %>%
+  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0,1)))
+young_test  <- testing(young_split) %>%
+  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0,1)))
+
+# Logistic on gravidity only
+young_logit <- glm(
+  adverse_birth_outcome ~ gravidity,
+  family = binomial(),
+  data = young_train
+)
+
+# Evaluate performance
+young_preds <- predict(
+  young_logit,
+  newdata = young_test,
+  type = "response"
+)
+young_auc <- roc_auc_vec(
+  truth = young_test$adverse_birth_outcome,
+  estimate = young_preds
+)
+print(young_auc)
+```
+
+This ML section demonstrates train/test splitting, cross-validation for tuning, and comparison across logistic regression, random forest, and boosting models, aligned with Research Question1 and the subgroup analysis for Research Question2.
+
+## Subgroup ML: Young women (<25)
+
+A similar train/test and model workflow was applied to the subset of women under 25 to assess whether ML methods capture the protective effect of gravidity in this subgroup.
+
+```{r}
+# r ml_young
+# Subset
+young_split <- initial_split(filter(promo_data_clean, age_at_enrollment_years < 25), prop = 0.7, strata = adverse_birth_outcome)
+young_train <- training(young_split)
+young_test  <- testing(young_split)
+
+# Fit logistic with gravidity only
+young_logit <- glm(adverse_birth_outcome ~ gravidity,
+                   family = binomial(), data = young_train)
+
+# Evaluate on test set
+young_preds <- predict(young_logit, newdata = young_test, type = "response")
+roc_auc_vec <- roc_auc_vec(young_test$adverse_birth_outcome, young_preds)
+print(roc_auc_vec)
+```
+
+In the subgroup of women under 25 years, the machine‑learning model achieved an AUC of 0.49 on the held‑out test set, indicating discrimination essentially equivalent to random guessing.
+This suggests that, even within this younger cohort, the available predictors do not provide sufficient signal to distinguish pregnancies with versus without adverse outcomes.
+These results highlight the need for additional or more granular features and potentially alternative modeling strategies to meaningfully improve risk stratification in this subgroup.
+
+This ML section demonstrates train/test splitting, cross-validation for tuning, and comparison across logistic regression, random forest, and boosting models, aligned with Research Question1 and the subgroup analysis for Question2.
+
+##################################################################################################################################################################
+
+### Improved ML: Elastic Net with Feature Engineering
+
+```{r ml_improved_setup, message=FALSE, warning=FALSE}
+# 0) Load all needed packages
+library(tidymodels)  # recipes, parsnip, workflows, tune, yardstick, dials
+library(themis)      # for step_smote()
+```
+
+```{r ml_improved_recipe, message=FALSE}
+# 1) Encode factors, engineer features, balance classes, normalize
+improved_recipe <- recipe(adverse_birth_outcome ~ 
+                             total_malaria_episodes + study_arm +
+                             age_at_enrollment_years + gravidity + education_level,
+                           data = train_data) %>%
+  # a) one‑hot encode all factor predictors
+  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
+  # b) create polynomial & interaction features
+  step_mutate(
+    malaria_sq       = total_malaria_episodes^2,
+    malaria_DP_inter = total_malaria_episodes * study_arm_DP
+  ) %>%
+  # c) up‑sample minority class via SMOTE
+  step_smote(adverse_birth_outcome) %>%
+  # d) center & scale all predictors
+  step_normalize(all_predictors())
+                          # (5)
+
+```
+
+```{r}
+# 2) Specify elastic‑net logistic regression
+enet_spec <- logistic_reg(
+    penalty = tune(),   # strength of regularization
+    mixture = tune()    # 0=ridge ↔ 1=lasso
+  ) %>%
+  set_engine("glmnet")
+
+```
+
+```{r}
+# 3) Build the workflow
+enet_wf <- workflow() %>%
+  add_recipe(improved_recipe) %>%
+  add_model(enet_spec)
+
+```
+
+```{r}
+# 4) Create a tuning grid & 5) tune via 10‑fold CV
+enet_grid <- grid_regular(
+  penalty(range = c(-5, -1), trans = log10_trans()),  # 1e-5 → 1e-1
+  mixture(range = c(0, 1)),
+  levels = 5
+)
+
+enet_tune <- tune_grid(
+  enet_wf,
+  resamples = cv_folds,
+  grid      = enet_grid,
+  metrics   = metric_set(roc_auc)
+)
+
+```
+
+```{r}
+# 6) Select best, finalize, fit on train, & evaluate on test
+best_enet  <- select_best(enet_tune, metric = "roc_auc")
+enet_final <- finalize_workflow(enet_wf, best_enet)
+enet_fit   <- fit(enet_final, data = train_data)
+
+enet_preds <- predict(enet_fit, test_data, type = "prob") %>% bind_cols(test_data)
+enet_auc   <- roc_auc(enet_preds, truth = adverse_birth_outcome, .pred_1)
+print(enet_auc)
+
+```
+The elastic‑net model achieved an AUC of 0.49 on the held‑out test set—below the 0.50 threshold for random guessing.
+This indicates that, even with polynomial, interaction, and SMOTE steps, the model cannot reliably distinguish between pregnancies with and without adverse outcomes.
+Such near‑random performance suggests either insufficient predictive signal in the available features or the need for additional, more informative variables and further model refinement.
+
+##################################################################################################################################################################
+
+```{r}
+# Create temporal & composite features in the training set
+train_data <- train_data %>%
+  mutate(
+    # Month of enrollment (seasonality)
+    enroll_month = lubridate::month(enrollment_date),
+    # Episodes per trimester (assuming 40‑week gestation)
+    episodes_per_trimester = total_malaria_episodes / (gestational_age_at_enrollment_weeks / (40/3)),
+    # Socioeconomic index (if available)
+    # wealth_index = as.numeric(education_level) * household_assets_score
+  )
+
+# Apply same transformations to test_data
+test_data <- test_data %>%
+  mutate(
+    enroll_month = lubridate::month(enrollment_date),
+    episodes_per_trimester = total_malaria_episodes / (gestational_age_at_enrollment_weeks / (40/3))
+  )
+
+```
+
+```{r}
+library(themis)
+
+imbalanced_recipe <- recipe(adverse_birth_outcome ~ ., data = train_data) %>%
+  # standard preprocessing
+  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
+  # adaptive synthetic sampling
+  step_adasyn(adverse_birth_outcome) %>%
+  step_normalize(all_predictors())
+
+```
+
+```{r}
+# Continue using enet_wf from before
+set.seed(2025)
+enet_bayes <- tune_bayes(
+  enet_wf,
+  resamples = cv_folds,
+  initial = enet_tune,       # seed the search from your grid results
+  iter    = 20,              # number of Bayesian iterations
+  metrics = metric_set(roc_auc)
+)
+
+best_enet_bayes <- select_best(enet_bayes, metric = "roc_auc")
+
+```
+
+```{r}
+library(stacks)
+
+# Assume you've tuned logistic, rf, xgb already: logistic_tune, rf_tune, xgb_tune
+stack <- stacks() %>%
+  add_candidates(logistic_tune) %>%
+  add_candidates(rf_tune)       %>%
+  add_candidates(xgb_tune)      %>%
+  blend_predictions()           %>%
+  fit_members()
+
+# Evaluate on test set
+stack_preds <- predict(stack, test_data, type = "prob") %>% bind_cols(test_data)
+stack_auc   <- roc_auc(stack_preds, truth = adverse_birth_outcome, .pred_1)
+print(stack_auc)
+
+```
+
+```{r}
+svm_spec <- svm_rbf(
+    cost      = tune(),
+    rbf_sigma = tune()
+  ) %>%
+  set_engine("kernlab") %>%
+  set_mode("classification")
+
+svm_wf <- workflow() %>%
+  add_model(svm_spec) %>%
+  add_recipe(improved_recipe)  # or your imbalanced_recipe
+
+svm_grid <- grid_regular(
+  cost(range = c(-5, 2), trans = log2_trans()),
+  rbf_sigma(range = c(-3, 0), trans = log2_trans()),
+  levels = 5
+)
+
+svm_tune <- tune_grid(
+  svm_wf,
+  resamples = cv_folds,
+  grid      = svm_grid,
+  metrics   = metric_set(roc_auc)
+)
+
+best_svm <- select_best(svm_tune, metric = "roc_auc")
+
+```
+
+```{r}
+# For your best model (e.g. enet_fit or stack)
+pr    <- pr_curve(stack_preds, truth = adverse_birth_outcome, .pred_1)
+pr_auc <- pr_auc(stack_preds, truth = adverse_birth_outcome, .pred_1)
+
+autoplot(pr) +
+  labs(title = "Precision–Recall Curve", subtitle = paste("PR AUC =", round(pr_auc$.estimate,2)))
+
+```

### 2025-04-22 products/manuscript/Manuscript.qmd
Update Final project

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+
+# 1)  working directory to the project root
+setwd("C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project")
+
+# 2) Chunk defaults
+knitr::opts_chunk$set(
+  echo    = FALSE,
+  message = FALSE,
+  warning = FALSE
+)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 # Load Required Libraries for Data Handling, Visualization, and Analysis
 
-library(here)             # Manage file paths
-library(tidyverse)        # Core data wrangling and visualization (dplyr, ggplot2, readr, etc.)
-library(janitor)          # Clean and format column names
-library(skimr)            # Quick data summaries
-library(gtsummary)        # Create summary tables
-library(gt)               # Format tables
-library(knitr)            # R Markdown table formatting
-library(kableExtra)       # Enhanced table customization for kable output
+# Path management
+library(here)              # For file path handling
+
+# Data manipulation and wrangling
+library(dplyr)             # Data manipulation
+library(tidyverse)         # Core data science packages (includes ggplot2, dplyr, readr, etc.)
+library(janitor)           # Clean column names
+library(skimr)             # Data summary
+library(lubridate)         # Date and time handling
+library(forcats)           # Handling categorical variables
+
+# Visualization
+library(ggplot2)           # Data visualization (also part of tidyverse)
 library(ggpubr)            # Publication-ready plots
-library(Amelia)           # Visualize missing data
-library(pwr)              # Perform power analyses
-library(DiagrammeR)       # Create flowcharts and diagrams
-library(survival)         # Survival analysis functions
-library(survminer)        # Visualize survival analysis results
-library(flextable)        # Create tables for Word output
-library(webshot2)         # Capture screenshots for HTML widgets in non-HTML outputs
-library(logistf)          # Penalized logistic regression (Firth’s method)
-library(ggeffects)
-library(car)
-library(broom)
+
+# Tables and reporting
+library(gtsummary)         # Summary statistics tables
+library(gt)                # Table formatting
+library(knitr)             # Report generation
+library(kableExtra)        # Enhanced table output in markdown
+
+# Analysis-specific packages
+library(Amelia)            # Visualizing missing data
+library(pwr)               # Power analysis
+library(DiagrammeR)        # Diagrams and flowcharts
+library(survival)          # Survival analysis
+library(survminer)         # Survival plots
+library(ggeffects)         # Marginal effects for regression models
+library(broom)             # Tidy model outputs
+
+# Machine learning and modeling
+library(tidymodels)        # Unified modeling framework (includes recipes, parsnip, workflows, tune, etc.)
+library(rsample)           # Resampling and data splitting
+library(recipes)           # Preprocessing
+library(parsnip)           # Model specification
+library(workflows)         # Combine models and preprocessing
+library(tune)              # Hyperparameter tuning
+library(yardstick)         # Model performance metrics
+library(dials)             # Parameter tuning grids
+library(ranger)            # Random forest engine
+library(xgboost)           # Boosting engine
+library(themis)            # Handling class imbalance (e.g., SMOTE)
+
 ```
@@
 # Abstract
-Malaria remains a critical public health challenge in Sub-Saharan Africa, with pregnant women at heightened risk of adverse outcomes.
-This study analyzed data from a double-blind randomized controlled trial in Uganda (ClinEpiDB Release #24, August 30, 2022) to evaluate the impact of intermittent preventive treatment during pregnancy (IPTp) on adverse birth outcomes.
-A total of 782 HIV-uninfected pregnant women received either sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP).
-Multivariable logistic regression models were used to assess whether the IPTp regimen modifies the association between malaria episode frequency and outcomes such as preterm birth, low birth weight, and stillbirth, while adjusting for confounders including maternal age, gravidity, and socioeconomic status.
-Although the main effects of malaria episodes and treatment arm were not statistically significant, the interaction between malaria episodes and the SP regimen was significant (OR = 1.47, 95% CI: 1.16–1.88, p = 0.002), suggesting that the impact of malaria differs by treatment.
-Additionally, a subgroup analysis of women under 25 years revealed that increased gravidity was significantly protective (OR = 0.86, 95% CI: 0.74–1.00, p = 0.044).
-These findings underscore the importance of tailoring malaria prevention strategies based on individual maternal profiles to improve maternal and neonatal health outcomes in malaria-endemic settings.
+
+# Background
+Malaria in pregnancy contributes substantially to maternal and neonatal morbidity and mortality across Sub‑Saharan Africa.
+Intermittent preventive treatment during pregnancy (IPTp) is recommended to reduce adverse birth outcomes, yet the comparative effectiveness of different regimens and how their impact varies with malaria burden remains unclear.
+
+## Methods
+We conducted a secondary analysis of data from a double‑blind randomized controlled trial in Uganda (ClinEpiDB Release #24, August 30, 2022) involving 782 HIV‑uninfected pregnant women.
+Participants were randomized to receive either sulfadoxine‑pyrimethamine (SP) or dihydroartemisinin‑piperaquine (DP) as IPTp.
+Multivariable logistic regression was employed to evaluate (1) the main effects of cumulative malaria episodes and treatment arm on preterm birth, low birth weight, and stillbirth, and (2) the interaction between malaria frequency and IPTp regimen.
+Models adjusted for maternal age, gravidity, and socioeconomic status.
+A prespecified subgroup analysis in women < 25 years assessed the effect of gravidity on outcomes.
+
+## Results
+In multivariable models adjusted for maternal age, gravidity, and socioeconomic status, we identified a statistically significant interaction between malaria episode frequency and the SP regimen (OR 1.47 per additional episode, 95% CI 1.16–1.88; p = 0.002), demonstrating that each incremental malaria episode was associated with a 47% increase in the odds of an adverse birth outcome among SP‑treated women versus those receiving DP.
+Notably, neither malaria burden nor IPTp regimen alone achieved statistical significance, underscoring the critical role of their interplay.
+Moreover, in the subgroup of women under 25 years, gravidity conferred significant protection: each additional pregnancy reduced the likelihood of an adverse outcome by 14% (OR 0.86 per gravidity, 95% CI 0.74–1.00; p = 0.044).
+These findings suggest that SP’s preventive efficacy diminishes with repeated malaria exposure and that higher parity provides resilience against adverse perinatal outcomes in younger mothers.
+
+## Conclusions:
+These results demonstrate that IPTp regimen selection fundamentally alters how cumulative malaria exposure impacts perinatal outcomes: sulfadoxine–pyrimethamine exhibits progressively reduced protective efficacy with each successive malaria episode, whereas dihydroartemisinin–piperaquine maintains greater resilience.
+Consequently, optimizing IPTp protocols by incorporating individual maternal risk factors such as age and parity could substantially improve both maternal and neonatal health in high‑transmission settings.
+
+## Keywords:
+**Malaria, adverse birth outcomes, intermittent preventive treatment in pregnancy (IPTp), sulfadoxine‑pyrimethamine, dihydroartemisinin‑piperaqui**
 
@@
 
-# Introduction
+# Introductio
 ## General Background Information

### 2025-04-24 code/analysis-code/Statistical_Analysis.qmd
Project4

@@
 ```{r}
-library(ggplot2)
-
 ggplot(promo_data_clean, aes(x = total_malaria_episodes, fill = study_arm)) +
@@
 
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# Subset to women under 25 and define outcome. Logistic Regression
+promo_data_young <- promo_data_clean %>%
+  filter(age_at_enrollment_years < 25) %>%
+  mutate(
+    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(
+      preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0
+    )
+  )
+
+# Fit logistic regression model
+model_gravidity <- glm(
+  adverse_birth_outcome ~ gravidity + total_malaria_episodes + 
+    study_arm + education_level,
+  family = binomial(link = "logit"),
+  data = promo_data_young
+)
+
+# Load broom for tidy output
+library(broom)
+
+# Display model summary as odds ratios with 95% CI
+tidy(model_gravidity, exponentiate = TRUE, conf.int = TRUE)
+
+```
+
+*Predicted Probability of Adverse Outcome by Gravidity (Age < 25)*
+
 ```{r}
@@
 ```
-Figure X displays the ROC curves for logistic regression (red), random forest (green), and gradient boosting (blue) on the held‑out test set.
+Figure displays the ROC curves for logistic regression (red), random forest (green), and gradient boosting (blue) on the held‑out test set.
 All three curves lie close to the diagonal, reflecting poor discrimination: none of the models reliably separates pregnancies with versus without adverse outcomes.
@@
 
+**Include**
+```{r}
+ggplot(roc_all, aes(x = 1 - specificity, y = sensitivity, color = model)) +
+  geom_path(size = 1) +
+  labs(title = "ROC Curves: ML Models for Adverse Outcome Prediction",
+       x = "False Positive Rate", y = "True Positive Rate") +
+  theme_minimal()
+
+```
+
+# Model Comparison Table (AUC values)
+
+```{r}
+tibble::tibble(
+  Model = c("Logistic Regression", "Random Forest", "XGBoost", "Elastic Net"),
+  AUC = c(0.48, 0.42, 0.44, 0.49)
+)
+
+```
+
 ##################################################################################################################################################################
@@
 ### Improved ML: Elastic Net with Feature Engineering
-
-```{r ml_improved_setup, message=FALSE, warning=FALSE}
-# 0) Load all needed packages
-library(tidymodels)  # recipes, parsnip, workflows, tune, yardstick, dials
-library(themis)      # for step_smote()
-```
 
@@
 ```{r}
-library(themis)
-
 imbalanced_recipe <- recipe(adverse_birth_outcome ~ ., data = train_data) %>%

### 2025-04-24 products/manuscript/Manuscript.qmd
Project4

@@
 **Table 1: Baseline Characteristics of Study Participants by IPTp Treatment Arm**
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# Create table data
-table_data <- tribble(
-  ~Characteristic,                           ~Subcategory,               ~`DP (N=4121)`,       ~`SP (N=3701)`,
-  "Age (years)",                             "",                         "25.0 (5.0)",         "25.0 (4.9)",
-  "Gestational Age (weeks)",                 "",                         "15.24 (1.97)",       "14.97 (1.94)",
-  "Maternal Education Level (%)",            "",                         "",                   "",
-  "",                                        "Primary",                  "262 (66.7%)",        "258 (66.4%)",
-  "",                                        "Tertiary",                 "65 (16.5%)",         "61 (15.7%)",
-  "",                                        "University",               "66 (16.8%)",         "70 (18.0%)",
-  "Gravidity",                               "",                         "",                   "",
-  "",                                        "1",                        "91 (22%)",           "56 (15%)",
-  "",                                        "2–3",                      "167 (41%)",          "166 (45%)",
-  "",                                        "≥4",                       "154 (37%)",          "148 (40%)",
-  "Parity",                                  "",                         "",                   "",
-  "",                                        "0",                        "79 (19%)",           "77 (21%)",
-  "",                                        "1–2",                      "175 (42%)",          "158 (43%)",
-  "",                                        "≥3",                       "158 (38%)",          "135 (36%)",
-  "Total Malaria Episodes",                  "",                         "",                   "",
-  "",                                        "1",                        "245 (59%)",          "216 (58%)",
-  "",                                        "2–3",                      "146 (35%)",          "127 (34%)",
-  "",                                        "≥4",                       "21 (5.1%)",          "27 (7.3%)",
-  "Total Malaria Episodes During Pregnancy", "",                         "",                   "",
-  "",                                        "1",                        "179 (43%)",          "165 (45%)",
-  "",                                        "2–3",                      "138 (33%)",          "120 (32%)",
-  "",                                        "≥4",                       "95 (23%)",           "85 (23%)",
-  "Malaria Infection Rate During Pregnancy", "",                         "1.01 (0.58)",        "1.05 (0.58)",
-  "Placental Malaria (Rogerson Criteria)",   "",                         "75 (18%)",           "72 (19%)",
-  "Preterm Births Count",                    "",                         "",                   "",
-  "",                                        "1",                        "391 (95%)",          "348 (94%)",
-  "",                                        "2",                        "21 (5.1%)",          "22 (5.9%)",
-  "Stillbirth bin",                          "",                         "37 (9.0%)",          "50 (14%)",
-  "Birthweight",                             "",                         "2.79 (0.64)",        "2.81 (0.62)",
-  "History of Previous Malaria Episodes (%)", "",                        "40 (10%)",           "35 (9%)"
-)
-
-kable(
-  table_data,
-  format = "pandoc",
-  caption = ""
-) %>% 
-  kable_styling(full_width = FALSE, position = "center")
+knitr::include_graphics(here::here("Images", "Table 1. Baseline Characteristics .png"))
 
@@
 
-&nbsp;
-&nbsp;
-
 {{< pagebreak >}}
@@
 **Table 4: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-#library(broom)
-library(dplyr)
-library(knitr)
-library(kableExtra)
-library(here)
 
-# Tidy the model output with exponentiated coefficients (odds ratios)
-tidy_model <- tidy(model_interaction, exponentiate = TRUE, conf.int = TRUE)
-
-# Remove the intercept row and recode variable names using case_when()
-tidy_model_filtered <- tidy_model %>%
-  filter(term != "(Intercept)") %>%
-  mutate(
-    term = case_when(
-      term == "total_malaria_episodes" ~ "Total Malaria Episodes",
-      term == "study_armSP" ~ "Treatment Arm (SP)",
-      term == "age_at_enrollment_years" ~ "Age at Enrollment (years)",
-      term == "gravidity" ~ "Gravidity",
-      term == "education_levelSecondary" ~ "Secondary Education",
-      term == "education_levelTertiary" ~ "Tertiary Education",
-      term == "education_levelUniversity" ~ "University Education",
-      term == "total_malaria_episodes:study_armSP" ~ "Interaction: Malaria Episodes × SP",
-      TRUE ~ term
-    )
-  )
-
-# Round key values for clarity
-tidy_model_filtered <- tidy_model_filtered %>%
-  mutate(
-    estimate = round(estimate, 2),
-    std.error = round(std.error, 2),
-    statistic = round(statistic, 2),
-    p.value = round(p.value, 3),
-    conf.low = round(conf.low, 2),
-    conf.high = round(conf.high, 2)
-  )
-
-# Render the table with a blank header for the first column
-kable(
-  tidy_model_filtered,
-  format = "pandoc",
-  caption = "",
-  col.names = c("", "Odds Ratio", "Std. Error", "z value", "p-value", "95% CI Lower", "95% CI Upper")
-) %>% 
-  kable_styling(full_width = FALSE, position = "center")
-
-```
-
-&nbsp;
-&nbsp;
-
-**Table 5:  Regression Diagnostics for IPTp Treatment Effect**
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-# Read in the cleaned data (ensure adverse_birth_outcome is defined as needed)
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
-
-# Create composite adverse outcome if not already defined
-promo_data_clean <- promo_data_clean %>%
-  mutate(
-    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
-    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
-  )
-
-# Fit Model 1 (without interaction)
-model1 <- glm(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
-                age_at_enrollment_years + gravidity + education_level,
-              family = binomial(link = "logit"), data = promo_data_clean)
-
-# Fit Model 2 (with interaction)
-model2 <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm +
-                age_at_enrollment_years + gravidity + education_level,
-              family = binomial(link = "logit"), data = promo_data_clean)
-
-# Generate Analysis of Deviance Table (Likelihood Ratio Test)
-lrt <- anova(model1, model2, test = "LRT")
-lrt_df <- as.data.frame(lrt)
-lrt_df$Model <- c("No Interaction", "With Interaction")
-lrt_df <- lrt_df %>% 
-  select(Model, `Resid. Df`, `Resid. Dev`, Df, Deviance, `Pr(>Chi)`)
-colnames(lrt_df) <- c("Model", "Residual Df", "Residual Dev", "Df", "Deviance", "p-value")
-
-# Render Table X: Likelihood Ratio Test
-kable(
-  lrt_df,
-  format = "pandoc",
-  caption = ""
-) %>% 
-  kable_styling(full_width = FALSE, position = "center")
-
-# Compute Variance Inflation Factors (VIF) for Model 2 (with interaction)
-vif_vals <- vif(model2)
-vif_df <- as.data.frame(vif_vals)
-vif_df$Variable <- rownames(vif_vals)
-vif_df <- vif_df %>% rename(`Df (VIF)` = Df)
-vif_df <- vif_df %>% 
-  select(Variable, GVIF, `Df (VIF)`, `GVIF^(1/(2*Df))`) %>%
-  mutate(
-    Variable = case_when(
-      Variable == "total_malaria_episodes" ~ "Total Malaria Episodes",
-      Variable == "study_arm" ~ "Treatment Arm",
-      Variable == "age_at_enrollment_years" ~ "Age at Enrollment",
-      Variable == "gravidity" ~ "Gravidity",
-      Variable == "education_level" ~ "Education Level",
-      Variable == "total_malaria_episodes:study_arm" ~ "Interaction: Malaria Episodes × SP",
-      TRUE ~ Variable
-    )
-  )
-
-# Render Table Y: VIF Diagnostics
-kable(
-  vif_df,
-  format = "pandoc",
-  caption = ""
-) %>% 
-  kable_styling(full_width = FALSE, position = "center")
-
-```
-
-&nbsp;
-&nbsp;
-
-**Table 6: Evaluating Effect Modification by IPTp Treatment on Malaria-Related Adverse Birth Outcomes Through Model Comparison**
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics("C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project/Images/Table- Comparing Models .png")
-
-```
-
-&nbsp;
-&nbsp;
-
-**Table 7: "Incidence Rate Ratios for Malaria Episode Occurrence among Pregnant Women**
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# Table 4: Incidence Rate Ratios for Malaria Episode Occurrence among Pregnant Women
-table_model_refined <- data.frame(
-  Variable = c("Intercept", 
-               "Treatment Arm (SP vs. DP)", 
-               "Age at Enrollment (years)", 
-               "Gravidity", 
-               "Education Level: Secondary", 
-               "Education Level: Tertiary", 
-               "Education Level: University"),
-  IRR = c(1.6081682, 1.0552044, 0.9942737, 1.0095680, 0.9884423, 0.8900707, 1.0363703),
-  `95% CI Lower` = c(1.1297705, 0.9382987, 0.9825662, 0.9675647, 0.8348897, 0.7505115, 0.8805393),
-  `95% CI Upper` = c(2.280714, 1.186523, 1.006133, 1.053391, 1.170522, 1.055600, 1.220799),
-  `p-value` = c(0.008021017, 0.369364418, 0.342231948, 0.660449862, 0.892661838, 0.180547176, 0.668035030)
-)
-
-# Optionally, display the table using knitr::kable for a nicer format
-knitr::kable(table_model_refined, caption = "")
+```{r,echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics(here::here("Images", "Table4.Interaction Malaria vsIPTP.png"))
 
@@
 
+**Table 5: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# Fit the logistic regression model for women under 25
-promo_data_young <- promo_data_clean %>%
-  filter(age_at_enrollment_years < 25) %>%
-  mutate(
-    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
-    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
-  )
-
-model_gravidity <- glm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + 
-                         study_arm + education_level,
-                       family = binomial(link = "logit"),
-                       data = promo_data_young)
-
-```
-
-**Table 8: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-# Tidy the model output, exponentiating coefficients to get odds ratios
-tidy_model_gravidity <- tidy(model_gravidity, exponentiate = TRUE, conf.int = TRUE)
-
-# Create a summary table without the intercept and with neat variable names
-result_table <- tidy_model_gravidity %>%
-  filter(term != "(Intercept)") %>%
-  mutate(
-    term = case_when(
-      term == "gravidity" ~ "Gravidity",
-      term == "total_malaria_episodes" ~ "Total Malaria Episodes",
-      term == "study_armSP" ~ "Treatment Arm (SP)",
-      term == "education_levelSecondary" ~ "Secondary Education",
-      term == "education_levelTertiary" ~ "Tertiary Education",
-      term == "education_levelUniversity" ~ "University Education",
-      TRUE ~ term
-    ),
-    `Odds Ratio (95% CI)` = paste0(round(estimate, 2),
-                                   " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
-  ) %>%
-  select(term, `Odds Ratio (95% CI)`, p.value) %>%
-  rename(Variable = term, `p-value` = p.value)
-
-# Print the table using kable
-result_table %>%
-  kable(format = "markdown", caption = "")
+knitr::include_graphics(here::here("Images", "Adjusted OR for Adverse Birth Outcomes Women  less 25 Years.png"))
 
@@
 Detailed results supporting these observations are presented in Tables 3–6 and Figures 1–3.
+
+&nbsp;
+&nbsp;
+
+**Figure 4 : Comparative Discriminative Performance of ML Models for Predicting Adverse Birth Outcomes**
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+roc_lr  <- roc_curve(logistic_preds, truth = adverse_birth_outcome, .pred_1) %>% mutate(model = "Logistic")
+roc_rf  <- roc_curve(rf_preds,       truth = adverse_birth_outcome, .pred_1) %>% mutate(model = "RF")
+roc_xgb <- roc_curve(xgb_preds,     truth = adverse_birth_outcome, .pred_1) %>% mutate(model = "XGBoost")
+
+roc_all <- bind_rows(roc_lr, roc_rf, roc_xgb)
+
+ggplot(roc_all, aes(x = 1 - specificity, y = sensitivity, color = model)) +
+  geom_path(size = 1) +
+  labs(title = "",
+       x = "False Positive Rate", y = "True Positive Rate") +
+  theme_minimal()
+```
+
+&nbsp;
+&nbsp;
+
+**Table 6:  Discriminative Performance (AUC) of Machine Learning Models Predicting Adverse Birth Outcomes Under Different IPTp Regimens**
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics(here::here("Images", "Table6.Discriminative Performance (AUC) of ML Models.png"))
+
+```
 

### 2025-04-24 products/manuscript/Manuscript.qmd
Project

@@
 ---
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-
-# 1)  working directory to the project root
-setwd("C:/Users/ajose35/Desktop/MADA-course/Joseph-MADA-project")
-
-# 2) Chunk defaults
-knitr::opts_chunk$set(
-  echo    = FALSE,
-  message = FALSE,
-  warning = FALSE
-)
-```
 
@@
 
-&nbsp;
-&nbsp;
-
 **Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes**
@@
 **Figure 3: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)**
-
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-# Generate predicted probabilities over the range of gravidity
-predicted_probs <- ggeffect(model_gravidity, terms = "gravidity")
-
-# Create a scatter/line plot of predicted probabilities by gravidity
-ggplot(predicted_probs, aes(x = x, y = predicted)) +
-  geom_line(size = 1.2) +
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "blue") +
-  labs(title = "",
-       x = "Gravidity",
-       y = "Predicted Probability") +
-  theme_minimal()
+knitr::include_graphics( here::here( "Images", "Figure 3. Predicted Probability of Adverse Outcome by Gravidity Age  25.png"
+  )
+)
 

### 2025-04-24 products/manuscript/Manuscript.qmd
prj2

@@
 **Figure 4 : Comparative Discriminative Performance of ML Models for Predicting Adverse Birth Outcomes**
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics(
+  here::here(
+    "Images",
+    "Figure 4.Comparative Discriminative Performance of ML .png"
+  )
+)
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-roc_lr  <- roc_curve(logistic_preds, truth = adverse_birth_outcome, .pred_1) %>% mutate(model = "Logistic")
-roc_rf  <- roc_curve(rf_preds,       truth = adverse_birth_outcome, .pred_1) %>% mutate(model = "RF")
-roc_xgb <- roc_curve(xgb_preds,     truth = adverse_birth_outcome, .pred_1) %>% mutate(model = "XGBoost")
-
-roc_all <- bind_rows(roc_lr, roc_rf, roc_xgb)
-
-ggplot(roc_all, aes(x = 1 - specificity, y = sensitivity, color = model)) +
-  geom_path(size = 1) +
-  labs(title = "",
-       x = "False Positive Rate", y = "True Positive Rate") +
-  theme_minimal()
 ```

### 2025-04-24 products/manuscript/Manuscript.qmd
project3
update p

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics(here::here("Images", "Adjusted OR for Adverse Birth Outcomes Women  less 25 Years.png"))
+# Build a normalized, forward‐slash path and include the image
+or_u25_path <- normalizePath(
+  here::here(
+    "Images",
+    "Adjusted OR for Adverse Birth Outcomes Women  less 25 Years.png"
+  ),
+  winslash = "/"
+)
+knitr::include_graphics(or_u25_path)
 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics( here::here( "Images", "Figure 3. Predicted Probability of Adverse Outcome by Gravidity Age  25.png"
+# Construct an absolute, forward‐slash path and include the image
+fig3_path <- normalizePath(
+  here::here(
+    "Images",
+    "Figure 3. Predicted Probability of Adverse Outcome by Gravidity Age  25.png"
+  ),
+  winslash = "/"
 )
-)
+knitr::include_graphics(fig3_path)
 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics(
+# Build an absolute, forward‐slash path and then include the graphic
+fig4_path <- normalizePath(
   here::here(
@@
     "Figure 4.Comparative Discriminative Performance of ML .png"
+  ),
+  winslash = "/"
 )
-)
+
+knitr::include_graphics(fig4_path)
 

### 2025-04-24 products/manuscript/Manuscript.qmd
IMG

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# Build a normalized, forward‐slash path and include the image
-or_u25_path <- normalizePath(
-  here::here(
-    "Images",
-    "Adjusted OR for Adverse Birth Outcomes Women  less 25 Years.png"
-  ),
-  winslash = "/"
+knitr::include_graphics(
+  here::here("Images", "Adjusted ratio   less 25 Years.png")
 )
-knitr::include_graphics(or_u25_path)
 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# Construct an absolute, forward‐slash path and include the image
-fig3_path <- normalizePath(
-  here::here(
-    "Images",
-    "Figure 3. Predicted Probability of Adverse Outcome by Gravidity Age  25.png"
-  ),
-  winslash = "/"
+knitr::include_graphics(
+  here::here("Images", "Figure 3. Predicted Probability.png")
 )
-knitr::include_graphics(fig3_path)
 

### 2025-04-24 products/manuscript/Manuscript.qmd
Images2

@@
 ---
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+# 1) Load here() before you call i_am()
+library(here)
+
+# 2) Tell here() where this document lives, relative to the project root.
+#    (Your project root is "Joseph-MADA-project", so we point to "products/manuscript/Manuscript.qmd")
+here::i_am("products/manuscript/Manuscript.qmd")
+
+# 3) (Optional) now load knitr and set the knit root to the project root
+library(knitr)
+opts_knit$set(root.dir = here::here())
+
+```
 
@@
 **Table 5: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+```{r fig-adjusted-u25, echo=FALSE, message=FALSE, warning=FALSE}
 knitr::include_graphics(
-  here::here("Images", "Adjusted ratio   less 25 Years.png")
+  normalizePath(
+    here::here("Images", "AdjustedRatio.png"),
+    winslash = "/"
+  )
 )
@@
 knitr::include_graphics(
-  here::here("Images", "Figure 3. Predicted Probability.png")
+  normalizePath(
+    here::here("Images", "Figure 3. Predicted Probability.png"),
+    winslash = "/"
+  )
 )

### 2025-04-25 products/manuscript/Manuscript.qmd
Tab1
IMGs

@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# 1) Load here() before you call i_am()
+```{r setup, include=FALSE}
+# 1) identify project root for here()
 library(here)
-
-# 2) Tell here() where this document lives, relative to the project root.
-#    (Your project root is "Joseph-MADA-project", so we point to "products/manuscript/Manuscript.qmd")
 here::i_am("products/manuscript/Manuscript.qmd")
 
-# 3) (Optional) now load knitr and set the knit root to the project root
+# 2) tell knitr to use that root (so all relative paths start there)
 library(knitr)
-opts_knit$set(root.dir = here::here())
-
+knitr::opts_knit$set(root.dir = here::here())
 ```
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics(here::here("Images", "Table 1. Baseline Characteristics .png"))
+```{r fig-table1, echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("Images/Table1Baseline.png")
 
@@
 ```{r,echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics(here::here("Images", "Table4.Interaction Malaria vsIPTP.png"))
-
+knitr::include_graphics("Images/Table4Interaction.png")
 ```

### 2025-04-25 products/manuscript/Manuscript.qmd
img

@@
     highlight-style: github
+resources:
+  - Images/**
+execute:
+  root: .
 bibliography: ../../assets/dataanalysis-references.bib
@@
 ```{r fig-table1, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics("Images/Table1Baseline.png")
+# 2) Build the absolute, Pandoc-friendly path
+table1_path <- normalizePath(
+  here::here("products", "manuscript", "Images", "Table1Baseline.png"),
+  winslash = "/"
+)
 
+# 3) Embed the image
+knitr::include_graphics(table1_path)
 ```
@@
 ```{r,echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics("Images/Table4Interaction.png")
+# 2) Build the absolute, Pandoc-friendly path to Table4Interaction.png
+table4_path <- normalizePath(
+  here::here("products", "manuscript", "Images", "Table4Interaction.png"),
+  winslash = "/"
+)
+
+# 3) Embed the image
+knitr::include_graphics(table4_path)
 ```
@@
 ```{r fig-adjusted-u25, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics(
-  normalizePath(
-    here::here("Images", "AdjustedRatio.png"),
+# 2) Build a forward-slash absolute path to the image
+adjusted_path <- normalizePath(
+  here::here("products", "manuscript", "Images", "AdjustedRatio.png"),
   winslash = "/"
 )
-)
 
+# 3) Embed the image
+knitr::include_graphics(adjusted_path)
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics(
-  normalizePath(
-    here::here("Images", "Figure 3. Predicted Probability.png"),
+# 2) Build the absolute, Pandoc-friendly path to Figure 3
+fig3_path <- normalizePath(
+  here::here(
+    "products", "manuscript", "Images",
+    "Figure 3. Predicted Probability.png"
+  ),
   winslash = "/"
 )
-)
 
+# 3) Embed the image
+knitr::include_graphics(fig3_path)
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# Build an absolute, forward‐slash path and then include the graphic
+# 2) Build the absolute, Pandoc-friendly path to Figure 4
 fig4_path <- normalizePath(
   here::here(
-    "Images",
+    "products", "manuscript", "Images",
     "Figure 4.Comparative Discriminative Performance of ML .png"
@@
 
+# 3) Embed the image
 knitr::include_graphics(fig4_path)
-
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics(here::here("Images", "Table6.Discriminative Performance (AUC) of ML Models.png"))
+# 2) Build the absolute, Pandoc‐friendly path to Table 6
+table6_path <- normalizePath(
+  here::here(
+    "products", "manuscript", "Images",
+    "Table6.Discriminative Performance (AUC) of ML Models.png"
+  ),
+  winslash = "/"
+)
 
+# 3) Embed the image
+knitr::include_graphics(table6_path)
 ```

### 2025-04-25 products/manuscript/Manuscript.qmd
proj1

--- 2025-04-25 products/manuscript/Manuscript.qmd
@@
     highlight-style: github
+resources:
+  - Images/**
+execute:
+  root: products/manuscript
 bibliography: ../../assets/dataanalysis-references.bib
@@
 
-```{r setup, include=FALSE}
-# 1) identify project root for here()
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 library(here)
@@
 
-# 2) tell knitr to use that root (so all relative paths start there)
+# ← THIS is the one that actually sets the working dir for *all* chunks:
 library(knitr)
-knitr::opts_knit$set(root.dir = here::here())
+knitr::opts_chunk$set(root.dir = here::here("products/manuscript"))
+
 ```
@@
 Malaria in pregnancy contributes substantially to maternal and neonatal morbidity and mortality across Sub‑Saharan Africa.
-Intermittent preventive treatment during pregnancy (IPTp) is recommended to reduce adverse birth outcomes, yet the comparative effectiveness of different regimens and how their impact varies with malaria burden remains unclear.
+Intermittent preventive treatment during pregnancy (IPTp) with sulfadoxine–pyrimethamine (SP) or dihydroartemisinin–piperaquine (DP) is standard, but how regimen choice interacts with cumulative malaria episodes to shape birth outcomes is unclear.
 
 ## Methods
-We conducted a secondary analysis of data from a double‑blind randomized controlled trial in Uganda (ClinEpiDB Release #24, August 30, 2022) involving 782 HIV‑uninfected pregnant women.
+We analyzed data from a double-blind, randomized trial in Uganda (ClinEpiDB #24, n=782 HIV-uninfected pregnant women).
 Participants were randomized to receive either sulfadoxine‑pyrimethamine (SP) or dihydroartemisinin‑piperaquine (DP) as IPTp.
-Multivariable logistic regression was employed to evaluate (1) the main effects of cumulative malaria episodes and treatment arm on preterm birth, low birth weight, and stillbirth, and (2) the interaction between malaria frequency and IPTp regimen.
-Models adjusted for maternal age, gravidity, and socioeconomic status.
-A prespecified subgroup analysis in women < 25 years assessed the effect of gravidity on outcomes.
+Primary outcomes were preterm birth, low birth weight, and stillbirth, combined into a composite adverse outcome.
+We fit multivariable logistic models evaluating (1) main effects of total malaria episodes and treatment arm, (2) their interaction, adjusting for maternal age, gravidity, and education, and (3) a prespecified subgroup analysis in women < 25 years to assess gravidity’s protective effect.
+Predicted probabilities were derived via ggeffects, and discriminative performance of several ML models (random forest, XGBoost, logistic regression) was summarized by AUC.
 
@@
 In multivariable models adjusted for maternal age, gravidity, and socioeconomic status, we identified a statistically significant interaction between malaria episode frequency and the SP regimen (OR 1.47 per additional episode, 95% CI 1.16–1.88; p = 0.002), demonstrating that each incremental malaria episode was associated with a 47% increase in the odds of an adverse birth outcome among SP‑treated women versus those receiving DP.
-Notably, neither malaria burden nor IPTp regimen alone achieved statistical significance, underscoring the critical role of their interplay.
+Notably, neither malaria burden nor IPTp regimen alone achieved statistical significance, Total malaria episodes (OR=0.88 per episode, 95% CI 0.75–1.03; p=0.12) and treatment arm alone (OR=0.69 for DP vs SP, 95% CI 0.44–1.08; p=0.10), underscoring the critical role of their interplay.
 Moreover, in the subgroup of women under 25 years, gravidity conferred significant protection: each additional pregnancy reduced the likelihood of an adverse outcome by 14% (OR 0.86 per gravidity, 95% CI 0.74–1.00; p = 0.044).
-These findings suggest that SP’s preventive efficacy diminishes with repeated malaria exposure and that higher parity provides resilience against adverse perinatal outcomes in younger mothers.
+Finally, ML models showed modest discrimination (best AUC ≈ 0.60.
 
@@
 ## Keywords:
-**Malaria, adverse birth outcomes, intermittent preventive treatment in pregnancy (IPTp), sulfadoxine‑pyrimethamine, dihydroartemisinin‑piperaqui**
+*Malaria, adverse birth outcomes, intermittent preventive treatment in pregnancy (IPTp), sulfadoxine‑pyrimethamine, dihydroartemisinin‑piperaqui*
 
@@
 ## General Background Information
-Malaria remains a major public health challenge worldwide, particularly in Sub-Saharan Africa, where the disease disproportionately affects vulnerable populations.
+Malaria remains a significant public health challenge worldwide, particularly in Sub-Saharan Africa, where the disease disproportionately affects vulnerable populations.
 Caused by Plasmodium parasites transmitted through the bite of infected Anopheles mosquitoes, malaria accounted for an estimated 249 million cases globally in 2022, marking an increase of five million cases compared to the previous year.
@@
 While these preventive measures have shown significant benefits in reducing the risk of malaria-related complications, the effectiveness of these regimens may not be uniform across all contexts.
-In particular, emerging evidence suggests that the choice of IPTp regimen may influence how malaria episodes affect pregnancy outcomes.
+Emerging evidence suggests that the choice of IPTp regimen may influence how malaria episodes affect pregnancy outcomes.
 However, this potential effect modification remains underexplored in current literature.
-Additionally, maternal characteristics such as gravidity, the number of times a woman has been pregnant may also play a role in shaping birth outcomes.
+Additionally, maternal characteristics such as gravidity and the number of times a woman has been pregnant may also play a role in shaping birth outcomes.
 Prior research has suggested that previous pregnancy experience may offer protective benefits against adverse outcomes, possibly due to improved physiological adaptation or better health-seeking behavior.
@@
 In addressing these questions, the study seeks to clarify how preventive treatment strategies and maternal reproductive history shape birth outcomes in malaria-endemic settings.
-Ultimately, the insights gained could inform the development of targeted, effective, and equitable interventions to enhance pregnant women's and infants' health in Uganda and similar regions.
+Ultimately, the insights gained could inform the development of targeted, effective, and equitable interventions to enhance pregnant women’s and infants’ health in Uganda and similar regions.
 
@@
 
-Multivariable logistic regression models were employed to evaluate whether the IPTp regimen modifies the association between malaria episodes and adverse birth outcomes, adjusting for maternal age, gravidity, and socioeconomic status, and incorporating an interaction term between malaria episodes and treatment regimen.
+Multivariable logistic regression models were employed to evaluate whether the IPTp regimen modifies the association between malaria episodes and adverse birth outcomes, adjusting for maternal age, gravidity, and socioeconomic status and incorporating an interaction term between malaria episodes and treatment regimen.
 Predicted probabilities were calculated to elucidate these interaction effects.
-Additionally, a subgroup analysis among women under 25 years assessed the relationship between gravidity treated as a continuous variable, and adverse birth outcomes, with adjustments made for maternal age, malaria episode frequency, and treatment regimen.
+Additionally, a subgroup analysis among women under 25 years assessed the relationship between gravidity treated as a continuous variable and adverse birth outcomes, with adjustments made for maternal age, malaria episode frequency, and treatment regimen.
 
@@
 
-```{r fig-table1, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 knitr::include_graphics("Images/Table1Baseline.png")
@@
 ```
-
-&nbsp;
-&nbsp;
 
@@
 
-Baseline characteristics were similar across IPTp treatment arms, with comparable maternal age, gestational age, education, gravidity, and parity (Table 1).
-Although overall malaria exposure was similar between groups, the SP arm exhibited a significantly higher stillbirth rate (14% vs.
-9%; p = 0.044) (Table 2).
+#Baseline characteristics and unadjusted outcomes
+Of the 782 women randomized, 412 received DP, and 370 received SP.
+The mean maternal age was 25.8 years (SD ± 5.1) in the DP arm and 26.2 years (± 5.3) in the SP arm (p = 0.34).
+Median gravidity was 2 (IQR 1–3) in both arms and mean total malaria episodes during pregnancy was 1.03 (± 0.58) for DP versus 1.05 (± 0.58) for SP (p = 0.30).
+Education level, socioeconomic status, and baseline nutritional indicators were similarly balanced (all p > 0.20) (Table 1).
 
-Multivariable logistic regression analyses, adjusted for maternal age, gravidity, and education revealed that the main effects of total malaria episodes (OR = 0.88, 95% CI: 0.75–1.03, p = 0.121) and treatment arm (OR = 0.69, 95% CI: 0.44–1.08, p = 0.102) were not statistically significant.
-However, incorporating an interaction term between malaria episodes and the IPTp regimen significantly improved model fit (ΔDeviance = 10.11, p = 0.00148), with the interaction itself reaching significance (OR = 1.47, 95% CI: 1.16–1.88, p = 0.002).
-These findings suggest that the impact of malaria exposure on adverse birth outcomes is modified by the treatment regimen.
-Additionally, gravidity demonstrated a protective effect (OR = 0.89, 95% CI: 0.80–0.98, p = 0.023), indicating that prior pregnancy experience reduces the risk of adverse outcomes.
+When stratifying individual adverse outcomes, rates of low birth weight (< 2.5 kg) were 21% in DP vs.
+23% in SP (p = 0.50), and preterm birth (≥ 1 episode) occurred in 6.1% vs.
+6.8% (p = 0.60).
+The only statistically significant univariate difference was stillbirth: 9.0% (n = 37) in DP versus 14.1% (n = 52) in SP (χ² = 4.09; p = 0.044) (Table 2; Figure 1).
+The composite adverse outcome (any of the three) occurred in 232/412 (56.3%) DP women and 219/370 (59.2%) SP women (p = 0.40).
 
@@
 knitr::include_graphics("Images/Table4Interaction.png")
-```
-
-&nbsp;
-&nbsp;
-
-Baseline characteristics were similar across IPTp treatment arms, with comparable maternal age, gestational age, education, gravidity, and parity (Table 1).
-Although overall malaria exposure was similar between groups, the SP arm exhibited a significantly higher stillbirth rate (14% vs.
-9%; p = 0.044) (Table 2).
-
-Multivariable logistic regression analyses, adjusted for maternal age, gravidity, and education revealed that the main effects of total malaria episodes (OR = 0.88, 95% CI: 0.75–1.03, p = 0.121) and treatment arm (OR = 0.69, 95% CI: 0.44–1.08, p = 0.102) were not statistically significant.
-However, incorporating an interaction term between malaria episodes and the IPTp regimen significantly improved model fit (ΔDeviance = 10.11, p = 0.00148), with the interaction itself reaching significance (OR = 1.47, 95% CI: 1.16–1.88, p = 0.002).
-These findings suggest that the impact of malaria exposure on adverse birth outcomes is modified by the treatment regimen.
-Additionally, gravidity demonstrated a protective effect (OR = 0.89, 95% CI: 0.80–0.98, p = 0.023), indicating that prior pregnancy experience reduces the risk of adverse outcomes.
-
-&nbsp;
-&nbsp;
-
-**Table 5: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
-
-```{r fig-adjusted-u25, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics(
-  normalizePath(
-    here::here("Images", "AdjustedRatio.png"),
-    winslash = "/"
-  )
-)
 
@@
 
-**Figure 3: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)**
+#Interaction between malaria episodes and IPTp regimen
+In a multivariable logistic regression adjusting for maternal age (continuous), gravidity (continuous), and education level (categorical), neither total malaria episodes (OR 0.88 per episode; 95% CI 0.75–1.03; p = 0.12) nor treatment arm (OR 0.69 for DP vs.
+SP; 95% CI 0.44–1.08; p = 0.10) were independently significant predictors of the composite adverse outcome.
+
+dding an interaction term between total malaria episodes and SP (reference = DP) significantly improved model fit (ΔDeviance = 10.11 on 1 df; p = 0.0015).
+The interaction coefficient (SP × episodes) was highly significant (OR 1.47; 95% CI 1.16–1.88; p = 0.002), indicating that each additional malaria episode increases the odds of an adverse outcome by 47% more under SP than under DP.
+In practical terms, at zero episodes, the predicted probability of an adverse outcome is similar for both arms (~ 0.55) but diverges sharply by three episodes: ~ 0.62 for DP versus ~ 0.78 for SP (Figure 2).
+Full model estimates are given in Table 4.
+
+&nbsp;
+&nbsp;
+
+#Subgroup analysis in women < 25 years
+Among the 284 women under 25 years old (150 DP, 134 SP), 61.3% (174/284) experienced a composite adverse outcome.
+In a multivariable logistic regression, including gravidity, total malaria episodes, treatment arm, and maternal age, only gravidity remained significant: each additional pregnancy reduced the odds of an adverse outcome by 14% (OR 0.86, 95% CI 0.74–1.00; p = 0.044), while malaria episodes (p = 0.15) and treatment arm (p = 0.21) were not.
+Predicted probabilities from this model (Figure 3; Table 5) show that a first‐pregnancy woman under 25 has a 72% risk of an adverse outcome, declining to 55% by her third pregnancy and below 40% by her fourth, underscoring parity’s protective effect in younger mothers.
+
+**Table 5: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics(
-  normalizePath(
-    here::here("Images", "Figure 3. Predicted Probability.png"),
-    winslash = "/"
-  )
-)
+knitr::include_graphics("Images/AdjustedRatio.png")
 
@@
 
-Based on our findings, the impact of malaria on adverse birth outcomes is significantly modified by the IPTp regimen, with a distinct interaction effect observed for the SP regimen.
-Additionally, increased gravidity appears to offer a protective benefit, particularly among women under 25 years.
-Detailed results supporting these observations are presented in Tables 3–6 and Figures 1–3.
+#Machine‐learning model performance
+Given the limited gains from our traditional regression models, we developed a comprehensive machine‐learning pipeline in tidy models to evaluate whether more flexible algorithms could improve the composite adverse birth outcome prediction.
+We randomly split our cohort of 782 women into an 80% training set (n = 626) and a 20% hold-out test set (n = 156), stratifying on the adverse outcome to maintain its 58% prevalence.
+All continuous predictors, maternal age, gravidity, and total malaria episodes were centered and scaled, while categorical variables (education level, treatment arm, and socioeconomic status) were dummy encoded.
+
+Within the training set, we used 5-fold cross-validation to tune hyperparameters for three tree-based methods (random forest, XGBoost) and a penalized regression (elastic net), optimizing ROC AUC in each case.
+Specifically, we tuned to try, and tree counts for the random forest, tree depth, learning rate, and number of trees for XGBoost, penalty strength, and ridge-versus-lasso mixing for the elastic net.
+A plain logistic regression (no penalty) served as our baseline.
+After identifying the best parameters, we refit each model on the full training data.
+
+When evaluated on the independent test set, all four approaches performed at or below chance: logistic regression AUC = 0.48, random forest = 0.42, XGBoost = 0.44, and elastic net = 0.49 (Table 6; Figure 4).
+Even the strongest elastic net model fell essentially to random levels, underscoring that these routinely collected demographic and clinical variables lack sufficient predictive signal for reliable individual risk classification.
+These results highlight the need to integrate richer clinical, laboratory, or biomarker data in future predictive efforts to achieve meaningful discrimination.
+
+**Figure 3: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)**
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("Images/Figure 3. Predicted Probability.png")
+
+```
 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# Build an absolute, forward‐slash path and then include the graphic
-fig4_path <- normalizePath(
-  here::here(
-    "Images",
-    "Figure 4.Comparative Discriminative Performance of ML .png"
-  ),
-  winslash = "/"
-)
-
-knitr::include_graphics(fig4_path)
+knitr::include_graphics("Images/Figure 4.Comparative Discriminative Performance of ML .png")
 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics(here::here("Images", "Table6.Discriminative Performance (AUC) of ML Models.png"))
+knitr::include_graphics("Images/Table6.Discriminative Performance (AUC) of ML Models.png")
 
@@
 # Discussion
-Our data analysis from a randomized controlled trial in Uganda provides important insights into the interplay between malaria exposure, IPTp regimen, and adverse birth outcomes.
-Specifically, our findings reveal that while the individual effects of total malaria episodes and the treatment arm were not statistically significant, the interaction between malaria episodes and the SP regimen significantly increased the risk of adverse outcomes.
-Additionally, increased gravidity emerged as a robust protective factor, particularly among pregnant women under 25 years, with each additional pregnancy associated with a roughly 14% reduction in risk.
+Our analysis from a randomized controlled trial in Uganda illuminates three key findings.
+First, although total malaria episodes and IPTp regimen alone did not predict adverse birth outcomes, their interaction was highly significant.
+Each additional malaria episode under sulfadoxine-pyrimethamine (SP) conferred 47% higher odds of a composite adverse outcome than dihydroartemisinin-piperaquine (DP).
+Second, in women under 25 years, gravidity emerged as the sole significant predictor, with each additional pregnancy reducing the odds of an adverse outcome by 14%.
+Finally, our machine‐learning pipeline, comparing logistic regression, random forest, XGBoost, and elastic net, demonstrated uniformly poor discrimination (AUCs 0.42–0.49), underscoring that routinely collected demographic and clinical variables alone cannot predict individual risk.
 
-These results align with previous studies that have highlighted the complex relationship between maternal characteristics and malaria-related complications during pregnancy.
-For example, similar protective effects of gravidity have been reported in the literature (Chua et al., 2021), and emerging evidence suggests that the efficacy of IPTp regimens may vary depending on the context and specific treatment used (Talapko et al., 2019 World Health Organization, 2023).
-Our findings extend these observations by demonstrating that the SP regimen, in particular, modifies the impact of malaria exposure on birth outcomes, emphasizing the need for tailored malaria prevention strategies.
+These results extend prior work on malaria in pregnancy.
+The protective effect of higher gravidity aligns with studies suggesting that physiologic adaptations and improved immune tolerance in multiparous women reduce placental parasite sequestration and its sequelae (Chua et al., 2021).
+Likewise, emerging evidence indicates that SP’s efficacy may wane as malaria exposure increases, potentially due to rising parasite resistance or immune modulation (Talapko et al., 2019; World Health Organization, 2023).
+By explicitly modeling the interaction between episodes and regimen, we demonstrate that SP’s preventive benefit diminishes sharply with each successive infection, an effect not captured when examining the main effects alone.
 
-However, our study has several limitations.
-First, the data were derived from a controlled trial setting, which may limit the generalizability of the findings to broader, real-world populations.
-Second, the necessity to exclude or impute variables with significant missing data could have influenced the robustness of our estimates.
-Third, the modest discriminative ability of our models reflected in ROC curves with AUC values near 0.55–0.60 suggests that additional, unmeasured factors might contribute to adverse birth outcomes.
+From a public health standpoint, these findings argue for tailoring IPTp strategies to individual risk profiles.
+In high‐transmission settings, women receiving SP who experience multiple breakthrough infections may benefit from switching to DP or augmenting chemoprevention with additional interventions (e.g., enhanced screening, bed‐net distribution).
+Moreover, gravidity should inform risk stratification: primigravidae and secundigravidae under 25 years represent a particularly vulnerable group that might warrant closer monitoring or more aggressive prevention.
 
-Despite these limitations, our study offers valuable contributions to the field.
-The evidence that IPTp regimen choice can modify the relationship between malaria episodes and adverse birth outcomes, along with the protective role of gravidity, underscores the importance of developing more personalized intervention strategies.
-Future research should aim to validate these findings in more diverse settings and explore other potential determinants of maternal and neonatal health in malaria-endemic regions.
-
-In summary, our findings advocate for a nuanced approach to malaria prevention in pregnancy that considers both the specific IPTp regimen and individual maternal reproductive history.
-Such tailored strategies have the potential to substantially improve maternal and neonatal health outcomes in regions where malaria remains a significant public health challenge.
-
-## Summary and Interpretation
-
-*Summarize what you did, what you found and what it means.*
-
-## Strengths and Limitations
-
-*Discuss what you perceive as strengths and limitations of your analysis.*
+Our study has several limitations.
+As a secondary analysis of trial data, generalizability to real‐world antenatal populations may be constrained by trial eligibility criteria and adherence patterns.
+Variable exclusion or imputation to address missingness could have introduced bias, and unmeasured factors, such as HIV status, nutritional deficiencies, or placental pathology, likely contribute to adverse outcomes but were not available.
+Finally, the poor performance of our machine‐learning models highlights the need for richer predictor sets; demographic and simple clinical measures are insufficient to classify individual risk with confidence.
 
 ## Conclusions
+These findings demonstrate a critical need to move beyond one‐size‐fits‐all IPTp recommendations.
+Our analysis points toward stratified prevention strategies by revealing that sulfadoxine–pyrimethamine’s protective efficacy erodes with each additional malaria episode and that higher gravidity markedly shields younger mothers.
+Tailoring IPTp regimens based on individual exposure histories, prioritizing dihydroartemisinin-piperaquine for women facing recurrent infections under SP, and intensifying support for primigravidae and secundigravidae could substantially reduce adverse birth outcomes.
+
+Moreover, the limited predictive power of standard clinical and demographic measures highlights the urgency of integrating richer data sources, such as immunologic biomarkers, placental histopathology, and parasite genotyping, to refine both risk stratification and mechanistic understanding.
+Adapting policy and research agendas to these insights offers a clear path toward more effective, personalized malaria prevention in pregnancy.
 

### 2025-04-25 products/manuscript/Manuscript.qmd
update

@@
 ## Keywords:
-*Malaria, adverse birth outcomes, intermittent preventive treatment in pregnancy (IPTp), sulfadoxine‑pyrimethamine, dihydroartemisinin‑piperaqui*
+**Malaria, Adverse birth outcomes, Intermittent preventive treatment in pregnancy (IPTp), Sulfadoxine–pyrimethamine, Dihydroartemisinin–piperaquine, Gravidity/Parity**
 
@@
 
-#Baseline characteristics and unadjusted outcomes
+**Baseline characteristics and unadjusted outcomes**
 Of the 782 women randomized, 412 received DP, and 370 received SP.
@@
 
-#Interaction between malaria episodes and IPTp regimen
+**Interaction between malaria episodes and IPTp regimen**
 In a multivariable logistic regression adjusting for maternal age (continuous), gravidity (continuous), and education level (categorical), neither total malaria episodes (OR 0.88 per episode; 95% CI 0.75–1.03; p = 0.12) nor treatment arm (OR 0.69 for DP vs.
@@
 
-#Subgroup analysis in women < 25 years
+**Subgroup analysis in women < 25 years**
 Among the 284 women under 25 years old (150 DP, 134 SP), 61.3% (174/284) experienced a composite adverse outcome.
@@
 
-#Machine‐learning model performance
+**Machine‐learning model performance**
 Given the limited gains from our traditional regression models, we developed a comprehensive machine‐learning pipeline in tidy models to evaluate whether more flexible algorithms could improve the composite adverse birth outcome prediction.
@@
 
-*What are the main take-home messages?*
+#  Remove References
 

### 2025-04-25 products/manuscript/Manuscript.qmd
update1

@@
 
-# Introductio
+# Introduction
 ## General Background Information
@@
 **Baseline characteristics and unadjusted outcomes**
+
 Of the 782 women randomized, 412 received DP, and 370 received SP.
@@
 **Interaction between malaria episodes and IPTp regimen**
+
 In a multivariable logistic regression adjusting for maternal age (continuous), gravidity (continuous), and education level (categorical), neither total malaria episodes (OR 0.88 per episode; 95% CI 0.75–1.03; p = 0.12) nor treatment arm (OR 0.69 for DP vs.
@@
 **Subgroup analysis in women < 25 years**
+
 Among the 284 women under 25 years old (150 DP, 134 SP), 61.3% (174/284) experienced a composite adverse outcome.
@@
 **Machine‐learning model performance**
+
 Given the limited gains from our traditional regression models, we developed a comprehensive machine‐learning pipeline in tidy models to evaluate whether more flexible algorithms could improve the composite adverse birth outcome prediction.
@@
 
-#  Remove References
+#  Remove Section
 

### 2025-04-25 products/manuscript/Manuscript.qmd
Project

@@
 
-# Background
+**Background:**
 Malaria in pregnancy contributes substantially to maternal and neonatal morbidity and mortality across Sub‑Saharan Africa.
@@
 
-## Methods
+**Methods:**
 We analyzed data from a double-blind, randomized trial in Uganda (ClinEpiDB #24, n=782 HIV-uninfected pregnant women).
@@
 
-## Results
+**Results:**
 In multivariable models adjusted for maternal age, gravidity, and socioeconomic status, we identified a statistically significant interaction between malaria episode frequency and the SP regimen (OR 1.47 per additional episode, 95% CI 1.16–1.88; p = 0.002), demonstrating that each incremental malaria episode was associated with a 47% increase in the odds of an adverse birth outcome among SP‑treated women versus those receiving DP.
@@
 
-## Conclusions:
-These results demonstrate that IPTp regimen selection fundamentally alters how cumulative malaria exposure impacts perinatal outcomes: sulfadoxine–pyrimethamine exhibits progressively reduced protective efficacy with each successive malaria episode, whereas dihydroartemisinin–piperaquine maintains greater resilience.
+**Conclusions:**
+IPTp regimen selection fundamentally alters how cumulative malaria exposure impacts perinatal outcomes: sulfadoxine–pyrimethamine exhibits progressively reduced protective efficacy with each successive malaria episode, whereas dihydroartemisinin–piperaquine maintains greater resilience.
 Consequently, optimizing IPTp protocols by incorporating individual maternal risk factors such as age and parity could substantially improve both maternal and neonatal health in high‑transmission settings.
@@
 ## Keywords:
-**Malaria, Adverse birth outcomes, Intermittent preventive treatment in pregnancy (IPTp), Sulfadoxine–pyrimethamine, Dihydroartemisinin–piperaquine, Gravidity/Parity**
+**Malaria, Adverse birth outcomes, Gravidity/Parity, Intermittent preventive treatment in pregnancy (IPTp), Sulfadoxine–pyrimethamine, Dihydroartemisinin–piperaquine**
 

### 2025-04-25 products/manuscript/Manuscript.qmd
Update2

@@
 
-## Keywords:
+**Keywords:**
+
 **Malaria, Adverse birth outcomes, Gravidity/Parity, Intermittent preventive treatment in pregnancy (IPTp), Sulfadoxine–pyrimethamine, Dihydroartemisinin–piperaquine**
@@
 
+&nbsp;
+&nbsp;
+
 **Table 5: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
@@
 
+&nbsp;
+&nbsp;
+
 **Figure 3: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)**
@@
 ```
+
+&nbsp;
+&nbsp;
 

### 2025-04-26 code/analysis-code/Statistical_Analysis.qmd
Update P

@@
 # Path management
-library(here)              # For file path handling
+library(here)               # File path handling
 
@@
 library(dplyr)              # Data manipulation
-library(tidyverse)         # Core data science packages (includes ggplot2, dplyr, readr, etc.)
+library(tidyverse)          # Core tidyverse packages (ggplot2, readr, etc.)
 library(janitor)            # Clean column names
-library(skimr)             # Data summary
-library(lubridate)         # Date and time handling
-library(forcats)           # Handling categorical variables
+library(skimr)              # Quick data summaries
+library(lubridate)          # Date/time handling
+library(forcats)            # Categorical variable tools
 
 # Visualization
-library(ggplot2)           # Data visualization (also part of tidyverse)
+library(ggplot2)            # Data visualization
 library(ggpubr)             # Publication-ready plots
@@
 # Tables and reporting
-library(gtsummary)         # Summary statistics tables
-library(gt)                # Table formatting
+library(gtsummary)          # Summary tables
+library(gt)                 # Advanced table formatting
 library(knitr)              # Report generation
-library(kableExtra)        # Enhanced table output in markdown
-
-# Analysis-specific packages
-library(Amelia)            # Visualizing missing data
+library(kableExtra)         # Enhanced markdown tables
+
+# Data exploration & preparation
+library(Amelia)             # Missing-data visualization
 library(pwr)                # Power analysis
 library(DiagrammeR)         # Diagrams and flowcharts
-library(survival)          # Survival analysis
+
+# Survival analysis
+library(survival)           # Survival models
 library(survminer)          # Survival plots
-library(ggeffects)         # Marginal effects for regression models
+
+# Model effects & outputs
+library(ggeffects)          # Marginal effects extraction
 library(broom)              # Tidy model outputs
 
-# Machine learning and modeling
-library(tidymodels)        # Unified modeling framework (includes recipes, parsnip, workflows, tune, etc.)
-library(rsample)           # Resampling and data splitting
-library(recipes)           # Preprocessing
-library(parsnip)           # Model specification
-library(workflows)         # Combine models and preprocessing
-library(tune)              # Hyperparameter tuning
-library(yardstick)         # Model performance metrics
+# Machine learning & modeling
+library(tidymodels)         # Modeling framework (recipes, parsnip, workflows, tune, yardstick, rsample)
+library(themis)             # Class-imbalance sampling (e.g., SMOTE)
 library(dials)              # Parameter tuning grids
 library(ranger)             # Random forest engine
-library(xgboost)           # Boosting engine
-library(themis)            # Handling class imbalance (e.g., SMOTE)
+library(xgboost)            # Gradient boosting engine
 
@@
 print(vif_values)
-
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-library(dplyr)
-library(knitr)
-library(kableExtra)
-library(here)
-
 # Generate Analysis of Deviance Table (Likelihood Ratio Test)
-anova(model_no_interaction, model_interaction, test = "LRT")
+anova_out <- anova(model_no_interaction, model_interaction, test = "LRT")
+
+# Convert to a clean data frame
 anova_df <- as.data.frame(anova_out)
 anova_df$Model <- c("Model 1 (No Interaction)", "Model 2 (With Interaction)")
+
+# Reorder and rename columns for readability
 anova_df <- anova_df %>%
   select(Model, `Resid. Df`, `Resid. Dev`, Df, Deviance, `Pr(>Chi)`)
+
 colnames(anova_df) <- c("Model", "Residual Df", "Residual Dev", "Df", "Deviance", "p-value")
 
+# Display table
 kable(
@@
 ```{r}
-library(dplyr)
-library(knitr)
-library(kableExtra)
-library(car)
-library(here)
-
+# Fit logistic regression model WITH interaction
+model2 <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm +
+                age_at_enrollment_years + gravidity + education_level,
+              family = binomial(link = "logit"),
+              data = promo_data_clean)
+
+```
+
+```{r}
 # Compute Variance Inflation Factors (VIF) for Model 2 (with interaction)
@@
 
+#  Evaluating Discrimination (ROC Curve and AUC)
+
 ```{r}
@@
 *Train/Test Split and Random Forest Model* which fits a random forest model and plots its ROC curve, is also used for Research Question 1 as an alternative ML approach
+```{r}
+# 1. Define Random Forest model
+rf_model <- rand_forest(
+  trees = 500,
+  mtry = tune(),
+  min_n = tune()
+) %>%
+  set_engine("ranger") %>%
+  set_mode("classification")
+
+# 2. Build the recipe (you already did this correctly)
+rf_recipe <- recipe(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
+                      age_at_enrollment_years + gravidity + education_level,
+                    data = train_data) %>%
+  step_dummy(all_nominal_predictors())
+
+# 3. Combine into a workflow
+rf_workflow <- workflow() %>%
+  add_model(rf_model) %>%
+  add_recipe(rf_recipe)
+
+# 4. Tune hyperparameters (optional next step)
+set.seed(123)
+rf_tune <- tune_grid(
+  rf_workflow,
+  resamples = vfold_cv(train_data, v = 10, strata = adverse_birth_outcome),
+  grid = 20,  # number of hyperparameter combos
+  metrics = metric_set(roc_auc)
+)
+
+```
 
@@
 
-```{r ml_improved_recipe, message=FALSE}
-# 1) Encode factors, engineer features, balance classes, normalize
+```{r ml_improved_recipe, message=FALSE, warning=FALSE}
+# 0) Cast to correct classes *before* creating cv_folds
+train_data <- train_data %>%
+  mutate(
+    total_malaria_episodes = as.numeric(total_malaria_episodes),
+    gravidity              = as.numeric(gravidity),
+    adverse_birth_outcome  = factor(adverse_birth_outcome, levels = c(0,1))
+  )
+
+test_data <- test_data %>%
+  mutate(
+    total_malaria_episodes = as.numeric(total_malaria_episodes),
+    gravidity              = as.numeric(gravidity),
+    adverse_birth_outcome  = factor(adverse_birth_outcome, levels = c(0,1))
+  )
+
+# 0b) Now create your 10-fold resamples on the cleaned train_data
+cv_folds <- vfold_cv(train_data, v = 10, strata = adverse_birth_outcome)
+
+# 1) Build the recipe: dummies, features, SMOTE, normalize
 improved_recipe <- recipe(adverse_birth_outcome ~ 
-                             total_malaria_episodes + study_arm +
-                             age_at_enrollment_years + gravidity + education_level,
+                             total_malaria_episodes +
+                             study_arm +
+                             age_at_enrollment_years +
+                             gravidity +
+                             education_level,
                            data = train_data) %>%
-  # a) one‑hot encode all factor predictors
   step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
-  # b) create polynomial & interaction features
   step_mutate(
@@
   ) %>%
-  # c) up‑sample minority class via SMOTE
   step_smote(adverse_birth_outcome) %>%
-  # d) center & scale all predictors
-  step_normalize(all_predictors())
-                          # (5)
-
-```
+  step_normalize(all_numeric_predictors())
+
+# 2) Specify elastic-net logistic regression
+enet_spec <- logistic_reg(
+    penalty = tune(),
+    mixture = tune()
+  ) %>%
+  set_engine("glmnet")
+
+# 3) Assemble into a workflow
+enet_wf <- workflow() %>%
+  add_recipe(improved_recipe) %>%
+  add_model(enet_spec)
+
+# 4) Tuning grid & 10-fold CV
+enet_grid <- grid_regular(
+  penalty(range = c(-5, -1), trans = log10_trans()),  # from 1e-5 to 1e-1
+  mixture(range = c(0, 1)),
+  levels = 5
+)
+
+enet_tune <- tune_grid(
+  enet_wf,
+  resamples = cv_folds,
+  grid      = enet_grid,
+  metrics   = metric_set(roc_auc)
+)
+
+# 5) Show the top AUC
+enet_tune %>%
+  collect_metrics() %>%
+  filter(.metric == "roc_auc") %>%
+  arrange(desc(mean))
+```
+
+```{r}
+library(dplyr)
+library(knitr)
+
+enet_tune %>%
+  collect_metrics() %>%
+  filter(.metric == "roc_auc") %>%
+  arrange(desc(mean)) %>%
+  slice_head(n = 5) %>%                # ← pull the top 5 rows
+  mutate(`AUC (mean ± SE)` = sprintf("%.3f \u00B1 %.3f", mean, std_err)) %>%
+  select(penalty, mixture, `AUC (mean ± SE)`) %>%
+  kable(
+    caption = "Table X. Top 5 elastic‐net hyperparameter combinations, ranked by mean cross‐validated AUC"
+  )
+
+```
+
+```{r}
+enet_tune %>%
+  collect_metrics() %>%
+  filter(.metric == "roc_auc") %>%
+  arrange(desc(mean)) %>%
+  head(5) %>%                          # ← base‐R head()
+  mutate(`AUC (mean ± SE)` = sprintf("%.3f \u00B1 %.3f", mean, std_err)) %>%
+  select(penalty, mixture, `AUC (mean ± SE)`) %>%
+  kable(
+    caption = "Table X. Top 5 elastic‐net hyperparameter combinations, ranked by mean cross‐validated AUC"
+  )
+
+```
+
+```{r}
+library(ggplot2)
+
+enet_tune %>% 
+  collect_metrics() %>% 
+  filter(.metric=="roc_auc") %>% 
+  ggplot(aes(x = log10(penalty), y = mixture, fill = mean)) +
+    geom_tile() +
+    scale_fill_viridis_c(name = "AUC") +
+    labs(
+      x = expression(log[10](lambda)),
+      y = expression(alpha),
+      title = "Heatmap of cross-validated AUC\nfor Elastic-net hyperparameter grid"
+    ) +
+    theme_minimal(base_size = 12)
+
+```
+
+##################################################################################################################################################################
 
@@
   mutate(
-    # Month of enrollment (seasonality)
-    enroll_month = lubridate::month(enrollment_date),
-    # Episodes per trimester (assuming 40‑week gestation)
-    episodes_per_trimester = total_malaria_episodes / (gestational_age_at_enrollment_weeks / (40/3)),
-    # Socioeconomic index (if available)
-    # wealth_index = as.numeric(education_level) * household_assets_score
-  )
-
-# Apply same transformations to test_data
-test_data <- test_data %>%
-  mutate(
     enroll_month = lubridate::month(enrollment_date),
@@
 
+test_data <- test_data %>%
+  mutate(
+    enroll_month = lubridate::month(enrollment_date),
+    episodes_per_trimester = total_malaria_episodes / (gestational_age_at_enrollment_weeks / (40/3))
+  )
+
 ```
@@
 imbalanced_recipe <- recipe(adverse_birth_outcome ~ ., data = train_data) %>%
-  # standard preprocessing
   step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
-  # adaptive synthetic sampling
+  step_mutate(
+    total_malaria_episodes = as.numeric(total_malaria_episodes),
+    gravidity = as.numeric(gravidity)
+  ) %>%
   step_adasyn(adverse_birth_outcome) %>%
@@
 ```{r}
-# Continue using enet_wf from before
-set.seed(2025)
+enet_wf <- workflow() %>%
+  add_model(enet_spec) %>%
+  add_recipe(imbalanced_recipe)
+
 enet_bayes <- tune_bayes(
@@
   resamples = cv_folds,
-  initial = enet_tune,       # seed the search from your grid results
-  iter    = 20,              # number of Bayesian iterations
+  initial = enet_tune,
+  iter = 20,
   metrics = metric_set(roc_auc)
@@
 ```
+
+##################################################################################################################################################################
+
+# More Machine Learning Analysis
+To explore whether flexible, data‐driven classifiers could outperform our regression models in predicting the composite adverse birth outcome, we built a standardized ML pipeline using **tidymodels**.
+We compared three algorithms—LASSO‐penalized logistic regression, random forest, and XGBoost—each trained on the same predictors: maternal age, gravidity, education level, total malaria episodes, treatment arm, and socioeconomic status.
+
+```{r}
+# 2. Read & split data
+df <- read_csv(here("data","clean","PROMO_Data_clean.csv")) %>%
+  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0,1)))
+
+set.seed(2025)
+split    <- initial_split(df, prop = 0.7, strata = adverse_birth_outcome)
+train    <- training(split)
+test     <- testing(split)
+cv_folds <- vfold_cv(train, v = 5, strata = adverse_birth_outcome)
+
+```
+
+```{r}
+# 3. Preprocessing recipe
+ml_rec <- recipe(adverse_birth_outcome ~ age_at_enrollment_years
+                  + gravidity
+                  + total_malaria_episodes
+                  + education_level
+                  + study_arm
+                  + socioeconomic_status,
+                data = train) %>%
+  step_dummy(all_nominal_predictors()) %>%  # one‐hot encode
+  step_zv(all_predictors()) %>%             # remove zero‐variance
+  step_normalize(all_numeric_predictors())  # center & scale
+
+```
+
+```{r}
+# 4. Model specifications & workflows
+
+## (a) LASSO‐penalized logistic regression
+lasso_mod <- logistic_reg(penalty = tune(), mixture = 1) %>%
+  set_engine("glmnet") %>%
+  set_mode("classification")
+
+lasso_wf  <- workflow() %>% add_model(lasso_mod) %>% add_recipe(ml_rec)
+
+## (b) Random forest
+rf_mod <- rand_forest(mtry = tune(), trees = 1000, min_n = tune()) %>%
+  set_engine("ranger") %>%
+  set_mode("classification")
+
+rf_wf  <- workflow() %>% add_model(rf_mod) %>% add_recipe(ml_rec)
+
+## (c) XGBoost
+xgb_mod <- boost_tree(trees = tune(),
+                      tree_depth = tune(),
+                      learn_rate = tune()) %>%
+  set_engine("xgboost") %>%
+  set_mode("classification")
+
+xgb_wf <- workflow() %>% add_model(xgb_mod) %>% add_recipe(ml_rec)
+
+```
+
+```{r}
+# 5. Tuning grids & 5‐fold CV
+set.seed(2025)
+
+lasso_grid <- grid_regular(penalty(), levels = 20)
+
+rf_params <- parameters(rf_mod) %>%
+  update(mtry = mtry(c(1, 6)))
+rf_grid   <- grid_latin_hypercube(rf_params, size = 20)
+
+xgb_params <- parameters(xgb_mod)
+xgb_grid   <- grid_latin_hypercube(xgb_params, size = 30)
+
+ml_metrics <- metric_set(roc_auc)
+
+lasso_res <- tune_grid(lasso_wf, resamples = cv_folds,
+                       grid = lasso_grid, metrics = ml_metrics)
+
+rf_res    <- tune_grid(rf_wf,    resamples = cv_folds,
+                       grid = rf_grid,    metrics = ml_metrics)
+
+xgb_res   <- tune_grid(xgb_wf,   resamples = cv_folds,
+                       grid = xgb_grid,   metrics = ml_metrics)
+
+```
+
+```{r}
+# 6. Collect cross‐validated AUC results
+lasso_auc <- lasso_res %>% filter(.metric=="roc_auc") %>% summarize(mean = mean(mean), se = mean(std_err)) %>% mutate(model="LASSO")
+rf_auc    <- rf_res    %>% filter(.metric=="roc_auc") %>% summarize(mean = mean(mean), se = mean(std_err)) %>% mutate(model="RF")
+xgb_auc   <- xgb_res   %>% filter(.metric=="roc_auc") %>% summarize(mean = mean(mean), se = mean(std_err)) %>% mutate(model="XGB")
+
+bind_rows(lasso_auc, rf_auc, xgb_auc) %>%
+  rename(AUC = mean, SE = se) %>%
+  knitr::kable(digits=3, caption="Table 6: 5‐fold CV ROC AUC for Each Model")
+
+```

### 2025-04-26 products/manuscript/Manuscript.qmd
Update3
Update P

@@
 
+&nbsp;
+&nbsp;
+
 **Figure 1: Distribution of Total Malaria Episodes by IPTp Treatment Arm**
@@
 
+&nbsp;
+&nbsp;
+
 **Interaction between malaria episodes and IPTp regimen**
@@
 
-When evaluated on the independent test set, all four approaches performed at or below chance: logistic regression AUC = 0.48, random forest = 0.42, XGBoost = 0.44, and elastic net = 0.49 (Table 6; Figure 4).
-Even the strongest elastic net model fell essentially to random levels, underscoring that these routinely collected demographic and clinical variables lack sufficient predictive signal for reliable individual risk classification.
-These results highlight the need to integrate richer clinical, laboratory, or biomarker data in future predictive efforts to achieve meaningful discrimination.
+The best elastic-net configuration (penalty = 0.01, mixture = 0) achieved a cross-validated AUC of 0.606 on the training folds.
+After selecting optimal hyperparameters, we refit each model on the full training data and evaluated all four on the independent test set (Table 6; Figure 4).
+Discrimination was uniformly poor: logistic regression AUC = 0.48, random forest = 0.42, XGBoost = 0.44, and elastic net = 0.49—essentially at chance levels.
+These findings underscore that with routinely collected demographic and clinical predictors alone, flexible machine-learning methods do not meaningfully improve risk stratification, and that future efforts should incorporate richer clinical, laboratory, or biomarker data to achieve robust prediction.
 

### 2025-04-26 code/analysis-code/Statistical_Analysis.qmd
PPUpdate

@@
 library(xgboost)            # Gradient boosting engine
+library(generalhoslem)
+library(ResourceSelection)
+library(rms)
 
@@
 
+# Hosmer–Lemeshow test
+```{r}
+library(ResourceSelection)
+
+# get predicted probabilities on the exact same rows as your observed outcome
+phat <- predict(
+  model_gravidity,
+  newdata = train_data,
+  type    = "response"
+)
+
+# convert the factor 0/1 outcome to numeric 0/1
+yobs <- as.numeric(train_data$adverse_birth_outcome) - 1
+
+# now run Hosmer–Lemeshow
+hl <- hoslem.test(
+  x = yobs,
+  y = phat,
+  g = 10
+)
+
+print(hl)
+
+```
+
+# Calibration curve
+
+```{r calibration-plot, fig.width=6, fig.height=6, warning=FALSE}
+# assume `cal` is your calibrate() object from rms::calibrate(lrm_fit, method="boot", B=200)
+
+# 1) Plot without default legend
+plot(cal,
+     lty    = c(3, 1, 2),                             # dashed, solid, dashed
+     col    = c("#1f78b4", "#33a02c", "#e31a1c"),     # blue, green, red
+     lwd    = 2,
+     xlab   = "Predicted Probability of Adverse Outcome",
+     ylab   = "Observed Proportion of Adverse Outcome",
+     main   = "Calibration of Gravidity Model\nin Women Aged < 25",
+     legend = FALSE,                                  # <-- turn off built‐in legend
+     xlim   = c(0.4, 0.8),
+     ylim   = c(0.4, 0.8)
+)
+
+# 2) Add a single, custom legend
+legend("bottomright",
+       legend = c("Apparent", "Bias–Corrected", "Ideal"),
+       col    = c("#1f78b4", "#33a02c", "#e31a1c"),
+       lty    = c(3, 1, 2),
+       lwd    = 2,
+       bty    = "n"
+)
+
+```
+
+```{r}
+# Fit logistic regression with rms
+dd <- datadist(train_data); options(datadist = "dd")
+rms_model <- lrm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + study_arm + education_level,
+                 data = train_data, x = TRUE, y = TRUE)
+# Calibration plot
+cal <- calibrate(rms_model, method="boot", B = 200)
+plot(cal, xlab="Predicted probability", ylab="Observed proportion")
+
+```
+
+```{r}
+library(dplyr); library(ggplot2)
+
+# 1) Get predictions on the train set
+train_data <- train_data %>%
+  mutate(
+    pred = predict(model_gravidity, type = "response"),
+    decile = ntile(pred, 10)
+  )
+
+# 2) Summarize observed vs. predicted in each decile
+calib_df <- train_data %>%
+  group_by(decile) %>%
+  summarize(
+    mean_pred = mean(pred),
+    obs_rate  = mean(as.numeric(adverse_birth_outcome) - 1)
+  )
+
+# 3) Plot
+ggplot(calib_df, aes(x = mean_pred, y = obs_rate)) +
+  geom_point(size = 3) +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  labs(
+    x = "Mean predicted probability",
+    y = "Observed event rate",
+    title = "Calibration plot (deciles)"
+  ) +
+  theme_minimal()
+
+```
+
 ##################################################################################################################################################################

### 2025-04-26 products/manuscript/Manuscript.qmd
PPUpdate

@@
 **Results:**
-In multivariable models adjusted for maternal age, gravidity, and socioeconomic status, we identified a statistically significant interaction between malaria episode frequency and the SP regimen (OR 1.47 per additional episode, 95% CI 1.16–1.88; p = 0.002), demonstrating that each incremental malaria episode was associated with a 47% increase in the odds of an adverse birth outcome among SP‑treated women versus those receiving DP.
-Notably, neither malaria burden nor IPTp regimen alone achieved statistical significance, Total malaria episodes (OR=0.88 per episode, 95% CI 0.75–1.03; p=0.12) and treatment arm alone (OR=0.69 for DP vs SP, 95% CI 0.44–1.08; p=0.10), underscoring the critical role of their interplay.
-Moreover, in the subgroup of women under 25 years, gravidity conferred significant protection: each additional pregnancy reduced the likelihood of an adverse outcome by 14% (OR 0.86 per gravidity, 95% CI 0.74–1.00; p = 0.044).
-Finally, ML models showed modest discrimination (best AUC ≈ 0.60.
+Adjusting for maternal age, gravidity, and socioeconomic status, our multivariable models revealed a significant interaction between malaria episode frequency and the SP regimen (OR 1.47 per additional episode, 95% CI 1.16–1.88; p = 0.002), indicating that each extra malaria episode increased the odds of an adverse birth outcome by 47% in SP‐treated women compared to those on DP.
+In contrast, neither malaria burden alone (OR 0.88 per episode, 95% CI 0.75–1.03; p = 0.12) nor treatment arm alone (OR 0.69 for DP vs SP, 95% CI 0.44–1.08; p = 0.10) reached significance, underscoring the importance of their interplay.
+In women under 25, each additional pregnancy conferred a 14% reduction in adverse‐outcome risk (OR 0.86, 95% CI 0.74–1.00; p = 0.044).
+The interaction model calibrated well (Hosmer–Lemeshow p = 0.37; MAE = 0.03), and on the held-out test set, the AUCs were 0.48 for logistic regression, 0.42 for the random forest, 0.44 for XGBoost, and 0.49 for the elastic net.
 
@@
 **Keywords:**
-
-**Malaria, Adverse birth outcomes, Gravidity/Parity, Intermittent preventive treatment in pregnancy (IPTp), Sulfadoxine–pyrimethamine, Dihydroartemisinin–piperaquine**
+Malaria, Adverse birth outcomes, Gravidity, Intermittent preventive treatment in pregnancy, Sulfadoxine–pyrimethamine and Dihydroartemisinin–piperaquine
 
@@
 # Methods
-This study analyzed data from a double-blind, randomized controlled trial conducted in Uganda, sourced from ClinEpiDB (Release #24, August 30, 2022).
-The trial evaluated intermittent preventive treatment during pregnancy (IPTp) using either sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP) in HIV-uninfected pregnant women.
-The dataset comprised 782 observations detailing maternal, pregnancy, and infant health information, with multiple births recorded separately according to the original trial protocols.
-
-Data preparation involved standardizing date fields, converting categorical variables into factors, and addressing missing data by excluding variables with over 20% missingness and imputing those with lower levels.
-Descriptive statistics and visual inspections ensured the integrity of the data.
-
-The analysis was designed to evaluate whether the IPTp regimen modifies the association between malaria episode frequency and adverse birth outcomes, specifically, preterm birth, low birth weight, and stillbirth, and to assess whether increased gravidity reduces the risk of these outcomes among women under 25 years.
-Multivariable logistic regression models were employed to examine the relationship between malaria episodes and adverse outcomes, incorporating an interaction term for the IPTp regimen and adjusting for confounders such as maternal age, gravidity, and socioeconomic status.
-Predicted probabilities were calculated to facilitate the interpretation of interaction effects.
-A subgroup analysis among women under 25 years used logistic regression models with gravidity treated as a continuous variable and adjustments for maternal age, malaria episode frequency, and treatment regimen, with sensitivity analyses performed to verify robustness.
-All statistical analyses were conducted using R (version 4.3.2).
+This study analyzed 782 maternal–infant records from a double-blind RCT in Uganda (ClinEpiDB Release #24, 30 Aug 2022) comparing IPTp-SP versus DP in HIV-negative pregnant women.
+We standardized dates, recoded categoricals, dropped variables with >20% missingness, and imputed the rest (<10%) via five-cycle MICE; a post-hoc power analysis (α = 0.05) showed 80% power to detect an interaction OR ≥ 1.45.
+Multivariable logistic models, with a regimen×malaria-frequency interaction and adjustment for maternal age, gravidity, and SES, evaluated preterm birth, low birth weight, and stillbirth; predicted probabilities aided interpretation and calibration using the Hosmer–Lemeshow test plus 200-replicate bootstrap curves.
+In women <25 years, gravidity was modeled continuously with sensitivity analyses.
+The machine-learning pipeline split data 80/20 (stratified by outcome), centered/scaled continuous predictors, dummy-encoded categoricals, then used 5-fold CV to tune random forest (mtry, trees), XGBoost (max depth, η, trees), and elastic net (penalty, mixture) for ROC AUC, with an unpenalized logistic regression as baseline.
+The final model performance was assessed on the test set.
+All analyses were conducted in R 4.3.2.
 
@@
 
+**Model calibration and goodness‑of‑fit**
+
+Hosmer–Lemeshow χ²=8.73 (df=8; p=0.366), indicating adequate fit.
+Bootstrap calibration (Figure 5) showed MAE=0.03, MSE=0.0017, 0.9‑quantile AE=0.055, confirming good agreement between predicted and observed risks.
+
+**Figure 3. Calibration curve (bootstrap)**
+```{r calibration-plot, echo=FALSE, out.width='70%', fig.cap="Calibration plot for the logistic regression model in women < 25 years (Hosmer–Lemeshow deciles vs. observed)"}
+knitr::include_graphics("Images/Figure5Calibration.png")
+```
+
 **Machine‐learning model performance**
@@
 
-**Figure 3: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)**
+**Figure 4: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
 knitr::include_graphics("Images/Figure 3. Predicted Probability.png")
-
-```
-
-&nbsp;
-&nbsp;
-
-&nbsp;
-&nbsp;
-
-**Figure 4 : Comparative Discriminative Performance of ML Models for Predicting Adverse Birth Outcomes**
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics("Images/Figure 4.Comparative Discriminative Performance of ML .png")
 
@@
 knitr::include_graphics("Images/Table6.Discriminative Performance (AUC) of ML Models.png")
+
+```
+
+&nbsp;
+&nbsp;
+
+&nbsp;
+&nbsp;
+
+**Figure 5 : Comparative Discriminative Performance of ML Models for Predicting Adverse Birth Outcomes**
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("Images/Figure 4.Comparative Discriminative Performance of ML .png")
 
@@
 
-Our study has several limitations.
-As a secondary analysis of trial data, generalizability to real‐world antenatal populations may be constrained by trial eligibility criteria and adherence patterns.
-Variable exclusion or imputation to address missingness could have introduced bias, and unmeasured factors, such as HIV status, nutritional deficiencies, or placental pathology, likely contribute to adverse outcomes but were not available.
-Finally, the poor performance of our machine‐learning models highlights the need for richer predictor sets; demographic and simple clinical measures are insufficient to classify individual risk with confidence.
+## Implications for Policy and Practice
+Our findings suggest that IPTp guidelines should move beyond a one‐size‐fits‐all approach and be tailored to individual risk profiles, incorporating both infection history and parity.
+In particular, women who experience recurrent malaria episodes while on sulfadoxine-pyrimethamine may benefit from a switch to dihydroartemisinin-piperaquine, alongside intensified screening and reinforced distribution of insecticide-treated bed nets.
+Moreover, primigravidae and secundigravidae, especially those under 25, represent a high‐risk subgroup that warrants closer antenatal monitoring and additional support services.
+Finally, to develop clinically actionable risk stratification tools, future predictive efforts must extend beyond demographic and basic clinical variables by integrating immunologic biomarkers and placental histopathology data.
+
+## Strengths and Limitations
+
+**Strengths**
+ This analysis leverages a double-blind, randomized controlled trial, which minimizes selection bias and ensures balanced baseline characteristics across SP and DP arms.
+By prespecifying and explicitly modeling the interaction between cumulative malaria episodes and IPTp regimen, we uncover effect modification that would be obscured by the main effects alone.
+We further bolster our inference through rigorous model diagnostics, reporting Hosmer–Lemeshow goodness-of-fit, calibration curves, and cross-validated hyperparameter tuning, demonstrating statistical robustness and transparency in model performance.
+Finally, complementing traditional regression with a comprehensive machine-learning pipeline, we assess the limits of routinely collected demographic and clinical predictors for individual risk stratification.
+
+**Limitations**
+ As a secondary analysis of trial data, external validity may be constrained by the original study’s eligibility criteria, adherence protocols, and controlled context.
+Our composite adverse outcome, combining preterm birth, low birth weight, and stillbirth, may mask heterogeneity in drivers and could benefit from separate analyses of each component.
+Missingness in key covariates required exclusion or imputation, introducing potential bias and untestable assumptions.
+Crucially, we lacked data on important confounders such as HIV status, nutritional deficiencies, immunologic biomarkers, and placental histopathology, leaving open the possibility of residual confounding.
+Lastly, the poor discrimination of our machine-learning models underscores that richer, more granular clinical or laboratory data are needed for meaningful individual‐level prediction.
 
 ## Conclusions
-These findings demonstrate a critical need to move beyond one‐size‐fits‐all IPTp recommendations.
-Our analysis points toward stratified prevention strategies by revealing that sulfadoxine–pyrimethamine’s protective efficacy erodes with each additional malaria episode and that higher gravidity markedly shields younger mothers.
-Tailoring IPTp regimens based on individual exposure histories, prioritizing dihydroartemisinin-piperaquine for women facing recurrent infections under SP, and intensifying support for primigravidae and secundigravidae could substantially reduce adverse birth outcomes.
-
-Moreover, the limited predictive power of standard clinical and demographic measures highlights the urgency of integrating richer data sources, such as immunologic biomarkers, placental histopathology, and parasite genotyping, to refine both risk stratification and mechanistic understanding.
-Adapting policy and research agendas to these insights offers a clear path toward more effective, personalized malaria prevention in pregnancy.
+Cumulative malaria exposure and the IPTp regimen interact to shape birth outcomes: SP’s protective efficacy wanes with each successive infection, while DP maintains resilience.
+Higher gravidity markedly reduces risk in younger mothers.
+Tailoring IPTp protocols by infection history and parity, prioritizing DP for recurrent infections under SP, and intensifying support for primigravidae and secundigravidae could substantially reduce adverse outcomes.
+Moreover, integrating immunologic and histopathologic data is essential to advance personalized malaria prevention in pregnancy.
 

### 2025-04-26 products/manuscript/Manuscript.qmd
Update1

@@
 **Strengths**
+
  This analysis leverages a double-blind, randomized controlled trial, which minimizes selection bias and ensures balanced baseline characteristics across SP and DP arms.
@@
 **Limitations**
+
  As a secondary analysis of trial data, external validity may be constrained by the original study’s eligibility criteria, adherence protocols, and controlled context.

### 2025-04-26 products/manuscript/Manuscript.qmd
new
update
update5

@@
 
+# Load Required Libraries
+
 # Path management
-library(here)              # For file path handling
+library(here)               # File path handling
 
@@
 # Tables and reporting
-library(gtsummary)         # Summary statistics tables
-library(gt)                # Table formatting
+library(gtsummary)          # Summary tables
+library(gt)                 # Advanced table formatting
 library(knitr)              # Report generation
-library(kableExtra)        # Enhanced table output in markdown
+library(kableExtra)         # Enhanced markdown tables
 
-# Analysis-specific packages
-library(Amelia)            # Visualizing missing data
+# Data exploration & preparation
+library(Amelia)             # Missing-data visualization
 library(pwr)                # Power analysis
 library(DiagrammeR)         # Diagrams and flowcharts
-library(survival)          # Survival analysis
+
+# Survival analysis
+library(survival)           # Survival models
 library(survminer)          # Survival plots
-library(ggeffects)         # Marginal effects for regression models
+
+# Model effects & outputs
+library(ggeffects)          # Marginal effects extraction
 library(broom)              # Tidy model outputs
 
-# Machine learning and modeling
-library(tidymodels)        # Unified modeling framework (includes recipes, parsnip, workflows, tune, etc.)
-library(rsample)           # Resampling and data splitting
-library(recipes)           # Preprocessing
-library(parsnip)           # Model specification
-library(workflows)         # Combine models and preprocessing
-library(tune)              # Hyperparameter tuning
-library(yardstick)         # Model performance metrics
+# Machine learning & modeling
+library(tidymodels)         # Modeling framework (recipes, parsnip, workflows, tune, yardstick, rsample)
+library(themis)             # Class-imbalance sampling (e.g., SMOTE)
 library(dials)              # Parameter tuning grids
 library(ranger)             # Random forest engine
-library(xgboost)           # Boosting engine
-library(themis)            # Handling class imbalance (e.g., SMOTE)
-
+library(xgboost)            # Gradient boosting engine
+library(generalhoslem)
+library(ResourceSelection)
+library(rms)
 ```
@@
 Malaria remains a significant public health challenge worldwide, particularly in Sub-Saharan Africa, where the disease disproportionately affects vulnerable populations.
-Caused by Plasmodium parasites transmitted through the bite of infected Anopheles mosquitoes, malaria accounted for an estimated 249 million cases globally in 2022, marking an increase of five million cases compared to the previous year.
-Uganda alone reported over 597,000 malaria cases during this period, reflecting the country’s substantial malaria burden (World Health Organization, 2023; Talapko et al., 2019).
+Caused by Plasmodium parasites transmitted through the bite of infected Anopheles mosquitoes, malaria accounted for an estimated 249 million cases globally in 2022 [@talapko2019], marking an increase of five million cases compared to the previous year.
+Uganda alone reported over 597,000 malaria cases during this period, reflecting the country’s substantial malaria burden [@WHO2023].
 Pregnant women represent one of the most vulnerable groups to malaria infection, facing an increased risk of severe clinical symptoms and poor pregnancy outcomes.
-Malaria during pregnancy has been associated with a range of adverse outcomes, including miscarriage, fetal loss, preterm birth, low birth weight, and neonatal mortality (Chua et al., 2021).
+Malaria during pregnancy has been associated with a range of adverse outcomes, including miscarriage, fetal loss, preterm birth, low birth weight, and neonatal mortality [@chua2021].
 
 Efforts to mitigate the impact of malaria in pregnancy have centered on preventive strategies such as the use of insecticide-treated bed nets (ITNs) and chemoprophylaxis through intermittent preventive treatment during pregnancy (IPTp).
-Two widely used IPTp regimens are sulfadoxine-pyrimethamine (SP) and dihydroartemisinin-piperaquine (DP).
-While these preventive measures have shown significant benefits in reducing the risk of malaria-related complications, the effectiveness of these regimens may not be uniform across all contexts.
+The two most commonly used IPTp regimens are sulfadoxine-pyrimethamine (SP) and dihydroartemisinin-piperaquine (DP).
+Although these interventions have proven effective at reducing malaria-related complications, their real-world impact can vary by setting.
+Moreover, coverage remains unacceptably low: as of 2017, only 22% of pregnant women in sub-Saharan Africa received the full IPTp-SP course, highlighting persistent gaps in access and uptake [@bauserman2019].
 Emerging evidence suggests that the choice of IPTp regimen may influence how malaria episodes affect pregnancy outcomes.
@@
 
-&nbsp;
-&nbsp;
-
 **Interaction between malaria episodes and IPTp regimen**
@@
 
+&nbsp;
+&nbsp;
+
 **Figure 3. Calibration curve (bootstrap)**
@@
 # Discussion
-Our analysis from a randomized controlled trial in Uganda illuminates three key findings.
-First, although total malaria episodes and IPTp regimen alone did not predict adverse birth outcomes, their interaction was highly significant.
-Each additional malaria episode under sulfadoxine-pyrimethamine (SP) conferred 47% higher odds of a composite adverse outcome than dihydroartemisinin-piperaquine (DP).
-Second, in women under 25 years, gravidity emerged as the sole significant predictor, with each additional pregnancy reducing the odds of an adverse outcome by 14%.
-Finally, our machine‐learning pipeline, comparing logistic regression, random forest, XGBoost, and elastic net, demonstrated uniformly poor discrimination (AUCs 0.42–0.49), underscoring that routinely collected demographic and clinical variables alone cannot predict individual risk.
+Our analysis from a randomized controlled trial in Uganda demonstrates three principal findings.
+First, although total malaria episodes and IPTp regimen alone did not predict adverse birth outcomes, their interaction was highly significant: each additional malaria episode under sulfadoxine-pyrimethamine (SP) conferred 47% higher odds of a composite adverse outcome compared with dihydroartemisinin-piperaquine (DP) (OR 1.47 per episode, 95% CI 1.16–1.88; p = 0.002).
+This mirrors findings from a recent East African trial, in which monthly IPTp-DP significantly reduced placental malaria relative to IPTp-SP yet failed to translate into overall improvements in birth outcomes when studied in isolation, highlighting that breakthrough infections under SP carry an escalating risk not captured by main-effects analyses alone [@kakuru2020].
+Moreover, the increasing prevalence of pfdhfr–pfdhps “quintuple” and emerging “sextuple” mutations have been linked to diminished SP efficacy, explaining why each successive SP failure amplifies adverse-risk (Desai et al., 2016).
 
-These results extend prior work on malaria in pregnancy.
-The protective effect of higher gravidity aligns with studies suggesting that physiologic adaptations and improved immune tolerance in multiparous women reduce placental parasite sequestration and its sequelae (Chua et al., 2021).
-Likewise, emerging evidence indicates that SP’s efficacy may wane as malaria exposure increases, potentially due to rising parasite resistance or immune modulation (Talapko et al., 2019; World Health Organization, 2023).
-By explicitly modeling the interaction between episodes and regimen, we demonstrate that SP’s preventive benefit diminishes sharply with each successive infection, an effect not captured when examining the main effects alone.
+Second, in women under 25 years, gravidity emerged as the sole significant predictor: each additional pregnancy reduced the odds of an adverse outcome by 14% (OR 0.86, 95% CI 0.76–0.98; p = 0.02).
+This protective effect aligns with systematic evidence that multiparous women acquire parity-dependent immunity, mediated by anti-VAR2CSA antibodies and robust Th1 responses [@cutts2020], which lowers placental parasite densities and improves birth outcomes compared with primigravidae[@kajubi2019].
+Kirosingh and colleagues further elucidated the cellular basis for this phenomenon, showing that primigravidae harbor higher frequencies of malaria-specific IL-10⁺IFNγ⁺ Type 1 regulatory (Tr1) CD4⁺ T cells—cells whose abundance correlates with increased parasitemia and placental malaria [@kirosingh2023].
+In contrast, multigravidae exhibit elevated TNFα⁺CD4⁺ T cell responses linked to protection [@doritchamou2022].
 
-From a public health standpoint, these findings argue for tailoring IPTp strategies to individual risk profiles.
-In high‐transmission settings, women receiving SP who experience multiple breakthrough infections may benefit from switching to DP or augmenting chemoprevention with additional interventions (e.g., enhanced screening, bed‐net distribution).
-Moreover, gravidity should inform risk stratification: primigravidae and secundigravidae under 25 years represent a particularly vulnerable group that might warrant closer monitoring or more aggressive prevention.
+Finally, our machine-learning pipeline (logistic regression, random forest, XGBoost, elastic net) exhibited uniformly poor discrimination (AUC 0.42–0.49), underscoring the limitations of routine demographic and clinical variables alone.
+This finding is consistent with Kubahoniyesu and colleagues whose models using standard covariates plateaued at AUCs < 0.85 for adverse pregnancy outcome prediction in Rwanda [@kubahoniyesu2024].
+By contrast, integrating metabolomic biomarkers into an SVM model achieved an AUC of 0.91, demonstrating that targeted “omics” data can dramatically enhance risk stratification over traditional approaches
 
 ## Implications for Policy and Practice
-Our findings suggest that IPTp guidelines should move beyond a one‐size‐fits‐all approach and be tailored to individual risk profiles, incorporating both infection history and parity.
-In particular, women who experience recurrent malaria episodes while on sulfadoxine-pyrimethamine may benefit from a switch to dihydroartemisinin-piperaquine, alongside intensified screening and reinforced distribution of insecticide-treated bed nets.
-Moreover, primigravidae and secundigravidae, especially those under 25, represent a high‐risk subgroup that warrants closer antenatal monitoring and additional support services.
-Finally, to develop clinically actionable risk stratification tools, future predictive efforts must extend beyond demographic and basic clinical variables by integrating immunologic biomarkers and placental histopathology data.
+Our results advocate a move beyond one-size-fits-all IPTp guidelines toward individualized strategies that account for infection history, drug resistance, and parity.
+In high-transmission, high-resistance settings, women who experience multiple breakthrough infections on SP should be offered IPTp-DP or layered prevention, such as intensified parasitological screening and strengthened insecticide-treated net distribution in line with [@who2022; @cutts2020].
+Equally, primigravidae and secundigravidae under 25 years represent a particularly vulnerable subgroup requiring enhanced antenatal surveillance and targeted support services.
+Finally, translating these recommendations into clinically actionable risk-stratification tools will demand the integration of immunologic biomarkers and placental histopathology alongside routine demographic and clinical variables.
 
@@
 
+&nbsp;
+&nbsp;
+
 #  Remove Section

### 2025-04-26 Norris-Review.qmd

---
title: Asmith Joseph Project Review 
author: Connor Norris
date: 2025-04-25
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Impact of IPTp Regimen on Pregnancy Outcomes in a Malaria-Endemic Setting

Name of project author(s): Asmith Joseph

Name of project reviewer: Connor Norris

# Instructions

Write your comments and feedback below for each section/component of the project. The goal should be to help the author improve their project. Make comments as constructive and actionable as possible. You can provide both criticism and praise.

For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.

Make sure your final document compiles/renders into a readable, well-formatted html document.

Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)

# Specific project content evaluation

## Background, Context and Motivation

How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

Overall, I think this section is very strong. There are a couple of places where I think you could cite more sources. I think a source would be good in the second paragraph after the sentences starting with "Emerging evidence suggests...", "Additionally, maternal characteristics such as gravidity...", and "Prior research has suggested that previous pregnancy experience." This section is believable and flows well into your study aims, but it would be good to know where you are getting your information here. I also think you explain very well what new information you hope to gain from your analysis. Other than that, make sure the scientific names *Plasmodium* and *Anopholes* in the first paragraph (and elsewhere) are italicized. It may also be good to define what gravidity is.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

-   strong contextualization and motivation

## Question description

How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

See comments above about defining gravidity. Other than that, I think your questions are very well explained.

### Summary assessment

-   question/hypotheses fully clear

## Data description

How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

I think you explain this well.

### Summary assessment

-   source and overall structure of data well explained

## Data wrangling and exploratory analysis

How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

I think what you did was pretty good, though I don't entirely understand some of the things you did. Why did you choose the specific variables you did? You mention converting categorical variables into factors - which variables were those? What does "imputing those with lower levels" mean? I don't think you need to make major changes, just add a little bit more to explain further.

### Summary assessment

-   some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

In your last methods paragraph, make it very clear which variables are you predicting with your models and which others are the predictors. For example, are you modeling malaria events using adverse health outcomes as predictors, or the other way around? I think your choice to use logistic regression was valid, but I am not entirely sure because it isn't the most clear what you're modeling from your methods.

### Summary assessment

-   defensible but not optimal analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

I like the layout of Table 1, but include the statistics you mention at the beginning of the results section. For Figure 1, I think the choice to use a histogram was good, but I don't like how the bars for each group alternate. I would either do the two groups on separate plots and then facet wrap them together into one figure, or I would overlay the histograms for the two groups on the same plot, whatever would work best for your data. I would also move your figure code around to place the figures nearest to where they are discussed. Also, move the part in your results where you talk about what you did for your machine learning pipeline to the methods.

### Summary assessment

-   results are presented ok, with room for improvement

## Discussion/Conclusions

Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

I think your conclusions were correct and well explained. I also think you explained your limitations well.

### Summary assessment

-   strong, complete and clear discussion

## Further comments

*Add any other comments regarding the different aspects of the project here. Write anything you think can help your classmate improve their project.*

# Overall project content evaluation

Overall, your prose was very easy to follow, I just think that there are a couple of places where you could add a little bit more to fully explain what you did and why you did it. I think your analysis was well done, and I appreciated the number of models that you fit and how you presented your work to compare their performance.

## Structure

Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Make sure that you either edit or remove the README files from your repo, they are the same as in the template. Otherwise your directly is well set-up.

### Summary assessment

-   mostly clear, but some confusing parts (e.g. useless files, things in the wrong folders)

## Documentation

### Feedback and Comments

I think that your EDA and analysis code is well documented and I can understand what you did. However, I cannot find your processing code, and it looks like there are a lot of redundant files in your processing code folder.

### Summary assessment

-   decently documented with some gaps

## Reproducibility

Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

Other than installing packages, I was able to run all of your code without problem.

### Summary assessment

-   fully reproducible without issues

## Thoroughness

How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

I think you were very thorough with the number of models that you tried, and I think you answered your questions well, even though your models did not give you much to go on.

### Summary assessment

-   strong level of thorougness

## Further comments

Great job overall! I think you only need to make minor changes from this point.

### 2025-04-26 products/manuscript/Manuscript.qmd
update6

@@
 
-dding an interaction term between total malaria episodes and SP (reference = DP) significantly improved model fit (ΔDeviance = 10.11 on 1 df; p = 0.0015).
+Adding an interaction term between total malaria episodes and SP (reference = DP) significantly improved model fit (ΔDeviance = 10.11 on 1 df; p = 0.0015).
 The interaction coefficient (SP × episodes) was highly significant (OR 1.47; 95% CI 1.16–1.88; p = 0.002), indicating that each additional malaria episode increases the odds of an adverse outcome by 47% more under SP than under DP.
@@
 
-&nbsp;
-&nbsp;
-
-#  Remove Section
-
-*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like.
-
 {{< pagebreak >}}

### 2025-04-26 code/analysis-code/Statistical_Analysis.qmd
Supp

@@
 ---
-title: "The Impact of Malaria Episodes and Treatment Regimens on Adverse Pregnancy Outcomes in Ugandan Women"
+title: "Impact of IPTp Regimen on Pregnancy Outcomes in a Malaria-Endemic Setting"
 author: "Asmith Joseph"
@@
 library(rms)
+library(car)
 
@@
 
-*Bar Graph of Total Malaria Episodes During Pregnancy by Treatment Arm*
-
-```{r}
-# Load necessary libraries
-library(dplyr)
-library(ggplot2)
-library(here)
-
-# Read in the cleaned data
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
-
-# Recode 'total_malaria_episodes_during_pregnancy' into a categorical variable
-promo_data_clean <- promo_data_clean %>%
-  mutate(
-    MalariaEpisodesPreg_cat = case_when(
-      total_malaria_episodes_during_pregnancy %in% c(0, 1) ~ "1",
-      total_malaria_episodes_during_pregnancy %in% c(2, 3) ~ "2–3",
-      total_malaria_episodes_during_pregnancy >= 4 ~ "≥4"
-    ),
-    MalariaEpisodesPreg_cat = factor(MalariaEpisodesPreg_cat, levels = c("1", "2–3", "≥4"))
-  )
-
-# Create the bar graph
-ggplot(promo_data_clean, aes(x = MalariaEpisodesPreg_cat, fill = study_arm)) +
-  geom_bar(position = "dodge") +
-  labs(title = "Total Malaria Episodes During Pregnancy by Treatment Arm",
-       x = "Total Malaria Episodes During Pregnancy (Categorical)",
-       y = "Count",
-       fill = "Treatment Arm") +
-  theme_minimal()
-
-```
-
-*Bar Graphs for Gravidity and Parity*
-
-```{r}
-# Load necessary libraries
-
-# Read in the cleaned data
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
-
-# Recode Gravidity into categories: "1", "2–3", "≥4"
-promo_data_clean <- promo_data_clean %>%
-  mutate(
-    Gravidity_cat = case_when(
-      gravidity == 1 ~ "1",
-      gravidity %in% c(2, 3) ~ "2–3",
-      gravidity >= 4 ~ "≥4"
-    ),
-    Gravidity_cat = factor(Gravidity_cat, levels = c("1", "2–3", "≥4"))
-  )
-
-# Create the bar graph for Gravidity by Treatment Arm
-ggplot(promo_data_clean, aes(x = Gravidity_cat, fill = study_arm)) +
-  geom_bar(position = "dodge") +
-  labs(title = "Gravidity Distribution by Treatment Arm",
-       x = "Gravidity Category (1, 2–3, ≥4)",
-       y = "Count",
-       fill = "Treatment Arm") +
-  theme_minimal()
-
-```
-
-*Parity Distribution by Treatment Arm*
-
-```{r}
-# Load necessary libraries
-library(dplyr)
-library(ggplot2)
-library(here)
-
-# Read in the cleaned data
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
-
-# Recode Parity into categories: "0", "1–2", "≥3"
-promo_data_clean <- promo_data_clean %>%
-  mutate(
-    Parity_cat = case_when(
-      parity == 0 ~ "0",
-      parity %in% c(1, 2) ~ "1–2",
-      parity >= 3 ~ "≥3"
-    ),
-    Parity_cat = factor(Parity_cat, levels = c("0", "1–2", "≥3"))
-  )
-
-# Create the bar graph for Parity by Treatment Arm
-ggplot(promo_data_clean, aes(x = Parity_cat, fill = study_arm)) +
-  geom_bar(position = "dodge") +
-  labs(title = "Parity Distribution by Treatment Arm",
-       x = "Parity Category (0, 1–2, ≥3)",
-       y = "Count",
-       fill = "Treatment Arm") +
-  theme_minimal()
-
-```
-
 *Figure 1: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes*
@@
 
-*Fit a logistic regression model without the interaction term*
-
-```{r}
-# Fit a logistic regression model without the interaction term
-model_no_interaction <- glm(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
-                              age_at_enrollment_years + gravidity + education_level,
+**Model 1: Main‐Effects Logistic Regression (No Interaction)**
+
+```{r compare-models-lrt-vif, message=FALSE, warning=FALSE}
+# 1) Subset to the exact variables and drop any rows with missingness
+analysis_data <- promo_data_clean %>%
+  select(
+    adverse_birth_outcome,
+    total_malaria_episodes,
+    study_arm,
+    age_at_enrollment_years,
+    gravidity,
+    education_level
+  ) %>%
+  drop_na()
+
+# 2) Fit the model WITHOUT the interaction
+model_no_interaction <- glm(
+  adverse_birth_outcome ~ 
+    total_malaria_episodes +
+    study_arm +
+    age_at_enrollment_years +
+    gravidity +
+    education_level,
   family = binomial(link = "logit"),
-                            data = promo_data_clean)
-
-# Perform a Likelihood Ratio Test comparing models with and without the interaction
-lrt <- anova(model_no_interaction, model_interaction, test = "LRT")
-print(lrt)
-
-# Check for multicollinearity using the VIF (Variance Inflation Factor)
-library(car)
-vif_values <- vif(model_interaction)
-print(vif_values)
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# Generate Analysis of Deviance Table (Likelihood Ratio Test)
-anova_out <- anova(model_no_interaction, model_interaction, test = "LRT")
-
-# Convert to a clean data frame
-anova_df <- as.data.frame(anova_out)
-anova_df$Model <- c("Model 1 (No Interaction)", "Model 2 (With Interaction)")
-
-# Reorder and rename columns for readability
-anova_df <- anova_df %>%
-  select(Model, `Resid. Df`, `Resid. Dev`, Df, Deviance, `Pr(>Chi)`)
-
-colnames(anova_df) <- c("Model", "Residual Df", "Residual Dev", "Df", "Deviance", "p-value")
-
-# Display table
-kable(
-  anova_df,
-  format = "pandoc",
-  caption = "*Table X: Analysis of Deviance Comparing Models With and Without Interaction*"
-) %>%
-  kable_styling(full_width = FALSE, position = "center")
-
-```
-
-```{r}
-# Fit logistic regression model WITH interaction
-model2 <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm +
-                age_at_enrollment_years + gravidity + education_level,
+  data   = analysis_data
+)
+
+# 3) Fit the model WITH the regimen × malaria interaction
+model_interaction <- glm(
+  adverse_birth_outcome ~ 
+    total_malaria_episodes * study_arm +
+    age_at_enrollment_years +
+    gravidity +
+    education_level,
   family = binomial(link = "logit"),
-              data = promo_data_clean)
-
-```
-
-```{r}
-# Compute Variance Inflation Factors (VIF) for Model 2 (with interaction)
-vif_vals <- vif(model2)
-vif_df <- as.data.frame(vif_vals)
-vif_df$Variable <- rownames(vif_vals)
-vif_df <- vif_df %>% rename(`Df (VIF)` = Df)
-vif_df <- vif_df %>% 
-  select(Variable, GVIF, `Df (VIF)`, `GVIF^(1/(2*Df))`) %>%
+  data   = analysis_data
+)
+
+# 4) Likelihood–Ratio Test (Analysis of Deviance)
+anova_out <- anova(
+  model_no_interaction,
+  model_interaction,
+  test = "LRT"
+)
+
+# 5) Tidy LRT into a data frame
+anova_df <- as.data.frame(anova_out) %>%
+  mutate(
+    Model = c("Model 1 (No Interaction)", 
+              "Model 2 (With Interaction)")
+  ) %>%
+  select(
+    Model,
+    `Resid. Df`,
+    `Resid. Dev`,
+    Df,
+    Deviance,
+    `Pr(>Chi)`
+  )
+
+# 6) Print the LRT results
+print("Analysis of Deviance (LRT) Comparing Models With and Without Interaction:")
+print(anova_df)
+
+# 7) Compute VIFs for the interaction model
+vif_vals <- car::vif(model_interaction)
+
+if (is.matrix(vif_vals)) {
+  # Use the orthonormalized GVIF: GVIF^(1/(2*Df))
+  vif_eff <- vif_vals[, "GVIF^(1/(2*Df))"]
+  terms   <- rownames(vif_vals)
+} else {
+  vif_eff <- vif_vals
+  terms   <- names(vif_vals)
+}
+
+vif_df <- data.frame(
+  Term = terms,
+  VIF  = vif_eff,
+  row.names = NULL
+)
+
+# 8) Print the VIF table
+print("Variance Inflation Factors for the Interaction Model:")
+print(vif_df)
+
+```
+
+ **Fit Model 2 with the regimen × malaria‐episodes interaction**
+
+```{r vif-model2, message=FALSE, warning=FALSE}
+# 1) Fit Model 2 with the regimen × malaria-episodes interaction
+model2 <- glm(
+  adverse_birth_outcome ~ 
+    total_malaria_episodes * study_arm +
+    age_at_enrollment_years +
+    gravidity +
+    education_level,
+  family = binomial(link = "logit"),
+  data   = promo_data_clean
+)
+
+# 2) Compute VIFs treating interaction terms as separate predictors
+vif_mat <- vif(model2, type = "predictor")
+
+# 3) Tidy into a data frame
+vif_df <- as.data.frame(vif_mat) %>%
+  rownames_to_column("Variable") %>%
+  rename(
+    GVIF           = GVIF,
+    `Df (VIF)`     = Df,
+    `GVIF^(1/(2*Df))` = `GVIF^(1/(2*Df))`
+  ) %>%
   mutate(
     Variable = case_when(
-      Variable == "total_malaria_episodes" ~ "Total Malaria Episodes",
-      Variable == "study_arm" ~ "Treatment Arm",
-      Variable == "age_at_enrollment_years" ~ "Age at Enrollment",
+      Variable == "total_malaria_episodes"              ~ "Total malaria episodes",
+      Variable == "study_armSP"                         ~ "Treatment arm (SP vs DP)",
+      Variable == "age_at_enrollment_years"             ~ "Age at enrollment",
       Variable == "gravidity"                           ~ "Gravidity",
-      Variable == "education_level" ~ "Education Level",
-      Variable == "total_malaria_episodes:study_arm" ~ "Interaction: Malaria Episodes × SP",
+      grepl("^education_level", Variable)               ~ 
+        paste("Education:", sub("education_level", "", Variable)),
+      Variable == "total_malaria_episodes:study_armSP"  ~ "Interaction: episodes × SP",
       TRUE                                              ~ Variable
     )
-  )
-
-kable(
-  vif_df,
-  format = "pandoc",
-  caption = "*Table Y: Variance Inflation Factors for Model 2 (With Interaction)*"
-) %>% 
-  kable_styling(full_width = FALSE, position = "center")
-
-```
-
-##################################################################################################################################################################
+  ) %>%
+  select(Variable, GVIF, `Df (VIF)`, `GVIF^(1/(2*Df))`)
+
+# 4) Print the VIF table to the console
+print(vif_df)
+
+```
+
+#############################################################################################################################################################################################################
 
@@
 ```{r}
-# Subset the data to include only women under 25 years of age
+# 1) Subset to women <25 and build the composite outcome
 promo_data_young <- promo_data_clean %>%
-  filter(age_at_enrollment_years < 25)
-
-# Create the composite adverse birth outcome variable if not already done
-promo_data_young <- promo_data_young %>%
+  filter(age_at_enrollment_years < 25) %>%
   mutate(
     low_birth_weight      = ifelse(birth_weight < 2.5, 1, 0),
-    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
-  )
-
-# Fit the logistic regression model
-model_gravidity <- glm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + study_arm + education_level,
+    adverse_birth_outcome = ifelse(preterm_births_count > 0 |
+                                   stillbirth_bin == 1    |
+                                   low_birth_weight   == 1, 1, 0)
+  )
+
+# 2) Fit the logistic regression
+model_gravidity <- glm(
+  adverse_birth_outcome ~ 
+    gravidity +
+    total_malaria_episodes +
+    study_arm +
+    education_level,
   family = binomial(link = "logit"),
-                       data = promo_data_young)
-
-# Display the model summary
+  data   = promo_data_young
+)
+
+# 3) Quick look at the coefficients
 summary(model_gravidity)
 
-# Tidy the model output, exponentiating coefficients to yield odds ratios
-tidy_model_gravidity <- tidy(model_gravidity, exponentiate = TRUE, conf.int = TRUE)
+# 4) Tidy up the output into odds ratios + 95% CI
+tidy_model_gravidity <- broom::tidy(
+  model_gravidity,
+  exponentiate = TRUE,
+  conf.int     = TRUE
+)
 print(tidy_model_gravidity)
 
-# Optionally, plot predicted probabilities for gravidity
-library(ggeffects)
-predicted_probs <- ggeffect(model_gravidity, terms = "gravidity")
+# 5) Plot predicted probability vs gravidity
+predicted_probs <- ggeffects::ggeffect(
+  model_gravidity,
+  terms = "gravidity"
+)
+
 ggplot(predicted_probs, aes(x = x, y = predicted)) +
   geom_line(size = 1.2) +
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "blue") +
-  labs(title = "Predicted Probability of Adverse Outcome by Gravidity (Age < 25)",
-       x = "Gravidity",
-       y = "Predicted Probability") +
+  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), 
+              alpha = 0.2, fill = "steelblue") +
+  labs(
+    title = "Predicted Probability of Adverse Outcome by Gravidity (Age < 25)",
+    x     = "Gravidity (number of prior pregnancies)",
+    y     = "Predicted Probability"
+  ) +
   theme_minimal()
@@
 ```{r}
-# Load required libraries
-library(broom)
-library(dplyr)
-library(knitr)
-library(kableExtra)
-
-# Assume model_gravidity is your fitted logistic regression model for women < 25
-# For example:
-# model_gravidity <- glm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + 
-#                          study_arm + education_level,
-#                        family = binomial(link = "logit"),
-#                        data = promo_data_young)
-
-# Tidy the model output, exponentiating coefficients to get odds ratios
-tidy_model_gravidity <- tidy(model_gravidity, exponentiate = TRUE, conf.int = TRUE)
-
-# Create a summary table without the intercept and with neat variable names
+# 2) Subset to women <25 and create the composite outcome
+promo_data_young <- promo_data_clean %>%
+  filter(age_at_enrollment_years < 25) %>%
+  mutate(
+    low_birth_weight      = ifelse(birth_weight < 2.5, 1, 0),
+    adverse_birth_outcome = ifelse(
+      preterm_births_count > 0 |
+      stillbirth_bin == 1    |
+      low_birth_weight == 1,
+      1, 0
+    )
+  )
+
+# 3) Fit the logistic regression model
+model_gravidity <- glm(
+  adverse_birth_outcome ~
+    gravidity +
+    total_malaria_episodes +
+    study_arm +
+    education_level,
+  family = binomial(link = "logit"),
+  data   = promo_data_young
+)
+
+# 4) Tidy and exponentiate coefficients (to get odds ratios + 95% CIs)
+tidy_model_gravidity <- broom::tidy(
+  model_gravidity,
+  exponentiate = TRUE,
+  conf.int     = TRUE
+)
+
+# 5) Build a neat summary table (dropping the intercept and renaming terms)
 result_table <- tidy_model_gravidity %>%
@@
     ),
-    `Odds Ratio (95% CI)` = paste0(round(estimate, 2),
-                                   " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
-  ) %>%
-  select(term, `Odds Ratio (95% CI)`, p.value) %>%
-  rename(Variable = term, `p-value` = p.value)
-
-# Print the table using kable
-result_table %>%
-  kable(format = "markdown", caption = "")
+    `Odds Ratio (95% CI)` = sprintf(
+      "%.2f (%.2f, %.2f)",
+      estimate, conf.low, conf.high
+    )
+  ) %>%
+  # first select the raw column names
+  dplyr::select(term, `Odds Ratio (95% CI)`, p.value) %>%
+  # then rename for presentation
+  dplyr::rename(
+    Variable = term,
+    `p-value` = p.value
+  )
+
+# 6) Print the final table
+print(result_table)
 
@@
 
-```{r}
-# Tidy the model output and exponentiate coefficients to get odds ratios
-tidy_model_gravidity <- tidy(model_gravidity, exponentiate = TRUE, conf.int = TRUE)
-
-# Create a summary table, remove the intercept, and recode variable names
+# Summarize gravidity model: build table of odds ratios (95% CI) and p-values
+```{r}
+# 1) Subset to <25 and define outcome
+promo_data_young <- promo_data_clean %>%
+  filter(age_at_enrollment_years < 25) %>%
+  mutate(
+    low_birth_weight      = as.integer(birth_weight < 2.5),
+    adverse_birth_outcome = as.integer(
+      preterm_births_count > 0 |
+      stillbirth_bin == 1    |
+      low_birth_weight == 1
+    )
+  )
+
+# 2) Fit logistic regression
+model_gravidity <- glm(
+  adverse_birth_outcome ~ 
+    gravidity +
+    total_malaria_episodes +
+    study_arm +
+    education_level,
+  family = binomial(link = "logit"),
+  data   = promo_data_young
+)
+
+# 3) Tidy into ORs + 95% CIs
+tidy_model_gravidity <- tidy(
+  model_gravidity,
+  exponentiate = TRUE,
+  conf.int     = TRUE
+)
+
+# 4) Build a clean summary table
 result_table <- tidy_model_gravidity %>%
@@
   mutate(
-    term = dplyr::recode(term,
-                         "gravidity" = "Gravidity",
-                         "total_malaria_episodes" = "Total Malaria Episodes",
-                         "study_armSP" = "Treatment Arm (SP)",
-                         "education_levelSecondary" = "Secondary Education",
-                         "education_levelTertiary" = "Tertiary Education",
-                         "education_levelUniversity" = "University Education"),
-    `Odds Ratio (95% CI)` = paste0(round(estimate, 2),
-                                   " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
-  ) %>%
-  select(term, `Odds Ratio (95% CI)`, p.value) %>%
-  rename(Variable = term, `p-value` = p.value)
-
-# Print the table using kable with a descriptive caption
-result_table %>%
-  kable(format = "markdown", 
-        caption = "Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)")
+    Variable = case_when(
+      term == "gravidity"                ~ "Gravidity",
+      term == "total_malaria_episodes"   ~ "Total Malaria Episodes",
+      term == "study_armSP"              ~ "Treatment Arm (SP)",
+      term == "education_levelSecondary" ~ "Secondary Education",
+      term == "education_levelTertiary"  ~ "Tertiary Education",
+      term == "education_levelUniversity"~ "University Education",
+      TRUE                                ~ term
+    ),
+    `Odds Ratio (95% CI)` = sprintf(
+      "%.2f (%.2f–%.2f)",
+      estimate, conf.low, conf.high
+    ),
+    `p-value` = p.value
+  ) %>%
+  # explicitly call dplyr::select to avoid masking
+  dplyr::select(Variable, `Odds Ratio (95% CI)`, `p-value`)
+
+# 5) Print
+print(result_table)
 
@@
 )
-
-# Load broom for tidy output
-library(broom)
 
@@
 
+## Random Forest Model Specification and Hyperparameter Tuning
 *Train/Test Split and Random Forest Model* which fits a random forest model and plots its ROC curve, is also used for Research Question 1 as an alternative ML approach
 ```{r}
-# 1. Define Random Forest model
+# 0) Make sure the outcome is a factor
+train_data <- train_data %>%
+  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0, 1)))
+
+# 1) Define the random‐forest spec
 rf_model <- rand_forest(
@@
 
-# 2. Build the recipe (you already did this correctly)
-rf_recipe <- recipe(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
+library(dials)   # for finalize()
+
+# 2) Build the recipe 
+rf_recipe <- recipe(adverse_birth_outcome ~ 
+                      total_malaria_episodes + study_arm + 
                       age_at_enrollment_years + gravidity + education_level,
@@
 
-# 3. Combine into a workflow
+# 3) Assemble workflow
 rf_workflow <- workflow() %>%
@@
 
-# 4. Tune hyperparameters (optional next step)
+# 4) Extract and finalize parameters
+rf_params <- parameters(rf_model) %>%
+  finalize(train_data)    # now mtry knows how many predictors
+
+# 5) Create a random tuning grid of size 20
+rf_grid <- grid_random(rf_params, size = 20)
+
+# 6) 10-fold stratified CV
+rf_folds <- vfold_cv(train_data, v = 10, strata = adverse_birth_outcome)
+
+# 7) Tune
 set.seed(123)
@@
   rf_workflow,
-  resamples = vfold_cv(train_data, v = 10, strata = adverse_birth_outcome),
-  grid = 20,  # number of hyperparameter combos
+  resamples = rf_folds,
+  grid      = rf_grid,
   metrics   = metric_set(roc_auc)
@@
 
-```
-
-```{r}
-# Ensure the outcome is a factor in train_data
+# 8) View results
+rf_tune %>% 
+  collect_metrics() %>% 
+  filter(.metric == "roc_auc")
+
+```
+
+# Random Forest model training and OOB Brier score output
+```{r}
+# 1)  outcome is a factor
 train_data <- train_data %>%
-  mutate(adverse_birth_outcome = factor(adverse_birth_outcome))
-
-# Build a recipe for the random forest model (no inline interaction here)
-rf_recipe <- recipe(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
-                      age_at_enrollment_years + gravidity + education_level,
-                    data = train_data) %>%
+  mutate(
+    adverse_birth_outcome = factor(
+      adverse_birth_outcome,
+      levels = c(0, 1),
+      labels = c("no", "yes")
+    )
+  )
+
+# 2) Define a Random Forest model (with fixed hyperparameters for illustration)
+rf_model <- rand_forest(
+    trees = 500,
+    mtry  = 3,    # you can choose a sensible default or tune later
+    min_n = 5     # likewise
+  ) %>%
+  set_engine("ranger") %>%
+  set_mode("classification")
+
+# 3) Build the recipe
+rf_recipe <- recipe(
+    adverse_birth_outcome ~
+      total_malaria_episodes +
+      study_arm +
+      age_at_enrollment_years +
+      gravidity +
+      education_level,
+    data = train_data
+  ) %>%
   step_dummy(all_nominal_predictors())
 
-# Create a workflow for the random forest model
+# 4) Assemble into a workflow
 rf_workflow <- workflow() %>%
@@
 
-# Fit the random forest model on training data
-rf_fit <- rf_workflow %>% fit(data = train_data)
-
-# Generate predicted probabilities on the test set
-rf_preds <- predict(rf_fit, new_data = test_data, type = "prob")
-
-# Compute ROC for the random forest using the probability for class "1"
-library(pROC)
-roc_rf <- roc(test_data$adverse_birth_outcome, rf_preds$.pred_1)
-plot(roc_rf, col = "red", lwd = 2, main = "ROC Curve: Random Forest")
-auc_rf <- auc(roc_rf)
-print(paste("Random Forest AUC:", round(auc_rf, 2)))
-
-```
-
--   Model Comparison and Interpretation\*
-
+# 5) Fit the random forest
+rf_fit <- rf_workflow %>% 
+  fit(data = train_data)
+
+# 6) Inspect the fitted workflow
+rf_fit
+
+```
+- Random forest fit ran without errors, but the OOB Brier score shows it’s essentially no better than random. 
+- That echoes what you saw with your logistic and other ML models (AUCs ~0.4–0.5): with these variables alone, none of the algorithms can predict the composite adverse birth outcome.
+
+# Model Comparison and Interpretation
 ```{r}
@@
 
-```{r}
-# Load necessary packages (if not already loaded)
-library(tidymodels)
-library(xgboost)
-
+# Define a boosting model for classification (using boost_tree)
+```{r}
 # Define a boosting model for classification (using boost_tree)

### 2025-04-26 code/eda-code/eda_manuscript.qmd
Supp

@@
 # Load Required Libraries
-# Essential libraries for data handling, visualization, and analysis
 
-library(here)             # Manage file paths
+# Path management
+library(here)               # File path handling
+
+# Data manipulation and wrangling
 library(dplyr)              # Data manipulation
-library(tidyverse)        # Data wrangling and visualization
+library(tidyverse)          # Core tidyverse packages (ggplot2, readr, etc.)
 library(janitor)            # Clean column names
-library(skimr)            # Quick data summary
-library(lubridate)        # Working with dates
-library(readr)            # Reading CSV files
-library(ggplot2)          # Visualization
+library(skimr)              # Quick data summaries
+library(lubridate)          # Date/time handling
+library(forcats)            # Categorical variable tools
+
+# Visualization
+library(ggplot2)            # Data visualization
 library(ggpubr)             # Publication-ready plots
-library(forcats)          # Categorical variables handling
 
-# Table and reporting packages
+# Tables and reporting
 library(gtsummary)          # Summary tables
-library(gt)               # Table formatting
-library(knitr)            # Table rendering
-library(kableExtra)       # Table customization
+library(gt)                 # Advanced table formatting
+library(knitr)              # Report generation
+library(kableExtra)         # Enhanced markdown tables
 
-# Analysis-specific packages
-library(Amelia)           # Missing data visualization
+# Data exploration & preparation
+library(Amelia)             # Missing-data visualization
 library(pwr)                # Power analysis
 library(DiagrammeR)         # Diagrams and flowcharts
-library(survival)         # Survival analysis
+
+# Survival analysis
+library(survival)           # Survival models
 library(survminer)          # Survival plots
-library(ggeffects)        # Marginal effects for regression models
+
+# Model effects & outputs
+library(ggeffects)          # Marginal effects extraction
 library(broom)              # Tidy model outputs
-library(rsample)          # Resampling and splitting datasets
+
+# Machine learning & modeling
+library(tidymodels)         # Modeling framework (recipes, parsnip, workflows, tune, yardstick, rsample)
+library(themis)             # Class-imbalance sampling (e.g., SMOTE)
+library(dials)              # Parameter tuning grids
+library(ranger)             # Random forest engine
+library(xgboost)            # Gradient boosting engine
+library(generalhoslem)
+library(ResourceSelection)
+library(rms)
 

### 2025-04-26 products/manuscript/supplement/Supplementary-Material.qmd
Supp

---
title: "Supplement Materials of Impact of IPTp Regimen on Pregnancy Outcomes in a Malaria-Endemic Setting"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

This shows some materials that could go into a supplementary file. Often you want/need references here too. You can use the same reference bib file for this and the main text (as done here) or have separate bib files.

For illustrative purposes, I'm doing the supplement as pdf. For this to work, you need a (La)TeX system installed. It's easy. Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).

Of course you would choose the format based on needs.

I'm also using a different style for the references here. (vancouver vs apa in the main manuscript). Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header. We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.

{{< pagebreak >}}

# Overview
The Supplementary Appendix begins with comprehensive methodological details, including variable‐by‐variable missingness (Table S1), analysis of deviance comparing models with and without the malaria×SP interaction (Table S2), and variance inflation factors for the final interaction model (Table S3). It also presents machine‐learning tuning results with the top five elastic-net hyperparameter combinations ranked by mean cross-validated AUC (Table S4) alongside a heatmap illustrating AUC across the penalty–mixture grid (Figure S1). Full code excerpts document our data-cleaning steps, rsample splits, recipe definitions, and tune_grid workflows. The appendix then moves on to additional results: a detailed stratification of outcome measures and malaria-exposure variables by IPTp arm (Table S5), bar graphs of total malaria episodes during pregnancy (Figure S2), gravidity (Figure S3), and parity (Figure S4) distributions by treatment arm, and a bootstrap calibration curve for the interaction model (Figure S5). Finally, it compares model discrimination with overlaid ROC curves for logistic regression, random forest, and XGBoost (Figure S6), presents a precision-recall curve for the top‐performing machine-learning model (Figure S7), and reports the test-set AUC for the gravidity-only model in women under 25 years (Table S6).

# Code and file information

Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
Essentially, give a full set of instructions to re-generate everything.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
library(here)
here::i_am("products/manuscript/Manuscript.qmd")

# ← THIS is the one that actually sets the working dir for *all* chunks:
library(knitr)
knitr::opts_chunk$set(root.dir = here::here("products/manuscript"))

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
# Load Required Libraries for Data Handling, Visualization, and Analysis

# Path management
library(here)               # File path handling

# Data manipulation and wrangling
library(dplyr)              # Data manipulation
library(tidyverse)          # Core tidyverse packages (ggplot2, readr, etc.)
library(janitor)            # Clean column names
library(skimr)              # Quick data summaries
library(lubridate)          # Date/time handling
library(forcats)            # Categorical variable tools

# Visualization
library(ggplot2)            # Data visualization
library(ggpubr)             # Publication-ready plots

# Tables and reporting
library(gtsummary)          # Summary tables
library(gt)                 # Advanced table formatting
library(knitr)              # Report generation
library(kableExtra)         # Enhanced markdown tables

# Data exploration & preparation
library(Amelia)             # Missing-data visualization
library(pwr)                # Power analysis
library(DiagrammeR)         # Diagrams and flowcharts

# Survival analysis
library(survival)           # Survival models
library(survminer)          # Survival plots

# Model effects & outputs
library(ggeffects)          # Marginal effects extraction
library(broom)              # Tidy model outputs

# Machine learning & modeling
library(tidymodels)         # Modeling framework (recipes, parsnip, workflows, tune, yardstick, rsample)
library(themis)             # Class-imbalance sampling (e.g., SMOTE)
library(dials)              # Parameter tuning grids
library(ranger)             # Random forest engine
library(xgboost)            # Gradient boosting engine
library(generalhoslem)
library(ResourceSelection)
library(rms)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
# Data Import & Initial Inspection 
# load the Dataset
PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"))

promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
```

{{< pagebreak >}}

# Additional Method Details

Often, the main manuscript only allows for an overview description of the methods. Use the supplement to describe all your methods, models and approaches in a lot of detail. Reference specific parts of your code as needed.

{{< pagebreak >}}

# Additional results

Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.

*Bar Graph of Total Malaria Episodes During Pregnancy by Treatment Arm*

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(here)

# Read in the cleaned data
promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))

# Recode 'total_malaria_episodes_during_pregnancy' into a categorical variable
promo_data_clean <- promo_data_clean %>%
  mutate(
    MalariaEpisodesPreg_cat = case_when(
      total_malaria_episodes_during_pregnancy %in% c(0, 1) ~ "1",
      total_malaria_episodes_during_pregnancy %in% c(2, 3) ~ "2–3",
      total_malaria_episodes_during_pregnancy >= 4 ~ "≥4"
    ),
    MalariaEpisodesPreg_cat = factor(MalariaEpisodesPreg_cat, levels = c("1", "2–3", "≥4"))
  )

# Create the bar graph
ggplot(promo_data_clean, aes(x = MalariaEpisodesPreg_cat, fill = study_arm)) +
  geom_bar(position = "dodge") +
  labs(title = "Total Malaria Episodes During Pregnancy by Treatment Arm",
       x = "Total Malaria Episodes During Pregnancy (Categorical)",
       y = "Count",
       fill = "Treatment Arm") +
  theme_minimal()

```

*Bar Graphs for Gravidity and Parity*

```{r}
# Load necessary libraries

# Read in the cleaned data
promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))

# Recode Gravidity into categories: "1", "2–3", "≥4"
promo_data_clean <- promo_data_clean %>%
  mutate(
    Gravidity_cat = case_when(
      gravidity == 1 ~ "1",
      gravidity %in% c(2, 3) ~ "2–3",
      gravidity >= 4 ~ "≥4"
    ),
    Gravidity_cat = factor(Gravidity_cat, levels = c("1", "2–3", "≥4"))
  )

# Create the bar graph for Gravidity by Treatment Arm
ggplot(promo_data_clean, aes(x = Gravidity_cat, fill = study_arm)) +
  geom_bar(position = "dodge") +
  labs(title = "Gravidity Distribution by Treatment Arm",
       x = "Gravidity Category (1, 2–3, ≥4)",
       y = "Count",
       fill = "Treatment Arm") +
  theme_minimal()

```

*Parity Distribution by Treatment Arm*

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(here)

# Read in the cleaned data
promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))

# Recode Parity into categories: "0", "1–2", "≥3"
promo_data_clean <- promo_data_clean %>%
  mutate(
    Parity_cat = case_when(
      parity == 0 ~ "0",
      parity %in% c(1, 2) ~ "1–2",
      parity >= 3 ~ "≥3"
    ),
    Parity_cat = factor(Parity_cat, levels = c("0", "1–2", "≥3"))
  )

# Create the bar graph for Parity by Treatment Arm
ggplot(promo_data_clean, aes(x = Parity_cat, fill = study_arm)) +
  geom_bar(position = "dodge") +
  labs(title = "Parity Distribution by Treatment Arm",
       x = "Parity Category (0, 1–2, ≥3)",
       y = "Count",
       fill = "Treatment Arm") +
  theme_minimal()

```

## Example additional result

@tbl-resulttable1 shows an additional table summarizing a model fit.

```{r}
#| label: tbl-resulttable1
#| tbl-cap: "Another fit table."
#| echo: FALSE
resulttable1 = readRDS(here("results","tables","resulttable1.rds"))
knitr::kable(resulttable1)
```

@fig-result2 shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result2
#| fig-cap: "Height and weight."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight.png"))
```

#  Evaluating Discrimination (ROC Curve and AUC)

```{r}
# Load necessary package for ROC analysis
library(pROC)

# Generate predicted probabilities using the logistic regression model (model_gravidity)
promo_data_young$predicted_prob <- predict(model_gravidity, type = "response")

# Create the ROC curve
roc_obj <- roc(promo_data_young$adverse_birth_outcome, promo_data_young$predicted_prob)
plot(roc_obj, col = "blue", lwd = 2, main = "ROC Curve for Model: Gravidity in Women < 25")
auc_value <- auc(roc_obj)
print(paste("AUC:", round(auc_value, 2)))

# Load necessary package for calibration plot
library(caret)

# Create a calibration plot
calibration_data <- data.frame(
  observed = factor(promo_data_young$adverse_birth_outcome, levels = c(0,1)),
  predicted = promo_data_young$predicted_prob
)

# Use the calibration function from the caret package
cal_plot <- calibration(observed ~ predicted, data = calibration_data, class = "1")
plot(cal_plot, main = "Calibration Plot for Model: Gravidity in Women < 25")

```

I see that the ROC curve for my model is close to the diagonal, with an AUC only slightly above 0.5. This tells me that the model doesn't have strong discriminative ability for predicting adverse outcomes in women under 25. Additionally, the calibration plot shows that my predicted probabilities often stray from the ideal diagonal—especially in the mid-range—indicating that my model’s risk estimates don't consistently match the observed rates. Overall, while gravidity is statistically significant, my model as a whole isn't very effective at distinguishing between those who experience adverse outcomes and those who don't, and its probability estimates need improvement.

# Heatmap of cross-validated AUC\nfor Elastic-net hyperparameter grid
```{r}
library(ggplot2)

enet_tune %>% 
  collect_metrics() %>% 
  filter(.metric=="roc_auc") %>% 
  ggplot(aes(x = log10(penalty), y = mixture, fill = mean)) +
    geom_tile() +
    scale_fill_viridis_c(name = "AUC") +
    labs(
      x = expression(log[10](lambda)),
      y = expression(alpha),
      title = "Heatmap of cross-validated AUC\nfor Elastic-net hyperparameter grid"
    ) +
    theme_minimal(base_size = 12)

```

```

{{< pagebreak >}}

# Discussion

Any additional discussion regarding the supplementary material/findings.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

{{< pagebreak >}}

# References

### 2025-04-28 Review template/Shaun.project-review.qmd
Project Update

---
title: "Impact of IPTp Regimen on Pregnancy Outcomes in a Malaria-Endemic Setting" 
author: Shaun van den Hurk
date: 25 April 2025
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Impact of IPTp Regimen on Pregnancy Outcomes in a Malaria-Endemic Setting

Name of project author(s): Asmith Joseph

Name of project reviewer: Shaun van den Hurk

# Specific project content evaluation

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments
The background is very strong. The project offers a comprehensive overview of malaria in pregnancy, current preventive strategies (IPTp with SP and DP), and knowledge gaps around regimen-specific effects. Previous work is properly referenced, and the study is well-placed within the broader context of malaria research and maternal health. The motivation is clear and compelling. The value of this work and the authors familiarity wit the topic is also clear.

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments
The research questions are clearly defined: (1) Does IPTp regimen modify the association between malaria episodes and adverse outcomes? (2) Does gravidity protect against adverse outcomes among younger women? Both are explicitly linked to the data and study goals and throroughly investigated.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments
The dataset source (ClinEpiDB) is clearly cited, and the structure of the data (variables, missingness) is explained well.

### Summary assessment
* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments
The data cleaning and preprocessing steps are reasonable and clearly described (e.g. missing data handling, recoding of variables, exploratory summaries). Exploratory figures and tables are provided and meaningful. One small weakness: the processing files can be removed since cleaning and eda both took place in the eda files and this could be made clearer. Otherwise, very good.

### Summary assessment
* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments
The analysis approach is appropriate: multivariable logistic regression with interaction terms, subgroup analysis, and appropriate model adjustments. The use of predicted probabilities to interpret interaction effects is a good approach. There appears to be good fundamental statistical basis for the approaches and in the analysis.

### Summary assessment
* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments
I did not get to view the final rendered Word document of the manuscript (the current version appears a bit outdated). However, the version available looks good and mthe figures look good. There are quite a few figures and tables though and maybe this can be reviewed to make sure that they are all adding value and necessary or if any can be moved to the supplementary document.
Tables and figures are clean, clear, and mostly publication-quality.

### Summary assessment
* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments
The discussion is thorough, acknowledges limitations appropriately, and interprets findings in light of existing literature. Limitations are realistic (e.g. trial setting, missing data, model performance). The conclusions are logical and grounded in the results. There is also a clear link to the public health benefits of this data which is positive and strong.

### Summary assessment
* strong, complete and clear discussion

## Further comments
- Strong writing quality overall.
- Excellent use of predicted probabilities for interaction interpretation.
- You could consider adding brief notes on clinical relevance of findings (e.g. potential policy implications for IPTp protocols)?

# Overall project content evaluation
Evaluate overall features of the project by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments
The structure of the project and the flow is very good and clear.
- You can remove the processing folder and files since you seem to have done your processing in your eda file.
- I recommend updating the ReadME files.

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments
Documentation is good overall — analysis steps are explained in the R scripts, and major coding decisions are annotated.
I appreciatye the inclusion of summaries of the results/interpretations within the statistical analysis file.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

There are a few issues with reproducibility. However, the actual issues are small and should be quick to fix.

_ You have a lot of different packages in your code that are loaded in different places. I would recommend putting all of the loading of packages together at the start, because I did not have many of them installed and so I kept having to go back and install and rerun scripts which was jarring. If this is all done in the beginning then it would help the flow.
- There are issues in the statistical analysis document that do not allow it to be reproduced. I have tried to run chunks separately but it seems that some variables are missing from multiple chunks. Here are some of the issues: Error: object 'anova_out' not found, Error: object 'rf_model' not found, Error: object 'logistic_tune' not found, Error: object 'stack_preds' not found
Maybe just go back and see if some variables are not being created somewhere.
- In your manuscript you set your working directory to your own computer which limits the reproducibility. It would be better to stick to using the "here" function and removing that setwd.
- The manuscript document does not render a Word document - there are issues from line 417 (the code does run though) with a missing object. I would recommend removing the raw code that produces these tables and figures and rather create them in the statistical analysis and import completed figures. That way you have fewer lines of code that can go wrong and you have a cleaner looking manuscript.

### Summary assessment
* major parts not reproducible

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments
The main research questions were thoroughly addressed. Interaction effects were carefully explored. Minor improvements could include testing alternative modeling approaches, but overall, the study is thorough and well thought out. This is a complete project.

### Summary assessment
* strong level of thorougness

## Further comments
You did a good job! Very polished project overall. Impressive handling of interaction effects, a relatively advanced topic.
There are a couple things that can be worked upon that I outlined under other points which should be simple to fix and will help complete the product.
Remember to update your references as well.

### 2025-04-28 products/manuscript/Manuscript.qmd
Project Update

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+# 1) Tell here where this file lives:
 library(here)
@@
 
-# ← THIS is the one that actually sets the working dir for *all* chunks:
-library(knitr)
-knitr::opts_chunk$set(root.dir = here::here("products/manuscript"))
+# 2) Set your global chunk options (no setwd or root.dir needed):
+knitr::opts_chunk$set(
+  echo    = FALSE,
+  message = FALSE,
+  warning = FALSE
+)
 
@@
 **Figure 1: Distribution of Total Malaria Episodes by IPTp Treatment Arm**
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r,echo=FALSE, message=FALSE, warning=FALSE}}
+knitr::include_graphics("Images/Figure1.DistribTotal Malaria.png")
 
-ggplot(promo_data_clean, aes(x = total_malaria_episodes, fill = study_arm)) +
-  geom_histogram(binwidth = 1, alpha = 0.6, position = "dodge") +
-  labs(title = "",
-       x = "Total Malaria Episodes",
-       y = "Frequency",
-       fill = "Treatment Arm") +
-  theme_minimal()
 ```
@@
 **Table 2: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
-```{r,  echo=FALSE, message=FALSE, warning=FALSE}
-# Load necessary libraries
-library(dplyr)
-library(gtsummary)
-library(here)
+```{r}
+knitr::include_graphics("Images/Table2.Outcome Measures Malaria .png")
 
-# Read in the cleaned data (if not already read)
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
-
-# Recode variables as needed for outcomes analysis:
-promo_data_clean <- promo_data_clean %>%
-  mutate(
-    # Create a composite adverse outcome variable: 1 if any of preterm birth, stillbirth, or low birth weight (<2.5 kg) occurs
-    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
-    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0),
-
-    # Recode Preterm Births Count: combine 0 and 1 as "1", and 2 as "2"
-    PretermBirths_cat = case_when(
-      preterm_births_count %in% c(0, 1) ~ "1",
-      preterm_births_count == 2 ~ "2"
-    ),
-    PretermBirths_cat = factor(PretermBirths_cat, levels = c("1", "2"))
-  )
-
-# Create Table Z: Outcome Variables by IPTp Treatment Arm with p-values
-table_outcomes <- promo_data_clean %>%
-  dplyr::select(
-    study_arm,
-    `Malaria Infection Rate During Pregnancy` = malaria_infection_rate_during_pregnancy,
-    `Placental Malaria (Rogerson Criteria)` = placental_malaria_by_rogerson_criteria,
-    `Preterm Births Count` = PretermBirths_cat,
-    `Stillbirth bin` = stillbirth_bin,
-    `Birthweight` = birth_weight,
-    `Composite Adverse Outcome` = adverse_birth_outcome
-  ) %>%
-  tbl_summary(
-    by = study_arm,
-    missing = "no",
-    statistic = list(
-      all_continuous() ~ "{mean} ({sd})",
-      all_categorical() ~ "{n} ({p}%)"
-    )
-  ) %>%
-  add_p() %>%
-  modify_header(label = "") %>%
-  modify_caption("")
-
-# Print Table Z
-table_outcomes
 ```
@@
 **Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes**
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# Create the composite adverse outcome variable
-promo_data_clean <- promo_data_clean %>%
-  mutate(
-    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
-    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
-  )
-
-# Fit the interaction model (ensuring that adverse_birth_outcome now exists)
-model_interaction <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm + 
-                           age_at_enrollment_years + gravidity + education_level,
-                         family = binomial(link = "logit"),
-                         data = promo_data_clean)
-
-# Generate predicted probabilities over the range of total malaria episodes by study arm
-pred <- ggeffect(model_interaction, terms = c("total_malaria_episodes [all]", "study_arm"))
-
-# Plot the predicted probabilities
-interaction_plot <- ggplot(pred, aes(x = x, y = predicted, color = group)) +
-  geom_line(size = 1.2) +
-  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
-  labs(title = "",
-       x = "Total Malaria Episodes",
-       y = "Predicted Probability of Adverse Outcome",
-       color = "Treatment Arm",
-       fill = "Treatment Arm") +
-  theme_minimal()
-
-interaction_plot
+```{r}
+knitr::include_graphics("Images/Figure2.Differential Impact of IPTp.png")
 ```
@@
 &nbsp;
-
-**Table 3: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
-```{r,  echo=FALSE, message=FALSE, warning=FALSE}
-# Load necessary libraries
-library(dplyr)
-library(gtsummary)
-library(here)
-
-# Read in the cleaned data (if not already read)
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
-
-# Recode variables as needed for outcomes analysis:
-promo_data_clean <- promo_data_clean %>%
-  mutate(
-    # Create a composite adverse outcome variable: 1 if any of preterm birth, stillbirth, or low birth weight (<2.5 kg) occurs
-    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
-    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0),
-
-    # Recode Preterm Births Count: combine 0 and 1 as "1", and 2 as "2"
-    PretermBirths_cat = case_when(
-      preterm_births_count %in% c(0, 1) ~ "1",
-      preterm_births_count == 2 ~ "2"
-    ),
-    PretermBirths_cat = factor(PretermBirths_cat, levels = c("1", "2"))
-  )
-
-# Create Table Z: Outcome Variables by IPTp Treatment Arm with p-values
-table_outcomes <- promo_data_clean %>%
-  dplyr::select(
-    study_arm,
-    `Malaria Infection Rate During Pregnancy` = malaria_infection_rate_during_pregnancy,
-    `Placental Malaria (Rogerson Criteria)` = placental_malaria_by_rogerson_criteria,
-    `Preterm Births Count` = PretermBirths_cat,
-    `Stillbirth bin` = stillbirth_bin,
-    `Birthweight` = birth_weight,
-    `Composite Adverse Outcome` = adverse_birth_outcome
-  ) %>%
-  tbl_summary(
-    by = study_arm,
-    missing = "no",
-    statistic = list(
-      all_continuous() ~ "{mean} ({sd})",
-      all_categorical() ~ "{n} ({p}%)"
-    )
-  ) %>%
-  add_p() %>%
-  modify_header(label = "") %>%
-  modify_caption("")
-
-# Print Table Z
-table_outcomes
-```
 
@@
 
-**Table 4: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**
+**Table 3: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**
 
@@
 
-**Table 5: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
+**Table 4: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-**Table 6:  Discriminative Performance (AUC) of Machine Learning Models Predicting Adverse Birth Outcomes Under Different IPTp Regimens**
+**Table 5:  Discriminative Performance (AUC) of Machine Learning Models Predicting Adverse Birth Outcomes Under Different IPTp Regimens**
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}

### 2025-04-28 code/analysis-code/Statistical_Analysis.qmd
P3

@@
 library(car)
-
 ```
@@
 names(promo_data_clean)
+```
+
+```{r}
+# Load the cleaned data
+promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+
+# 1. Check the class of the variable
+class(promo_data_clean$preterm_births_count)
+
+# 2. View its internal structure
+str(promo_data_clean$preterm_births_count)
+
+# 3. Get basic summary statistics
+summary(promo_data_clean$preterm_births_count)
+
+# 4. Tabulate all observed values (including NAs)
+table(promo_data_clean$preterm_births_count, useNA = "always")
+
+# 5. Quick glimpse using dplyr (avoids calling select on a vector)
+glimpse(promo_data_clean %>% select(preterm_births_count))
 ```
@@
 ```{r}
-# Load necessary libraries
-library(dplyr)
-library(gtsummary)
-library(here)
-
 # Read in the cleaned data (if not already read)

### 2025-04-28 code/analysis-code/Statistical_Analysis.qmd
Update

--- 2025-04-26 code/analysis-code/Statistical_Analysis.qmd
@@
 
-# Data manipulation and wrangling
-library(dplyr)              # Data manipulation
-library(tidyverse)          # Core tidyverse packages (ggplot2, readr, etc.)
+# Tidyverse core for data import, manipulation & visualization
+library(tidyverse)          # includes ggplot2, dplyr, tidyr, readr, purrr, etc.
+
+# Data cleaning & summaries
 library(janitor)            # Clean column names
 library(skimr)              # Quick data summaries
+
+# Date/time and categorical helpers
 library(lubridate)          # Date/time handling
 library(forcats)            # Categorical variable tools
-
-# Visualization
-library(ggplot2)            # Data visualization
-library(ggpubr)             # Publication-ready plots
 
@@
 
+# Visualization enhancements
+library(ggpubr)             # Publication-ready plots
+
 # Data exploration & preparation
@@
 # Machine learning & modeling
-library(tidymodels)         # Modeling framework (recipes, parsnip, workflows, tune, yardstick, rsample)
+library(tidymodels)         # Modeling framework (recipes, workflows, tune, yardstick, rsample)
 library(themis)             # Class-imbalance sampling (e.g., SMOTE)
@@
 library(xgboost)            # Gradient boosting engine
-library(generalhoslem)
-library(ResourceSelection)
-library(rms)
-library(car)
+
+# Goodness-of-fit & diagnostics
+library(generalhoslem)      # Hosmer–Lemeshow tests
+library(ResourceSelection)  # Additional GOF tests
+library(rms)                # Regression modeling strategies
+library(car)                # Companion to applied regression
 
@@
 # load the Dataset
-PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"))
-
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+# ────────────────────────────────────────────────────────────────────────────────
+# Project setup: anchor here() at the project root
+# ────────────────────────────────────────────────────────────────────────────────
+library(here)
+here::i_am("data/raw-data/PROMO_Data.csv")
+
+# ────────────────────────────────────────────────────────────────────────────────
+# Load data using project‐relative paths
+# ────────────────────────────────────────────────────────────────────────────────
+library(readr)
+
+PROMO_Data <- read_csv(
+  here("data", "raw-data", "PROMO_Data.csv")
+)
+
+promo_data_clean <- read_csv(
+  here("data", "clean", "PROMO_Data_clean.csv")
+)
 ```
@@
 
-```{r}
-names(promo_data_clean)
-```
-
 *Table 1*
-
 ```{r}
@@
 analysis_data <- promo_data_clean %>%
-  select(
+  dplyr::select(
     adverse_birth_outcome,
@@
   ) %>%
-  drop_na()
+  tidyr::drop_na()
 
@@
 anova_df <- as.data.frame(anova_out) %>%
-  mutate(
-    Model = c("Model 1 (No Interaction)", 
-              "Model 2 (With Interaction)")
-  ) %>%
-  select(
+  dplyr::mutate(
+    Model = c("Model 1 (No Interaction)", "Model 2 (With Interaction)")
+  ) %>%
+  dplyr::select(
     Model,
@@
 # 2) Compute VIFs treating interaction terms as separate predictors
-vif_mat <- vif(model2, type = "predictor")
-
-# 3) Tidy into a data frame
+vif_mat <- car::vif(model2, type = "predictor")
+
+# 3) Tidy into a data frame, explicitly using tibble::rownames_to_column() and dplyr:: verbs
 vif_df <- as.data.frame(vif_mat) %>%
-  rownames_to_column("Variable") %>%
-  rename(
+  tibble::rownames_to_column("Variable") %>%
+  dplyr::rename(
     GVIF               = GVIF,
@@
   ) %>%
-  mutate(
-    Variable = case_when(
+  dplyr::mutate(
+    Variable = dplyr::case_when(
       Variable == "total_malaria_episodes"             ~ "Total malaria episodes",
@@
       Variable == "gravidity"                          ~ "Gravidity",
-      grepl("^education_level", Variable)               ~ 
-        paste("Education:", sub("education_level", "", Variable)),
+      grepl("^education_level", Variable)              ~ paste0("Education:", sub("education_level", "", Variable)),
       Variable == "total_malaria_episodes:study_armSP" ~ "Interaction: episodes × SP",
@@
   ) %>%
-  select(Variable, GVIF, `Df (VIF)`, `GVIF^(1/(2*Df))`)
-
-# 4) Print the VIF table to the console
+  dplyr::select(Variable, GVIF, `Df (VIF)`, `GVIF^(1/(2*Df))`)
+
+# 4) View
 print(vif_df)
@@
 
-## Random Forest Model Specification and Hyperparameter Tuning
-*Train/Test Split and Random Forest Model* which fits a random forest model and plots its ROC curve, is also used for Research Question 1 as an alternative ML approach
-```{r}
-# 0) Make sure the outcome is a factor
+## Random Forest: Specification, Tuning & ROC Comparison
+This section performs a 70/30 stratified train/test split, tunes a 500-tree random forest via 10-fold cross-validation to maximize ROC-AUC, fits the final model on the training set, and plots its ROC curve (alongside the logistic regression) as an alternative ML approach for Research Question 1.
+```{r model-comparison, echo=TRUE}
+# 0) Load required packages
+library(dplyr)
+library(rsample)     # for initial_split()
+library(tidymodels)  # recipes, workflows, tuning
+library(dials)       # for parameter grids
+library(pROC)        # for ROC curves
+
+# 1) Train/test split (70/30 stratified on outcome)
+set.seed(1234)
+split_data <- initial_split(promo_data_clean, prop = 0.7, strata = adverse_birth_outcome)
+train_data <- training(split_data)
+test_data  <- testing(split_data)
+
+# 2) Convert outcome to a two‐level factor (no/yes)
 train_data <- train_data %>%
-  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0, 1)))
-
-# 1) Define the random‐forest spec
-rf_model <- rand_forest(
+  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0,1), labels = c("no","yes")))
+test_data <- test_data %>%
+  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0,1), labels = c("no","yes")))
+
+# --------------------------------------------------
+# 3) Fit logistic regression and build its ROC curve
+# --------------------------------------------------
+logit_fit <- glm(
+  adverse_birth_outcome ~ 
+    total_malaria_episodes * study_arm + 
+    age_at_enrollment_years + 
+    gravidity + 
+    education_level,
+  family = binomial(link = "logit"),
+  data   = train_data
+)
+
+# Predict and construct ROC for logistic model
+logit_probs <- predict(logit_fit, newdata = test_data, type = "response")
+roc_logit  <- roc(test_data$adverse_birth_outcome, logit_probs, levels = c("no","yes"), direction = "<")
+
+# --------------------------------------------------
+# 4) Hyperparameter tuning for Random Forest
+# --------------------------------------------------
+rf_spec <- rand_forest(
     trees = 500,
@@
 
-library(dials)   # for finalize()
-
-# 2) Build the recipe 
-rf_recipe <- recipe(adverse_birth_outcome ~ 
-                      total_malaria_episodes + study_arm + 
-                      age_at_enrollment_years + gravidity + education_level,
-                    data = train_data) %>%
-  step_dummy(all_nominal_predictors())
-
-# 3) Assemble workflow
-rf_workflow <- workflow() %>%
-  add_model(rf_model) %>%
-  add_recipe(rf_recipe)
-
-# 4) Extract and finalize parameters
-rf_params <- parameters(rf_model) %>%
-  finalize(train_data)    # now mtry knows how many predictors
-
-# 5) Create a random tuning grid of size 20
-rf_grid <- grid_random(rf_params, size = 20)
-
-# 6) 10-fold stratified CV
-rf_folds <- vfold_cv(train_data, v = 10, strata = adverse_birth_outcome)
-
-# 7) Tune
-set.seed(123)
-rf_tune <- tune_grid(
-  rf_workflow,
-  resamples = rf_folds,
-  grid      = rf_grid,
-  metrics   = metric_set(roc_auc)
-)
-
-# 8) View results
-rf_tune %>% 
-  collect_metrics() %>% 
-  filter(.metric == "roc_auc")
-
-```
-
-# Random Forest model training and OOB Brier score output
-```{r}
-# 1)  outcome is a factor
-train_data <- train_data %>%
-  mutate(
-    adverse_birth_outcome = factor(
-      adverse_birth_outcome,
-      levels = c(0, 1),
-      labels = c("no", "yes")
-    )
-  )
-
-# 2) Define a Random Forest model (with fixed hyperparameters for illustration)
-rf_model <- rand_forest(
-    trees = 500,
-    mtry  = 3,    # you can choose a sensible default or tune later
-    min_n = 5     # likewise
-  ) %>%
-  set_engine("ranger") %>%
-  set_mode("classification")
-
-# 3) Build the recipe
 rf_recipe <- recipe(
@@
 
-# 4) Assemble into a workflow
-rf_workflow <- workflow() %>%
-  add_model(rf_model) %>%
+rf_wf <- workflow() %>%
+  add_model(rf_spec) %>%
   add_recipe(rf_recipe)
 
-# 5) Fit the random forest
-rf_fit <- rf_workflow %>% 
-  fit(data = train_data)
-
-# 6) Inspect the fitted workflow
-rf_fit
-
-```
-- Random forest fit ran without errors, but the OOB Brier score shows it’s essentially no better than random. 
-- That echoes what you saw with your logistic and other ML models (AUCs ~0.4–0.5): with these variables alone, none of the algorithms can predict the composite adverse birth outcome.
+rf_params <- parameters(rf_spec) %>%
+  finalize(train_data)
+
+set.seed(123)
+rf_tune <- tune_grid(
+  rf_wf,
+  resamples = vfold_cv(train_data, v = 10, strata = adverse_birth_outcome),
+  grid      = grid_random(rf_params, size = 20),
+  metrics   = metric_set(roc_auc)
+)
+
+# Select best hyperparameters and finalize
+best_rf <- select_best(rf_tune, metric = "roc_auc")
+rf_final_spec <- finalize_model(rf_spec, best_rf)
+
+# --------------------------------------------------
+# 5) Fit final Random Forest and build its ROC curve
+# --------------------------------------------------
+rf_final_wf <- workflow() %>%
+  add_model(rf_final_spec) %>%
+  add_recipe(rf_recipe)
+
+rf_fit <- fit(rf_final_wf, data = train_data)
+
+rf_probs <- predict(rf_fit, new_data = test_data, type = "prob") %>%
+  pull(.pred_yes)
+
+roc_rf <- roc(test_data$adverse_birth_outcome, rf_probs, levels = c("no","yes"), direction = "<")
+
+# --------------------------------------------------
+# 6) Plot both ROC curves together
+# --------------------------------------------------
+plot(
+  roc_logit,
+  col  = "blue",
+  lwd  = 2,
+  main = "ROC Comparison: Logistic Regression vs. Random Forest"
+)
+lines(roc_rf, col = "red", lwd = 2)
+legend(
+  "bottomright",
+  legend = c("Logistic Regression", "Random Forest"),
+  col    = c("blue", "red"),
+  lwd    = 2
+)
+```
 
@@
 
-I started by performing a 70/30 train/test split, stratified on the adverse birth outcome.
-For Research Question 1, I fit a logistic regression model with an interaction term on the training set, generated predictions on the test set, and plotted the ROC curve (with an AUC computed to assess discriminative performance).
-Then, I fit a random forest classifier (converting factor predictors to dummies in the recipe) and similarly evaluated its performance on the test set using ROC analysis.
-Finally, I overlaid both ROC curves to compare the models directly.
-This approach provides an "honest" assessment of model performance on unseen data and allows me to explore additional machine learning methods for addressing my research question.
+I performed a 70/30 train/test split stratified on adverse birth outcome.
+On the training set I fit a logistic regression (including our pre-specified interaction term), then generated test-set predictions and computed its ROC curve (AUC = 0.57).
+In parallel, I tuned a 500-tree random forest via 10-fold CV (stratified on outcome), fit the final model, and similarly evaluated its ROC (AUC = 0.54).
+Finally, I overlaid both ROC curves against the 45° no-skill line.
+Both models lie only slightly above that baseline, indicating poor discrimination with our current predictors—though the logistic regression edges out the random forest by a slim margin.
 
@@
 # Define a boosting model for classification (using boost_tree)
-```{r}
-# Define a boosting model for classification (using boost_tree)
-boost_model <- boost_tree(
+
+```{r boost-model, echo=TRUE}
+library(dplyr)
+library(rsample)
+library(tidymodels)  # recipes, workflows, tune_grid
+library(pROC)
+
+# 0) Subset & drop any rows with missing outcome/predictors
+ml_data <- promo_data_clean %>%
+  dplyr::select(
+    adverse_birth_outcome,
+    total_malaria_episodes,
+    study_arm,
+    age_at_enrollment_years,
+    gravidity,
+    education_level
+  ) %>%
+  tidyr::drop_na()
+
+# 1) 70/30 split, stratified on the now‐clean outcome
+set.seed(1234)
+split_obj  <- initial_split(ml_data, prop = 0.7, strata = adverse_birth_outcome)
+train_data  <- training(split_obj)
+test_data   <- testing(split_obj)
+
+# 2) Make sure the outcome is a 2‐level factor
+train_data <- train_data %>%
+  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0,1), labels = c("no","yes")))
+test_data  <- test_data  %>%
+  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0,1), labels = c("no","yes")))
+
+# 3) Boosting spec + recipe + workflow
+boost_spec <- boost_tree(
     mode       = "classification",
@@
 
-# Build a recipe (convert factor predictors to dummy variables)
-boost_recipe <- recipe(adverse_birth_outcome ~ total_malaria_episodes + study_arm +
-                         age_at_enrollment_years + gravidity + education_level,
-                       data = train_data) %>%
+boost_rec <- recipe(
+    adverse_birth_outcome ~
+      total_malaria_episodes +
+      study_arm +
+      age_at_enrollment_years +
+      gravidity +
+      education_level,
+    data = train_data
+  ) %>%
   step_dummy(all_nominal_predictors())
 
-# Create a workflow
-boost_workflow <- workflow() %>%
-  add_model(boost_model) %>%
-  add_recipe(boost_recipe)
-
-# Create a tuning grid for learning rate and tree depth
+boost_wf <- workflow() %>%
+  add_model(boost_spec) %>%
+  add_recipe(boost_rec)
+
+# 4) Tuning grid + CV + tune_grid (with roc_auc)
 boost_grid <- grid_regular(
@@
 
-# Create 10-fold cross-validation folds from the training data
-cv_folds <- vfold_cv(train_data, v = 10)
-
-# Perform 10-fold cross-validation tuning using the boosting model workflow
+cv_folds <- vfold_cv(train_data, v = 10, strata = adverse_birth_outcome)
+
+set.seed(2345)
 boost_tune <- tune_grid(
-  boost_workflow,
+  boost_wf,
   resamples = cv_folds,
-  grid = boost_grid
-)
-
-# Select best parameters based on ROC AUC
+  grid      = boost_grid,
+  metrics   = metric_set(roc_auc)
+)
+
+# 5) Finalize, fit & evaluate
 best_boost <- select_best(boost_tune, metric = "roc_auc")
-
-# Finalize the workflow with best parameters
-final_boost <- finalize_workflow(boost_workflow, best_boost)
-
-# Fit the final boosting model on the training data
-boost_fit <- final_boost %>% fit(data = train_data)
-
-# Generate predicted probabilities on the test set
-boost_preds <- predict(boost_fit, new_data = test_data, type = "prob")
-
-# Compute ROC for the boosting model using the probability for class "1"
-roc_boost <- roc(test_data$adverse_birth_outcome, boost_preds$.pred_1)
-plot(roc_boost, col = "green", lwd = 2, main = "ROC Curve: Boosting Model")
+final_boost_spec <- finalize_model(boost_spec, best_boost)
+
+boost_final_wf <- workflow() %>%
+  add_model(final_boost_spec) %>%
+  add_recipe(boost_rec)
+
+boost_fit <- fit(boost_final_wf, data = train_data)
+
+# ROC on the test set
+boost_probs <- predict(boost_fit, new_data = test_data, type = "prob") %>% pull(.pred_yes)
+roc_boost  <- roc(test_data$adverse_birth_outcome, boost_probs, levels = c("no","yes"))
+plot(roc_boost, col = "darkgreen", lwd = 2, main = "ROC Curve: XGBoost")
 auc_boost <- auc(roc_boost)
 print(paste("Boosting Model AUC:", round(auc_boost, 2)))
-
-```
-
+```
 The ROC curve for the boosting model rises more clearly above the diagonal line than the previous models, suggesting it does a better job distinguishing between those who experience adverse birth outcomes and those who do not.
@@
 ```{r}
-# Fit logistic regression with rms
-dd <- datadist(train_data); options(datadist = "dd")
-rms_model <- lrm(adverse_birth_outcome ~ gravidity + total_malaria_episodes + study_arm + education_level,
-                 data = train_data, x = TRUE, y = TRUE)
-# Calibration plot
-cal <- calibrate(rms_model, method="boot", B = 200)
-plot(cal, xlab="Predicted probability", ylab="Observed proportion")
-
-```
-
-```{r}
-library(dplyr); library(ggplot2)
+# only these variables are needed for the logistic model
+dd <- datadist(
+  train_data[, c(
+    "gravidity",
+    "total_malaria_episodes",
+    "study_arm",
+    "education_level",
+    "adverse_birth_outcome"
+  )]
+)
+options(datadist = "dd")
+
+# now fit with rms
+rms_model <- lrm(
+  adverse_birth_outcome ~ 
+    gravidity + total_malaria_episodes + study_arm + education_level,
+  data = train_data,
+  x = TRUE, y = TRUE
+)
+
+# bootstrap‐calibrated calibration curve
+cal <- calibrate(rms_model, method = "boot", B = 200)
+plot(
+  cal,
+  xlab = "Predicted probability",
+  ylab = "Observed proportion"
+)
+
+```
+
+```{r}
+library(dplyr)
+library(ggplot2)
 
@@
   mutate(
-    pred = predict(model_gravidity, type = "response"),
+    pred   = predict(model_gravidity, newdata = ., type = "response"),
     decile = ntile(pred, 10)
@@
 
-# 2) Summarize observed vs. predicted in each decile
+# 2) Summarize observed vs. predicted in each decile (explicitly using dplyr)
 calib_df <- train_data %>%
-  group_by(decile) %>%
-  summarize(
-    mean_pred = mean(pred),
-    obs_rate  = mean(as.numeric(adverse_birth_outcome) - 1)
-  )
-
-# 3) Plot
+  dplyr::group_by(decile) %>%
+  dplyr::summarize(
+    mean_pred = mean(pred, na.rm = TRUE),
+    obs_rate  = mean(as.numeric(adverse_birth_outcome) - 1, na.rm = TRUE)
+  )
+
+# 3) Plot calibration curve
 ggplot(calib_df, aes(x = mean_pred, y = obs_rate)) +
@@
 library(dplyr)
-library(knitr)
-
-enet_tune %>%
+
+# Get top 5 elastic‐net hyperparameter combinations and print as a regular tibble
+top5_enet <- enet_tune %>%
   collect_metrics() %>%
@@
   arrange(desc(mean)) %>%
-  slice_head(n = 5) %>%                # ← pull the top 5 rows
-  mutate(`AUC (mean ± SE)` = sprintf("%.3f \u00B1 %.3f", mean, std_err)) %>%
-  select(penalty, mixture, `AUC (mean ± SE)`) %>%
-  kable(
-    caption = "Table X. Top 5 elastic‐net hyperparameter combinations, ranked by mean cross‐validated AUC"
-  )
-
-```
-
-```{r}
-enet_tune %>%
+  slice_head(n = 5) %>%
+  mutate(`AUC (mean ± SE)` = sprintf("%.3f ± %.3f", mean, std_err)) %>%
+  dplyr::select(penalty, mixture, `AUC (mean ± SE)`)
+
+# Print the table
+print(top5_enet)
+
+```
+
+```{r}
+library(dplyr)
+
+# Extract and print the top 5 elastic-net tuning results as a regular tibble
+top5_enet <- enet_tune %>%
   collect_metrics() %>%
@@
   arrange(desc(mean)) %>%
-  head(5) %>%                          # ← base‐R head()
-  mutate(`AUC (mean ± SE)` = sprintf("%.3f \u00B1 %.3f", mean, std_err)) %>%
-  select(penalty, mixture, `AUC (mean ± SE)`) %>%
-  kable(
-    caption = "Table X. Top 5 elastic‐net hyperparameter combinations, ranked by mean cross‐validated AUC"
-  )
+  slice_head(n = 5) %>%
+  mutate(`AUC (mean ± SE)` = sprintf("%.3f ± %.3f", mean, std_err)) %>%
+  dplyr::select(penalty, mixture, `AUC (mean ± SE)`)
+
+print(top5_enet)
 
@@
 
-```{r}
-# 2) Specify elastic‑net logistic regression
-enet_spec <- logistic_reg(
-    penalty = tune(),   # strength of regularization
-    mixture = tune()    # 0=ridge ↔ 1=lasso
-  ) %>%
-  set_engine("glmnet")
-
-```
-
-```{r}
-# 3) Build the workflow
-enet_wf <- workflow() %>%
-  add_recipe(improved_recipe) %>%
-  add_model(enet_spec)
-
-```
-
-```{r}
-# 4) Create a tuning grid & 5) tune via 10‑fold CV
-enet_grid <- grid_regular(
-  penalty(range = c(-5, -1), trans = log10_trans()),  # 1e-5 → 1e-1
-  mixture(range = c(0, 1)),
-  levels = 5
-)
-
+## Elastic Net Logistic Regression: Specification, Tuning & Evaluation
+
+```{r feature-engineering}
+
+library(dplyr)
+library(tidyr)
+library(lubridate)
+library(rsample)
+library(tidymodels)
+
+# 0) Engineer features and impute their NAs
+full_data <- promo_data_clean %>%
+  mutate(
+    enroll_month = month(enrollment_date),
+    episodes_per_trimester = total_malaria_episodes /
+      (gestational_age_at_enrollment_weeks / (40/3))
+  )
+
+# compute medians for imputation
+median_enroll <- median(full_data$enroll_month, na.rm = TRUE)
+median_eps    <- median(full_data$episodes_per_trimester, na.rm = TRUE)
+
+full_data <- full_data %>%
+  mutate(
+    enroll_month           = if_else(is.na(enroll_month),    median_enroll,    enroll_month),
+    episodes_per_trimester = if_else(is.na(episodes_per_trimester),
+                                     median_eps,
+                                     episodes_per_trimester)
+  )
+
+# 1) Safe stratification check
+stratum_counts <- full_data %>% count(adverse_birth_outcome)
+can_stratify   <- nrow(stratum_counts) >= 2 && all(stratum_counts$n >= 2)
+
+# 2) 70/30 split (stratified if safe)
+set.seed(1234)
+split_obj <- if (can_stratify) {
+  initial_split(full_data, prop = 0.7, strata = adverse_birth_outcome)
+} else {
+  warning("Not enough data to stratify; performing unstratified split.")
+  initial_split(full_data, prop = 0.7)
+}
+
+raw_train <- training(split_obj)
+raw_test  <- testing(split_obj)
+
+# 3) Drop rows with NAs in the *core* six predictors or outcome
+train_data <- raw_train %>%
+  tidyr::drop_na(
+    adverse_birth_outcome,
+    total_malaria_episodes,
+    study_arm,
+    age_at_enrollment_years,
+    gravidity,
+    education_level
+  )
+
+test_data <- raw_test %>%
+  tidyr::drop_na(
+    adverse_birth_outcome,
+    total_malaria_episodes,
+    study_arm,
+    age_at_enrollment_years,
+    gravidity,
+    education_level
+  )
+
+# 4) Build dynamic v-fold CV (up to 10, but ≤ nrow(train_data))
+n_train <- nrow(train_data)
+if (n_train < 2) stop("Not enough rows for cross‐validation (need ≥ 2).")
+v <- min(10, n_train)
+
+cv_folds <- if (can_stratify) {
+  vfold_cv(train_data, v = v, strata = adverse_birth_outcome)
+} else {
+  vfold_cv(train_data, v = v)
+}
+
+# 5) Finally, tune your workflow—no more stray NAs in enroll_month
 enet_tune <- tune_grid(
@@
 ```{r}
-# 6) Select best, finalize, fit on train, & evaluate on test
-best_enet  <- select_best(enet_tune, metric = "roc_auc")
-enet_final <- finalize_workflow(enet_wf, best_enet)
-enet_fit   <- fit(enet_final, data = train_data)
-
-enet_preds <- predict(enet_fit, test_data, type = "prob") %>% bind_cols(test_data)
-enet_auc   <- roc_auc(enet_preds, truth = adverse_birth_outcome, .pred_1)
+library(tidymodels)
+
+# (A) Tune hyperparameters via grid search
+enet_tune <- tune_grid(
+  enet_wf,               # your workflow
+  resamples = cv_folds,  # the vfold_cv object you just built
+  grid      = enet_grid, # the grid of penalty/mixture values
+  metrics   = metric_set(roc_auc)
+)
+
+# (B) Look at the tuning results
+collect_metrics(enet_tune) %>% 
+  filter(.metric == "roc_auc") %>%
+  arrange(desc(mean))
+
+# (C) Pick the best parameters
+best_params <- select_best(enet_tune, "roc_auc")
+
+# (D) Finalize & refit on all of train_data
+final_enet_wf <- finalize_workflow(enet_wf, best_params)
+enet_fit      <- fit(final_enet_wf, data = train_data)
+
+# (E) Evaluate on the test set
+enet_preds <- predict(enet_fit, new_data = test_data, type = "prob") %>%
+  bind_cols(test_data)
+
+enet_auc <- roc_auc(enet_preds, truth = adverse_birth_outcome, .pred_yes)
 print(enet_auc)
-
-```
-The elastic‑net model achieved an AUC of 0.49 on the held‑out test set—below the 0.50 threshold for random guessing.
-This indicates that, even with polynomial, interaction, and SMOTE steps, the model cannot reliably distinguish between pregnancies with and without adverse outcomes.
-Such near‑random performance suggests either insufficient predictive signal in the available features or the need for additional, more informative variables and further model refinement.
-
-##################################################################################################################################################################
-
-```{r}
-# Create temporal & composite features in the training set
-train_data <- train_data %>%
-  mutate(
-    enroll_month = lubridate::month(enrollment_date),
-    episodes_per_trimester = total_malaria_episodes / (gestational_age_at_enrollment_weeks / (40/3))
-  )
-
-test_data <- test_data %>%
-  mutate(
-    enroll_month = lubridate::month(enrollment_date),
-    episodes_per_trimester = total_malaria_episodes / (gestational_age_at_enrollment_weeks / (40/3))
-  )
-
-```
-
-```{r}
-imbalanced_recipe <- recipe(adverse_birth_outcome ~ ., data = train_data) %>%
-  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
-  step_mutate(
-    total_malaria_episodes = as.numeric(total_malaria_episodes),
-    gravidity = as.numeric(gravidity)
-  ) %>%
-  step_adasyn(adverse_birth_outcome) %>%
-  step_normalize(all_predictors())
-
-```
-
-```{r}
-enet_wf <- workflow() %>%
-  add_model(enet_spec) %>%
-  add_recipe(imbalanced_recipe)
-
-enet_bayes <- tune_bayes(
-  enet_wf,
-  resamples = cv_folds,
-  initial = enet_tune,
-  iter = 20,
-  metrics = metric_set(roc_auc)
-)
-
-best_enet_bayes <- select_best(enet_bayes, metric = "roc_auc")
-
-```
-
-```{r}
-library(stacks)
-
-# Assume you've tuned logistic, rf, xgb already: logistic_tune, rf_tune, xgb_tune
-stack <- stacks() %>%
-  add_candidates(logistic_tune) %>%
-  add_candidates(rf_tune)       %>%
-  add_candidates(xgb_tune)      %>%
-  blend_predictions()           %>%
-  fit_members()
-
-# Evaluate on test set
-stack_preds <- predict(stack, test_data, type = "prob") %>% bind_cols(test_data)
-stack_auc   <- roc_auc(stack_preds, truth = adverse_birth_outcome, .pred_1)
-print(stack_auc)
-
-```
-
-```{r}
-svm_spec <- svm_rbf(
-    cost      = tune(),
-    rbf_sigma = tune()
-  ) %>%
-  set_engine("kernlab") %>%
-  set_mode("classification")
-
-svm_wf <- workflow() %>%
-  add_model(svm_spec) %>%
-  add_recipe(improved_recipe)  # or your imbalanced_recipe
-
-svm_grid <- grid_regular(
-  cost(range = c(-5, 2), trans = log2_trans()),
-  rbf_sigma(range = c(-3, 0), trans = log2_trans()),
-  levels = 5
-)
-
-svm_tune <- tune_grid(
-  svm_wf,
-  resamples = cv_folds,
-  grid      = svm_grid,
-  metrics   = metric_set(roc_auc)
-)
-
-best_svm <- select_best(svm_tune, metric = "roc_auc")
-
-```
-
-```{r}
-# For your best model (e.g. enet_fit or stack)
-pr    <- pr_curve(stack_preds, truth = adverse_birth_outcome, .pred_1)
-pr_auc <- pr_auc(stack_preds, truth = adverse_birth_outcome, .pred_1)
-
-autoplot(pr) +
-  labs(title = "Precision–Recall Curve", subtitle = paste("PR AUC =", round(pr_auc$.estimate,2)))
 

### 2025-04-28 products/manuscript/Manuscript.qmd
P3
Update

@@
 )
-
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Load Required Libraries for Data Handling, Visualization, and Analysis
-
 # Load Required Libraries
@@
 library(kableExtra)         # Enhanced markdown tables
+
+# Visualization enhancements
+library(ggpubr)             # Publication-ready plots
 
@@
 library(xgboost)            # Gradient boosting engine
-library(generalhoslem)
-library(ResourceSelection)
-library(rms)
+
+# Goodness-of-fit & diagnostics
+library(generalhoslem)      # Hosmer–Lemeshow tests
+library(ResourceSelection)  # Additional GOF tests
+library(rms)                # Regression modeling strategies
+library(car)                # Companion to applied regression
+
 ```

### 2025-04-30 products/manuscript/Manuscript.qmd
update

@@
 library(car)                # Companion to applied regression
-
 ```
@@
 knitr::include_graphics("Images/Table1Baseline.png")
-
 ```
@@
 knitr::include_graphics("Images/Figure1.DistribTotal Malaria.png")
-
 ```
@@
 knitr::include_graphics("Images/Table2.Outcome Measures Malaria .png")
-
 ```
@@
 knitr::include_graphics("Images/Table4Interaction.png")
-
 ```
@@
 knitr::include_graphics("Images/AdjustedRatio.png")
-
 ```
@@
 knitr::include_graphics("Images/Figure 3. Predicted Probability.png")
-
 ```
@@
 knitr::include_graphics("Images/Table6.Discriminative Performance (AUC) of ML Models.png")
-
 ```
@@
 knitr::include_graphics("Images/Figure 4.Comparative Discriminative Performance of ML .png")
-
 ```

### 2025-04-30 products/manuscript/Manuscript.qmd
update1

@@
 **Figure 1: Distribution of Total Malaria Episodes by IPTp Treatment Arm**
-```{r,echo=FALSE, message=FALSE, warning=FALSE}}
-knitr::include_graphics("Images/Figure1.DistribTotal Malaria.png")
+```{r,echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("Images/Figure1.DistribTotalMalaria.png")
 ```
@@
 **Table 2: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
-```{r}
-knitr::include_graphics("Images/Table2.Outcome Measures Malaria .png")
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("Images/Table2.OutcomeMeasuresandMalariaExposure.png")
 ```
@@
 **Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes**
-```{r}
-knitr::include_graphics("Images/Figure2.Differential Impact of IPTp.png")
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("Images/Figure2.DifferentialImpactofIPTp.png")
 ```
@@
 **Table 3: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**
-
-```{r,echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics("Images/Table4Interaction.png")
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("Images/Table3Interaction.png")
 ```
@@
 **Figure 4: Predicted Probability of Adverse Outcome by Gravidity (Age < 25)**
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics("Images/Figure 3. Predicted Probability.png")
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("Images/Figure3.PredictedProbability.png")
 ```
@@
 **Table 5:  Discriminative Performance (AUC) of Machine Learning Models Predicting Adverse Birth Outcomes Under Different IPTp Regimens**
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics("Images/Table6.Discriminative Performance (AUC) of ML Models.png")
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("Images/Table6.DiscriminativePerformance(AUC)ofMLModels.png")
 ```
@@
 **Figure 5 : Comparative Discriminative Performance of ML Models for Predicting Adverse Birth Outcomes**
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-knitr::include_graphics("Images/Figure 4.Comparative Discriminative Performance of ML .png")
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("Images/Figure5Calibration.png")
 ```

### 2025-05-02 code/analysis-code/Statistical_Analysis.qmd
Project 1

@@
 library(dplyr)
-library(tidyr)
 library(lubridate)
@@
 library(tidymodels)
-
-# 0) Engineer features and impute their NAs
+library(themis)    # for step_smote()
+library(scales)    # for log10_trans()
+
+set.seed(1234)
+
+# 1) PREPARE full_data (no complex imputation needed)
 full_data <- promo_data_clean %>%
   mutate(
-    enroll_month = month(enrollment_date),
-    episodes_per_trimester = total_malaria_episodes /
-      (gestational_age_at_enrollment_weeks / (40/3))
-  )
-
-# compute medians for imputation
-median_enroll <- median(full_data$enroll_month, na.rm = TRUE)
-median_eps    <- median(full_data$episodes_per_trimester, na.rm = TRUE)
-
-full_data <- full_data %>%
-  mutate(
-    enroll_month           = if_else(is.na(enroll_month),    median_enroll,    enroll_month),
-    episodes_per_trimester = if_else(is.na(episodes_per_trimester),
-                                     median_eps,
-                                     episodes_per_trimester)
-  )
-
-# 1) Safe stratification check
-stratum_counts <- full_data %>% count(adverse_birth_outcome)
-can_stratify   <- nrow(stratum_counts) >= 2 && all(stratum_counts$n >= 2)
-
-# 2) 70/30 split (stratified if safe)
-set.seed(1234)
-split_obj <- if (can_stratify) {
-  initial_split(full_data, prop = 0.7, strata = adverse_birth_outcome)
-} else {
-  warning("Not enough data to stratify; performing unstratified split.")
-  initial_split(full_data, prop = 0.7)
-}
-
-raw_train <- training(split_obj)
-raw_test  <- testing(split_obj)
-
-# 3) Drop rows with NAs in the *core* six predictors or outcome
-train_data <- raw_train %>%
-  tidyr::drop_na(
+    adverse_birth_outcome               = factor(adverse_birth_outcome, levels = c(0,1), labels = c("no","yes")),
+    total_malaria_episodes              = as.integer(total_malaria_episodes),
+    gravidity                           = as.integer(gravidity),
+    gestational_age_at_enrollment_weeks = as.numeric(gestational_age_at_enrollment_weeks),
+    malaria_infection_rate_during_pregnancy = as.numeric(malaria_infection_rate_during_pregnancy),
+    placental_malaria_by_rogerson_criteria = factor(placental_malaria_by_rogerson_criteria)
+  ) %>%
+  drop_na(
     adverse_birth_outcome,
     total_malaria_episodes,
+    gravidity,
+    gestational_age_at_enrollment_weeks,
+    malaria_infection_rate_during_pregnancy,
+    placental_malaria_by_rogerson_criteria,
     study_arm,
     age_at_enrollment_years,
-    gravidity,
     education_level
@@
 
-test_data <- raw_test %>%
-  tidyr::drop_na(
-    adverse_birth_outcome,
-    total_malaria_episodes,
-    study_arm,
-    age_at_enrollment_years,
-    gravidity,
-    education_level
-  )
-
-# 4) Build dynamic v-fold CV (up to 10, but ≤ nrow(train_data))
-n_train <- nrow(train_data)
-if (n_train < 2) stop("Not enough rows for cross‐validation (need ≥ 2).")
-v <- min(10, n_train)
-
-cv_folds <- if (can_stratify) {
-  vfold_cv(train_data, v = v, strata = adverse_birth_outcome)
-} else {
-  vfold_cv(train_data, v = v)
-}
-
-# 5) Finally, tune your workflow—no more stray NAs in enroll_month
+# 2) Split 70/30 (unstratified)
+split_obj  <- initial_split(full_data, prop = 0.7)
+train_data <- training(split_obj)
+test_data  <- testing(split_obj)
+
+# 3) Recipe (no more enroll_month or episodes_per_trimester)
+enet_recipe <- recipe(adverse_birth_outcome ~
+                        total_malaria_episodes +
+                        gravidity +
+                        gestational_age_at_enrollment_weeks +
+                        malaria_infection_rate_during_pregnancy +
+                        placental_malaria_by_rogerson_criteria +
+                        study_arm +
+                        age_at_enrollment_years +
+                        education_level,
+                      data = train_data) %>%
+  step_dummy(all_nominal_predictors()) %>%
+  step_smote(adverse_birth_outcome) %>%
+  step_normalize(all_numeric_predictors())
+
+# 4) Model spec & workflow
+enet_spec <- logistic_reg(penalty = tune(), mixture = tune()) %>%
+  set_engine("glmnet") %>%
+  set_mode("classification")
+
+enet_wf <- workflow() %>%
+  add_recipe(enet_recipe) %>%
+  add_model(enet_spec)
+
+# 5) Tuning grid
+enet_grid <- grid_regular(
+  penalty(range = c(-5, -1), trans = log10_trans()),
+  mixture(range = c(0, 1)),
+  levels = 5
+)
+
+# 6) CV folds
+cv_folds <- vfold_cv(train_data, v = 5)
+
+# 7) Tune
 enet_tune <- tune_grid(
@@
 
+# 8) Select best & refit
+best_params   <- select_best(enet_tune, metric = "roc_auc")
+final_enet_wf <- finalize_workflow(enet_wf, best_params)
+enet_fit      <- fit(final_enet_wf, data = train_data)
+
+# 9) Evaluate on test set
+enet_preds <- predict(enet_fit, new_data = test_data, type = "prob") %>%
+  bind_cols(test_data)
+
+print( roc_auc(enet_preds, truth = adverse_birth_outcome, .pred_yes) )
+
 ```
@@
   enet_wf,               # your workflow
-  resamples = cv_folds,  # the vfold_cv object you just built
-  grid      = enet_grid, # the grid of penalty/mixture values
+  resamples = cv_folds,  # the vfold_cv object
+  grid      = enet_grid, # grid of penalty/mixture values
   metrics   = metric_set(roc_auc)
@@
 
-# (B) Look at the tuning results
-collect_metrics(enet_tune) %>% 
+# (B) Look at the tuning results for ROC AUC
+roc_results <- collect_metrics(enet_tune) %>%
   filter(.metric == "roc_auc") %>%
@@
 
-# (C) Pick the best parameters
-best_params <- select_best(enet_tune, "roc_auc")
+print(roc_results)
+
+# (C) Pick the best parameters (note the named argument)
+best_params <- select_best(enet_tune, metric = "roc_auc")
 
@@
 ```{r}
+library(tidymodels)
+library(readr)
+library(here)
+
 # 2. Read & split data
@@
 set.seed(2025)
-split    <- initial_split(df, prop = 0.7, strata = adverse_birth_outcome)
-train    <- training(split)
-test     <- testing(split)
+splits    <- initial_split(df, prop = 0.7, strata = adverse_birth_outcome)
+train     <- training(splits)
+test      <- testing(splits)
 cv_folds  <- vfold_cv(train, v = 5, strata = adverse_birth_outcome)
 
-```
-
-```{r}
 # 3. Preprocessing recipe
-ml_rec <- recipe(adverse_birth_outcome ~ age_at_enrollment_years
-                  + gravidity
-                  + total_malaria_episodes
-                  + education_level
-                  + study_arm
-                  + socioeconomic_status,
+ml_rec <- recipe(adverse_birth_outcome ~ 
+                   age_at_enrollment_years +
+                   gravidity +
+                   total_malaria_episodes +
+                   education_level +
+                   study_arm +
+                   normalized_malaria_rate,   # ← replaced here
                  data = train) %>%
@@
 
-```
-
-```{r}
 # 4. Model specifications & workflows
@@
 
-lasso_wf  <- workflow() %>% add_model(lasso_mod) %>% add_recipe(ml_rec)
+lasso_wf  <- workflow() %>% 
+  add_model(lasso_mod) %>% 
+  add_recipe(ml_rec)
 
@@
 
-rf_wf  <- workflow() %>% add_model(rf_mod) %>% add_recipe(ml_rec)
+rf_wf  <- workflow() %>% 
+  add_model(rf_mod) %>% 
+  add_recipe(ml_rec)
 
@@
 
-xgb_wf <- workflow() %>% add_model(xgb_mod) %>% add_recipe(ml_rec)
-
-```
-
-```{r}
+xgb_wf <- workflow() %>% 
+  add_model(xgb_mod) %>% 
+  add_recipe(ml_rec)
+
 # 5. Tuning grids & 5‐fold CV
@@
 
-```
-
-```{r}
-# 6. Collect cross‐validated AUC results
-lasso_auc <- lasso_res %>% filter(.metric=="roc_auc") %>% summarize(mean = mean(mean), se = mean(std_err)) %>% mutate(model="LASSO")
-rf_auc    <- rf_res    %>% filter(.metric=="roc_auc") %>% summarize(mean = mean(mean), se = mean(std_err)) %>% mutate(model="RF")
-xgb_auc   <- xgb_res   %>% filter(.metric=="roc_auc") %>% summarize(mean = mean(mean), se = mean(std_err)) %>% mutate(model="XGB")
-
-bind_rows(lasso_auc, rf_auc, xgb_auc) %>%
-  rename(AUC = mean, SE = se) %>%
-  knitr::kable(digits=3, caption="Table 6: 5‐fold CV ROC AUC for Each Model")
-
-```
+library(dplyr)
+library(knitr)
+
+# 6) Pull the metrics into plain tibbles
+lasso_metrics <- collect_metrics(lasso_res)
+rf_metrics    <- collect_metrics(rf_res)
+xgb_metrics   <- collect_metrics(xgb_res)
+
+# 7) Summarize AUC for each model
+lasso_auc <- lasso_metrics %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean),
+    SE   = mean(std_err)
+  ) %>%
+  mutate(Model = "LASSO")
+
+rf_auc <- rf_metrics %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean),
+    SE   = mean(std_err)
+  ) %>%
+  mutate(Model = "Random Forest")
+
+xgb_auc <- xgb_metrics %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean),
+    SE   = mean(std_err)
+  ) %>%
+  mutate(Model = "XGBoost")
+
+library(dplyr)
+library(knitr)
+
+# 8) Combine into one data frame and print as a plain table
+table6 <- bind_rows(lasso_auc, rf_auc, xgb_auc) %>%
+  dplyr::select(Model, Mean, SE)
+
+print(table6)
+
+```
+
+```{r}
+library(tidymodels)  # for collect_metrics()
+
+# 1) Collect tuning metrics into plain tibbles
+lasso_metrics <- collect_metrics(lasso_res)
+rf_metrics    <- collect_metrics(rf_res)
+xgb_metrics   <- collect_metrics(xgb_res)
+
+# 2) Summarize ROC AUC for each model
+lasso_auc <- lasso_metrics %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean),
+    SE   = mean(std_err)
+  ) %>%
+  mutate(Model = "LASSO")
+
+rf_auc <- rf_metrics %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean),
+    SE   = mean(std_err)
+  ) %>%
+  mutate(Model = "Random Forest")
+
+xgb_auc <- xgb_metrics %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean),
+    SE   = mean(std_err)
+  ) %>%
+  mutate(Model = "XGBoost")
+
+# 3) Combine into one table
+auc_table <- bind_rows(lasso_auc, rf_auc, xgb_auc)
+
+# 4) Print Table 6
+auc_table %>%
+  dplyr::select(Model, Mean, SE) %>%
+  knitr::kable(
+    digits  = 3,
+    caption = "Table 6: 5‑fold CV ROC AUC for Each Model"
+  )
+# 4) Print as a plain data frame (regular table)
+print(auc_table)
+
+```
+
+```{r}
+ml_rec <- recipe(adverse_birth_outcome ~
+                   age_at_enrollment_years +
+                   gravidity +
+                   total_malaria_episodes +
+                   education_level +
+                   study_arm +
+                   normalized_malaria_rate,
+                 data = train) %>%
+  step_dummy(all_nominal_predictors()) %>%   # one‑hot encode factors
+  step_zv(all_predictors()) %>%              # drop zero‑variance
+  step_normalize(all_numeric_predictors())   # center & scale numerics
+
+```
+
+```{r}
+# 4. Model specifications & workflows
+
+## (a) LASSO‐penalized logistic regression
+lasso_mod <- logistic_reg(penalty = tune(), mixture = 1) %>%
+  set_engine("glmnet") %>%
+  set_mode("classification")
+
+lasso_wf  <- workflow() %>% add_model(lasso_mod) %>% add_recipe(ml_rec)
+
+## (b) Random forest
+rf_mod <- rand_forest(mtry = tune(), trees = 1000, min_n = tune()) %>%
+  set_engine("ranger") %>%
+  set_mode("classification")
+
+rf_wf  <- workflow() %>% add_model(rf_mod) %>% add_recipe(ml_rec)
+
+## (c) XGBoost
+xgb_mod <- boost_tree(trees = tune(),
+                      tree_depth = tune(),
+                      learn_rate = tune()) %>%
+  set_engine("xgboost") %>%
+  set_mode("classification")
+
+xgb_wf <- workflow() %>% add_model(xgb_mod) %>% add_recipe(ml_rec)
+
+```
+
+```{r}
+# 5. Tuning grids & 5‐fold CV
+set.seed(2025)
+
+lasso_grid <- grid_regular(penalty(), levels = 20)
+
+rf_params <- parameters(rf_mod) %>%
+  update(mtry = mtry(c(1, 6)))
+rf_grid   <- grid_latin_hypercube(rf_params, size = 20)
+
+xgb_params <- parameters(xgb_mod)
+xgb_grid   <- grid_latin_hypercube(xgb_params, size = 30)
+
+ml_metrics <- metric_set(roc_auc)
+
+lasso_res <- tune_grid(lasso_wf, resamples = cv_folds,
+                       grid = lasso_grid, metrics = ml_metrics)
+
+rf_res    <- tune_grid(rf_wf,    resamples = cv_folds,
+                       grid = rf_grid,    metrics = ml_metrics)
+
+xgb_res   <- tune_grid(xgb_wf,   resamples = cv_folds,
+                       grid = xgb_grid,   metrics = ml_metrics)
+
+```
+
+```{r}
+library(dplyr)
+
+# 6) Pull the AUC metrics out and summarize
+lasso_auc <- collect_metrics(lasso_res) %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean, na.rm = TRUE),
+    SE   = mean(std_err, na.rm = TRUE)
+  ) %>%
+  mutate(Model = "LASSO")
+
+rf_auc <- collect_metrics(rf_res) %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean, na.rm = TRUE),
+    SE   = mean(std_err, na.rm = TRUE)
+  ) %>%
+  mutate(Model = "Random Forest")
+
+xgb_auc <- collect_metrics(xgb_res) %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean, na.rm = TRUE),
+    SE   = mean(std_err, na.rm = TRUE)
+  ) %>%
+  mutate(Model = "XGBoost")
+
+# 7) Combine & print
+auc_table <- bind_rows(lasso_auc, rf_auc, xgb_auc) %>%
+  dplyr::select(Model, Mean, SE)
+
+print(auc_table)
+
+```
+
+##################################################################################################################################################################
+
+*END*

### 2025-05-02 products/manuscript/Manuscript.qmd
Project 1

@@
 **Methods:**
-We analyzed data from a double-blind, randomized trial in Uganda (ClinEpiDB #24, n=782 HIV-uninfected pregnant women).
-Participants were randomized to receive either sulfadoxine‑pyrimethamine (SP) or dihydroartemisinin‑piperaquine (DP) as IPTp.
+We analyzed data from a double‑blind, randomized trial in Uganda (ClinEpiDB#24, n=782 HIV‑uninfected pregnant women).
+Participants were randomized to receive either sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP) as IPTp.
 Primary outcomes were preterm birth, low birth weight, and stillbirth, combined into a composite adverse outcome.
-We fit multivariable logistic models evaluating (1) main effects of total malaria episodes and treatment arm, (2) their interaction, adjusting for maternal age, gravidity, and education, and (3) a prespecified subgroup analysis in women < 25 years to assess gravidity’s protective effect.
-Predicted probabilities were derived via ggeffects, and discriminative performance of several ML models (random forest, XGBoost, logistic regression) was summarized by AUC.
+We first fit multivariable logistic regression models to evaluate (1) the main effects of total malaria episodes and treatment arm, (2) their interaction, adjusting for maternal age, gravidity, and education level, and (3) a prespecified subgroup analysis in women <25 years to assess gravidity’s protective effect.
+
+To explore more flexible classifiers, we compared four algorithms, standard logistic regression, elastic‑net (LASSO‑penalized) logistic regression, random forest, and XGBoost.
+Hyperparameters were tuned by 5‑fold cross‑validation on a 70% training split, and mean cross‑validated AUC (±SE) was used to compare methods (elastic net:0.55±0.01; random forest:0.52±0.01; XGBoost:0.50±0.01; standard logistic:0.48±0.01).
+The best‑tuned models were then refitted on the full training set and evaluated on the remaining 30% hold‑out set; final test‑set AUCs are reported in Table6.
 
 **Results:**
-Adjusting for maternal age, gravidity, and socioeconomic status, our multivariable models revealed a significant interaction between malaria episode frequency and the SP regimen (OR 1.47 per additional episode, 95% CI 1.16–1.88; p = 0.002), indicating that each extra malaria episode increased the odds of an adverse birth outcome by 47% in SP‐treated women compared to those on DP.
-In contrast, neither malaria burden alone (OR 0.88 per episode, 95% CI 0.75–1.03; p = 0.12) nor treatment arm alone (OR 0.69 for DP vs SP, 95% CI 0.44–1.08; p = 0.10) reached significance, underscoring the importance of their interplay.
-In women under 25, each additional pregnancy conferred a 14% reduction in adverse‐outcome risk (OR 0.86, 95% CI 0.74–1.00; p = 0.044).
-The interaction model calibrated well (Hosmer–Lemeshow p = 0.37; MAE = 0.03), and on the held-out test set, the AUCs were 0.48 for logistic regression, 0.42 for the random forest, 0.44 for XGBoost, and 0.49 for the elastic net.
+A significant interaction between malaria episode frequency and the SP regimen emerged: each additional malaria episode under SP was associated with 47% higher odds of an adverse birth outcome than DP (OR 1.47 per episode, 95% CI 1.16–1.88; p=0.002).
+By itself, total malaria episodes did not significantly predict the composite outcome (OR 0.88 per episode, 95% CI 0.75–1.03; p=0.12), nor did the treatment arm alone (OR 0.69 for DP vs SP, 95% CI 0.44–1.08; p = 0.10), underscoring the critical role of their interplay.
+In women under 25, each additional pregnancy reduced the odds of an adverse outcome by 14% (OR 0.86, 95% CI 0.74–1.00; p=0.044).
+The final interaction model exhibited good calibration (Hosmer–Lemeshow p=0.37; mean absolute error=0.03).
+Among machine‑learning classifiers, elastic‑net (LASSO) achieved the highest mean cross‑validated AUC (0.55±0.01), followed by random forest (0.52±0.01) and XGBoost (0.50±0.01) on the 70% training split.
+On the independent 30% hold‑out test set, final AUCs were 0.48 for standard logistic regression, 0.42 for random forest, 0.44 for XGBoost, and 0.49 for elastic‑net (Table 6).
 

### 2025-05-02 products/manuscript/supplement/Supplementary-Material.qmd
Project 1

@@
 
+```{r}
+
+```{r}
+library(tidymodels)
+library(readr)
+library(here)
+
+# 2. Read & split data
+df <- read_csv(here("data","clean","PROMO_Data_clean.csv")) %>%
+  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0,1)))
+
+set.seed(2025)
+splits    <- initial_split(df, prop = 0.7, strata = adverse_birth_outcome)
+train     <- training(splits)
+test      <- testing(splits)
+cv_folds  <- vfold_cv(train, v = 5, strata = adverse_birth_outcome)
+
+# 3. Preprocessing recipe
+ml_rec <- recipe(adverse_birth_outcome ~ 
+                   age_at_enrollment_years +
+                   gravidity +
+                   total_malaria_episodes +
+                   education_level +
+                   study_arm +
+                   normalized_malaria_rate,   # ← replaced here
+                 data = train) %>%
+  step_dummy(all_nominal_predictors()) %>%  # one‐hot encode
+  step_zv(all_predictors()) %>%             # remove zero‐variance
+  step_normalize(all_numeric_predictors())  # center & scale
+
+# 4. Model specifications & workflows
+
+## (a) LASSO‐penalized logistic regression
+lasso_mod <- logistic_reg(penalty = tune(), mixture = 1) %>%
+  set_engine("glmnet") %>%
+  set_mode("classification")
+
+lasso_wf  <- workflow() %>% 
+  add_model(lasso_mod) %>% 
+  add_recipe(ml_rec)
+
+## (b) Random forest
+rf_mod <- rand_forest(mtry = tune(), trees = 1000, min_n = tune()) %>%
+  set_engine("ranger") %>%
+  set_mode("classification")
+
+rf_wf  <- workflow() %>% 
+  add_model(rf_mod) %>% 
+  add_recipe(ml_rec)
+
+## (c) XGBoost
+xgb_mod <- boost_tree(trees = tune(),
+                      tree_depth = tune(),
+                      learn_rate = tune()) %>%
+  set_engine("xgboost") %>%
+  set_mode("classification")
+
+xgb_wf <- workflow() %>% 
+  add_model(xgb_mod) %>% 
+  add_recipe(ml_rec)
+
+# 5. Tuning grids & 5‐fold CV
+set.seed(2025)
+
+lasso_grid <- grid_regular(penalty(), levels = 20)
+
+rf_params <- parameters(rf_mod) %>%
+  update(mtry = mtry(c(1, 6)))
+rf_grid   <- grid_latin_hypercube(rf_params, size = 20)
+
+xgb_params <- parameters(xgb_mod)
+xgb_grid   <- grid_latin_hypercube(xgb_params, size = 30)
+
+ml_metrics <- metric_set(roc_auc)
+
+lasso_res <- tune_grid(lasso_wf, resamples = cv_folds,
+                       grid = lasso_grid, metrics = ml_metrics)
+
+rf_res    <- tune_grid(rf_wf,    resamples = cv_folds,
+                       grid = rf_grid,    metrics = ml_metrics)
+
+xgb_res   <- tune_grid(xgb_wf,   resamples = cv_folds,
+                       grid = xgb_grid,   metrics = ml_metrics)
+
+library(dplyr)
+library(knitr)
+
+# 6) Pull the metrics into plain tibbles
+lasso_metrics <- collect_metrics(lasso_res)
+rf_metrics    <- collect_metrics(rf_res)
+xgb_metrics   <- collect_metrics(xgb_res)
+
+# 7) Summarize AUC for each model
+lasso_auc <- lasso_metrics %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean),
+    SE   = mean(std_err)
+  ) %>%
+  mutate(Model = "LASSO")
+
+rf_auc <- rf_metrics %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean),
+    SE   = mean(std_err)
+  ) %>%
+  mutate(Model = "Random Forest")
+
+xgb_auc <- xgb_metrics %>%
+  filter(.metric == "roc_auc") %>%
+  dplyr::summarise(
+    Mean = mean(mean),
+    SE   = mean(std_err)
+  ) %>%
+  mutate(Model = "XGBoost")
+
+library(dplyr)
+library(knitr)
+
+# 8) Combine into one data frame and print as a plain table
+table6 <- bind_rows(lasso_auc, rf_auc, xgb_auc) %>%
+  dplyr::select(Model, Mean, SE)
+
+print(table6)
+
 ```
+
+During 5‑fold cross‑validation, the elastic‐net model achieved the highest mean AUC (0.55±0.01), followed by random forest (0.52±0.01) and XGBoost (0.50±0.01).
 

### 2025-05-02 products/manuscript/supplement/Supplementary-Material.qmd
update project

@@
 format:
-  pdf:
+  docx:
     toc: false
@@
 ---
-
-This shows some materials that could go into a supplementary file.
-Often you want/need references here too.
-You can use the same reference bib file for this and the main text (as done here) or have separate bib files.
-
-For illustrative purposes, I'm doing the supplement as pdf.
-For this to work, you need a (La)TeX system installed.
-It's easy.
-Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).
-
-Of course you would choose the format based on needs.
-
-I'm also using a different style for the references here.
-(vancouver vs apa in the main manuscript).
-Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header.
-We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.
 
@@
 
-```{r}
-# Load necessary package for ROC analysis
+```{r roc-calibration, echo=TRUE, message=FALSE, warning=FALSE}
+# 1) Load required libraries
+library(dplyr)
 library(pROC)
+library(caret)
+library(here)
 
-# Generate predicted probabilities using the logistic regression model (model_gravidity)
-promo_data_young$predicted_prob <- predict(model_gravidity, type = "response")
+# 2) Read cleaned data and subset to women < 25
+promo_data_clean <- read.csv(here("data/clean/PROMO_Data_clean.csv"))
+promo_data_young <- promo_data_clean %>%
+  filter(age_at_enrollment_years < 25) %>%
+  mutate(
+    # Composite adverse outcome: 1 if preterm, stillbirth, or low birth weight
+    low_birth_weight      = as.integer(birth_weight < 2.5),
+    adverse_birth_outcome = as.integer(
+      preterm_births_count > 0 |
+      stillbirth_bin == 1        |
+      low_birth_weight == 1
+    )
+  )
 
-# Create the ROC curve
-roc_obj <- roc(promo_data_young$adverse_birth_outcome, promo_data_young$predicted_prob)
-plot(roc_obj, col = "blue", lwd = 2, main = "ROC Curve for Model: Gravidity in Women < 25")
-auc_value <- auc(roc_obj)
-print(paste("AUC:", round(auc_value, 2)))
+# 3) Fit logistic regression model
+model_gravidity <- glm(
+  adverse_birth_outcome ~ gravidity + total_malaria_episodes +
+    study_arm + education_level,
+  family = binomial(link = "logit"),
+  data   = promo_data_young
+)
 
-# Load necessary package for calibration plot
-library(caret)
+# 4) Generate predicted probabilities
+promo_data_young$predicted_prob <- predict(
+  model_gravidity,
+  newdata = promo_data_young,
+  type    = "response"
+)
 
-# Create a calibration plot
-calibration_data <- data.frame(
+# 5) ROC curve and AUC
+roc_obj  <- roc(
+  promo_data_young$adverse_birth_outcome,
+  promo_data_young$predicted_prob
+)
+plot(
+  roc_obj,
+  col   = "blue",
+  lwd   = 2,
+  main  = "ROC Curve: Gravidity Model (Age < 25)"
+)
+auc_val <- auc(roc_obj)
+cat("AUC:", round(auc_val, 2), "\n")
+
+# 6) Calibration plot
+data_cal <- data.frame(
   observed = factor(promo_data_young$adverse_birth_outcome, levels = c(0,1)),
@@
 )
-
-# Use the calibration function from the caret package
-cal_plot <- calibration(observed ~ predicted, data = calibration_data, class = "1")
-plot(cal_plot, main = "Calibration Plot for Model: Gravidity in Women < 25")
+cal_plot <- calibration(
+  observed ~ predicted,
+  data  = data_cal,
+  class = "1"
+)
+plot(cal_plot, main = "Calibration Plot: Gravidity Model (Age < 25)")
 
@@
 Overall, while gravidity is statistically significant, my model as a whole isn't very effective at distinguishing between those who experience adverse outcomes and those who don't, and its probability estimates need improvement.
-
-# Heatmap of cross-validated AUC\nfor Elastic-net hyperparameter grid
-```{r}
-library(ggplot2)
-
-enet_tune %>% 
-  collect_metrics() %>% 
-  filter(.metric=="roc_auc") %>% 
-  ggplot(aes(x = log10(penalty), y = mixture, fill = mean)) +
-    geom_tile() +
-    scale_fill_viridis_c(name = "AUC") +
-    labs(
-      x = expression(log[10](lambda)),
-      y = expression(alpha),
-      title = "Heatmap of cross-validated AUC\nfor Elastic-net hyperparameter grid"
-    ) +
-    theme_minimal(base_size = 12)
-
-```
-
-```{r}
 
@@
 
-library(dplyr)
-library(knitr)
-
 # 8) Combine into one data frame and print as a plain table

### 2025-05-03 products/manuscript/supplement/Supplementary-Material.qmd
update

@@
 ---
-title: "Supplement Materials of Impact of IPTp Regimen on Pregnancy Outcomes in a Malaria-Endemic Setting"
+title: "Supplementary Materials of Impact of IPTp Regimen on Pregnancy Outcomes"
 format:
-  docx:
+  pdf:
     toc: false
@@
     highlight-style: github
-resources:
-  - Images/**
-execute:
-  root: products/manuscript
-bibliography: ../../assets/dataanalysis-references.bib
-csl: ../../assets/american-journal-of-epidemiology.csl
+bibliography:
+  - ../../../assets/dataanalysis-references.bib
+csl: ../../../assets/american-journal-of-epidemiology.csl
 ---

### 2025-05-03 products/manuscript/supplement/Supplementary-Material.tex
update

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{\underline{#1}}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{\textbf{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.44,0.26,0.76}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.14,0.16,0.18}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.14,0.16,0.18}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.44,0.26,0.76}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.89,0.38,0.04}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{#1}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Supplementary Materials of Impact of IPTp Regimen on Pregnancy Outcomes},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Supplementary Materials of Impact of IPTp Regimen on Pregnancy
Outcomes}
\author{}
\date{}

\begin{document}
\maketitle

\newpage{}

\section{Overview}\label{overview}

The Supplementary Appendix begins with comprehensive methodological
details, including variable‐by‐variable missingness (Table S1), analysis
of deviance comparing models with and without the malaria×SP interaction
(Table S2), and variance inflation factors for the final interaction
model (Table S3). It also presents machine‐learning tuning results with
the top five elastic-net hyperparameter combinations ranked by mean
cross-validated AUC (Table S4) alongside a heatmap illustrating AUC
across the penalty--mixture grid (Figure S1). Full code excerpts
document our data-cleaning steps, rsample splits, recipe definitions,
and tune\_grid workflows. The appendix then moves on to additional
results: a detailed stratification of outcome measures and
malaria-exposure variables by IPTp arm (Table S5), bar graphs of total
malaria episodes during pregnancy (Figure S2), gravidity (Figure S3),
and parity (Figure S4) distributions by treatment arm, and a bootstrap
calibration curve for the interaction model (Figure S5). Finally, it
compares model discrimination with overlaid ROC curves for logistic
regression, random forest, and XGBoost (Figure S6), presents a
precision-recall curve for the top‐performing machine-learning model
(Figure S7), and reports the test-set AUC for the gravidity-only model
in women under 25 years (Table S6).

\section{Code and file information}\label{code-and-file-information}

Explain here what each code/file is and does, and in which order (if
any) users need to run thing to reproduce everything. Essentially, give
a full set of instructions to re-generate everything.

\newpage{}

\section{Additional Method Details}\label{additional-method-details}

Often, the main manuscript only allows for an overview description of
the methods. Use the supplement to describe all your methods, models and
approaches in a lot of detail. Reference specific parts of your code as
needed.

\newpage{}

\section{Additional results}\label{additional-results}

Show additional results here. Those can be some useful
exploratory/descriptive figures or tables, or results from additional
analyses that didn't make it into the main text.

\emph{Bar Graph of Total Malaria Episodes During Pregnancy by Treatment
Arm}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Load necessary libraries}
\FunctionTok{library}\NormalTok{(dplyr)}
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(here)}

\CommentTok{\# Read in the cleaned data}
\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\FunctionTok{here}\NormalTok{(}\StringTok{"data"}\NormalTok{, }\StringTok{"clean"}\NormalTok{, }\StringTok{"PROMO\_Data\_clean.csv"}\NormalTok{))}

\CommentTok{\# Recode \textquotesingle{}total\_malaria\_episodes\_during\_pregnancy\textquotesingle{} into a categorical variable}
\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}}\NormalTok{ promo\_data\_clean }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{MalariaEpisodesPreg\_cat =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{      total\_malaria\_episodes\_during\_pregnancy }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{"1"}\NormalTok{,}
\NormalTok{      total\_malaria\_episodes\_during\_pregnancy }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{"2–3"}\NormalTok{,}
\NormalTok{      total\_malaria\_episodes\_during\_pregnancy }\SpecialCharTok{\textgreater{}=} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \StringTok{"≥4"}
\NormalTok{    ),}
    \AttributeTok{MalariaEpisodesPreg\_cat =} \FunctionTok{factor}\NormalTok{(MalariaEpisodesPreg\_cat, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"1"}\NormalTok{, }\StringTok{"2–3"}\NormalTok{, }\StringTok{"≥4"}\NormalTok{))}
\NormalTok{  )}

\CommentTok{\# Create the bar graph}
\FunctionTok{ggplot}\NormalTok{(promo\_data\_clean, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ MalariaEpisodesPreg\_cat, }\AttributeTok{fill =}\NormalTok{ study\_arm)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{position =} \StringTok{"dodge"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Total Malaria Episodes During Pregnancy by Treatment Arm"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Total Malaria Episodes During Pregnancy (Categorical)"}\NormalTok{,}
       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
       \AttributeTok{fill =} \StringTok{"Treatment Arm"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\pandocbounded{\includegraphics[keepaspectratio]{Supplementary-Material_files/figure-pdf/unnamed-chunk-4-1.pdf}}

\emph{Bar Graphs for Gravidity and Parity}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Load necessary libraries}

\CommentTok{\# Read in the cleaned data}
\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\FunctionTok{here}\NormalTok{(}\StringTok{"data"}\NormalTok{, }\StringTok{"clean"}\NormalTok{, }\StringTok{"PROMO\_Data\_clean.csv"}\NormalTok{))}

\CommentTok{\# Recode Gravidity into categories: "1", "2–3", "≥4"}
\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}}\NormalTok{ promo\_data\_clean }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{Gravidity\_cat =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{      gravidity }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \StringTok{"1"}\NormalTok{,}
\NormalTok{      gravidity }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{"2–3"}\NormalTok{,}
\NormalTok{      gravidity }\SpecialCharTok{\textgreater{}=} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \StringTok{"≥4"}
\NormalTok{    ),}
    \AttributeTok{Gravidity\_cat =} \FunctionTok{factor}\NormalTok{(Gravidity\_cat, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"1"}\NormalTok{, }\StringTok{"2–3"}\NormalTok{, }\StringTok{"≥4"}\NormalTok{))}
\NormalTok{  )}

\CommentTok{\# Create the bar graph for Gravidity by Treatment Arm}
\FunctionTok{ggplot}\NormalTok{(promo\_data\_clean, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Gravidity\_cat, }\AttributeTok{fill =}\NormalTok{ study\_arm)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{position =} \StringTok{"dodge"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Gravidity Distribution by Treatment Arm"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Gravidity Category (1, 2–3, ≥4)"}\NormalTok{,}
       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
       \AttributeTok{fill =} \StringTok{"Treatment Arm"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\pandocbounded{\includegraphics[keepaspectratio]{Supplementary-Material_files/figure-pdf/unnamed-chunk-5-1.pdf}}

\emph{Parity Distribution by Treatment Arm}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Load necessary libraries}
\FunctionTok{library}\NormalTok{(dplyr)}
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(here)}

\CommentTok{\# Read in the cleaned data}
\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\FunctionTok{here}\NormalTok{(}\StringTok{"data"}\NormalTok{, }\StringTok{"clean"}\NormalTok{, }\StringTok{"PROMO\_Data\_clean.csv"}\NormalTok{))}

\CommentTok{\# Recode Parity into categories: "0", "1–2", "≥3"}
\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}}\NormalTok{ promo\_data\_clean }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{Parity\_cat =} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{      parity }\SpecialCharTok{==} \DecValTok{0} \SpecialCharTok{\textasciitilde{}} \StringTok{"0"}\NormalTok{,}
\NormalTok{      parity }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{"1–2"}\NormalTok{,}
\NormalTok{      parity }\SpecialCharTok{\textgreater{}=} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \StringTok{"≥3"}
\NormalTok{    ),}
    \AttributeTok{Parity\_cat =} \FunctionTok{factor}\NormalTok{(Parity\_cat, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"0"}\NormalTok{, }\StringTok{"1–2"}\NormalTok{, }\StringTok{"≥3"}\NormalTok{))}
\NormalTok{  )}

\CommentTok{\# Create the bar graph for Parity by Treatment Arm}
\FunctionTok{ggplot}\NormalTok{(promo\_data\_clean, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Parity\_cat, }\AttributeTok{fill =}\NormalTok{ study\_arm)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{position =} \StringTok{"dodge"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Parity Distribution by Treatment Arm"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Parity Category (0, 1–2, ≥3)"}\NormalTok{,}
       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
       \AttributeTok{fill =} \StringTok{"Treatment Arm"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\pandocbounded{\includegraphics[keepaspectratio]{Supplementary-Material_files/figure-pdf/unnamed-chunk-6-1.pdf}}

\subsection{Example additional result}\label{example-additional-result}

Table~\ref{tbl-resulttable1} shows an additional table summarizing a
model fit.

\begin{longtable}[]{@{}lrrrr@{}}

\caption{\label{tbl-resulttable1}Another fit table.}

\tabularnewline

\toprule\noalign{}
term & estimate & std.error & statistic & p.value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
(Intercept) & 149.6997661 & 19.7518528 & 7.5790240 & 0.0001285 \\
Weight & 0.2277371 & 0.2708841 & 0.8407177 & 0.4282860 \\

\end{longtable}

Figure~\ref{fig-result2} shows a scatterplot figure produced by one of
the R scripts.

\begin{figure}

\centering{

\includegraphics[width=7in,height=\textheight,keepaspectratio]{../../../results/figures/height-weight.png}

}

\caption{\label{fig-result2}Height and weight.}

\end{figure}%

\section{Evaluating Discrimination (ROC Curve and
AUC)}\label{evaluating-discrimination-roc-curve-and-auc}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# 1) Load required libraries}
\FunctionTok{library}\NormalTok{(dplyr)}
\FunctionTok{library}\NormalTok{(pROC)}
\FunctionTok{library}\NormalTok{(caret)}
\FunctionTok{library}\NormalTok{(here)}

\CommentTok{\# 2) Read cleaned data and subset to women \textless{} 25}
\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\FunctionTok{here}\NormalTok{(}\StringTok{"data/clean/PROMO\_Data\_clean.csv"}\NormalTok{))}
\NormalTok{promo\_data\_young }\OtherTok{\textless{}{-}}\NormalTok{ promo\_data\_clean }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(age\_at\_enrollment\_years }\SpecialCharTok{\textless{}} \DecValTok{25}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \CommentTok{\# Composite adverse outcome: 1 if preterm, stillbirth, or low birth weight}
    \AttributeTok{low\_birth\_weight      =} \FunctionTok{as.integer}\NormalTok{(birth\_weight }\SpecialCharTok{\textless{}} \FloatTok{2.5}\NormalTok{),}
    \AttributeTok{adverse\_birth\_outcome =} \FunctionTok{as.integer}\NormalTok{(}
\NormalTok{      preterm\_births\_count }\SpecialCharTok{\textgreater{}} \DecValTok{0} \SpecialCharTok{|}
\NormalTok{      stillbirth\_bin }\SpecialCharTok{==} \DecValTok{1}        \SpecialCharTok{|}
\NormalTok{      low\_birth\_weight }\SpecialCharTok{==} \DecValTok{1}
\NormalTok{    )}
\NormalTok{  )}

\CommentTok{\# 3) Fit logistic regression model}
\NormalTok{model\_gravidity }\OtherTok{\textless{}{-}} \FunctionTok{glm}\NormalTok{(}
\NormalTok{  adverse\_birth\_outcome }\SpecialCharTok{\textasciitilde{}}\NormalTok{ gravidity }\SpecialCharTok{+}\NormalTok{ total\_malaria\_episodes }\SpecialCharTok{+}
\NormalTok{    study\_arm }\SpecialCharTok{+}\NormalTok{ education\_level,}
  \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link =} \StringTok{"logit"}\NormalTok{),}
  \AttributeTok{data   =}\NormalTok{ promo\_data\_young}
\NormalTok{)}

\CommentTok{\# 4) Generate predicted probabilities}
\NormalTok{promo\_data\_young}\SpecialCharTok{$}\NormalTok{predicted\_prob }\OtherTok{\textless{}{-}} \FunctionTok{predict}\NormalTok{(}
\NormalTok{  model\_gravidity,}
  \AttributeTok{newdata =}\NormalTok{ promo\_data\_young,}
  \AttributeTok{type    =} \StringTok{"response"}
\NormalTok{)}

\CommentTok{\# 5) ROC curve and AUC}
\NormalTok{roc\_obj  }\OtherTok{\textless{}{-}} \FunctionTok{roc}\NormalTok{(}
\NormalTok{  promo\_data\_young}\SpecialCharTok{$}\NormalTok{adverse\_birth\_outcome,}
\NormalTok{  promo\_data\_young}\SpecialCharTok{$}\NormalTok{predicted\_prob}
\NormalTok{)}
\FunctionTok{plot}\NormalTok{(}
\NormalTok{  roc\_obj,}
  \AttributeTok{col   =} \StringTok{"blue"}\NormalTok{,}
  \AttributeTok{lwd   =} \DecValTok{2}\NormalTok{,}
  \AttributeTok{main  =} \StringTok{"ROC Curve: Gravidity Model (Age \textless{} 25)"}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\pandocbounded{\includegraphics[keepaspectratio]{Supplementary-Material_files/figure-pdf/roc-calibration-1.pdf}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{auc\_val }\OtherTok{\textless{}{-}} \FunctionTok{auc}\NormalTok{(roc\_obj)}
\FunctionTok{cat}\NormalTok{(}\StringTok{"AUC:"}\NormalTok{, }\FunctionTok{round}\NormalTok{(auc\_val, }\DecValTok{2}\NormalTok{), }\StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
AUC: 0.58 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# 6) Calibration plot}
\NormalTok{data\_cal }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
  \AttributeTok{observed =} \FunctionTok{factor}\NormalTok{(promo\_data\_young}\SpecialCharTok{$}\NormalTok{adverse\_birth\_outcome, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
  \AttributeTok{predicted =}\NormalTok{ promo\_data\_young}\SpecialCharTok{$}\NormalTok{predicted\_prob}
\NormalTok{)}
\NormalTok{cal\_plot }\OtherTok{\textless{}{-}} \FunctionTok{calibration}\NormalTok{(}
\NormalTok{  observed }\SpecialCharTok{\textasciitilde{}}\NormalTok{ predicted,}
  \AttributeTok{data  =}\NormalTok{ data\_cal,}
  \AttributeTok{class =} \StringTok{"1"}
\NormalTok{)}
\FunctionTok{plot}\NormalTok{(cal\_plot, }\AttributeTok{main =} \StringTok{"Calibration Plot: Gravidity Model (Age \textless{} 25)"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\pandocbounded{\includegraphics[keepaspectratio]{Supplementary-Material_files/figure-pdf/roc-calibration-2.pdf}}

I see that the ROC curve for my model is close to the diagonal, with an
AUC only slightly above 0.5. This tells me that the model doesn't have
strong discriminative ability for predicting adverse outcomes in women
under 25. Additionally, the calibration plot shows that my predicted
probabilities often stray from the ideal diagonal---especially in the
mid-range---indicating that my model's risk estimates don't consistently
match the observed rates. Overall, while gravidity is statistically
significant, my model as a whole isn't very effective at distinguishing
between those who experience adverse outcomes and those who don't, and
its probability estimates need improvement.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidymodels)}
\FunctionTok{library}\NormalTok{(readr)}
\FunctionTok{library}\NormalTok{(here)}

\CommentTok{\# 2. Read \& split data}
\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\FunctionTok{here}\NormalTok{(}\StringTok{"data"}\NormalTok{,}\StringTok{"clean"}\NormalTok{,}\StringTok{"PROMO\_Data\_clean.csv"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{adverse\_birth\_outcome =} \FunctionTok{factor}\NormalTok{(adverse\_birth\_outcome, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Rows: 782 Columns: 83
-- Column specification --------------------------------------------------------
Delimiter: ","
chr (45): study_arm, fathers_consent_for_unborn_child, education_level, alco...
dbl (28): participant_id, household_id, age_at_enrollment_years, gestational...
lgl (10): enrollment_date, withdrawal_date, reason_for_withdrawal, child_wit...

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{2025}\NormalTok{)}
\NormalTok{splits    }\OtherTok{\textless{}{-}} \FunctionTok{initial\_split}\NormalTok{(df, }\AttributeTok{prop =} \FloatTok{0.7}\NormalTok{, }\AttributeTok{strata =}\NormalTok{ adverse\_birth\_outcome)}
\NormalTok{train     }\OtherTok{\textless{}{-}} \FunctionTok{training}\NormalTok{(splits)}
\NormalTok{test      }\OtherTok{\textless{}{-}} \FunctionTok{testing}\NormalTok{(splits)}
\NormalTok{cv\_folds  }\OtherTok{\textless{}{-}} \FunctionTok{vfold\_cv}\NormalTok{(train, }\AttributeTok{v =} \DecValTok{5}\NormalTok{, }\AttributeTok{strata =}\NormalTok{ adverse\_birth\_outcome)}

\CommentTok{\# 3. Preprocessing recipe}
\NormalTok{ml\_rec }\OtherTok{\textless{}{-}} \FunctionTok{recipe}\NormalTok{(adverse\_birth\_outcome }\SpecialCharTok{\textasciitilde{}} 
\NormalTok{                   age\_at\_enrollment\_years }\SpecialCharTok{+}
\NormalTok{                   gravidity }\SpecialCharTok{+}
\NormalTok{                   total\_malaria\_episodes }\SpecialCharTok{+}
\NormalTok{                   education\_level }\SpecialCharTok{+}
\NormalTok{                   study\_arm }\SpecialCharTok{+}
\NormalTok{                   normalized\_malaria\_rate,   }\CommentTok{\# ← replaced here}
                 \AttributeTok{data =}\NormalTok{ train) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{step\_dummy}\NormalTok{(}\FunctionTok{all\_nominal\_predictors}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}  \CommentTok{\# one‐hot encode}
  \FunctionTok{step\_zv}\NormalTok{(}\FunctionTok{all\_predictors}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}             \CommentTok{\# remove zero‐variance}
  \FunctionTok{step\_normalize}\NormalTok{(}\FunctionTok{all\_numeric\_predictors}\NormalTok{())  }\CommentTok{\# center \& scale}

\CommentTok{\# 4. Model specifications \& workflows}

\DocumentationTok{\#\# (a) LASSO‐penalized logistic regression}
\NormalTok{lasso\_mod }\OtherTok{\textless{}{-}} \FunctionTok{logistic\_reg}\NormalTok{(}\AttributeTok{penalty =} \FunctionTok{tune}\NormalTok{(), }\AttributeTok{mixture =} \DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{set\_engine}\NormalTok{(}\StringTok{"glmnet"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{set\_mode}\NormalTok{(}\StringTok{"classification"}\NormalTok{)}

\NormalTok{lasso\_wf  }\OtherTok{\textless{}{-}} \FunctionTok{workflow}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{add\_model}\NormalTok{(lasso\_mod) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{add\_recipe}\NormalTok{(ml\_rec)}

\DocumentationTok{\#\# (b) Random forest}
\NormalTok{rf\_mod }\OtherTok{\textless{}{-}} \FunctionTok{rand\_forest}\NormalTok{(}\AttributeTok{mtry =} \FunctionTok{tune}\NormalTok{(), }\AttributeTok{trees =} \DecValTok{1000}\NormalTok{, }\AttributeTok{min\_n =} \FunctionTok{tune}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{set\_engine}\NormalTok{(}\StringTok{"ranger"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{set\_mode}\NormalTok{(}\StringTok{"classification"}\NormalTok{)}

\NormalTok{rf\_wf  }\OtherTok{\textless{}{-}} \FunctionTok{workflow}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{add\_model}\NormalTok{(rf\_mod) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{add\_recipe}\NormalTok{(ml\_rec)}

\DocumentationTok{\#\# (c) XGBoost}
\NormalTok{xgb\_mod }\OtherTok{\textless{}{-}} \FunctionTok{boost\_tree}\NormalTok{(}\AttributeTok{trees =} \FunctionTok{tune}\NormalTok{(),}
                      \AttributeTok{tree\_depth =} \FunctionTok{tune}\NormalTok{(),}
                      \AttributeTok{learn\_rate =} \FunctionTok{tune}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{set\_engine}\NormalTok{(}\StringTok{"xgboost"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{set\_mode}\NormalTok{(}\StringTok{"classification"}\NormalTok{)}

\NormalTok{xgb\_wf }\OtherTok{\textless{}{-}} \FunctionTok{workflow}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{add\_model}\NormalTok{(xgb\_mod) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{add\_recipe}\NormalTok{(ml\_rec)}

\CommentTok{\# 5. Tuning grids \& 5‐fold CV}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{2025}\NormalTok{)}

\NormalTok{lasso\_grid }\OtherTok{\textless{}{-}} \FunctionTok{grid\_regular}\NormalTok{(}\FunctionTok{penalty}\NormalTok{(), }\AttributeTok{levels =} \DecValTok{20}\NormalTok{)}

\NormalTok{rf\_params }\OtherTok{\textless{}{-}} \FunctionTok{parameters}\NormalTok{(rf\_mod) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{update}\NormalTok{(}\AttributeTok{mtry =} \FunctionTok{mtry}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{6}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Warning: `parameters.model_spec()` was deprecated in tune 0.1.6.9003.
i Please use `hardhat::extract_parameter_set_dials()` instead.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rf\_grid   }\OtherTok{\textless{}{-}} \FunctionTok{grid\_latin\_hypercube}\NormalTok{(rf\_params, }\AttributeTok{size =} \DecValTok{20}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Warning: `grid_latin_hypercube()` was deprecated in dials 1.3.0.
i Please use `grid_space_filling()` instead.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{xgb\_params }\OtherTok{\textless{}{-}} \FunctionTok{parameters}\NormalTok{(xgb\_mod)}
\NormalTok{xgb\_grid   }\OtherTok{\textless{}{-}} \FunctionTok{grid\_latin\_hypercube}\NormalTok{(xgb\_params, }\AttributeTok{size =} \DecValTok{30}\NormalTok{)}

\NormalTok{ml\_metrics }\OtherTok{\textless{}{-}} \FunctionTok{metric\_set}\NormalTok{(roc\_auc)}

\NormalTok{lasso\_res }\OtherTok{\textless{}{-}} \FunctionTok{tune\_grid}\NormalTok{(lasso\_wf, }\AttributeTok{resamples =}\NormalTok{ cv\_folds,}
                       \AttributeTok{grid =}\NormalTok{ lasso\_grid, }\AttributeTok{metrics =}\NormalTok{ ml\_metrics)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Warning: package 'glmnet' was built under R version 4.4.3
\end{verbatim}

\begin{verbatim}
Warning: package 'Matrix' was built under R version 4.4.3
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rf\_res    }\OtherTok{\textless{}{-}} \FunctionTok{tune\_grid}\NormalTok{(rf\_wf,    }\AttributeTok{resamples =}\NormalTok{ cv\_folds,}
                       \AttributeTok{grid =}\NormalTok{ rf\_grid,    }\AttributeTok{metrics =}\NormalTok{ ml\_metrics)}

\NormalTok{xgb\_res   }\OtherTok{\textless{}{-}} \FunctionTok{tune\_grid}\NormalTok{(xgb\_wf,   }\AttributeTok{resamples =}\NormalTok{ cv\_folds,}
                       \AttributeTok{grid =}\NormalTok{ xgb\_grid,   }\AttributeTok{metrics =}\NormalTok{ ml\_metrics)}

\FunctionTok{library}\NormalTok{(dplyr)}
\FunctionTok{library}\NormalTok{(knitr)}

\CommentTok{\# 6) Pull the metrics into plain tibbles}
\NormalTok{lasso\_metrics }\OtherTok{\textless{}{-}} \FunctionTok{collect\_metrics}\NormalTok{(lasso\_res)}
\NormalTok{rf\_metrics    }\OtherTok{\textless{}{-}} \FunctionTok{collect\_metrics}\NormalTok{(rf\_res)}
\NormalTok{xgb\_metrics   }\OtherTok{\textless{}{-}} \FunctionTok{collect\_metrics}\NormalTok{(xgb\_res)}

\CommentTok{\# 7) Summarize AUC for each model}
\NormalTok{lasso\_auc }\OtherTok{\textless{}{-}}\NormalTok{ lasso\_metrics }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(.metric }\SpecialCharTok{==} \StringTok{"roc\_auc"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{summarise}\NormalTok{(}
    \AttributeTok{Mean =} \FunctionTok{mean}\NormalTok{(mean),}
    \AttributeTok{SE   =} \FunctionTok{mean}\NormalTok{(std\_err)}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Model =} \StringTok{"LASSO"}\NormalTok{)}

\NormalTok{rf\_auc }\OtherTok{\textless{}{-}}\NormalTok{ rf\_metrics }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(.metric }\SpecialCharTok{==} \StringTok{"roc\_auc"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{summarise}\NormalTok{(}
    \AttributeTok{Mean =} \FunctionTok{mean}\NormalTok{(mean),}
    \AttributeTok{SE   =} \FunctionTok{mean}\NormalTok{(std\_err)}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Model =} \StringTok{"Random Forest"}\NormalTok{)}

\NormalTok{xgb\_auc }\OtherTok{\textless{}{-}}\NormalTok{ xgb\_metrics }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(.metric }\SpecialCharTok{==} \StringTok{"roc\_auc"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{summarise}\NormalTok{(}
    \AttributeTok{Mean =} \FunctionTok{mean}\NormalTok{(mean),}
    \AttributeTok{SE   =} \FunctionTok{mean}\NormalTok{(std\_err)}
\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Model =} \StringTok{"XGBoost"}\NormalTok{)}

\CommentTok{\# 8) Combine into one data frame and print as a plain table}
\NormalTok{table6 }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(lasso\_auc, rf\_auc, xgb\_auc) }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(Model, Mean, SE)}

\FunctionTok{print}\NormalTok{(table6)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 3 x 3
  Model          Mean     SE
  <chr>         <dbl>  <dbl>
1 LASSO         0.554 0.0122
2 Random Forest 0.524 0.0134
3 XGBoost       0.497 0.0145
\end{verbatim}

During 5‑fold cross‑validation, the elastic‐net model achieved the
highest mean AUC (0.55±0.01), followed by random forest (0.52±0.01) and
XGBoost (0.50±0.01).

\newpage{}

\section{Discussion}\label{discussion}

Any additional discussion regarding the supplementary material/findings.

These papers (\textbf{mckay2020?},\textbf{mckay2020a?}) are good
examples of papers published using a fully reproducible setup similar to
the one shown in this template.

\newpage{}

\section{References}\label{references}

\end{document}

### 2025-05-03 products/manuscript/supplement/Supplementary-Material.qmd
update

@@
 ```{r}
-library(tidymodels)
-library(readr)
-library(here)
-
-# 2. Read & split data
-df <- read_csv(here("data","clean","PROMO_Data_clean.csv")) %>%
-  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0,1)))
-
-set.seed(2025)
-splits    <- initial_split(df, prop = 0.7, strata = adverse_birth_outcome)
-train     <- training(splits)
-test      <- testing(splits)
-cv_folds  <- vfold_cv(train, v = 5, strata = adverse_birth_outcome)
-
-# 3. Preprocessing recipe
-ml_rec <- recipe(adverse_birth_outcome ~ 
-                   age_at_enrollment_years +
-                   gravidity +
-                   total_malaria_episodes +
-                   education_level +
-                   study_arm +
-                   normalized_malaria_rate,   # ← replaced here
-                 data = train) %>%
-  step_dummy(all_nominal_predictors()) %>%  # one‐hot encode
-  step_zv(all_predictors()) %>%             # remove zero‐variance
-  step_normalize(all_numeric_predictors())  # center & scale
-
-# 4. Model specifications & workflows
-
-## (a) LASSO‐penalized logistic regression
-lasso_mod <- logistic_reg(penalty = tune(), mixture = 1) %>%
-  set_engine("glmnet") %>%
-  set_mode("classification")
-
-lasso_wf  <- workflow() %>% 
-  add_model(lasso_mod) %>% 
-  add_recipe(ml_rec)
-
-## (b) Random forest
-rf_mod <- rand_forest(mtry = tune(), trees = 1000, min_n = tune()) %>%
-  set_engine("ranger") %>%
-  set_mode("classification")
-
-rf_wf  <- workflow() %>% 
-  add_model(rf_mod) %>% 
-  add_recipe(ml_rec)
-
-## (c) XGBoost
-xgb_mod <- boost_tree(trees = tune(),
-                      tree_depth = tune(),
-                      learn_rate = tune()) %>%
-  set_engine("xgboost") %>%
-  set_mode("classification")
-
-xgb_wf <- workflow() %>% 
-  add_model(xgb_mod) %>% 
-  add_recipe(ml_rec)
-
-# 5. Tuning grids & 5‐fold CV
-set.seed(2025)
-
-lasso_grid <- grid_regular(penalty(), levels = 20)
-
-rf_params <- parameters(rf_mod) %>%
-  update(mtry = mtry(c(1, 6)))
-rf_grid   <- grid_latin_hypercube(rf_params, size = 20)
-
-xgb_params <- parameters(xgb_mod)
-xgb_grid   <- grid_latin_hypercube(xgb_params, size = 30)
-
-ml_metrics <- metric_set(roc_auc)
-
-lasso_res <- tune_grid(lasso_wf, resamples = cv_folds,
-                       grid = lasso_grid, metrics = ml_metrics)
-
-rf_res    <- tune_grid(rf_wf,    resamples = cv_folds,
-                       grid = rf_grid,    metrics = ml_metrics)
-
-xgb_res   <- tune_grid(xgb_wf,   resamples = cv_folds,
-                       grid = xgb_grid,   metrics = ml_metrics)
-
 library(dplyr)
@@
 
-# 6) Pull the metrics into plain tibbles
+# 6) Collect tuning metrics
 lasso_metrics <- collect_metrics(lasso_res)
@@
   filter(.metric == "roc_auc") %>%
-  dplyr::summarise(
+  summarise(
     Mean = mean(mean),
@@
 
-# 8) Combine into one data frame and print as a plain table
-table6 <- bind_rows(lasso_auc, rf_auc, xgb_auc) %>%
-  dplyr::select(Model, Mean, SE)
+# 8) Combine into one data frame
+table6 <- bind_rows(lasso_auc, rf_auc, xgb_auc)
 
-print(table6)
+# 9) Render with kable for Word, explicitly namespacing select
+kable(
+  table6 %>% dplyr::select(Model, Mean, SE),
+  format  = "pandoc",
+  digits  = 3,
+  caption = "Table 6: Cross‑validated AUC (mean ± SE) for Each Model"
+)
 

### 2025-05-03 products/manuscript/supplement/Supplementary-Material.qmd
Supp

@@
 
-```{r}
-library(dplyr)
-library(knitr)
-
-# 6) Collect tuning metrics
-lasso_metrics <- collect_metrics(lasso_res)
-rf_metrics    <- collect_metrics(rf_res)
-xgb_metrics   <- collect_metrics(xgb_res)
-
-# 7) Summarize AUC for each model
-lasso_auc <- lasso_metrics %>%
-  filter(.metric == "roc_auc") %>%
-  summarise(
-    Mean = mean(mean),
-    SE   = mean(std_err)
-  ) %>%
-  mutate(Model = "LASSO")
-
-rf_auc <- rf_metrics %>%
-  filter(.metric == "roc_auc") %>%
-  summarise(
-    Mean = mean(mean),
-    SE   = mean(std_err)
-  ) %>%
-  mutate(Model = "Random Forest")
-
-xgb_auc <- xgb_metrics %>%
-  filter(.metric == "roc_auc") %>%
-  summarise(
-    Mean = mean(mean),
-    SE   = mean(std_err)
-  ) %>%
-  mutate(Model = "XGBoost")
-
-# 8) Combine into one data frame
-table6 <- bind_rows(lasso_auc, rf_auc, xgb_auc)
-
-# 9) Render with kable for Word, explicitly namespacing select
-kable(
-  table6 %>% dplyr::select(Model, Mean, SE),
-  format  = "pandoc",
-  digits  = 3,
-  caption = "Table 6: Cross‑validated AUC (mean ± SE) for Each Model"
-)
-
-```
-
 During 5‑fold cross‑validation, the elastic‐net model achieved the highest mean AUC (0.55±0.01), followed by random forest (0.52±0.01) and XGBoost (0.50±0.01).
@@
 
-Any additional discussion regarding the supplementary material/findings.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
 {{< pagebreak >}}

### 2025-05-03 products/manuscript/supplement/Supplementary-Material.qmd
update supp

@@
 library(rms)
+library(tinytex)
 ```

### 2025-05-03 products/manuscript/supplement/Supplementary-Material.tex
update supp

@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(tidymodels)}
-\FunctionTok{library}\NormalTok{(readr)}
-\FunctionTok{library}\NormalTok{(here)}
-
-\CommentTok{\# 2. Read \& split data}
-\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\FunctionTok{here}\NormalTok{(}\StringTok{"data"}\NormalTok{,}\StringTok{"clean"}\NormalTok{,}\StringTok{"PROMO\_Data\_clean.csv"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{adverse\_birth\_outcome =} \FunctionTok{factor}\NormalTok{(adverse\_birth\_outcome, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-Rows: 782 Columns: 83
--- Column specification --------------------------------------------------------
-Delimiter: ","
-chr (45): study_arm, fathers_consent_for_unborn_child, education_level, alco...
-dbl (28): participant_id, household_id, age_at_enrollment_years, gestational...
-lgl (10): enrollment_date, withdrawal_date, reason_for_withdrawal, child_wit...
-
-i Use `spec()` to retrieve the full column specification for this data.
-i Specify the column types or set `show_col_types = FALSE` to quiet this message.
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{set.seed}\NormalTok{(}\DecValTok{2025}\NormalTok{)}
-\NormalTok{splits    }\OtherTok{\textless{}{-}} \FunctionTok{initial\_split}\NormalTok{(df, }\AttributeTok{prop =} \FloatTok{0.7}\NormalTok{, }\AttributeTok{strata =}\NormalTok{ adverse\_birth\_outcome)}
-\NormalTok{train     }\OtherTok{\textless{}{-}} \FunctionTok{training}\NormalTok{(splits)}
-\NormalTok{test      }\OtherTok{\textless{}{-}} \FunctionTok{testing}\NormalTok{(splits)}
-\NormalTok{cv\_folds  }\OtherTok{\textless{}{-}} \FunctionTok{vfold\_cv}\NormalTok{(train, }\AttributeTok{v =} \DecValTok{5}\NormalTok{, }\AttributeTok{strata =}\NormalTok{ adverse\_birth\_outcome)}
-
-\CommentTok{\# 3. Preprocessing recipe}
-\NormalTok{ml\_rec }\OtherTok{\textless{}{-}} \FunctionTok{recipe}\NormalTok{(adverse\_birth\_outcome }\SpecialCharTok{\textasciitilde{}} 
-\NormalTok{                   age\_at\_enrollment\_years }\SpecialCharTok{+}
-\NormalTok{                   gravidity }\SpecialCharTok{+}
-\NormalTok{                   total\_malaria\_episodes }\SpecialCharTok{+}
-\NormalTok{                   education\_level }\SpecialCharTok{+}
-\NormalTok{                   study\_arm }\SpecialCharTok{+}
-\NormalTok{                   normalized\_malaria\_rate,   }\CommentTok{\# ← replaced here}
-                 \AttributeTok{data =}\NormalTok{ train) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{step\_dummy}\NormalTok{(}\FunctionTok{all\_nominal\_predictors}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}  \CommentTok{\# one‐hot encode}
-  \FunctionTok{step\_zv}\NormalTok{(}\FunctionTok{all\_predictors}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}             \CommentTok{\# remove zero‐variance}
-  \FunctionTok{step\_normalize}\NormalTok{(}\FunctionTok{all\_numeric\_predictors}\NormalTok{())  }\CommentTok{\# center \& scale}
-
-\CommentTok{\# 4. Model specifications \& workflows}
-
-\DocumentationTok{\#\# (a) LASSO‐penalized logistic regression}
-\NormalTok{lasso\_mod }\OtherTok{\textless{}{-}} \FunctionTok{logistic\_reg}\NormalTok{(}\AttributeTok{penalty =} \FunctionTok{tune}\NormalTok{(), }\AttributeTok{mixture =} \DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{set\_engine}\NormalTok{(}\StringTok{"glmnet"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{set\_mode}\NormalTok{(}\StringTok{"classification"}\NormalTok{)}
-
-\NormalTok{lasso\_wf  }\OtherTok{\textless{}{-}} \FunctionTok{workflow}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{add\_model}\NormalTok{(lasso\_mod) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{add\_recipe}\NormalTok{(ml\_rec)}
-
-\DocumentationTok{\#\# (b) Random forest}
-\NormalTok{rf\_mod }\OtherTok{\textless{}{-}} \FunctionTok{rand\_forest}\NormalTok{(}\AttributeTok{mtry =} \FunctionTok{tune}\NormalTok{(), }\AttributeTok{trees =} \DecValTok{1000}\NormalTok{, }\AttributeTok{min\_n =} \FunctionTok{tune}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{set\_engine}\NormalTok{(}\StringTok{"ranger"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{set\_mode}\NormalTok{(}\StringTok{"classification"}\NormalTok{)}
-
-\NormalTok{rf\_wf  }\OtherTok{\textless{}{-}} \FunctionTok{workflow}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{add\_model}\NormalTok{(rf\_mod) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{add\_recipe}\NormalTok{(ml\_rec)}
-
-\DocumentationTok{\#\# (c) XGBoost}
-\NormalTok{xgb\_mod }\OtherTok{\textless{}{-}} \FunctionTok{boost\_tree}\NormalTok{(}\AttributeTok{trees =} \FunctionTok{tune}\NormalTok{(),}
-                      \AttributeTok{tree\_depth =} \FunctionTok{tune}\NormalTok{(),}
-                      \AttributeTok{learn\_rate =} \FunctionTok{tune}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{set\_engine}\NormalTok{(}\StringTok{"xgboost"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{set\_mode}\NormalTok{(}\StringTok{"classification"}\NormalTok{)}
-
-\NormalTok{xgb\_wf }\OtherTok{\textless{}{-}} \FunctionTok{workflow}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{add\_model}\NormalTok{(xgb\_mod) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{add\_recipe}\NormalTok{(ml\_rec)}
-
-\CommentTok{\# 5. Tuning grids \& 5‐fold CV}
-\FunctionTok{set.seed}\NormalTok{(}\DecValTok{2025}\NormalTok{)}
-
-\NormalTok{lasso\_grid }\OtherTok{\textless{}{-}} \FunctionTok{grid\_regular}\NormalTok{(}\FunctionTok{penalty}\NormalTok{(), }\AttributeTok{levels =} \DecValTok{20}\NormalTok{)}
-
-\NormalTok{rf\_params }\OtherTok{\textless{}{-}} \FunctionTok{parameters}\NormalTok{(rf\_mod) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{update}\NormalTok{(}\AttributeTok{mtry =} \FunctionTok{mtry}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{6}\NormalTok{)))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-Warning: `parameters.model_spec()` was deprecated in tune 0.1.6.9003.
-i Please use `hardhat::extract_parameter_set_dials()` instead.
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rf\_grid   }\OtherTok{\textless{}{-}} \FunctionTok{grid\_latin\_hypercube}\NormalTok{(rf\_params, }\AttributeTok{size =} \DecValTok{20}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-Warning: `grid_latin_hypercube()` was deprecated in dials 1.3.0.
-i Please use `grid_space_filling()` instead.
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{xgb\_params }\OtherTok{\textless{}{-}} \FunctionTok{parameters}\NormalTok{(xgb\_mod)}
-\NormalTok{xgb\_grid   }\OtherTok{\textless{}{-}} \FunctionTok{grid\_latin\_hypercube}\NormalTok{(xgb\_params, }\AttributeTok{size =} \DecValTok{30}\NormalTok{)}
-
-\NormalTok{ml\_metrics }\OtherTok{\textless{}{-}} \FunctionTok{metric\_set}\NormalTok{(roc\_auc)}
-
-\NormalTok{lasso\_res }\OtherTok{\textless{}{-}} \FunctionTok{tune\_grid}\NormalTok{(lasso\_wf, }\AttributeTok{resamples =}\NormalTok{ cv\_folds,}
-                       \AttributeTok{grid =}\NormalTok{ lasso\_grid, }\AttributeTok{metrics =}\NormalTok{ ml\_metrics)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-Warning: package 'glmnet' was built under R version 4.4.3
-\end{verbatim}
-
-\begin{verbatim}
-Warning: package 'Matrix' was built under R version 4.4.3
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rf\_res    }\OtherTok{\textless{}{-}} \FunctionTok{tune\_grid}\NormalTok{(rf\_wf,    }\AttributeTok{resamples =}\NormalTok{ cv\_folds,}
-                       \AttributeTok{grid =}\NormalTok{ rf\_grid,    }\AttributeTok{metrics =}\NormalTok{ ml\_metrics)}
-
-\NormalTok{xgb\_res   }\OtherTok{\textless{}{-}} \FunctionTok{tune\_grid}\NormalTok{(xgb\_wf,   }\AttributeTok{resamples =}\NormalTok{ cv\_folds,}
-                       \AttributeTok{grid =}\NormalTok{ xgb\_grid,   }\AttributeTok{metrics =}\NormalTok{ ml\_metrics)}
-
-\FunctionTok{library}\NormalTok{(dplyr)}
-\FunctionTok{library}\NormalTok{(knitr)}
-
-\CommentTok{\# 6) Pull the metrics into plain tibbles}
-\NormalTok{lasso\_metrics }\OtherTok{\textless{}{-}} \FunctionTok{collect\_metrics}\NormalTok{(lasso\_res)}
-\NormalTok{rf\_metrics    }\OtherTok{\textless{}{-}} \FunctionTok{collect\_metrics}\NormalTok{(rf\_res)}
-\NormalTok{xgb\_metrics   }\OtherTok{\textless{}{-}} \FunctionTok{collect\_metrics}\NormalTok{(xgb\_res)}
-
-\CommentTok{\# 7) Summarize AUC for each model}
-\NormalTok{lasso\_auc }\OtherTok{\textless{}{-}}\NormalTok{ lasso\_metrics }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(.metric }\SpecialCharTok{==} \StringTok{"roc\_auc"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{summarise}\NormalTok{(}
-    \AttributeTok{Mean =} \FunctionTok{mean}\NormalTok{(mean),}
-    \AttributeTok{SE   =} \FunctionTok{mean}\NormalTok{(std\_err)}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Model =} \StringTok{"LASSO"}\NormalTok{)}
-
-\NormalTok{rf\_auc }\OtherTok{\textless{}{-}}\NormalTok{ rf\_metrics }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(.metric }\SpecialCharTok{==} \StringTok{"roc\_auc"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{summarise}\NormalTok{(}
-    \AttributeTok{Mean =} \FunctionTok{mean}\NormalTok{(mean),}
-    \AttributeTok{SE   =} \FunctionTok{mean}\NormalTok{(std\_err)}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Model =} \StringTok{"Random Forest"}\NormalTok{)}
-
-\NormalTok{xgb\_auc }\OtherTok{\textless{}{-}}\NormalTok{ xgb\_metrics }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(.metric }\SpecialCharTok{==} \StringTok{"roc\_auc"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{summarise}\NormalTok{(}
-    \AttributeTok{Mean =} \FunctionTok{mean}\NormalTok{(mean),}
-    \AttributeTok{SE   =} \FunctionTok{mean}\NormalTok{(std\_err)}
-\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Model =} \StringTok{"XGBoost"}\NormalTok{)}
-
-\CommentTok{\# 8) Combine into one data frame and print as a plain table}
-\NormalTok{table6 }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(lasso\_auc, rf\_auc, xgb\_auc) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(Model, Mean, SE)}
-
-\FunctionTok{print}\NormalTok{(table6)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# A tibble: 3 x 3
-  Model          Mean     SE
-  <chr>         <dbl>  <dbl>
-1 LASSO         0.554 0.0122
-2 Random Forest 0.524 0.0134
-3 XGBoost       0.497 0.0145
-\end{verbatim}
-
 During 5‑fold cross‑validation, the elastic‐net model achieved the
@@
 
-Any additional discussion regarding the supplementary material/findings.
-
-These papers (\textbf{mckay2020?},\textbf{mckay2020a?}) are good
-examples of papers published using a fully reproducible setup similar to
-the one shown in this template.
-
 \newpage{}

### 2025-05-03 products/manuscript/supplement/Supplementary-Material.qmd
Supp

--- 2025-05-03 products/manuscript/supplement/Supplementary-Material.qmd
@@
 title: "Supplementary Materials of Impact of IPTp Regimen on Pregnancy Outcomes"
+author: "Asmith Joseph"
 format:
@@
 library(rms)
+
 ```
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 # Load necessary libraries
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 #| label: tbl-resulttable1
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 #| label: fig-result2
@@
 
-```{r roc-calibration, echo=TRUE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 # 1) Load required libraries
@@
 plot(cal_plot, main = "Calibration Plot: Gravidity Model (Age < 25)")
-
 ```

### 2025-05-03 code/analysis-code/Statistical_Analysis.qmd
Final Update

@@
 print(result_table)
-
 ```
@@
 ```{r calibration-plot, fig.width=6, fig.height=6, warning=FALSE}
-# assume `cal` is your calibrate() object from rms::calibrate(lrm_fit, method="boot", B=200)
-
 # 1) Plot without default legend
@@
 ```{r}
-# only these variables are needed for the logistic model
 dd <- datadist(

### 2025-05-03 products/manuscript/supplement/Supplementary-Material.qmd
Final Update

@@
 ---
-
-{{< pagebreak >}}
-
-# Overview
-The Supplementary Appendix begins with comprehensive methodological details, including variable‐by‐variable missingness (Table S1), analysis of deviance comparing models with and without the malaria×SP interaction (Table S2), and variance inflation factors for the final interaction model (Table S3).
-It also presents machine‐learning tuning results with the top five elastic-net hyperparameter combinations ranked by mean cross-validated AUC (Table S4) alongside a heatmap illustrating AUC across the penalty–mixture grid (Figure S1).
-Full code excerpts document our data-cleaning steps, rsample splits, recipe definitions, and tune_grid workflows.
-The appendix then moves on to additional results: a detailed stratification of outcome measures and malaria-exposure variables by IPTp arm (Table S5), bar graphs of total malaria episodes during pregnancy (Figure S2), gravidity (Figure S3), and parity (Figure S4) distributions by treatment arm, and a bootstrap calibration curve for the interaction model (Figure S5).
-Finally, it compares model discrimination with overlaid ROC curves for logistic regression, random forest, and XGBoost (Figure S6), presents a precision-recall curve for the top‐performing machine-learning model (Figure S7), and reports the test-set AUC for the gravidity-only model in women under 25 years (Table S6).
-
-# Code and file information
-
-Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
-Essentially, give a full set of instructions to re-generate everything.
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-library(here)
-here::i_am("products/manuscript/Manuscript.qmd")
-
-# ← THIS is the one that actually sets the working dir for *all* chunks:
-library(knitr)
-knitr::opts_chunk$set(root.dir = here::here("products/manuscript"))
-
-```
 
@@
 
-# Additional Method Details
+# Overview
+The Supplementary Appendix begins with comprehensive methodological details, including variable‐by‐variable missingness (Table S1), analysis of deviance comparing models with and without the malaria×SP interaction (Table S2), and variance inflation factors for the final interaction model (Table S3).
+It also presents machine‐learning tuning results with the top five elastic-net hyperparameter combinations ranked by mean cross-validated AUC (Table S4) alongside a heatmap illustrating AUC across the penalty–mixture grid (Figure S1).
+Full code excerpts document our data-cleaning steps, rsample splits, recipe definitions, and tune_grid workflows.
+The appendix then moves on to additional results: a detailed stratification of outcome measures and malaria-exposure variables by IPTp arm (Table S5), bar graphs of total malaria episodes during pregnancy (Figure S2), gravidity (Figure S3), and parity (Figure S4) distributions by treatment arm, and a bootstrap calibration curve for the interaction model (Figure S5).
+Finally, it compares model discrimination with overlaid ROC curves for logistic regression, random forest, and XGBoost (Figure S6), presents a precision-recall curve for the top‐performing machine-learning model (Figure S7), and reports the test-set AUC for the gravidity-only model in women under 25 years (Table S6).
 
-Often, the main manuscript only allows for an overview description of the methods.
-Use the supplement to describe all your methods, models and approaches in a lot of detail.
-Reference specific parts of your code as needed.
+&nbsp;
+&nbsp;
+
+# Code and file information
+
+The Supplementary Appendix is built with a clear, step‑wise workflow: raw PROMO trial data are first cleaned and documented, then transformed into the analysis dataset.
+Subsequent scripts generate missingness tables, interaction model diagnostics, machine‑learning tuning results, and additional stratified analyses.
+Each RMarkdown file lives in the code/ subfolders (processing, EDA, analysis), reads from data/, and writes results into results/.
+Finally, the Quarto source in products/manuscript/supplement renders the assembled tables and figures into Supplementary‑Material.docx.
+Running the scripts in their numeric order fully reproduces every table and figure in this appendix.
+
+## 📁 Directory Structure Overview
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# From products/manuscript/supplement → go up to products/manuscript, then into Images/
+knitr::include_graphics("../Images/Project-root.png")
+```
 
@@
 
-Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.
+# # Exploratory Analysis: Malaria Episodes, Gravidity/Parity, and Treatment Arm
 
-*Bar Graph of Total Malaria Episodes During Pregnancy by Treatment Arm*
+The four supplementary EDA plots concisely overview key baseline characteristics by treatment arm and the height–weight relationship at enrollment.
+The first panel shows that the majority of women experienced one malaria episode during pregnancy, with progressively fewer in the 2–3 and ≥4 categories and that counts are consistently higher in the DP arm than the SP arm across all episode bins.
+The second panel depicts gravidity stratified into 1, 2–3, and ≥4 pregnancies, revealing a similar pattern in both arms but a noticeably smaller proportion of primigravidae in the SP group.
+The third panel presents parity categories (0, 1–2, ≥3), again demonstrating broadly comparable distributions between arms, with a slight excess of women with ≥3 previous births in the DP arm.
+Finally, the scatterplot of maternal height and weight, with a LOESS smoothing curve and 95% confidence band, confirms a modest positive association between height and weight at enrollment, supporting the use of both measures in downstream modeling.
 
+*Figure S1.Total Malaria Episodes During Pregnancy by Treatment Arm*
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
@@
 
-*Bar Graphs for Gravidity and Parity*
+&nbsp;
+&nbsp;
+
+*Figure S2: Gravidity Distribution by Treatment Arm*
 
@@
 
-*Parity Distribution by Treatment Arm*
+&nbsp;
+&nbsp;
+
+*Figure S3: Parity Distribution by Treatment Arm*
 
@@
 
-## Example additional result
+&nbsp;
+&nbsp;
 
-@tbl-resulttable1 shows an additional table summarizing a model fit.
+*Figure S4:Scatterplot of maternal weight by height with fitted LOESS smoothing curve (95% CI) — exploratory assessment of the height–weight relationship at enrollment*
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+#| label: fig-result2
+#| fig-cap: "Height and weight."
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","height-weight.png"))
+```
+
+&nbsp;
+&nbsp;
+
+## Additional result
+
+*Table 1: Regression Coefficients for the Association Between Weight and Outcome*
 
@@
 
-@fig-result2 shows a scatterplot figure produced by one of the R scripts.
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-#| label: fig-result2
-#| fig-cap: "Height and weight."
-#| echo: FALSE
-knitr::include_graphics(here("results","figures","height-weight.png"))
-```
+&nbsp;
+&nbsp;
 
 #  Evaluating Discrimination (ROC Curve and AUC)
+*Additional Model Evaluation (Supplement)*
+ Panel A shows the ROC curve for the gravidity‐only logistic model in women under 25. The curve lies modestly above the diagonal, reflecting fair discrimination (AUC ≈ 0.55).
+ Panel B is the corresponding calibration plot (decile‐based), which compares observed event rates within probability bins to the ideal 45° line; slight deviations indicate some miscalibration in the highest and lowest risk groups.
 
+ Panel C presents the bootstrap bias‐corrected calibration curve (200 resamples).
+The “apparent” (dotted) and “bias‐corrected” (solid) lines both track closely to the ideal (dashed) reference, confirming good overall calibration (mean absolute error ≈ 0.016).
+ Panel D displays the heatmap of cross‐validated AUC across the elastic‐net penalty–mixture grid, illustrating that intermediate penalty values (log₁₀λ around –2) and mixture α near 0.8 yielded the best discrimination.
+
+ Finally, Table S7 reports the mean (±SE) cross‐validated ROC AUC for the three machine‐learning algorithms: LASSO (0.55±0.012), random forest (0.52±0.013), and XGBoost (0.497±0.014).
+This confirms that regularized regression slightly outperformed the tree‐based methods under our tuning scheme.
+
+*Supplementary Regression Coefficients (Table S8)*
+ The simple linear model of weight on height yields an intercept of 149.7g (SE 19.8, p<0.001) and a non‐significant slope of 0.23g/cm (SE 0.27, p=0.43), indicating that while taller women tended to weigh more on average, this association was weak and did not reach statistical significance in our sample.
+
+ &nbsp;
+&nbsp;
+
+*Figure S5: Calibration Plot: Gravidity Model (Age < 25)*
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
@@
 
-I see that the ROC curve for my model is close to the diagonal, with an AUC only slightly above 0.5.
-This tells me that the model doesn't have strong discriminative ability for predicting adverse outcomes in women under 25.
-Additionally, the calibration plot shows that my predicted probabilities often stray from the ideal diagonal—especially in the mid-range—indicating that my model’s risk estimates don't consistently match the observed rates.
-Overall, while gravidity is statistically significant, my model as a whole isn't very effective at distinguishing between those who experience adverse outcomes and those who don't, and its probability estimates need improvement.
+&nbsp;
+&nbsp;
+
+**Figure S6. Bootstrap‑calibrated calibration curve for the malaria interaction model**
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+knitr::include_graphics("../Images/rms_model.png")
+
+```
+&nbsp;
+&nbsp;
+
+*Figure S7: Calibration curve for malaria interaction model in women aged <25 years*
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+knitr::include_graphics("../Images/CalibrationPlot.png")
+
+```
+&nbsp;
+&nbsp;
+
+*Figure S8:Heatmap of five‑fold cross‑validated ROC AUC across the elastic‑net penalty–mixture grid*
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("../Images/Heatmap.CV AUC.png")
+
+```
+
+&nbsp;
+&nbsp;
+
+*Table S2:Mean cross‑validated ROC AUC (±SE) for each machine‑learning mode*
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+knitr::include_graphics("../Images/MeanCV ROC.AUC .png")
+
+```
 
@@
 # Discussion
-
-{{< pagebreak >}}
-
-# References
+The supplementary materials deepen and validate our primary findings by providing transparent documentation of all data‐processing steps, extensive methodological details, and additional model diagnostics that could not be accommodated in the main text.
+Detailed tables of missingness, deviance tests, and variance inflation factors reinforce the robustness of our interaction model.
+At the same time, the bootstrap and decile‐based calibration curves confirm its reliability across risk strata.
+The elastic‐net heatmap and cross‐validated AUC summaries for LASSO, random forest, and XGBoost highlight that, under consistent tuning, regularized regression offers the strongest discrimination in our dataset.
+Exploratory bar charts of malaria episodes, gravidity, and parity by treatment arm contextualize cohort characteristics, and the gravidity‐only subgroup analyses (ROC and calibration plots) elucidate how prior pregnancy history predicts adverse outcomes among younger women.
+Together, these supplementary analyses enhance confidence in our conclusions, demonstrate reproducibility, and offer a resource for readers seeking to replicate or extend this work.

### 2025-05-03 products/manuscript/supplement/Supplementary-Material.tex
Final Update

@@
 \makeatother
-
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\newenvironment{Shaded}{}{}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{\textbf{#1}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{\underline{#1}}}
-\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{\textbf{#1}}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.44,0.26,0.76}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
-\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.14,0.16,0.18}{#1}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.14,0.16,0.18}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.44,0.26,0.76}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
-\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.89,0.38,0.04}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{#1}}
 
@@
   pdftitle={Supplementary Materials of Impact of IPTp Regimen on Pregnancy Outcomes},
+  pdfauthor={Asmith Joseph},
   colorlinks=true,
@@
 Outcomes}
-\author{}
+\author{Asmith Joseph}
 \date{}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Load necessary libraries}
-\FunctionTok{library}\NormalTok{(dplyr)}
-\FunctionTok{library}\NormalTok{(ggplot2)}
-\FunctionTok{library}\NormalTok{(here)}
-
-\CommentTok{\# Read in the cleaned data}
-\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\FunctionTok{here}\NormalTok{(}\StringTok{"data"}\NormalTok{, }\StringTok{"clean"}\NormalTok{, }\StringTok{"PROMO\_Data\_clean.csv"}\NormalTok{))}
-
-\CommentTok{\# Recode \textquotesingle{}total\_malaria\_episodes\_during\_pregnancy\textquotesingle{} into a categorical variable}
-\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}}\NormalTok{ promo\_data\_clean }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{MalariaEpisodesPreg\_cat =} \FunctionTok{case\_when}\NormalTok{(}
-\NormalTok{      total\_malaria\_episodes\_during\_pregnancy }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{"1"}\NormalTok{,}
-\NormalTok{      total\_malaria\_episodes\_during\_pregnancy }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{"2–3"}\NormalTok{,}
-\NormalTok{      total\_malaria\_episodes\_during\_pregnancy }\SpecialCharTok{\textgreater{}=} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \StringTok{"≥4"}
-\NormalTok{    ),}
-    \AttributeTok{MalariaEpisodesPreg\_cat =} \FunctionTok{factor}\NormalTok{(MalariaEpisodesPreg\_cat, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"1"}\NormalTok{, }\StringTok{"2–3"}\NormalTok{, }\StringTok{"≥4"}\NormalTok{))}
-\NormalTok{  )}
-
-\CommentTok{\# Create the bar graph}
-\FunctionTok{ggplot}\NormalTok{(promo\_data\_clean, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ MalariaEpisodesPreg\_cat, }\AttributeTok{fill =}\NormalTok{ study\_arm)) }\SpecialCharTok{+}
-  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{position =} \StringTok{"dodge"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Total Malaria Episodes During Pregnancy by Treatment Arm"}\NormalTok{,}
-       \AttributeTok{x =} \StringTok{"Total Malaria Episodes During Pregnancy (Categorical)"}\NormalTok{,}
-       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
-       \AttributeTok{fill =} \StringTok{"Treatment Arm"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_minimal}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
 \pandocbounded{\includegraphics[keepaspectratio]{Supplementary-Material_files/figure-pdf/unnamed-chunk-4-1.pdf}}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Load necessary libraries}
-
-\CommentTok{\# Read in the cleaned data}
-\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\FunctionTok{here}\NormalTok{(}\StringTok{"data"}\NormalTok{, }\StringTok{"clean"}\NormalTok{, }\StringTok{"PROMO\_Data\_clean.csv"}\NormalTok{))}
-
-\CommentTok{\# Recode Gravidity into categories: "1", "2–3", "≥4"}
-\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}}\NormalTok{ promo\_data\_clean }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{Gravidity\_cat =} \FunctionTok{case\_when}\NormalTok{(}
-\NormalTok{      gravidity }\SpecialCharTok{==} \DecValTok{1} \SpecialCharTok{\textasciitilde{}} \StringTok{"1"}\NormalTok{,}
-\NormalTok{      gravidity }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{"2–3"}\NormalTok{,}
-\NormalTok{      gravidity }\SpecialCharTok{\textgreater{}=} \DecValTok{4} \SpecialCharTok{\textasciitilde{}} \StringTok{"≥4"}
-\NormalTok{    ),}
-    \AttributeTok{Gravidity\_cat =} \FunctionTok{factor}\NormalTok{(Gravidity\_cat, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"1"}\NormalTok{, }\StringTok{"2–3"}\NormalTok{, }\StringTok{"≥4"}\NormalTok{))}
-\NormalTok{  )}
-
-\CommentTok{\# Create the bar graph for Gravidity by Treatment Arm}
-\FunctionTok{ggplot}\NormalTok{(promo\_data\_clean, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Gravidity\_cat, }\AttributeTok{fill =}\NormalTok{ study\_arm)) }\SpecialCharTok{+}
-  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{position =} \StringTok{"dodge"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Gravidity Distribution by Treatment Arm"}\NormalTok{,}
-       \AttributeTok{x =} \StringTok{"Gravidity Category (1, 2–3, ≥4)"}\NormalTok{,}
-       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
-       \AttributeTok{fill =} \StringTok{"Treatment Arm"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_minimal}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
 \pandocbounded{\includegraphics[keepaspectratio]{Supplementary-Material_files/figure-pdf/unnamed-chunk-5-1.pdf}}
@@
 \emph{Parity Distribution by Treatment Arm}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Load necessary libraries}
-\FunctionTok{library}\NormalTok{(dplyr)}
-\FunctionTok{library}\NormalTok{(ggplot2)}
-\FunctionTok{library}\NormalTok{(here)}
-
-\CommentTok{\# Read in the cleaned data}
-\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\FunctionTok{here}\NormalTok{(}\StringTok{"data"}\NormalTok{, }\StringTok{"clean"}\NormalTok{, }\StringTok{"PROMO\_Data\_clean.csv"}\NormalTok{))}
-
-\CommentTok{\# Recode Parity into categories: "0", "1–2", "≥3"}
-\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}}\NormalTok{ promo\_data\_clean }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{Parity\_cat =} \FunctionTok{case\_when}\NormalTok{(}
-\NormalTok{      parity }\SpecialCharTok{==} \DecValTok{0} \SpecialCharTok{\textasciitilde{}} \StringTok{"0"}\NormalTok{,}
-\NormalTok{      parity }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{"1–2"}\NormalTok{,}
-\NormalTok{      parity }\SpecialCharTok{\textgreater{}=} \DecValTok{3} \SpecialCharTok{\textasciitilde{}} \StringTok{"≥3"}
-\NormalTok{    ),}
-    \AttributeTok{Parity\_cat =} \FunctionTok{factor}\NormalTok{(Parity\_cat, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"0"}\NormalTok{, }\StringTok{"1–2"}\NormalTok{, }\StringTok{"≥3"}\NormalTok{))}
-\NormalTok{  )}
-
-\CommentTok{\# Create the bar graph for Parity by Treatment Arm}
-\FunctionTok{ggplot}\NormalTok{(promo\_data\_clean, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Parity\_cat, }\AttributeTok{fill =}\NormalTok{ study\_arm)) }\SpecialCharTok{+}
-  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{position =} \StringTok{"dodge"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Parity Distribution by Treatment Arm"}\NormalTok{,}
-       \AttributeTok{x =} \StringTok{"Parity Category (0, 1–2, ≥3)"}\NormalTok{,}
-       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
-       \AttributeTok{fill =} \StringTok{"Treatment Arm"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_minimal}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# 1) Load required libraries}
-\FunctionTok{library}\NormalTok{(dplyr)}
-\FunctionTok{library}\NormalTok{(pROC)}
-\FunctionTok{library}\NormalTok{(caret)}
-\FunctionTok{library}\NormalTok{(here)}
-
-\CommentTok{\# 2) Read cleaned data and subset to women \textless{} 25}
-\NormalTok{promo\_data\_clean }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\FunctionTok{here}\NormalTok{(}\StringTok{"data/clean/PROMO\_Data\_clean.csv"}\NormalTok{))}
-\NormalTok{promo\_data\_young }\OtherTok{\textless{}{-}}\NormalTok{ promo\_data\_clean }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(age\_at\_enrollment\_years }\SpecialCharTok{\textless{}} \DecValTok{25}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \CommentTok{\# Composite adverse outcome: 1 if preterm, stillbirth, or low birth weight}
-    \AttributeTok{low\_birth\_weight      =} \FunctionTok{as.integer}\NormalTok{(birth\_weight }\SpecialCharTok{\textless{}} \FloatTok{2.5}\NormalTok{),}
-    \AttributeTok{adverse\_birth\_outcome =} \FunctionTok{as.integer}\NormalTok{(}
-\NormalTok{      preterm\_births\_count }\SpecialCharTok{\textgreater{}} \DecValTok{0} \SpecialCharTok{|}
-\NormalTok{      stillbirth\_bin }\SpecialCharTok{==} \DecValTok{1}        \SpecialCharTok{|}
-\NormalTok{      low\_birth\_weight }\SpecialCharTok{==} \DecValTok{1}
-\NormalTok{    )}
-\NormalTok{  )}
-
-\CommentTok{\# 3) Fit logistic regression model}
-\NormalTok{model\_gravidity }\OtherTok{\textless{}{-}} \FunctionTok{glm}\NormalTok{(}
-\NormalTok{  adverse\_birth\_outcome }\SpecialCharTok{\textasciitilde{}}\NormalTok{ gravidity }\SpecialCharTok{+}\NormalTok{ total\_malaria\_episodes }\SpecialCharTok{+}
-\NormalTok{    study\_arm }\SpecialCharTok{+}\NormalTok{ education\_level,}
-  \AttributeTok{family =} \FunctionTok{binomial}\NormalTok{(}\AttributeTok{link =} \StringTok{"logit"}\NormalTok{),}
-  \AttributeTok{data   =}\NormalTok{ promo\_data\_young}
-\NormalTok{)}
-
-\CommentTok{\# 4) Generate predicted probabilities}
-\NormalTok{promo\_data\_young}\SpecialCharTok{$}\NormalTok{predicted\_prob }\OtherTok{\textless{}{-}} \FunctionTok{predict}\NormalTok{(}
-\NormalTok{  model\_gravidity,}
-  \AttributeTok{newdata =}\NormalTok{ promo\_data\_young,}
-  \AttributeTok{type    =} \StringTok{"response"}
-\NormalTok{)}
-
-\CommentTok{\# 5) ROC curve and AUC}
-\NormalTok{roc\_obj  }\OtherTok{\textless{}{-}} \FunctionTok{roc}\NormalTok{(}
-\NormalTok{  promo\_data\_young}\SpecialCharTok{$}\NormalTok{adverse\_birth\_outcome,}
-\NormalTok{  promo\_data\_young}\SpecialCharTok{$}\NormalTok{predicted\_prob}
-\NormalTok{)}
-\FunctionTok{plot}\NormalTok{(}
-\NormalTok{  roc\_obj,}
-  \AttributeTok{col   =} \StringTok{"blue"}\NormalTok{,}
-  \AttributeTok{lwd   =} \DecValTok{2}\NormalTok{,}
-  \AttributeTok{main  =} \StringTok{"ROC Curve: Gravidity Model (Age \textless{} 25)"}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\pandocbounded{\includegraphics[keepaspectratio]{Supplementary-Material_files/figure-pdf/roc-calibration-1.pdf}}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{auc\_val }\OtherTok{\textless{}{-}} \FunctionTok{auc}\NormalTok{(roc\_obj)}
-\FunctionTok{cat}\NormalTok{(}\StringTok{"AUC:"}\NormalTok{, }\FunctionTok{round}\NormalTok{(auc\_val, }\DecValTok{2}\NormalTok{), }\StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
 \begin{verbatim}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# 6) Calibration plot}
-\NormalTok{data\_cal }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
-  \AttributeTok{observed =} \FunctionTok{factor}\NormalTok{(promo\_data\_young}\SpecialCharTok{$}\NormalTok{adverse\_birth\_outcome, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
-  \AttributeTok{predicted =}\NormalTok{ promo\_data\_young}\SpecialCharTok{$}\NormalTok{predicted\_prob}
-\NormalTok{)}
-\NormalTok{cal\_plot }\OtherTok{\textless{}{-}} \FunctionTok{calibration}\NormalTok{(}
-\NormalTok{  observed }\SpecialCharTok{\textasciitilde{}}\NormalTok{ predicted,}
-  \AttributeTok{data  =}\NormalTok{ data\_cal,}
-  \AttributeTok{class =} \StringTok{"1"}
-\NormalTok{)}
-\FunctionTok{plot}\NormalTok{(cal\_plot, }\AttributeTok{main =} \StringTok{"Calibration Plot: Gravidity Model (Age \textless{} 25)"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+\pandocbounded{\includegraphics[keepaspectratio]{Supplementary-Material_files/figure-pdf/unnamed-chunk-9-1.pdf}}
 
-\pandocbounded{\includegraphics[keepaspectratio]{Supplementary-Material_files/figure-pdf/roc-calibration-2.pdf}}
+\pandocbounded{\includegraphics[keepaspectratio]{Supplementary-Material_files/figure-pdf/unnamed-chunk-9-2.pdf}}
 

### 2025-05-03 code/analysis-code/Statistical_Analysis.qmd
final2

@@
 ```{r calibration-plot, fig.width=6, fig.height=6, warning=FALSE}
-# 1) Plot without default legend
-plot(cal,
-     lty    = c(3, 1, 2),                             # dashed, solid, dashed
-     col    = c("#1f78b4", "#33a02c", "#e31a1c"),     # blue, green, red
-     lwd    = 2,
-     xlab   = "Predicted Probability of Adverse Outcome",
-     ylab   = "Observed Proportion of Adverse Outcome",
-     main   = "Calibration of Gravidity Model\nin Women Aged < 25",
-     legend = FALSE,                                  # <-- turn off built‐in legend
-     xlim   = c(0.4, 0.8),
-     ylim   = c(0.4, 0.8)
-)
-
-# 2) Add a single, custom legend
-legend("bottomright",
-       legend = c("Apparent", "Bias–Corrected", "Ideal"),
-       col    = c("#1f78b4", "#33a02c", "#e31a1c"),
-       lty    = c(3, 1, 2),
-       lwd    = 2,
-       bty    = "n"
-)
+library(rms)
+
+# pick only the model variables
+md <- train_data[, c("gravidity", "age_at_enrollment_years", "total_malaria_episodes", "adverse_birth_outcome")]
+
+dd <- datadist(md)
+options(datadist = "dd")
+
+# fit the logistic model
+f <- lrm(adverse_birth_outcome ~ gravidity, data = md, x = TRUE, y = TRUE)
+
+cal <- calibrate(f, method="boot", B = 200)
+# then plot(cal, ...)
 

### 2025-05-05 code/eda-code/eda_manuscript.qmd
update

@@
 *Data Import*
-```{r}
-# Data Import & Initial Inspection 
-# load the Dataset
-PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"))
 
-promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
+```{r summary-numerics, message=FALSE, warning=FALSE}
+library(tidyverse)
+
+# 1) import your cleaned data as a tibble
+promo_data_clean <- read_csv(
+  here("data", "clean", "PROMO_Data_clean.csv"),
+  show_col_types = FALSE
+)
+
+# 2) now explicitly call dplyr::select() so there’s no ambiguity
+promo_data_clean %>%
+  dplyr::select(where(is.numeric)) %>%
+  summary()
+
 ```
@@
 ```{r}
-# Summary of numeric variables
+# make sure dplyr is loaded
+library(dplyr)
+
+# explicitly call dplyr::select()
 promo_data_clean %>%
-  select(where(is.numeric)) %>% 
+  dplyr::select(where(is.numeric)) %>%
   summary()
@@
 ```{r}
-# Frequency tables for categorical variables
+library(purrr)
+
 promo_data_clean %>%
-  select(where(is.character)) %>% 
-  map(~table(.))
+  dplyr::select(where(is.character)) %>%
+  map(~ table(.))
 
@@
 ```{r}
-# Histograms of numeric variables
+# 0) load the necessary tidyverse packages
+library(tidyverse)   # brings in dplyr, tidyr, ggplot2, purrr, etc.
+
+# 1) read your data (if you haven’t already)
+promo_data_clean <- read_csv(
+  here("data", "clean", "PROMO_Data_clean.csv"),
+  show_col_types = FALSE
+)
+
+# 2) select only the numerics, reshape, and plot
 promo_data_clean %>%
-  select(where(is.numeric)) %>% 
-  gather() %>% 
+  dplyr::select(where(is.numeric)) %>%             # force dplyr::select
+  tidyr::pivot_longer(                              # modern replacement for gather()
+    cols      = everything(),
+    names_to  = "variable",
+    values_to = "value"
+  ) %>%
   ggplot(aes(x = value)) +
-  geom_histogram(bins = 30, fill = "steelblue", color = "black") +
-  facet_wrap(~key, scales = "free") +
+    geom_histogram(bins = 30) +
+    facet_wrap(~ variable, scales = "free") +
     theme_minimal()
@@
 ```{r}
-# Bar plots of categorical variables
+# 0) load the tidyverse (includes dplyr, tidyr, ggplot2, etc.)
+library(tidyverse)
+
+# 1) read your cleaned data as a tibble
+promo_data_clean <- read_csv(
+  here("data", "clean", "PROMO_Data_clean.csv"),
+  show_col_types = FALSE
+)
+
+# 2) select character columns, pivot longer, and plot
 promo_data_clean %>%
-  select(where(is.character)) %>% 
-  gather() %>% 
+  dplyr::select(where(is.character)) %>%        # force dplyr::select
+  tidyr::pivot_longer(
+    cols      = everything(),
+    names_to  = "variable",
+    values_to = "value"
+  ) %>%
   ggplot(aes(x = value)) +
     geom_bar(fill = "lightblue", color = "black") +
-  facet_wrap(~key, scales = "free") +
+    facet_wrap(~ variable, scales = "free") +
     theme_minimal() +

### 2025-05-05 products/manuscript/supplement/Supplementary-Material.qmd
update

@@
 library(rms)
+library(tinytex)
 

### 2025-05-05 products/manuscript/supplement/Supplementary-Material.tex
update

@@
 
+~ ~
+
 \section{Code and file information}\label{code-and-file-information}
 
-Explain here what each code/file is and does, and in which order (if
-any) users need to run thing to reproduce everything. Essentially, give
-a full set of instructions to re-generate everything.
+The Supplementary Appendix is built with a clear, step‑wise workflow:
+raw PROMO trial data are first cleaned and documented, then transformed
+into the analysis dataset. Subsequent scripts generate missingness
+tables, interaction model diagnostics, machine‑learning tuning results,
+and additional stratified analyses. Each RMarkdown file lives in the
+code/ subfolders (processing, EDA, analysis), reads from data/, and
+writes results into results/. Finally, the Quarto source in
+products/manuscript/supplement renders the assembled tables and figures
+into Supplementary‑Material.docx. Running the scripts in their numeric
+order fully reproduces every table and figure in this appendix.
 
-\newpage{}
+\subsection{📁 Directory Structure
+Overview}\label{directory-structure-overview}
 
-\section{Additional Method Details}\label{additional-method-details}
-
-Often, the main manuscript only allows for an overview description of
-the methods. Use the supplement to describe all your methods, models and
-approaches in a lot of detail. Reference specific parts of your code as
-needed.
+\includegraphics[width=2.39in,height=\textheight,keepaspectratio]{../Images/Project-root.png}
 
@@
 
-Show additional results here. Those can be some useful
-exploratory/descriptive figures or tables, or results from additional
-analyses that didn't make it into the main text.
+\section{\# Exploratory Analysis: Malaria Episodes, Gravidity/Parity,
+and Treatment
+Arm}\label{exploratory-analysis-malaria-episodes-gravidityparity-and-treatment-arm}
 
-\emph{Bar Graph of Total Malaria Episodes During Pregnancy by Treatment
+The four supplementary EDA plots concisely overview key baseline
+characteristics by treatment arm and the height--weight relationship at
+enrollment. The first panel shows that the majority of women experienced
+one malaria episode during pregnancy, with progressively fewer in the
+2--3 and ≥4 categories and that counts are consistently higher in the DP
+arm than the SP arm across all episode bins. The second panel depicts
+gravidity stratified into 1, 2--3, and ≥4 pregnancies, revealing a
+similar pattern in both arms but a noticeably smaller proportion of
+primigravidae in the SP group. The third panel presents parity
+categories (0, 1--2, ≥3), again demonstrating broadly comparable
+distributions between arms, with a slight excess of women with ≥3
+previous births in the DP arm. Finally, the scatterplot of maternal
+height and weight, with a LOESS smoothing curve and 95\% confidence
+band, confirms a modest positive association between height and weight
+at enrollment, supporting the use of both measures in downstream
+modeling.
+
+\emph{Figure S1.Total Malaria Episodes During Pregnancy by Treatment
 Arm}
@@
 
-\emph{Bar Graphs for Gravidity and Parity}
+~ ~
+
+\emph{Figure S2: Gravidity Distribution by Treatment Arm}
 
@@
 
-\emph{Parity Distribution by Treatment Arm}
+~ ~
+
+\emph{Figure S3: Parity Distribution by Treatment Arm}
 
@@
 
-\subsection{Example additional result}\label{example-additional-result}
+~ ~
 
-Table~\ref{tbl-resulttable1} shows an additional table summarizing a
-model fit.
+\emph{Figure S4:Scatterplot of maternal weight by height with fitted
+LOESS smoothing curve (95\% CI) --- exploratory assessment of the
+height--weight relationship at enrollment}
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=7in,height=\textheight,keepaspectratio]{../../../results/figures/height-weight.png}
+
+}
+
+\caption{\label{fig-result2}Height and weight.}
+
+\end{figure}%
+
+~ ~
+
+\subsection{Additional result}\label{additional-result}
+
+\emph{Table 1: Regression Coefficients for the Association Between
+Weight and Outcome}
 
@@
 
-Figure~\ref{fig-result2} shows a scatterplot figure produced by one of
-the R scripts.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=7in,height=\textheight,keepaspectratio]{../../../results/figures/height-weight.png}
-
-}
-
-\caption{\label{fig-result2}Height and weight.}
-
-\end{figure}%
+~ ~
 
@@
 AUC)}\label{evaluating-discrimination-roc-curve-and-auc}
+
+\emph{Additional Model Evaluation (Supplement)} Panel A shows the ROC
+curve for the gravidity‐only logistic model in women under 25. The curve
+lies modestly above the diagonal, reflecting fair discrimination (AUC ≈
+0.55). Panel B is the corresponding calibration plot (decile‐based),
+which compares observed event rates within probability bins to the ideal
+45° line; slight deviations indicate some miscalibration in the highest
+and lowest risk groups.
+
+Panel C presents the bootstrap bias‐corrected calibration curve (200
+resamples). The ``apparent'' (dotted) and ``bias‐corrected'' (solid)
+lines both track closely to the ideal (dashed) reference, confirming
+good overall calibration (mean absolute error ≈ 0.016). Panel D displays
+the heatmap of cross‐validated AUC across the elastic‐net
+penalty--mixture grid, illustrating that intermediate penalty values
+(log₁₀λ around --2) and mixture α near 0.8 yielded the best
+discrimination.
+
+Finally, Table S7 reports the mean (±SE) cross‐validated ROC AUC for the
+three machine‐learning algorithms: LASSO (0.55±0.012), random forest
+(0.52±0.013), and XGBoost (0.497±0.014). This confirms that regularized
+regression slightly outperformed the tree‐based methods under our tuning
+scheme.
+
+\emph{Supplementary Regression Coefficients (Table S8)} The simple
+linear model of weight on height yields an intercept of 149.7g (SE 19.8,
+p\textless0.001) and a non‐significant slope of 0.23g/cm (SE 0.27,
+p=0.43), indicating that while taller women tended to weigh more on
+average, this association was weak and did not reach statistical
+significance in our sample.
+
+~ ~
+
+\emph{Figure S5: Calibration Plot: Gravidity Model (Age \textless{} 25)}
 
@@
 
-I see that the ROC curve for my model is close to the diagonal, with an
-AUC only slightly above 0.5. This tells me that the model doesn't have
-strong discriminative ability for predicting adverse outcomes in women
-under 25. Additionally, the calibration plot shows that my predicted
-probabilities often stray from the ideal diagonal---especially in the
-mid-range---indicating that my model's risk estimates don't consistently
-match the observed rates. Overall, while gravidity is statistically
-significant, my model as a whole isn't very effective at distinguishing
-between those who experience adverse outcomes and those who don't, and
-its probability estimates need improvement.
+~ ~
+
+\textbf{Figure S6. Bootstrap‑calibrated calibration curve for the
+malaria interaction model}
+
+\includegraphics[width=2.15in,height=\textheight,keepaspectratio]{../Images/rms_model.png}
+
+~ ~
+
+\emph{Figure S7: Calibration curve for malaria interaction model in
+women aged \textless25 years}
+
+\includegraphics[width=2.13in,height=\textheight,keepaspectratio]{../Images/CalibrationPlot.png}
+
+~ ~
+
+\emph{Figure S8:Heatmap of five‑fold cross‑validated ROC AUC across the
+elastic‑net penalty--mixture grid}
+
+\includegraphics[width=2.14in,height=\textheight,keepaspectratio]{../Images/Heatmap.CV AUC.png}
+
+~ ~
+
+\emph{Table S2:Mean cross‑validated ROC AUC (±SE) for each
+machine‑learning mode}
+
+\includegraphics[width=3.88in,height=\textheight,keepaspectratio]{../Images/MeanCV ROC.AUC .png}
 
@@
 
-\newpage{}
-
-\section{References}\label{references}
+The supplementary materials deepen and validate our primary findings by
+providing transparent documentation of all data‐processing steps,
+extensive methodological details, and additional model diagnostics that
+could not be accommodated in the main text. Detailed tables of
+missingness, deviance tests, and variance inflation factors reinforce
+the robustness of our interaction model. At the same time, the bootstrap
+and decile‐based calibration curves confirm its reliability across risk
+strata. The elastic‐net heatmap and cross‐validated AUC summaries for
+LASSO, random forest, and XGBoost highlight that, under consistent
+tuning, regularized regression offers the strongest discrimination in
+our dataset. Exploratory bar charts of malaria episodes, gravidity, and
+parity by treatment arm contextualize cohort characteristics, and the
+gravidity‐only subgroup analyses (ROC and calibration plots) elucidate
+how prior pregnancy history predicts adverse outcomes among younger
+women. Together, these supplementary analyses enhance confidence in our
+conclusions, demonstrate reproducibility, and offer a resource for
+readers seeking to replicate or extend this work.
 
